### 2019-05-16 positiveRK/ideas.tex
Move work into a subdirectory.

\documentclass{article}

\usepackage{amsmath}

\title{Ideas for using Runge--Kutta methods to preserve positivity and similar properties}

\begin{document}
\maketitle

\section{Basic relaxation approach}

\section{Linear programming approach}
Given a Runge--Kutta method with coefficients $(A,b)$, after we have computed
the stages
$$
    y_j = u^n + h \sum_i a_{ij} f(y_i),
$$
the idea is to adaptively choose the weight $b_j$ in order to enforce positivity.
This leads to a linear program with the linear inequality constraints
$$
u^{n+1}_i = u^n_i + h \sum_j b_j f(y_j) \ge 0
$$
and the linear equality constraints that come from the order conditions.  The idea is
to enforce as many of these as possible.  For first order:
\begin{align}
\sum_j b_j = 1.
\end{align}
For second order:
\begin{align}
\sum_j c_j b_j = 1/2.
\end{align}
For third order:
\begin{align}
\sum_j c_j^2 b_j & = 1/3 \\
\sum_j \tau_{2j} b_j & = 0.
\end{align}
For fourth order:
\begin{align}
\sum_j c_j^3 b_j & = 1/4 \\
\sum_j \tau_{2j} c_j b_j & = 0 \\
\tau_{2}^T A^T b & = 0 \\
\tau_{3}^T b & = 0.
\end{align}
For fifth order:
\begin{align}
\tau_2^T A^T C b & = 0 \\
\tau_2^T C A^T b & = 0 \\
\tau_2^T (A^T)^2 b & = 0 \\
\tau_2^T C^2 b & = 0 \\
\tau_2^2 b & = 0 \\
\tau_3^T A^T b & = 0 \\
\tau_3^T C b & = 0 \\
(\tau_2^T)^2 b & = 0 \\
\tau_4^T b & = 0.
\end{align}
Here vector powers mean componentwise multiplication while
matrix powers mean matrix multiplication.  The matrix $C$ is the
diagonal matrix of the abscissas $c_i$ and
$\tau_k$ are the stage order residual vectors:
$$
    \tau_k - \frac{1}{k!} c^k - \frac{1}{(k-1)!}Ac^{k-1}.
$$
Since they only appear in equations where the right-hand-side is zero,
it's okay to multiply by $k!$ and use instead the vectors
$$
    \hat{\tau}_k - c^k - kAc^{k-1}.
$$

\end{document}

### 2019-05-16 polygonal_hydr_jump/review.tex
Add first iteration on literature review related to hydraulic jump

\documentclass[english,11pt]{article}

\usepackage{amsthm}
\usepackage[latin9]{inputenc}
\usepackage{babel}
\usepackage[hmargin=0.9in]{geometry}

%\usepackage{subfigure}
\usepackage{subfig}
\usepackage{graphicx}
\usepackage[colorlinks=true,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage[numbers]{natbib}
%\usepackage{cite}
\let\cite=\citet

\usepackage{float}
\usepackage{url}
\usepackage{bm}

\usepackage{amsmath}
\usepackage{amssymb,amsfonts}
\usepackage{upgreek}
\usepackage[textsize=footnotesize]{todonotes}
\usepackage{stmaryrd}
%\newcommand{\todo}[1]{{\bf \color{red} #1}\vskip 10pt}

\newtheorem{remark}{Remark}
\numberwithin{remark}{subsection}

\makeatletter
\providecommand{\tabularnewline}{\\}
\makeatother

\graphicspath{{./figures/}}

% Macros -- use them
\newcommand{\bfu}{\mathbf{u}}
\newcommand{\bfv}{\mathbf{v}}
\newcommand{\bff}{\mathbf{f}}
\newcommand{\bfn}{\mathbf{n}}
\newcommand{\bfx}{\mathbf{x}}
\newcommand{\order}{{\mathcal O}}
\newcommand{\yhat}{\hat{y}}
\newcommand{\Khat}{\hat{K}}
\newcommand{\rhohat}{\hat{\rho}}
\newcommand{\etanot}{\eta_0}
\def\avg[#1]{\left\langle #1 \right\rangle}
\def\intavg[#1]{\llbracket #1 \rrbracket}
\def\ceff{c_\text{eff}}
\def\bk{\bf k}
\newcommand{\bx}{{\bf x}}

\begin{document}
\title{Polygonal hydraulic jumps: 3D free-surface flow simulations}
%\author{
%  Manuel Quezada de Luna \thanks{KAUST CEMSE, manuel.quezada@kaust.edu.sa.} \and
%  David I. Ketcheson \thanks{KAUST CEMSE, david.ketcheson@kaust.edu.sa.}}
\maketitle

%\begin{abstract}
%\end{abstract}

%\keywords{}

\section{Literature review}

\subsection{General remarks}
\begin{itemize}
\item The hydraulic jump and different theoretical and experimental aspects of it have been studied by many
  authors. For instance, \cite{rayleigh1914theory} studied the inviscid hydraulic jump and proposed an
  estimation of the radius of the jump. 
  \cite{watson1964radial}, used momentum balanced theory and consider inviscid and viscous flows to also propose estimates
  for the radius of the jump. These studies didn't include surface tension effects. 
  Later, \cite{bush2003influence} proposed a correction to the results by Watson to include the effects of surface
  tension on the radius of the jump. Many more models and corrections have been proposed to estimate the radius of the
  jump in different regimes and using different tools.

\item According to \cite{labousse2013hydraulic}, \cite{bohr1996hydraulic} and \cite{watanabe2003integral}
  distinguised between hydraulic jumps of type I and II.
  \begin{itemize}
  \item[*] In type I there is only one vortex leading to the standard hydraulic jump.
  \item[*] In type II, a secondary vortex forms leading to the polygonal jump.
    Within type II, type IIa and IIb refer to polygonal jumps with one and two jumps respectively.
  \end{itemize}

\item The most up-to-date model for type II hydraulic jumps is given by \cite{martens2012model}.
\end{itemize}

\subsection{About the observation of instability}
\begin{itemize}
\item \cite{craik1981circular} was one of the first to observe the instability that leads to type II jumps.
  However, he used water and didn't observe the creation of polygons.
  He triggered the instability by increasing the downstream depth.

\item \cite{ishigai1977heat} observed and controlled the formation of the instability via the Froude number upstream.

\item According to \cite{liu1993hydraulic}, \cite{errico1986study} observed the formation of the instability
  by increasing the volume flow.

\item In \cite{liu1993hydraulic} the authors present more observations of the instability and remark that
  surface tension plays an important role. In particular, the authors describe the surface tension as a
  mechanism to stabilize the turbulent behavior of the instability. 
  They conclude that the height of the jump and the radius of the roller are critical for surface tension to act as
  a stabilization process. For exmple, in a hydraulic jump in a channel, the height of the jump and the radius of the roller
  are very large making the surface tension effects negligible.

\item \cite{kasimov2008stationary} proposed a theory based on a shallow water model to determine the radius of the jump
  based on far-field and jump conditions including surface tension effects. The author showed that the steady circular
  jump can't exist if the surface tension is larger than a critical value.
  The author also remarked that the critical surface tension decreases as the viscosity increases.
  Furthermore, he suggests this might explain why the polygonal jumps appear only in highly viscous fluids. 
\end{itemize}

\subsection{Experimental observation of hydraulic jump of type IIa}
\begin{itemize}
\item In \cite{ellegaard1998creating}, the authors present and explain the phenomenon.
  Given a fluid (ethylene glycol) they can control the shape of the jump by adjusting the height of the external wall
  (which increases the height of the jump).
  At some point (as the height of the external wall increases), the polygonal shape turns into a circular shape again and eventually the jumps closes.
  If the height of the external wall is too small then a circular hydraulic jump (of type I) is formed.
  The authors provide the simple dimensionless parameter $\phi=\nu Q/g(h_\text{ext}^2-h_\text{int}^2)$, where $\nu$ is the viscosity
  and $Q$ is the flow rate. They observe that as $\phi$ increases, the number of corners increase.
  %
  The authors present details about the derivation of $\phi$ in \cite{ellegaard1999cover} and details about the experiment in
  \cite{ellegaard1996experimental}.

\item In \cite{bohr1996hydraulic}, the authors explain further the experimental results and start to attempt a numerical simulation.
  The model to be numerically solved is based on the steady state incompressible 2D-NS.
  The free surface is modeled by a linear and rigid boundary close to one observed in experiments.
  They neglect the effects of surface tension in this model. Even with this simplified model, the authors are able
  to produce the vortex in the type II jumps.
  Here it is important to remark that even without surface tension they observed (numerically) the formation of the
  roller (the second vortex in the type II jumps). Therefore, the apperance of the roller might not depend entirely upon the
  surface tension. 
  The authors also work out in this paper a simplified model to start to develop a theory.

\item In \cite{bush2006experimental}, experiments were performed to produce the polygonal hydraulic
  (type II) jump with many different shapes.
  The goal of the authors was to identify important dimensionless parameters that characterize type IIa jumps. They
  also present experimental results of type IIb jumps.
  %
  In this work, the authors demonstrated experimentally the importance of surface tension by adding a surfactant to the flow.
  A type IIa jump transformed to a type I when a surfactant was added (see figure 6).
  %
  In section 6, the authors explain a (potential) reason for which surface tension is important to create type II jumps.
  The idea is that with high surface tension, the instability (a pinch-off) is triggered by a lower surface energy configuration
  given by the polygonal shape.
  This process is characterized by the Ohnesorge number $Oh=\sigma R/(\mu\nu)$
  At high $Oh$ the pinch-off is resisted by fluid intertia and at low $Oh$ the pinch off is resisted by viscosity.
  %
  This work gives some results related to Ohnesorge number and frequency in the jump.
  This and other quantitative results in this work are something we might be able to validate.

\item \cite{andersen2010separation} remarks about the importance of the separation to create the roller in the type II jumps.

\end{itemize}

\subsection{About hydraulic jumps of type IIb}
\begin{itemize}
\item \cite{craik1981circular} observed the second jump using water right before the jump becomes unstable.
  It is important to remark that in this work the authors used water. It is believed that the viscosity and surface tension
  of water is not strong enough to create the polygonal shapes. Indeed, the authors in this work didn't observe those shapes.
  However, they observed the instability and, moreover, the formation of the second jump in the type IIb hydraulic jump
  (see figure 2b).

\item \cite{bush2006experimental} has a good number of experiments with type IIb jumps with different shapes.

\item In \cite{liu1993hydraulic} the second jump was also observed (see figure 5).

\end{itemize}

\subsection{Other types of circular hydraulic jumps}
\begin{itemize}
\item{\bf Crown shape hydraulic jumps.} See figure 14 in \cite{bush2006experimental}.
  I have simulations of the SWEs with viscosity that seem to show a similar behavior.

\item
  {\bf Hydraulic bump.} In \cite{labousse2013hydraulic}, the authors introduce the idea of hydraulic bump,
  which happens below the hydraulic jump threshold. In this case the jet impinges a deep steady water and only 
  weakly perturbes the free surface creating a circular bump that can evolve into polygonal shapes
  (presumably due to the same reasons than with hydraulic jumps).

\end{itemize}

\subsection{Numerical simulations}

\begin{itemize}
\item In \cite{bohr1996hydraulic}, the authors assumed a fixed profile for the free surface based on experimental data.
  Then they solved the steady state 2D-NS equations and observed the two vortices.

\item In \cite{yokoi2002mechanism}, numerical simulations of the compressible NS equations are performed.
  They use axisymmetrical simulations (i.e. 2D) and consider type I and type II jumps.
  They concluded that the transition between type I to II jumps is heavily affected by surface tension.
  %
  It is important to remark that the authors marked a jump as type II if the roller is observed.
  Off course with 2D simulations it is not possible to observe the polygonal shapes. 
  %
  The numerical methods are based on a (non conservative) level-set approach.
  The results in this paper have been presented with more physics oriented details in
  \cite{yokoi2000relationships,yokoi1999numerical}.
  At this point (in time) the authors claimed that there is no model for polygonal hydraulic jumps.

\item In \cite{passandideh2011numerical} the authors solved the incompressible Navier-Stokes equations for two immiscible
  viscous flows coupled with a volume-of-fluid approach to track the free surface.
  The authors considered type I and II jumps.
  %
  It appears to me that only two-dimensional simulations are performed; however, the authors extruded
  the solution in the angular direction to obtain a three-dimensional visualization. Off course this is
  only valid if the jump is axisymmetric. Therefore, the authors didn't obtain polygonal jumps.
\end{itemize}

\section{Objectives}
\begin{itemize}
\item Simulate for the first time the 3D hydraulic jump of type II to observe the polygonal shapes.
\item If possible try to obtain other type of jumps like type IIb, crown hydraulic jumps and hydraulic bumps (of type I and II). 
\item After proper validation against experimental data, extend the experiments to other regimes that might be difficult
  to achieve in a laboratory. For example, highly viscous but without surface tension or vice versa,
  see the effect of slip BCs on the bottom of the tank (is friction necessary to have polygonal shapes?).
\item Confirm the role and importance on surface tension in the creation the polygonal shapes.
\item Test Kasimov's hypothesis that higher viscous effects might reduce the importance of surface tension in the
  creation of the polygonal jump.
\item As byproduct simulate (for the first time) the 3D hydraulic jump of type I.
\end{itemize}

\section{Two-dimensional simulations}

\section{Three-dimensional simulations}

%\bibliographystyle{plain}
\bibliographystyle{unsrtnat}
%\bibliographystyle{abbrvnat}
\bibliography{ref}

\end{document}

### 2019-05-16 polygonal_hydr_jump/clean.sh

@@ -1,0 +1,1 @@
+rm *aux *bbl *blg *out *pdf *~ *log

### 2019-06-01 MRK4-fluxLimiting/MRK4.py
make own folder for modified RK4 with limiting on last stage and other changes

@@ -1,0 +1,108 @@
+import numpy as np
+from matplotlib import pyplot as plt
+
+# These functions are designed for a 1D periodic domain from [0,1] #
+
+def compute_matrices(U,N):
+    h=1/(N-1)
+    D1=np.diag(np.ones(N-1),-1)-np.diag(np.ones(N-1),1)
+    D1[0,-1]=1.    
+    D1[-1,0]=-1.
+    D1*=U/2/h
+
+    # MPP artificial viscosity operator #
+    DL=D1*0
+    for i in range(N):
+        for j in np.arange(i+1,N):
+            DL[i,j]=np.max([-D1[i,j],0,-D1[j,i]])
+            DL[j,i]=DL[i,j]
+        DL[i,i]=-np.sum(DL[i,:])
+        
+    return D1,DL
+#
+
+def sdisc(u,D1,DL,compute_flux_matrices=True):
+    fG = np.dot(D1+DL,u)
+    fB = np.dot(-DL,u)
+    
+    fijG = D1*0
+    fijB = D1*0
+    
+    if compute_flux_matrices:
+        for i in range(len(D1)):
+            jVector = [len(D1)-1,i+1] if i==0 else ([0,i-1] if i==len(D1)-1 else [i-1,i+1])
+            for j in jVector:
+                fijG[i,j] = D1[i,j]*u[j]-D1[j,i]*u[i] + DL[i,j]*(u[j]-u[i])
+                fijB[i,j] = -DL[i,j]*(u[j]-u[i])                
+            #
+        #
+    #
+    return fG+fB,fG,fB,fijG,fijB
+#
+
+def limited_non_conservative_flux(dt,un,fG,fB,global_bounds=False,single_gamma=False):
+    gamma=0*fB
+    for i in range(len(fB)):
+        # compute bounds 
+        if global_bounds:
+            umax=1
+            umin=0
+        else:
+            jVector = [len(un)-1,i,i+1] if i==0 else ([i-1,i,0] if i==len(un)-1 else [i-1,i,i+1])
+            for j in jVector:
+                umax=max(umax,un[j])
+                umin=min(umin,un[j])
+            #
+        # compute non-conservative limiters 
+        gamma_neg = (umin-un[i]-dt*fG[i])/dt/fB[i] if (un[i]+dt*(fG[i]+fB[i])<umin and fB[i]!=0) else 1.0
[…]

### 2019-06-01 MRK4-fluxLimiting/clean.sh

@@ -1,0 +1,1 @@
+rm -r *~ *pyc *__*

### 2019-06-01 ideas.tex

\documentclass{article}

\usepackage{amsmath}
\usepackage{amssymb,amsfonts}
\usepackage[hmargin=0.9in]{geometry}

\usepackage{subfig}
\usepackage{graphicx}
\usepackage[colorlinks=true,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage{cite}

\newtheorem{remark}{Remark}
\numberwithin{remark}{subsection}

\newcommand{\bfv}{{\bf v}}
\newcommand{\bfn}{{\bf n}}
\newcommand{\bfx}{{\bf x}}

\title{Ideas for using Runge--Kutta methods to preserve positivity and similar properties}

\begin{document}
\maketitle

\section{Basic relaxation approach}

\section{Linear programming approach}
Given a Runge--Kutta method with coefficients $(A,b)$, after we have computed
the stages
$$
    y_j = u^n + h \sum_i a_{ij} f(y_i),
$$
the idea is to adaptively choose the weight $b_j$ in order to enforce positivity.
This leads to a linear program with the linear inequality constraints
$$
u^{n+1}_i = u^n_i + h \sum_j b_j f(y_j) \ge 0
$$
and the linear equality constraints that come from the order conditions.  The idea is
to enforce as many of these as possible.  For first order:
\begin{align}
\sum_j b_j = 1.
\end{align}
For second order:
\begin{align}
\sum_j c_j b_j = 1/2.
\end{align}
For third order:
\begin{align}
\sum_j c_j^2 b_j & = 1/3 \\
\sum_j \tau_{2j} b_j & = 0.
\end{align}
For fourth order:
\begin{align}
\sum_j c_j^3 b_j & = 1/4 \\
\sum_j \tau_{2j} c_j b_j & = 0 \\
\tau_{2}^T A^T b & = 0 \\
\tau_{3}^T b & = 0.
\end{align}
For fifth order:
\begin{align}
\tau_2^T A^T C b & = 0 \\
\tau_2^T C A^T b & = 0 \\
\tau_2^T (A^T)^2 b & = 0 \\
\tau_2^T C^2 b & = 0 \\
\tau_2^2 b & = 0 \\
\tau_3^T A^T b & = 0 \\
\tau_3^T C b & = 0 \\
(\tau_2^T)^2 b & = 0 \\
\tau_4^T b & = 0.
\end{align}
Here vector powers mean componentwise multiplication while
matrix powers mean matrix multiplication.  The matrix $C$ is the
diagonal matrix of the abscissas $c_i$ and
$\tau_k$ are the stage order residual vectors:
$$
    \tau_k = \frac{1}{k!} c^k - \frac{1}{(k-1)!}Ac^{k-1}.
$$
Since they only appear in equations where the right-hand-side is zero,
it's okay to multiply by $k!$ and use instead the vectors
$$
    \hat{\tau}_k = c^k - kAc^{k-1}.
$$

\clearpage
\section{Runge-Kutta methods with flux limiting}
Consider the transport equation
\begin{align}\label{transport_equation}
  \partial_t u + \bfv\cdot \nabla u = 0
\end{align}

\subsection{Forward Euler with low-order discrete upwinding}\label{sec:FE_discUpwind}

Let $m_i=\int_\Omega \phi_i dx$ be the $i$-th entry in the diagonal lumped mass matrix
and assume $m_i>0$.
Then, following \cite{kuzmin2012flux},
the full discretization with discrete upwinding is given by
\begin{align}\label{low-order}
  m_i\left(\frac{U_i^L-U_i^n}{\Delta t}\right)+\sum_j\left(K_{ij}+D_{ij}^L\right)U_j^n=0,
\end{align}
where $U^L$ is (the vector of DOFs of) the low-order solution at time $t^{n+1}$,
$U^n$ is the solution at time $t^n$ and 
\begin{align}\label{transport_and_discUpwinding_matrices}
  K_{ij} = \int_\Omega (\bfv\cdot\nabla \phi_j) \phi_i dx,
  & &
  D^L_{ij} =
  \begin{cases}
    -\max(K_{ij},0,K_{ji}), & \text{ if } i\neq j, \\
    -\sum_{j\neq i} D_{ij}^L, & \text{ if } i=j
  \end{cases}
\end{align}
are the transport and artificial dissipative matrices respectively.

\begin{remark}[About the discrete maximum principle (DMP).]
  The solution $U^L$ satisfies the following local bounds:
  \begin{align}\label{local-bounds}
    U_i^{\min}:= \min_{j\in N_i} U_j^n \leq U_i^L \leq \max_{j\in N_i} U_j^n=:U_i^{\max},
  \end{align}
  where $N_i$ is the set of DOFs that define the stencil of the discretization at the $i$-th node.
  Rewrite equation \eqref{low-order} as $U_i^L=\sum_j R_{ij}U_j^n$ with
  $R_{ij}:=\delta_{ij}-\Delta t m_i^{-1} (K_{ij}+D_{ij}^L)$, where $\delta_{ij}$ is
  the Kronecker delta function. Note that by construction
  $K_{ij}+D^L_{ij}\leq 0, \forall i\neq j\implies R_{ij}\geq 0, \forall i\neq j$.
  The diagonal entries of $R$ are positive provided $\Delta t$ is small
  enough such that $1-\Delta t m_i^{-1}(K_{ii}+D_{ii}^L)>0, \forall i$. 
  Note also that, due to partition of unity and by construction of the dissipative operator $D^L$,
  $\sum_j K_{ij}+D^L_{ij}=0\implies \sum_j R_{ij}=1$. Therefore,
  \begin{align*}
    U_i^L=\sum_j R_{ij}U_j^n\leq U_j^{\max}\sum_j R_{ij} = U_j^{\max},
  \end{align*}
  and similarly for the lower bound.
\end{remark}

\begin{remark}[About mass conservation.]\label{remark:conservation_low-order}
  Let $\bfn$ denote the unit normal vector to the boundary $\partial\Omega$.
  Assume the velocity field is divergence free $\nabla\cdot \bfv=0$ and that $\bfv\cdot \bfn=0$,
  then \eqref{low-order} is conservative in the sense that
  $\sum_i m_i U_i^L=\sum_i m_i U_i^n$. Since $\bfv\cdot \bfn=0$, via integration by parts,
  $K_{ij}=\int_\Omega (\bfv\cdot\nabla \phi_j)\phi_i d\bfx = -\int_\Omega\nabla\cdot(\bfv\phi_i)\phi_jd\bfx$. Therefore $\sum_j K_{ji}=-\sum_j\int_\Omega \nabla\cdot(\bfv\phi_j)\phi_i=0$
  (where we use the partition of unity property and
  the assumption that the velocity is divergence free).
  By construction $\sum_jD_{ij}^L=0\implies \sum_j D_{ij}^L U_j^n=\sum_jD_{ij}^L (U_j^n-U_i^n)$.
  Therefore,
  \begin{align*}
    \sum_j (K_{ij}+D_{ij}^L)U_j^n =
    \sum_j \left[(K_{ij}U_j^n - K_{ji}U_i^n) + D_{ij}^L(U_j^n-U_j^n)\right].
  \end{align*}
  Let $f_{ij}^L:=(K_{ij}U_j^n - K_{ji}U_i^n) + D_{ij}^L(U_j^n-U_j^n)$
  and note that (since $D^L_{ij}=D_{ji}^L$)
  $f_{ij}^L=-f_{ji}^L$; i.e., $f_{ij}^L$ forms a skew-symmetric matrix.
  Therefore, $\sum_i\sum_j f_{ij}^L=0\implies\sum_i m_i U_i^L=\sum_i m_i U_i^n$.
\end{remark}

\subsection{Modified RK methods with flux limiting on last stage}
Let $F_i(t)=\sum_j f_{ij}(t)$ be some high-order spatial discretization of
\eqref{transport_equation}
that doesn't necessarily satisfies positivity, local bounds or other monotonicity constraint
(like the one given by \eqref{local-bounds}).
%
We are interested on solving the semi-discretization
\begin{align}
  \frac{d}{dt} U_i(t)=:F_i(t) %-m_i^{-1}\sum_j(K_{ij}+D_{ij}^L)U_j(t)=:F_i(t)
\end{align}
via high-order Runge-Kutta methods. Moreover, we aim to modify the Runge-Kutta method
to impose the local bounds \eqref{local-bounds}.

\begin{remark}[About the spatial discretization.]
  In the following we consider a high-order discretization to be given by the pure
  continuous Galerkin scheme. This is off course highly oscillatory when the time discretization
  is given by a Runge-Kutta method and even unconditionally unstable for forward Euler.
  To remedy this problem, it is common in the literature of flux corrected transport
  for continuous Galerkin finite elements to pre-stabilize the high-order spatial discretization.
  See for example \cite{lohmann2017flux,guermond2014second}.
  %
  In addition, in this section we lump the mass matrix, 
  which introduces dispersive errors \cite{guermond2017effect}
  We remedy these two problems later.
  %
  Therefore, for now the high-order semi discretization of \eqref{transport_equation} is given by
  \begin{align}\label{high-order_semi-discretization}
    \frac{d}{dt}U_i^H(t)=F_i(t)=-m_i^{-1}\sum_j K_{ij}(t) U_j(t),
  \end{align}
  where $U^H$ denotes the high-order solution. See more details in \S\ref{sec:FE_discUpwind}.
\end{remark}

We apply a standard Runge-Kutta scheme to \eqref{high-order_semi-discretization} to get
\begin{align}\label{high-order_RK}
  U_i^H=U_i^n+\Delta t\sum_{m=1}^M b_k k_{m,i}.
\end{align}
Here we abuse the notation and let $U^H$ denote the high-order solution at time $t^{n+1}$.
The idea is to extract the forward Euler component from \eqref{high-order_RK}.
Since $k_{1,i} = F_i(t^n,U^n)=:F_i^n$, we get 
\begin{align}\label{last_stage_RK_aux1}
  U_i^H=U_i^n+\Delta t F^n_i + \Delta t (b_1-1)F^n_i + \Delta t \sum_{m=2}^M b_k k_{m,i}.
\end{align}
Moreover, we split the forward Euler flux as
\begin{align*}
  F_i^n=-m_i^{-1}\sum_j K^n_{ij}U_j^n = -m_i^{-1}\sum_j \left[(K^n_{ij}+D_{ij}^{L,n})U_j^n - D_{ij}^{L,n}U_j^n\right],
\end{align*}
where $D_{ij}^{L,n}$ is the discrete upwinding matrix
(given by \eqref{transport_and_discUpwinding_matrices}) at time $t^n$. 
Using the definition of the low-order method \eqref{low-order},
equation \eqref{last_stage_RK_aux1} becomes
\begin{align}\label{last_stage_RK_aux2}
  m_i(U_i^H-U_i^L) = \Delta t \sum_j D_{ij}^{L,n} U_j^n +
  \Delta t m_i (b_1-1)F^n_i + \Delta t m_i\sum_{m=2}^M b_k k_{m,i}.
\end{align}

Since $U^L$ satisfies the local bounds \eqref{local-bounds},
it is clear that the right hand side of \eqref{last_stage_RK_aux2}
is responsible not only for increasing the accuracy of $U^H$ in space and time
but also for creating violations of the local bounds.
%
We use the standard ideas in the Flux Corrected Transport (FCT) methodology
by \cite{boris1973flux} and \cite{zalesak1979fully}, see also \cite{kuzmin2012flux},
to limit the flux correction; i.e., the right hand side of \eqref{last_stage_RK_aux2}.
%
To do this we need to rewrite the right hand side of \eqref{last_stage_RK_aux2}
as an anti-diffusive flux correction $f_{ij}^C$.
To guarantee conservation of mass we need $f_{ij}^C=-f_{ji}^C$.
Let $K^m_{ij}$ be the entries of the transport operator corresponding to the $m$-th stage
in the RK method. Then 
\begin{align}
  k_{m,i} = F_{m,i}=\sum_j (K^m_{ij}U_j^m - K^m_{ji}U_i^m),
\end{align}
see remark \ref{remark:conservation_low-order} for more details.
Similarly $F_i^n=\sum_j(K^n_{ij}U_j^n - K^n_{ji}U_i^n)$.
Finally, since $\sum_j D_{ij}^{L,n}=0$, we get
$\sum_jD_{ij}^{L,n}U_j^n=\sum_jD_{ij}^{L,n}(U_j^n-U_i^n)$. Therefore, the flux correction is
given as follows: 
\begin{align}
  \begin{split}
  f_{ij}^C = \Delta t \sum_j D_{ij}^{L,n}(U_j^n-U_i^n)
  +  \Delta t m_i(b_1-1)\sum_j(K^n_{ij}U_j^n - K^n_{ji}U_i^n) 
  +  \Delta t m_i\sum_{m=2}^M b_k \sum_j (K^m_{ij}U_j^m - K^m_{ji}U_i^m).
  \end{split}
\end{align}
Since $D_{ij}^{L,n}=D_{ji}^{L,n}$, it is clear that $f_{ij}^C=-f_{ji}^C$.
Now we rewrite \eqref{last_stage_RK_aux2} as
\begin{align}
  m_i(U_i^H-U_i^L) = \sum_j f_{ij}^C.
\end{align}

Finally the solution at time $t^{n+1}$ is given by 
\begin{align}\label{MRK_last_stage}
  m_i(U_i^{n+1}-U_i^L)=\sum_j\mathcal{L}_{ij}f_{ij}^C,
\end{align}
where $\mathcal{L}_{ij}$ are the flux limiters given by
\begin{subequations}
\begin{align}
  \mathcal{L}_{ij}:=
  \begin{cases}
    \min(R_i^+,R_j^-), & \text{ if } f_{ij}^C\geq 0, \\
    \min(R_i^-,R_j^+), & \text{otherwise}, \\
  \end{cases}
\end{align}
where
\begin{align}
  R_i^+ &:=
  \begin{cases}
    \min\left[1, \frac{m_i(U_i^{\max}-U_i^L)}{f_i^{C,+}} \right], & f_i^{C,+}\neq 0, \\
    1, & \text{otherwise}
  \end{cases},
  & R_i^- &:=
  \begin{cases}
    \min\left[1, \frac{m_i(U_i^{\min}-U_i^L)}{f_i^{C,-}} \right], & f_i^{C,-}\neq 0, \\
    1, & \text{otherwise}
  \end{cases}, \\
  f_i^{C,+}&:=\sum_j \max(0,f_{ij}^C)  & f_i^{C,-}&:=\sum_j \min(0,f_{ij}^C).
\end{align}
\end{subequations}

\begin{remark}[About the discrete maximum principle (DMP).]
  The solution $U_i^{n+1}$ satisfies the following local bounds:
  \begin{align*}
    U_i^{\min} \leq U_i^{n+1} \leq U_i^{\max}.
  \end{align*}
  The proof is very standard (see for instance \cite{guermond2014second,kuzmin2012flux}).
  Assume $f_i^{C,+}\neq 0$, then 
  \begin{align*}
    \begin{split}
      m_i(U_i^{n+1}-U_i^L)
      &=\sum_j\mathcal{L}_{ij}f_{ij}^C
      \leq\sum_{j,f_{ij}^C\geq0}\mathcal{L}_{ij}f_{ij}^C
      =\sum_{j,f_{ij}^C\geq0}\min(R_i^+,R_j^-)f_{ij}^C 
      \leq R_i^+ f_i^{C,+}
      \leq \frac{m_i(U_i^{\max}-U_i^L)}{f_i^{C,+}}f_i^{C,+}\\
      &=m_i(U_i^{\max}-U_i^L)
      \implies U_i^{n+1}\leq U_i^{\max}. 
    \end{split}
  \end{align*}
  If $f_i^{C,+}=0$,
  \begin{align*}
    \begin{split}
      m_i(U_i^{n+1}-U_i^L)\leq R_i^+ f_i^{C,+}=0\implies U_i^{n+1}\leq U_i^L\leq U_i^{\max}.
    \end{split}
  \end{align*}
  The lower bound is proven similarly.
\end{remark}

\begin{remark}[About mass conservation.]
  The method \eqref{MRK_last_stage} is mass conservative in the sense that
  $\sum_i m_iU_i^{n+1}=\sum_i m_iU_i^n$.
  Recall that the limiters are symmetric and the flux correction is skew symmetric; i.e.,
  $\mathcal{L}_{ij}=\mathcal{L}_{ji}$ and $f_{ij}^C=-f_{ji}^C$, respectively.
  Therefore,
  \begin{align*}
    \begin{split}
    \mathcal{L}_{ij}f_{ij}^C=-\mathcal{L}_{ji}f_{ji}^C
    \implies \sum_i\sum_j \mathcal{L}_{ij}f_{ij}^C=0 
    \implies \sum_i m_iU_i^{n+1}=\sum_i m_iU_i^L.
    \end{split}
  \end{align*}
  Finally, due to remark \ref{remark:conservation_low-order}, $\sum_i m_iU_i^L=\sum_i m_iU_i^n$.
  Therefore, $\sum_i m_iU_i^{n+1}=\sum_i m_iU_i^n$.  
\end{remark}

\subsection{Modified RK methods with flux limiting on all stages}
\subsection{Hierarchical modified RK methods with flux limiting}

\bibliographystyle{plain}
%\bibliographystyle{abbrvnat}
\bibliography{References}

\end{document}

### 2019-06-01 MRK4-fluxLimiting/clean.sh
Add stand alone code in python to run different tests

@@ -1,1 +1,1 @@
-rm -r *~ *pyc *__*
+rm -r *~ *pyc *__* *png

### 2019-06-01 MRK4-fluxLimiting/run_tests.py

@@ -1,0 +1,168 @@
+import numpy as np
+from matplotlib import pyplot as plt
+from MRK4 import *
+
+global_bounds=False
+
+## ************************ ##
+## ***** STANDARD FCT ***** ##
+## ************************ ##
+def run_FCT(N,name,smooth=False):
+    USE_FCT=True
+    
+    # physical parameters #
+    U=1
+    T=1.0
+    DT=0.1
+    x=np.linspace(0,1,N)
+
+    # Numerical parameters #
+    h=1./(N-1);
+    C=0.75
+    dt=C*h*U
+    
+    # Initial conditions #
+    u0 = (0.5*(np.sin(2*np.pi*x)+1)) if smooth==True else (1.0*(x>0.45)*(x<0.55))
+    un=1.0*u0;
+
+    (D1,DL)=compute_matrices(U,N)
+    initial_mass=h*np.sum(u0)
+    
+    # Time loop #
+    NT=T/dt
+    times = np.linspace(0,T,int(NT)+1)
+    uu = np.zeros([len(u0),int(NT)+1])
+    i=0
+    for t in times:
+        # First stage #
+        f,fG,fB,fijG,fijB=sdisc(un,D1,DL) # spatial discretization
+        sf,_ = limited_conservative_flux(un,un+dt*fG,dt*fijB,global_bounds=global_bounds) # limit anti-diffusive fluxes 
+        sf/=dt
+        u1 = un + dt*(fG + (sf if USE_FCT else fB))
+        
+        # Second stage 
+        f,fG,fB,fijG,fijB=sdisc(u1,D1,DL)
+        sf,_ = limited_conservative_flux(u1,u1+dt*fG,dt*fijB,global_bounds=global_bounds)
+        sf/=dt
+        u2 = 3./4*un + 1./4*(u1 + dt*(fG + (sf if USE_FCT else fB)))
+        
+        # Third stage
+        f,fG,fB,fijG,fijB=sdisc(u2,D1,DL)
+        sf,_ = limited_conservative_flux(u2,u2+dt*fG,dt*fijB,global_bounds=global_bounds)
+        sf/=dt
+        u3 = 1./3*un + 2./3*(u2 + dt*(fG + (sf if USE_FCT else fB)))
+
+        # update solution at new time step
+        unp1 = u3
+    
[…]

### 2019-06-01 MRK4-fluxLimiting/MRK4.py
Minor fixes in python code for the MRK4

@@ -4,11 +4,11 @@ from matplotlib import pyplot as plt
 # These functions are designed for a 1D periodic domain from [0,1] #
 
 def compute_matrices(U,N):
-    h=1/(N-1)
+    h=1.0/(N-1)
     D1=np.diag(np.ones(N-1),-1)-np.diag(np.ones(N-1),1)
     D1[0,-1]=1.    
     D1[-1,0]=-1.
-    D1*=U/2/h
+    D1*=U/2.0/h
 
     # MPP artificial viscosity operator #
     DL=D1*0

### 2019-06-01 MRK4-fluxLimiting/run_tests.py

@@ -2,7 +2,7 @@ import numpy as np
 from matplotlib import pyplot as plt
 from MRK4 import *
 
-global_bounds=False
+global_bounds=True
 
 ## ************************ ##
 ## ***** STANDARD FCT ***** ##
@@ -17,7 +17,7 @@ def run_FCT(N,name,smooth=False):
     x=np.linspace(0,1,N)
 
     # Numerical parameters #
-    h=1./(N-1);
+    h=1.0/(N-1);
     C=0.75
     dt=C*h*U
     
@@ -83,10 +83,10 @@ def run_MRK4(N,name,smooth=False):
     x=np.linspace(0,1,N)
 
     # Numerical parameters #
-    h=1./(N-1);
+    h=1.0/(N-1);
     C=0.8
     dt=C*h*U
-    dt/=1
+    dt/=1.0
 
     # Initial conditions #
     u0 = (0.5*(np.sin(2*np.pi*x)+1)) if smooth==True else (1.0*(x>0.45)*(x<0.55))
@@ -110,7 +110,7 @@ def run_MRK4(N,name,smooth=False):
         f3,_,_,_,_=sdisc(un_RK4+0.5*dt*f2,D1,DL)
         f4,_,_,_,_=sdisc(un_RK4+dt*f3,D1,DL)
         unp1_RK4 = un_RK4 + dt/6.0*(f1+2*f2+2*f3+f4)
-        
+
         # MRK4: stabilization in space with limiters
         f1,f1G,f1B,f1ijG,f1ijB=sdisc(un_MRK4,D1,DL)
         f2,f2G,f2B,f2ijG,f2ijB=sdisc(un_MRK4+0.5*dt*f1,D1,DL)
@@ -137,8 +137,8 @@ def run_MRK4(N,name,smooth=False):
     loss_RK4 = (initial_mass - h*np.sum(unp1_RK4))/initial_mass
     loss_MRK4 = (initial_mass - h*np.sum(unp1_MRK4))/initial_mass
 
-    # print('(min, max) value of RK4=',np.min(uu_RK4),np.max(uu_RK4))
-    # print('(min, max) value of MRK4=',np.min(uu_MRK4),np.max(uu_MRK4))
+    #print('(min, max) value of RK4=',np.min(uu_RK4),np.max(uu_RK4))
+    #print('(min, max) value of MRK4=',np.min(uu_MRK4),np.max(uu_MRK4))
[…]

### 2019-06-01 ideas.tex
Add convergence related to MRK4 with smooth initial data

@@
   The diagonal entries of $R$ are positive provided $\Delta t$ is small
-  enough such that $1-\Delta t m_i^{-1}(K_{ii}+D_{ii}^L)>0, \forall i$. 
+  enough such that
+  \begin{align}\label{CFL_low_order_scheme}
+    1-\Delta t m_i^{-1}(K_{ii}+D_{ii}^L)>0,\quad \forall i.
+  \end{align}
   Note also that, due to partition of unity and by construction of the dissipative operator $D^L$,
@@
 
+\clearpage
+\section{Numerical experiments}
+
+{\it Advection with smooth initial data.} Consider the domain $\Omega=(0,1)$
+with periodic boundary conditions. The velocity is given by $\bfv=1$ and the initial
+condition by
+\begin{align}\label{smooth_initial_data}
+  u(\bfx,t=0)=\frac{1}{2}(\sin(2\pi x)+1), 
+\end{align}
+which is clearly bounded between $[0,1]$.
+
+\subsection{Modified RK methods with flux limiting on last stage}
+
+{\it Advection with smooth initial data.}
+We consider the one-dimensional problem with initial data given by \eqref{smooth_initial_data}
+and perform a convergence study with different methods.
+The spatial discretization is given by second order central finite differences.
+%
+In all cases, we solve the problem up to $t=1$ and report the minimum and maximum values
+and the error and convergence in the $L^1$-norm given by
+\begin{align}
+  E=h\sum_i |U_i-u_i^{\text{exact}}|.
+\end{align}
+
+%
+As a reference, we solve the problem using the standard 4th order Runge-Kutta method.
+We consider \eqref{MRK_last_stage} with global bounds given by $(u_i^{\min}, u_i^{\max})=(0,1), ~\forall i$
+and local bounds given by \eqref{local-bounds}.
+In all cases the time step is given by $\Delta t=C h/U$ with $C=0.8, U=1$ and $h=\frac{1}{N-1}$ where $N$ is the number of DOFs
+reported in the results. The results are shown in table \ref{table:MRK4_limiting_last_stage}.
+
+{\color{red} NOTE: The time step should be chosen via \eqref{CFL_low_order_scheme}.
+  However, I think $\Delta t=C h/U$ is enough for 1D with uniform meshes. I will make the change in the code and run again the tests.}
+
+\begin{table}[h]\scriptsize
+\begin{center}
+\begin{tabular}{|c||c|c|c|c||c|c|c|c||c|c|c|c||} \cline{1-13}
+\multicolumn{1}{|c||}{} &
+\multicolumn{4}{|c||}{Standard RK4} &
+\multicolumn{4}{|c||}{Modified RK4 with global bounds} &
+\multicolumn{4}{|c||}{Modified RK4 with local bounds} \\ \hline
+$N$ & $ \min$ &  $1-\max$ & $ E $ &  & $\min$ &  $\max$ & $E$ &   & $\min$ & $\max$ & $E$ &  \\ \hline
+21  & 3.52E-2 & -1.46E-2 & 1.26E-2 & --   & 3.79E-2 & 0.9996 & 1.17E-2 & --   & 4.30E-2 & 0.9745 & 1.04E-2 & --    \\ \hline
+41  & -1.58E-2 & -7.66E-3 & 2.90E-3 & 2.12 & 0       & 0.9986 & 2.64E-3 & 2.14 & 1.13E-2 & 0.9905 & 2.58E-3 & 2.01 \\ \hline
+81  & -2.61E-3 & -2.88E-3 & 6.74E-4 & 2.10 & 7.03E-4 & 1      & 6.44E-4 & 2.03 & 5.38E-3 & 0.9966 & 6.42E-4 & 2.00 \\ \hline
+161 & -7.78E-4 & -7.89E-4 & 1.61E-4 & 2.05 & 0       & 1      & 1.59E-4 & 2.01 & 2.02E-3 & 0.9987 & 1.59E-4 & 2.00 \\ \hline
+321 & -3.39E-4 & -2.87E-4 & 3.98E-5 & 2.02 & 0       & 1      & 3.95E-5 & 2.01 & 7.82E-4 & 0.9995 & 3.96E-5 & 2.01 \\ \hline
+641 & -1.30E-4 & -3.56E-5 & 9.85E-6 & 2.01 & 0       & 1      & 9.83E-6 & 2.00 & 3.11E-4 & 0.9998 & 9.85E-6 & 2.00 \\ \hline
+\end{tabular}
+\caption{Grid convergence with second order central finite differences in space.
+  The time discretization is done via the standard RK4
+  and the modified RK4 \eqref{MRK_last_stage} using global and local bounds.}
+\label{table:MRK4_limiting_last_stage}
+\end{center}
+\end{table}
+
 \bibliographystyle{plain}

### 2019-06-03 MRK4-fluxLimiting/clean.sh
add convergence.ods file

@@ -1,1 +1,1 @@
-rm -r *~ *pyc *__* *png
+rm -r *~ *pyc *__* *png *csv

### 2019-11-21 project.tex
Add project description.

\documentclass{article}
\usepackage[utf8]{inputenc}

\title{Positivity-preserving adaptive Runge-Kutta methods: project context and outline}
\author{}

\begin{document}

\maketitle

\section{Main idea}

The goal of this project is to develop a class of Runge--Kutta methods that enforce positivity or similar bound constraints (such as a maximum principle) on the numerical solution by adaptively choosing the method coefficients.

If one considers the stages and their coefficients to be fixed,
then choosing the weights in a way that guarantees a positive
solution is a linear programming problem.

\subsection{Additional potential directions}
\begin{enumerate}
    \item Adapt subquadrature weights ($A$) as well
    \item Adapt time step size as well
    \item Linear multistep methods
\end{enumerate}

\section{Challenges and potential solutions}
\begin{enumerate}
    \item Solving a linear programming problem at each step may be expensive.  This can be alleviated by pruning the set of positivity constraints.  Or, by focusing on implicit methods,
    this cost is likely to be less significant in relative terms.
    \item It is difficult to ensure stability of the resulting method,
    since stability properties depend in a highly nonlinear way
    on the weights.  One way to potentially avoid instabilities is
    to ensure that the weights are not too far from those of a known
    good method.
    \item There may be no solution to the linear programming problem.
    In this case we need a fallback.  This could involve resorting
    to a lower-order embedded method (and possibly modifying its
    weights as well).
\end{enumerate}

\section{Existing related approaches and their limitations}
\subsection{Strong stability preserving (SSP) methods}
Guarantee positivity under a step size restriction proportional
to the explicit Euler positive step size.  If that step size
is very small, SSP methods are not so useful.  Explicit
one-step SSP methods can only be fourth-order accurate.  Also,
implicit SSP methods do not guarantee positivity under
large step sizes.  See e.g. \cite{gottlieb2009high,gottlieb2011strong}.

\subsection{Modified Patankar methods}
Guarantee positivity for certain types of ODEs
(conservative production-destruction systems).
Not applicable outside that problem class.  These methods
are implicit, though only in a linear way.
It's difficult to achieve higher than 3rd-order
accuracy with them.  See e.g. \cite{ortleb2017patankar,kopecz2018unconditionally}.

\subsection{Spatial adaptation of the time discretization}
Yet another approach is to use different time integration methods
in a localized spatial region where positivity is an issue.
For instance, one can use a convex combination of high-
and low-order schemes, with the combination chosen locally
to ensure positivity \cite{duraisamy2003concepts}.
See also \cite{ketcheson2013spatially,duraisamy2007implicit}.

\subsection{Local adaptation of the spatial discretization}
There are many works that take this approach, but we won't
review them here as they are quite different in principle
from the present focus.

\section{Applications and test problems}

\subsection{Explicit methods}

\begin{enumerate}
    \item Scalar conservation laws (advection, Burgers): solution bounded by extrema of initial data
    \item Shallow water equations: enforce positive depth
    \item Euler equations: enforce positive density, pressure
    \item Reactive Euler equations or multiphase equations: box bounds
\end{enumerate}

\subsection{Implicit methods}

\begin{enumerate}
    \item Heat equation with non-smooth initial data: enforce positivity and/or maximum principle
    \item Robertson problem
    \item HV atmospheric pollution problem
\end{enumerate}

\section{Methods to work with}
\subsection{Explicit methods}

\begin{enumerate}
    \item  SSP(10,4): lots of degrees of freedom in choosing weights
    \item Other high-order explicit methods with extra stages
    \item  Explicit extrapolation or deferred correction methods: have a built-in hierarchy of methods of increasing order, to use as fallbacks.  In particular, extrapolation methods based on explicit Euler can be guaranteed to have positive intermediate stages.
\end{enumerate}

\subsection{Implicit methods}
\begin{enumerate}
    \item  Implicit extrapolation or deferred correction methods: have a built-in hierarchy of methods of increasing order, to use as fallbacks.  In particular, extrapolation methods based on implicit Euler can be guaranteed to have positive intermediate stages.
\end{enumerate}
\end{document}

\section{Research tasks}
The research for this project mainly consists of implementing
different approaches and testing the implementation on
appropriate problems.  At each step, depending on the test
results, various additional steps may be necessary.

\begin{itemize}
    \item Develop basic implementation of the idea and test it on simple problems.  Identify any problems that arise. (done)
    \item Test fallback to lower-order embedded methods.
    \item Implement implicit Euler extrapolation methods.
\end{itemize}

### 2019-12-03 latex/overview.tex
Added a tex document with description and as a place to condense the results

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}

\title{Positivity-preserving adaptive Runge-Kutta methods: results}
\author{Stephan Nuesslein}

\begin{document}

\maketitle

\section{Main Idea}
The goal of the project is to develop a positivity preserving method that uses adaptive b-Coefficients for every step.  This means that the computed stage values can be used and don't have to be recomputed. 
The resulting b should comply with the order conditions as far as possible.

Flowchart to evaluate a step:

\begin{itemize}
\item Calculate the stage values $k_1,\cdots,k_s$
\item Calculate the $u_{n+1} = u_n + \Delta t K b_{d}$ with a default b
\item Test it $u_{n+1} \geq 0$ if not:
\item find a new b with $u_n + \Delta t K b \geq 0$
\item Calculate the $u_{n+1} = u_n + \Delta t K b$ with the new b
\end{itemize}

\section{Applicable Problems}
The approach can be used with ODE systems $u' = f(t,u)$ that ensure  $u(t) \geq 0 \forall_t \forall_{  u_0 \geq 0}$

This can be tested with the property 
\begin{equation}
u_i=0 \Rightarrow f_i(t,u_1,\cdots,u_i,\cdots,u_n) \geq 0  \forall_{u_c \geq 0} \forall_{t}
\end{equation}

For problems where only the combination of $f(t,u)$ and $u_0$ leads to positifity tests did not show promising results, additional it is not sure if the computed solution is reasonable.

\section{Details}

\subsection{Positifity}
For solving an PDE with the method of lines an spartial grid  with $p$ points is introduced. 
From this we get an ODE with an $u \in R^p$.

To enshure that the solution is positive

\begin{align}
 u_i^{n+1} &\geq 0   \;   \forall_{i \in \{0,p-1 \}}  \\
 u_i^n + h \sum_{j=0}^{s-1} b_j h_i^j  &\geq 0   \;   \forall_{i \in \{0,p-1 \}}  
\end{align}

has to be fulfilled.
This infers $p$ positivity constraints to the optimisation problem. These can be written as

\begin{equation}
u_i + h K  b \geq 0     
\end{equation}

where $K = \big[k^1 , \cdots k^{s-1}\big]$.

\subsection{Order Conditions}

The Order Conditions are an linear equations system and can be written as $O b = r$.

The vector $b \in R^s$ contains the b coefficients. $s$ is the number of stages. $O$ is an $n \times s$ matrix, where $g$ is the number of Order Conditions. $r$ is an vector containing the right hand side of the Order Conditions

We if we use a $A$ from a known method we know that there is at least one solution for the linear equation system.
In order to get an optimization problem the equation system has to be underdetermined.
This implies that $rank{O} \leq s$. 
Because the Quadrature Conditions are linear independent $p \leq s$ has to be fulfilled.
This implies that the number of stages has to be higher as the wanted required order.

\subsection{Objective function}
Different types of objective functions have been tried. (Notebook: objectives.ipynb)

\begin{description}

\item[L1 norm] The optimisation Problem opimises for $min(|b-rkm.b|_1)$

\item[Quadrature]   The optimisation Problem opimises for $min(|bTq-r|)$
                            where q is the quadrature condition of the next higher order 
                            and r is the expected right hand side

\item[Hom. Order]    The optimisation Problem opimises for $min(|bTO|)$ where O are the homogenus Order Conditions of the next higher orders

\end{description}

The different approaches have been tested for a test problem from Kopecz and Meister 2018 (Two substances reacting). The optimization problem was used at every step.
The magnitude of $b$ can get large if the Order Conditions are used as objective function.
By changing the $b$ the stability function is altered. This can lead to oscillations if the poles of the rhs are no longer in the stability region of the resulting method.
For this reason $||b_{optim}-b_{orig}||_1$ seems like a good choice
It is important to only adapt the b if needed (also for performance reasons).

\section{Results of experiments}
\subsection{Explicit methods}
For most of the test problems problems with stability occur before getting negative values. This means that we need methods with large stability region.
A test problem that works good is the advection‐diffusion‐production‐destruction systems. 
For the production‐destruction systems negative values at the intermediate steps occurred. These results of these could be suppressed by the optimization problem  except for some timesteps where these yield to glitches in the solution.

Conclusion: it is important to control for negative Values in the intermediate stages.
Possible solutions: Solve an positifity ensuring optimization Problem at every intermediate step.
Can be formulated as choosing a new $b$ for every intermediate stage which then gets part of $A$ in the next step.

Another idea to explore (Is kind of obvious but could also be stupid): Replace $f(t,u)$ with an $f_{pos}(t,u)$ with 
$f(t,u)=f_{pos}(t,u) \forall u \geq 0 $
Whereas $f_{pos}(t,u)$ is welldefined for $u \leq 0$

\subsection{Implicit methods}
Implicit methods seem like an advantageous choice for a couple of reasons.

\begin{itemize}
\item The computational effort of solving the optimisation Problem is relatively small compared to the overall cost of the method.
\item Implicit methods are stable for large step sizes.
\item Implicit methods ensure positifity for the backward euler step, even for large step sizes.
\item There are no methods for ensuring positifity for higher order methods (Which we also can not get around for the general case)
\end{itemize}

Used for the ADP model the model yields good results, the main limit for the stepsize is the accuracy of the solver for the stageequations.

When tested on the diffusion equation with an spike as initial data the $b$ is only changed for the first step. The change of the $b$ can lead to reasonable changes of the hight of the maximum.

\section{Useful proofs}

\subsection{Existence of b}
Ensuring a feasible $b$ exists for a certain order condition is in general very difficult.
If we drop the order conditions it is known that there is a positive solution as long as there is a $k_c$ with $k_c \geq 0$.
This is no problem when using a implicit method because a backward euler step always gives a positive solution.

\subsection{Stability function of linear combinations}
Let $R_b(z)$ denote the stability function of a RKM method with the butcher tableau

$$
\begin{array}
{c|c}
c & A\\
\hline
& b^T
\end{array}
$$

We want to prove that $R_{\alpha b_a+\beta b_b}(z)$ = $\alpha R_{ b_a}(z)+\beta R_{b_b}(z)$ when $\alpha + \beta = 1$ and $A$ is the same for all methods.

We start with the definition of the RKM

\begin{align}
u_{n+1} &= u_n + \Delta t K (\alpha b_a+\beta b_b) \label{eq:u_n+1}  \\
u_{n+1}^{sum} &= \alpha (u_n + \Delta t K b_a) + \beta (u_n + \Delta t K b_b) \label{eq:u_n+1_sum} \\
& = (\alpha + \beta) u_n + \Delta t \alpha K b_a + \Delta t \beta K b_b \\
&= u_n + \Delta t K (\alpha b_a+\beta b_b) \\
&=u_{n+1} 
\end{align}

Now we insert the Stability functions for \ref{eq:u_n+1} and \ref{eq:u_n+1_sum} which we already know to be equal. We get

\begin{align}
u_{n+1} &= u_{n+1}^{sum} \\
R_{\alpha b_a+\beta b_b}(z) u_n &= \alpha R_{\alpha b_a}(z) u_n + \beta R_{\beta b_b}(z) u_n \\
&= (\alpha R_{\alpha b_a}(z) + \beta R_{\beta b_b}(z)) u_n
\end{align}

because this has to be true for all $u_n$ we get

\begin{equation}
R_{\alpha b_a+\beta b_b}(z) = \alpha R_{\alpha b_a}(z) + \beta R_{\beta b_b}(z) 
\end{equation}

\subsection{Stability of convex combination}

With this we can proof that the convex combination of two methods is absolutely stable at the points where both original methods were absolutely stable.

We want to show that $|R_{b_a}(z)|  < 1 ^ |R_{b_b}(z)| < 1\Rightarrow |R_{\chi b_a +(1- \chi) b_b}(z)| < 1$ with $\chi \in [0,1]$.
We write $\alpha = \chi$ and $\beta = 1-\chi$ for convenience.

\begin{align}
|R_{\alpha b_a +\beta b_b}(z)| &= |\alpha R_{b_a}(z) + \beta R_{b_a}(z)| \leq |\alpha R_{b_a}(z)| + |\beta R_{b_a}(z)|\\
 &=| \alpha| \underbrace{|R_{b_a}(z)|}_{<1} + |\beta| \underbrace{|R_{b_a}(z)|}_{<1} < \alpha + \beta = 1
\end{align}

\end{document}

### 2019-12-03 OrderCondition.py
Improved the LP solver
The solving of the LP problem was moved into a seperate function to keep it consistant in the explicit and implicit implementation. A option to reduce the number of positifity constraints on the LP-Problem was introduced. 
A interface for an alternate LP solver form scipy was implemented. The options for the cvxpy solvers are now accesible.

Update Description of OrderCondition.py

@@ -12,7 +12,20 @@ def tau_hat(k,A,c):
 
 
 def OrderCond(A,c,order = 1):
+    """
     #Generates Order Condition Matrix O and right side vector r for Linear Equation System O@b=r
+    
+    example:
+    
+    A: A matrix of RKM
+    c: c vector of RKM
+    order: Order to compute the order condition
+    
+    (O,rhs) = OrderCond(rkm.A,rkm.c,order = 1)
+    
+    O@b - rhs #should return 0
+    
+    """
     
     s = len(c) #number of stages
     

### 2019-12-03 RKimple.py

@@ -1,14 +1,186 @@
-#This file defines Order conditions for Runge Kutta methods in the form O@b = r
+#This file conains implementation for an genreric RKM method with adaptive b control.
+#It contains a seperate implementation for explicit RK methods and DIRK methods 
 
 
 import numpy as np
 from scipy.optimize import fsolve
 from OrderCondition import *
 import cvxpy as cp
+from scipy.optimize import linprog
+
+
+def setup_LP(solver,rkm,O,rhs,s):
+    """
+    This method sets up the necesarry variables for the LP Problem
+    
+    Parameters:
+    solver:  solver that should be used
+    rkm: RKM used
+    O: Order conditin matrix
+    s: Number od stages
+    
+    returns:
+    LP_vars Tuple with variables for the LP solver
+    
+    """
+    
+    if solver == 'scipy_ip' or solver == 'scipy_sim':
+        bounds = (0, None)
+        A_eq = np.concatenate((O,-O),axis = 1)
+        b_eq = rhs - O@rkm.b
+            
+        e = np.ones(2*s)
+        
+        LP_vars = (A_eq,b_eq,e,bounds)
+            
+    else:
+        ap_op =cp.Variable(s)
+        an_op =cp.Variable(s)
+        e = np.ones(s) #vector for goal Function, just generates the 1-Norm of b
+        
+        LP_vars = (ap_op,an_op,e,O,rhs)
+
+    return LP_vars
+
+def solve_LP(solver,LP_vars,rkm,u,K,dt,reduce = False,**options):
+    """
+    This method solves the LP Problem
+    
+    Parameters:
+    
+    solver:  a string with the solver that should be used
+    LP_vars: the set of variables generated by setup_LP
+    
+    reduce: If set, the LP problem is first solved with a reduced set of constriants
+    
+    Returns:
+    status: integer representing the status of the algorithm.
+      For scipy
+        0 : Optimization proceeding nominally.
+        1 : Iteration limit reached.
+        2 : Problem appears to be infeasible.
+        3 : Problem appears to be unbounded.
+        4 : Numerical difficulties encountered.
+      For cvypy
+        0 : Optimization proceeding nominally.
+        2 : Problem appears to be infeasible.
+        4 : Numerical difficulties encountered.
+        5 : solver crashed
+
+    b: found b
+    
+    """
+    
+    
+                
[…]

### 2019-12-05 RKimple.py
Tests for explicit Methods, Convergenc
The developement is focused on two main topics: 
- Explicit methods:
	For which Problems and timesteps is the adaptation approach usefull
	Which basis methods are good choiches (in terms of stability and ensured positifity)
	
- Implicit Methods:
	Measure and improve the performance of the implicit solver. In terms of convergenc and speed

utils.py
Added a file with utility functions:
Functions for finding important dt for explicit methods
Function for plotting the convergenc

ADP.ipynb
Further testing on feasible dt for the explicit method
Convergenc plot

@@ -3,7 +3,7 @@
 
 
 import numpy as np
-from scipy.optimize import fsolve
+import scipy.optimize as opt
 from OrderCondition import *
 import cvxpy as cp
 from scipy.optimize import linprog
@@ -14,10 +14,10 @@ def setup_LP(solver,rkm,O,rhs,s):
     This method sets up the necesarry variables for the LP Problem
     
     Parameters:
-    solver:  solver that should be used
-    rkm: RKM used
-    O: Order conditin matrix
-    s: Number od stages
+    solver:     solver that should be used
+    rkm:        RKM used
+    O:          Order condition matrix
+    s:          Number od stages
     
     returns:
     LP_vars Tuple with variables for the LP solver
@@ -42,16 +42,18 @@ def setup_LP(solver,rkm,O,rhs,s):
 
     return LP_vars
 
-def solve_LP(solver,LP_vars,rkm,u,K,dt,reduce = False,**options):
+def solve_LP(solver,LP_vars,rkm,u,K,dt,reduce = False,verbose_LP = False, **options):
     """
     This method solves the LP Problem
     
     Parameters:
     
-    solver:  a string with the solver that should be used
-    LP_vars: the set of variables generated by setup_LP
+    solver:     a string with the solver that should be used
+    LP_vars:    the set of variables generated by setup_LP
     
-    reduce: If set, the LP problem is first solved with a reduced set of constriants
+    reduce:     If set, the LP problem is first solved with a reduced set of constriants
+    verbose_LP: prints additional messages
+    options:    additional Options, are passed through to the used solver
     
     Returns:
     status: integer representing the status of the algorithm.
@@ -67,10 +69,10 @@ def solve_LP(solver,LP_vars,rkm,u,K,dt,reduce = False,**options):
         4 : Numerical difficulties encountered.
         5 : solver crashed
 
-    b: found b
+    b: found b, if solver failed rkm.b
     
     """
-    
+
     
                 
     if solver == 'scipy_ip' or solver == 'scipy_sim':
@@ -87,7 +89,7 @@ def solve_LP(solver,LP_vars,rkm,u,K,dt,reduce = False,**options):
             i = np.zeros_like(u,dtype = bool)
             while np.any(u_<0):
                 i = (u_<0)|i #update indecies for conditions
-                print(np.sum(i),'constraints')
[…]

### 2019-12-05 utils.py

@@ -1,0 +1,195 @@
+"""
+This file contains utility functions for investigating properties of the rk methods
+
+plot_cnvergence: function to plot the convergence of methods
+
+find_dt: method for investigating certain dts using a bisect algorithm
+
+"""
+
+import numpy as np
+import matplotlib.pyplot as plt
+
+
+def plot_convergence(time_integrator,rkm,f,u0,dt,refernce,step=1,error='abs',dx='1',Norm = 2,Params = {}):
+    """"
+    Parameters:
+    time_integrator:    function used to integrate the ODE
+    rkm,f,u0:           Arguments for time integrator
+    dt:                 dt array with dts
+    reference:          Array with reference solutions to compare the computet solution against 
+    error:              Definition of error computation, one of 'abs','rel','grid'
+    dx:                 Discretisation for grid norm
+    Norm:               Norm to use for Error calculation ||u-u'||_Norm
+    Params:             Parameters for time integrator
+    
+    Return:
+    sol:                Array with the solutions used for calculationg the errors
+    err:                Array with errors
+
+    """
+    
+    err = np.zeros_like(dt)
+    
+    sol = []
+    
+    
+    for i in range(dt.size):
+        print('dt='+str(dt[i]))
+        t,u,b = time_integrator(rkm,dt[i],f,u0,**Params)
+        dif = refernce[:,i]-u[:,step]
+        if error == 'abs':
+            err[i] = np.linalg.norm(dif,ord=Norm)
+        elif error == 'rel':
+            err[i] = np.linalg.norm(dif,ord=Norm)/np.linalg.norm(refernce[:,i],ord=Norm)
+        elif error == 'grid': #Grid function Norm (LeVeque Appendix A.5)
+            error[i] = dx**(1/Norm)*np.linalg.norm(dif,ord=Norm)
+        else:
+            print('Error not defined')
+            print(error)
+            raise ValueError
+        sol.append(u[:,step])
+        
+    plt.plot(dt,err,'o-')
+    plt.yscale('log')
+    plt.xscale('log')
+    plt.grid()
+    plt.ylabel('Error')
+    plt.xlabel('dt')
+    
+    return sol,err
+    
+        
+
+def find_dt(time_integrator,rkm,f,u0,dt_start,cond = '',tol = 0.001,Params = {}):
+    
+    """"
+    The function searchs for the upper bound of dt that sattisfies a condition.
+    A bisect search approach is used
+
+    Parameters:
+    time_integrator:    function used to integrate the ODE
[…]

### 2020-01-13 latex/overview.tex
Updated the .tex document describing the project.
overview.tex: 
Added new result and added some notes on topics that have open questions or need further investigation

@@
 
-\section{Main Idea}
+With increasing complexity of the right hand side a step requect has a high cost. 
 The goal of the project is to develop a positivity preserving method that uses adaptive b-Coefficients for every step.
@@
 The resulting b should comply with the order conditions as far as possible.
+
+\section{Main Idea}
+
+When computing the solution of an ODE $u ' = f(t,u) $ using a RK Method with $s$ Stages we first compute the stage values
+
+\begin{equation}
+k^{j} =  f(t^n + \Delta t c_j, u^n + \Delta t \sum_{l = 1}^{s} a_{jl} k^l  \quad j = 1,\cdots,s)
+\end{equation}
+
+Afterwards we calculate the solution
+
+\begin{equation}
+u^{n+1} = u^n + \Delta t \sum_{j  = 1}^s k^{j} b_j .
+\end{equation}
+
+This equation can be formulated to
+
+\begin{equation}\label{eq:Combination}
+u^{n+1} = u^n + \Delta t K b
+\end{equation}
+
+where $K$ is a matrix containing the stage values $k_1,\cdots,k_s$.
+
+One reason why steps are requected is that the new solution does not comply with positifity. 
+So it may be adventagious to adapt Equation\,\ref{eq:Combination} in such a way that $u^{n+1}$ is positive. This is done by adapting the weights $b$.
+As we can see in Equation\,\ref{eq:Combination} $u^{n+1}$ depends linearly on the stage values.
+By altering the weights we can violate the Order Conditions. 
+Because $A$ is already fixed the Order Conditions reduce to linear equations.
+This means that finding the $b$ can be done by solving a linear Programming problem.
 
@@
 \item Calculate the $u_{n+1} = u_n + \Delta t K b_{d}$ with a default b
-\item Test it $u_{n+1} \geq 0$ if not:
+\item Test $u_{n+1} \geq 0$, if not:
 \item find a new b with $u_n + \Delta t K b \geq 0$
@@
 \subsection{Positifity}
-For solving an PDE with the method of lines an spartial grid  with $p$ points is introduced. 
-From this we get an ODE with an $u \in R^p$.
+For solving an PDE with the method of lines an spartial grid  with $m$ points is introduced. 
+From this we get an ODE with an $u \in R^m$.
 
@@
 \begin{align}
- u_i^{n+1} &\geq 0   \;   \forall_{i \in \{0,p-1 \}}  \\
- u_i^n + h \sum_{j=0}^{s-1} b_j h_i^j  &\geq 0   \;   \forall_{i \in \{0,p-1 \}}  
+ u_i^{n+1} &\geq 0   \;   \forall_{i \in \{1, \cdots,m \}}  \\
+ u_i^n + h \sum_{j=0}^{s-1} b_j h_i^j  &\geq 0   \;   \forall_{i \in \{1,\cdots,m \}}  
 \end{align}
@@
 where $K = \big[k^1 , \cdots k^{s-1}\big]$.
+
+This causes unnecessary many inequality constraints on the LP-Problem. To simplify the problem we want to reduce the positifity constraints at places where there is no issue with positifity.
+
+For this we write
+\begin{equation}
+u_i^n + h \sum_{j=0}^{s-1} b_j h_i^j  \geq 0   \;   \forall_{i \in k \subseteq \{1,\cdot,m \}} 
+\end{equation}
+
+Now we have to choose a $k \subseteq \{1,\cdots,m \}$ that include all indecies that are needed for determining the $b$.
+
+A reasonable approach is to set $k_0 = {i \in \{1,\cdots,m \} |  u_i^{n+1}  < 0}$.
+It is possible that this set changes with another b that was computed using the $k_0$.
+To account for this we generate $k_1 = {i \in \{1,\cdots,m \}|  u_i^{n+1}  < 0} \cup k_0$ using the new $u_i^{n+1}$.
+This ensures that we cannot get stuck in a loop.
+In the worst case $|k_a|$ is growing very slowly until $k = \{1,\cdots,m \}$.
+
+This effect did not occur so far.
+A approach on this would be to make a educated guess which indecies could get a problem with positifity.
+A possible estimate would be $\frac{u_i^{n+1}}{max(K_{(i,0)}, \cdots ,K_{(i,0)})} $
+
+When enforcing a maximum value we can do the same.
+When both enforcing maximum and minimum values we have two separate sets of active constraints $k_{min}$ and $k_{max}$ where $k_{max}$ denotes the set of active maximum constraints and $k_{min}$ denotes the set of active minimum constraints.
+Here it is important to update these sets simultaneously.
 
@@
 This implies that $rank{O} \leq s$. 
-Because the Quadrature Conditions are linear independent $p \leq s$ has to be fulfilled.
+Because the Quadrature Conditions are linear independent, $p \leq s$ has to be fulfilled.
 This implies that the number of stages has to be higher as the wanted required order.
+
+Accomplished using equality constraints for the LP-Problem.
 
@@
 
+A reasonable property of the objective function would be that for a set of b only constrained by the order condition the LP-Problem gives the original b as optimal solution.
+$argmin(f_{optim}(b)) = b_{orig}$.
+In practice this also means that we do not have to solve the optimization Problem if no constraints are violated because we already know the solution of this case.
+This is adventagious for performance reasons.
 The different approaches have been tested for a test problem from Kopecz and Meister 2018 (Two substances reacting). The optimization problem was used at every step.
@@
 By changing the $b$ the stability function is altered. This can lead to oscillations if the poles of the rhs are no longer in the stability region of the resulting method.
-For this reason $||b_{optim}-b_{orig}||_1$ seems like a good choice
-It is important to only adapt the b if needed (also for performance reasons).
+For this reason $||b_{optim}-b_{orig}||$ seems like a good choice.
+The Norm $||x||_1$ can be implement the with Linear Programming using slack variables.
+We choose $||b_{optim}-b_{orig}||_1$ as objectie function.
+
+Large values of b can lead to numerical instability.
+Because $\sum_{i  = 1}^s b_i = 1$, a increase of $|b|$ leads to a combination of negative and positive values.
+Because the computation only is done with a finite precision (and usually in floating point) large $||b||$ can lead to absorption effects.
+
+\subsection{Adaption of Timestep using the dense output formual}
+
+If the original step $\Delta t$ was to large and there is no suitable $b$, we would like to calculate a $u^(n+1)$ using the stage values instead of rejecting the step.
+For this the time step is reduced after calculation the stage values.
+This can be done using the Order Conditions for dense Output (Source for this?).
+We introduce $\theta \in [0,1]$. The new step taken has the length $\theta \Delta t$ whereas $\Delta t$ is the time step used when calculating the Stagevalues.
+To calculate a reduced step the Order Conditions have to be adapted and a new $b_{orig}$ for the objective function (and more important the Error detection afterwards) has to be generated. 
+(TODO: the $b_{orig}$ is still a open question)
+
+When adapting the Order Condition to a new $\theta$ the Order Condition Matrix $Q$ stays unchanged whereas the right hand side depends on $\theta$.
+
+The dependency of $r(\theta)$ is nonlinear.
+This means that the maximum $\theta$ with a suitable $b$ can not be calculated by using a LP-Problem.
+This means that different $\theta$ have to be tested independently.
+
+\subsection{Choice of Baseline method}
+
+Existence of embedded methods and corresponding degrees of freedom for the b
+
+\# TODO We could introduce some table here with the dimensions of the embedded methods for different methods and orders.
+
+Stability region in comparison with the maximum timestep that ensures positifity.
+
+Existence of a positive solution.
+This boils down to the question if there is a embedded first order method that ensures positivity.
+This is particularly interesting for Implicit methods, because the $\Delta t$ is not limited by the stability.
+If we have a embedded first Order Method we can ensure that there is always a positive solution, which might be of first Order.
+
+\subsection{Error detection/Approximation}
+The main question is: We see that for certain $b$ that comply with the Order Conditions we get results that do not resemble the solution. (some type of glitches, artefacts)
+
+Note: in general it is difficult to give some mathematically rigorous statement for this statement because -by definition- it only applies for $\delta t \gg 0$ where the proofs using the Taylor Series do not give valid results any more.
+We could say that the main idea of the adaption of $b$ is to alter the error in a certain way that we know that the numerical solution gets 'better' in the sense that it only leads to positive solutions.
+When taking the Taylor Series of the numeric solution the coefficients up to the Order of the method are fixed by the Order Conditions.
+The only degrees of freedom are the coefficents above the Order Conditions.
+This requires that $\Delta t > 0$.
+(More like $\Delta t^n \approx \Delta t^(n+1) $)
+The main way is to ensure that for $\Delta t \to 0$ : $b \to b_{orig}$. For this case we already know that the numeric solution converges to the exact solution.  
+This is made sure by property of the Objective function that $argmin(f_{optim}(b)) = b_{orig}$.
+
+To make sure that the new adapted solution is still reasonable we measure the deviation form the Original method.
+In the worst case the errors add up.
+(todo: here some math formula...
+(We know that can not be the case because we already know that we will reduce the error for the quantity we will make positive, still we can use it as a upper bound.))
+
+\subsection{Stability region}
+
+By adapting the $b$ we change the used RK Method. This also alters the stability function and therefore changes the region of absolute stability. 
+It depends on the solved problem and the used $\Delta t$ on how many steps the stability function is altered.
+For Problems where only a small number of step are affected a bigger change in the stability function would be acceptable.
+For Problems with require a adaptation of the weights for every step one needs to make sure that the resulting method is stable.
+
+In \ref{proof:combiningb} it is stated that one can construct the stability function of a RKM  $R_b$  with the weights $b = \alpha b_1 + \beta b_2$ by adding the weighted stability functions $R_{\alpha b_a+\beta b_b}(z) = \alpha R_{\alpha b_a}(z) + \beta R_{\beta b_b}(z) $ where all stability functions have the same $A$ and $\alpha + \beta = 1$.
+
+(Todo:It is not straightforward if there are no pole zero cancellations in the stability function.
+For explicit methods the stability function is a polynomial o degree $\leq s$ (Hundsdorfer).
+Pole zero cancellations can not happen for these.
+For implicit functions that are A-stable the poles have to be in the right halve plane.
+Here is starts to get interesting.
+Because the A-Stability of one method with a certain $A$ is not a sufficient  condition that there are no Poles for all RKM with this A-Matrix.
+But it is possible (or easy) to prove that if $a_{ii} > 0 \forall i = 1,\cdots , s$ then there are only poles in the right halve plane where they do not matter for the further considerations.)
+
+(Note:They do not behave like a Vector space because there is no scalar Product, and all the vector have to add up to 1... Is there something simmilar so we don not have to explain everything here?)
+
+Alternative 1:
+
+If the resulting method can be written as a convex combination of different methods it is easy to prove that the resulting stability function is at least the Intersection of the stability region of the methods that define the convex combination.
+
+Alternative2:
+
+A useful property of the resulting stability region is that it is similar to the stability region of the baseline method. 
+(Note: is it possible to give some statement like: If the stability regions of the rkm used for multiple steps are similar, then the overall method is stable for the intersection of the stability regions)
+
+Maybe place here some results on the stability region in dependence of the $b$
+(Note: I have some proof handy that states that for certain $A$s the boundary of the stability region does only change in regard to a changing b in a limited way.
+But the proof is a bit convoluted.
+I have to look if there is a more compact way to prove it)
+
+\section{Implementation aspects of the Algorithm}
+
+(Todo: Maybe a flowchart and a description on how we stitch everything together.)
+
+The algorithm on adapting the b consists of two main loops.
+The outer loop loops through all the $\theta$, staring with $\theta = 1$ down to the lowest $\theta$.
+The inner loop loops through the orders, starting with the highest Order down to the lowest Order.
+Until a acceptable $b$ the algorithm tests combiantions of $\theta$ and $p$.
+At first the Order Conditions are constructed.
+On these a LP-Solver is invoiced.
+If the LP-Problem is infeasible the algorithm immediately tries the next combination.
+If the LP-Problem is feasible we test if the $b$ is acceptable.
+At first it is tested weather the new $u^{n+1}$ indeed complies with the constraints.
+This is necessary because sometimes the used LP-solvers incorrectly identify the problem as feasible.
+(Note: because of the chosen objective function it can not be unbounded) As second test the deviation from the old solution (Compare Error estimation) is calculated.
+If the error is bigger than a set maximum error the $b$ is also rejected and the algorithm also switch to the next combination.
+As soon as the algorithm finds a suitable $b$ the loops are aborted and the $b$ is used for the update step.
+If all the combinations of $\theta$ and $p$ did not lead to a acceptable $b$ the step has to be rejected and a new set of stage values has to bee computed.
+
+(Note: It would also be possible to use another strategy for traversing all the combinations of $(p,\theta)$. Is it worth exploring these?)
+
+(Note: Is here a good place to give a note on where supplementary code can be found?)
+
+(Note: Should we discuss which solver is used. Probably it is not worth using space for some plot etc. on this... Maybe a short note?)
 
@@
 For most of the test problems problems with stability occur before getting negative values. This means that we need methods with large stability region.
+
+\subsubsection{ADP}
 A test problem that works good is the advection‐diffusion‐production‐destruction systems. 
@@
 Whereas $f_{pos}(t,u)$ is welldefined for $u \leq 0$
+
+We are going to test the solvers for multiple $\Delta t$. For the explicit method we want to show the important times.
+
+$$ \Delta t \; \underset{0}{+}---------\underset{\Delta t_{pos}}{+}\underbrace{--------}_{\gamma}\underset{\Delta t_{feasible}}{+}---\underset{\Delta t_{stable}}{+}--->  $$
+
+\begin{itemize}
+\item
+$\Delta t_{pos}$ is the largest stepsize that yields to positive results without adapting the $b$
+
+\item
+$\Delta t_{feasible}$ is the largest stepsize that can be used with the adaptive $b$
+
+\item
+$\Delta t_{stable}$ is the largest stepsize for that the method is stable
+\end{itemize}
+
+The timesteps where the adaption of the b is usefull is denoted with $\gamma$ idealy this interval is as large as possible.
+This means that $\Delta t_{pos} \ll \Delta t_{feasible} \approx \Delta t_{stable}$
+
+With stepsizes $\Delta t > \Delta t_{stable} $ we cann not expect the solution to be reasonable
+
+$\Delta t$ for the ADP Problem
+
+\begin{tabular}{|c|c|c|c|c|}
+\hline 
+Method & $\Delta t_{pos}$ &$\Delta t_{stable}$& $\Delta t_{feasible}$  \\ 
+\hline
+fe & 0.0015 & --- &  0.0051 \\
+\hline 
+ssp104 fallback =1 & 0.0134 & 0.0309 & 0.0311 \\ 
+\hline 
+dp5 &0.0031 &0.0031 & 0.0031\\ 
+\hline 
+ck5 & 0.0044&  0.0044 & 0.0044\\ 
+\hline 
+merson4 fallback= 1& 0.0038& 0.0055 & 0.0084\\
+\hline
+\end{tabular}
+
+We can see that for the ssp104 method is a good choice for this approach when solving this testproblem.  The merson4 method is also suitable. We can not get benefits form dp5 and ck5.
+
+\subsubsection{Quantum Simulations}
+Shows following shortcomings:
+
+To ensure positifity we do not only need $\rho_{i,i} \geq 0$ but also the additional constraint $|\rho_{j,k}|^2 \leq \rho_{j,j} \rho_{k,k}$ at least for the elements with $j = k +1$ and $j = k-1$.
+This conditions can not be formulated as LP-Problem.
 
@@
 
-\subsection{Stability function of linear combinations}
+\subsection{Stability function of linear combinations}\label{proof:combiningb}
 Let $R_b(z)$ denote the stability function of a RKM method with the butcher tableau
@@
 
-\subsection{Stability of convex combination}
+\subsection{Stability of convex combination}\label{proof:convex_comb}
 

### 2020-01-23 latex/overview.tex
Update overview.tex
Added a section regarding adaption using convex combinations. Some detail changes.

@@
 
+Important things to point out:
+Approach preserves linear invariants
+Not cost anything if there are no negative components (exept for the test)
+
+(Note: The problem is connected to the theory about embedded methods, in the sense that one have different RKM with the same set of stage values.
+It may be interesting to have some literature regarding this)
+
 \section{Main Idea}
@@
 \begin{equation}
-k^{j} =  f(t^n + \Delta t c_j, u^n + \Delta t \sum_{l = 1}^{s} a_{jl} k^l  \quad j = 1,\cdots,s)
+k^{(j)} =  f(t^n + \Delta t c_j, u^n + \Delta t \sum_{l = 1}^{s} a_{jl} k^{(l)}  \quad j = 1,\cdots,s)
 \end{equation}
@@
 \begin{equation}
-u^{n+1} = u^n + \Delta t \sum_{j  = 1}^s k^{j} b_j .
+u^{n+1} = u^n + \Delta t \sum_{j  = 1}^s k^{(j)} b_j .
 \end{equation}
@@
 As we can see in Equation\,\ref{eq:Combination} $u^{n+1}$ depends linearly on the stage values.
-By altering the weights we can violate the Order Conditions. 
+When altering the weights the Order Conditions should be still meet. 
 Because $A$ is already fixed the Order Conditions reduce to linear equations.
@@
 \begin{equation}
-u_i + h K  b \geq 0     
+u_i + \Delta t K  b \geq 0     
 \end{equation}
@@
 \begin{equation}
-u_i^n + h \sum_{j=0}^{s-1} b_j h_i^j  \geq 0   \;   \forall_{i \in k \subseteq \{1,\cdot,m \}} 
+u_i^n + \Delta b \sum_{j=0}^{s-1} b_j h_i^j  \geq 0   \;   \forall_{i \in h \subseteq \{1,\cdot,m \}} 
 \end{equation}
 
-Now we have to choose a $k \subseteq \{1,\cdots,m \}$ that include all indecies that are needed for determining the $b$.
+Now we have to choose a $h \subseteq \{1,\cdots,m \}$ that include all indecies that are needed for determining the $b$.
 
-A reasonable approach is to set $k_0 = {i \in \{1,\cdots,m \} |  u_i^{n+1}  < 0}$.
-It is possible that this set changes with another b that was computed using the $k_0$.
-To account for this we generate $k_1 = {i \in \{1,\cdots,m \}|  u_i^{n+1}  < 0} \cup k_0$ using the new $u_i^{n+1}$.
-This ensures that we cannot get stuck in a loop.
-In the worst case $|k_a|$ is growing very slowly until $k = \{1,\cdots,m \}$.
+A reasonable approach is to set $h_0 = \{ i \in \{1,\cdots,m \} |  u_i^{n+1}  < 0 \}$. 
+After the solution with the reduced set is computed, the algorithm has to check if the positifity of the other values $i \notin h \subseteq \{1,\cdot,m \}$ is still meet. 
+If the conditions is not meet a new set $h_{a+1} = \{ i \in \{1,\cdots,m \}|  u_i^{n+1}  < 0 \} \cup h_{a}$ using the new $u_i^{n+1}$ is generated and the LP-Problem is solved again.
+This process id repeated until a solution is found that leads to a $u_{n+1} \geq 0$.
+The choice of $h_{a+1}$ based on the $h_{a}$ ensures that the algorithm cannot get stuck in a loop.
+In the worst case $h_a$ is growing very slowly until $h_a = \{1,\cdots,m \}$.
 
@@
 
-The Order Conditions are an linear equations system and can be written as $O b = r$.
+The Order Conditions are an linear equations system and can be written as $Q b = r$.
 
@@
 $s$ is the number of stages.
-$O$ is an $n \times s$ matrix, where $g$ is the number of Order Conditions.
+$Q$ is an $n \times s$ matrix, where $g$ is the number of Order Conditions.
 $r$ is an vector containing the right hand side of the Order Conditions
 
-We if we use a $A$ from a known method we know that there is at least one solution for the linear equation system.
+If the $A$ of a  preexisting RK-method is used, there is at least one solution for the linear equation system.
 In order to get an optimization problem the equation system has to be underdetermined.
-This implies that $rank{O} \leq s$. 
-Because the Quadrature Conditions are linear independent, $p \leq s$ has to be fulfilled.
+This implies that $rank{Q} < s$. 
+Because the Quadrature Conditions are linear independent, there are at least $q$ independent vectors in $Q$. This leads to the $p < s$.
 This implies that the number of stages has to be higher as the wanted required order.
 
-Accomplished using equality constraints for the LP-Problem.
+\subsection{Ways to formulate the Optimization Problem}
 
-\subsection{Objective function}
+There are two possible ways to write down the adaptation of the $b$. The first way is to directly adapt the $b$ by adding a correction  $\Delta b$ that ensures positivity. 
+This is the easiest way to adapt the $b$. An advantageous property is that the step size can be reduced after calculation the stage values by using the dense output order conditions.  
+(Note: Recommendation for explicit methods)
+
+The second approach is to choose $b$ as a convex combination of a set of $b$s. This approach is more useful if some embedded methods with certain properties are know. 
+This approach is also shows more  predictable behavior because the resulting $u^{n+1}_b$ is also a convex combination of the solutions $u^{n+1}_{b_1},u^{n+1}_{b_2},\cdots$ 
+(Note: Recommendation for implicit methods)
+
+Both approaches are explained in more detail in the following sections.
+
+\subsection{Direct Adaptation}
+
+The Direct Adaptation approach leads to two set of constraints on the $b$.
+
+The first set of constraints is the positifity and can be written as 
+$$u_{n+1}=u_n+\Delta K \vec{b} \geq 0$$
+
+The second set of constraints are the Order Conditions
+$$Qb=r$$
+
+The weights $b = b_{orig} + \Delta b$ consist of the weights $b_{orig}$ of the original method and an adaption $\Delta b$. This $\Delta b \in ker \{ Q \}$.
+
+To determine the optimal $b$ a objective function is needed.
 Different types of objective functions have been tried. (Notebook: objectives.ipynb)
@@
 
-A reasonable property of the objective function would be that for a set of b only constrained by the order condition the LP-Problem gives the original b as optimal solution.
+A reasonable property of the objective function would be that for a set of $b$ only constrained by the order condition the LP-Problem gives $b = b_{orig}$ as optimal solution.
 $argmin(f_{optim}(b)) = b_{orig}$.
-In practice this also means that we do not have to solve the optimization Problem if no constraints are violated because we already know the solution of this case.
+In practice this also means that there is no need to solve solve the optimization Problem if no constraints are violated because the solution of this problem is already known.
 This is adventagious for performance reasons.
@@
 By changing the $b$ the stability function is altered. This can lead to oscillations if the poles of the rhs are no longer in the stability region of the resulting method.
-For this reason $||b_{optim}-b_{orig}||$ seems like a good choice.
+For this reason $||b-b_{orig}||$ seems like a good choice.
 The Norm $||x||_1$ can be implement the with Linear Programming using slack variables.
-We choose $||b_{optim}-b_{orig}||_1$ as objectie function.
+We choose $||b_{optim}-b_{orig}||_1$ as objective function.
 
@@
 
-\subsection{Adaption of Timestep using the dense output formual}
+\subsubsection{Adaption of Timestep using the dense output formual}
 
@@
 For this the time step is reduced after calculation the stage values.
-This can be done using the Order Conditions for dense Output (Source for this?).
+This can be done using the Order Conditions for dense Output (todo: Source for this).
 We introduce $\theta \in [0,1]$. The new step taken has the length $\theta \Delta t$ whereas $\Delta t$ is the time step used when calculating the Stagevalues.
@@
 This means that different $\theta$ have to be tested independently.
+(Note: we could include some algorithm to search for it)
+
+\subsection{Convex combination of existing methods}
+For some test problems (e.g. the Heat Equation solved with an implicit method), large distortions occur, even though a norm of the distortion of the b is used as a objective function. 
+For implicit methods it is known that implicit euler steps lead to a positive solution for any $\Delta t$.
+The backward Euler is still only first Order but shows more exact results in certain cases than a method of higher Order that shows distortions of the solution caused by the larger error coefficients ($\Delta t \gg 0$, so they do not disappear because of the nature of the polynomial).
+A possible method to fix this is to generate the new b based on a convex combination of embedded methods of RKM.
+
+todo: write this down in more detail with proofs
+$b$ is in a convex hull of al b
+If all the used b complie with an order conditions the new method also has this order
+
+$$ q^T b = q^T (\chi b_{\alpha} + (1-\chi) b_{\beta}) = \chi q^T  b_{\alpha} + (1-\chi) q^T  b_{\beta}) = \chi rhs + (1-\chi) rhs = rhs$$
+
+The optimisation problem is changed to:
+
+$$u_{n+1}=u_n+\Delta K \vec{b} \geq 0$$
+
+with
+
+$$ \vec{b} = \sum_i^d a_i \vec{b_i}$$
+
+$$ \forall_{i \in \{1, \cdots d \}} 0 \leq a_1 \leq 1$$
+
+and the objective
+
+$$ min \left(\sum_i^d w_i a_i \right) $$
+
+with some weighting factors $w_1,\cdots,w_d$. The weighting factors determine which is the preferred b to use.
+
+Because the optimal solution of the unconstrained problem has to be the original $b$ the weight $w_0$ corresponding to $b_{orig}$ should be the smallest $w$.
+This weight is simply set to $w_{b_{orig}} = 0$.
+For the other weights $w_i = \frac{1}{Order\{b_i\}}$ is used.
+
+For methods with $s \geq p$ we would still like to have some degrees of freedom for the highest order.
+For this we have to introduce at least one $b$ that complies with the order Condition of the highest order and $b \neq b_{orig}$.
+These additional $b$ can be constructed by calculating a Span $S$ of the Kernel $ker Q$ of the Order Condition Matrix $Q$.
+For every vector $s \in S$ one set of weights with $b = b_{orig} + a s$ and one with $b = b_{orig} - a s$ is added.
+The $a$ is a parameter that determines how big a change is allowed.
+
+As a first order method a $b$ is added that corresponds to a series of backward euler steps. These are know to be yield to a positive result (Todo: source).
+
+An important property of the embedded methods is that they should show the desired stability characteristics.
+
+A Reduction of the step size suing the order Conditions for dense output would also be possible. For this 
+Doing dense Output here is a bit complicated, but it would not help ensure positifity only (maybe) increase exactness.
+(Note: is it worth exploring this or is it better to just leave this open for further research it may be also interesting regardless the study of positifity)
+Everything has downsides...
 
@@
 
-Alternative 1:
+If the new $b$ is a chosen as a convex combination of $b$ it is easy to prove that the resulting  the Region of absolute stability is at least the Intersection of the region of absolute stability of the methods that define the convex combination.
 
-If the resulting method can be written as a convex combination of different methods it is easy to prove that the resulting stability function is at least the Intersection of the stability region of the methods that define the convex combination.
-
-Alternative2:
-
+If the $b$ is adapted directly there is no such proof.
 A useful property of the resulting stability region is that it is similar to the stability region of the baseline method. 

### 2020-01-23 latex/overview.tex
Update overview.tex
Derivative of stability region in regard to changing the b at the direct adaptation

@@
 \usepackage{amsmath}
+\usepackage{amsfonts}
 
@@
 
+\subsection{Stability of direct adaptation}
+
+We would like to calculate the derivative of the border in respect to a change of $b$.
+
+The stability function depending on the $b$ is expressed (Note: it is a bit complicated her because we cannot let it depend on the b directly, but only give b after a coordinate transformation, which also makes the expression easier because we can already introduce the order conditions here)
+$m$ is the number of dimensions of the Kernel of the Order Condition $m = dim\{Ker Q \}$
+The $b_1,\cdots,b_m$  are the Basis vectors of $Ker Q$. By this we can get the vector $w$ by $w = \left[b_1,\cdots,b_m\right]^{-1}(b-b_{orig})$.
+
+The General Stability function of all embedded RKM that satisfy the Order Conditions $Q$ is given by
+
+\begin{equation}\label{eq:gen_stabilityf}
+R(w,z) = (1-(w_1,\cdots,w_m))R_{b_{orig}}(z) + w_1 R_{b_1}(z) + \cdots + w_m R_{b_m}(z)
+\end{equation}
+
+The border is defined by the implicit function 
+\begin{equation}\label{eq:border}
+|R(w,u+iv)|^2 -1 = 0
+\end{equation}
+This function can be vied either as $\mathbb{R}^m \times  \mathbb{C} \rightarrow \mathbb{R}$ or $\mathbb{R}^m \times  \mathbb{R}^2 \rightarrow \mathbb{R}$.
+We are only concerned with the expansion or contraction. Because of this we are only interested in the change perpendicular to the Border of the stability region.
+
+For a point $z_0= u_0 +i v_0 $ 
+with $ |R(w,z_0)|^2 -1 = 0 $ the gradient of
+ $|R(w,u,v)|^2$ is calculated. 
+If the gradient does not vanish at $z = z_0$ Equation\,\ref{eq:border} Can be transformed in a function $\mathbb{R} \rightarrow \mathbb{R}$ depending on a single variable $a$ by setting $z = z_0 + n a$.
+The complex number $n$ denotes the normal vector $\vec{n} = \frac{\nabla |R(w,u,v))|^2}{\left| \nabla |R(w,u,v))|^2 \right|}$ of the border of the stability region written in complex notation.
+This defines the function $a(w)$ with $|R(w,z_0 + n a(w))|^2 -1 = 0$.
+This gives the condition that this approach is only applicable if the gradient does not vanish on the border of the stability function. This is not true in general but for the most common RKM.
+The derivative $\frac{\mathrm d}{\mathrm d w} (a(w))$ can be calculated using the implicit function theorem.
+The defining function is $f: \mathbb{R} \rightarrow \mathbb{R}$ $f(a) = |R(b,z_0 + n a(w))|^2 -1 $
+If explicit methods are used this function is known to be continuously differentiable because it can be written as a Polynomial of $w_1,\cdots,w_m,u$ and $v$.
+This is a necessary condition for the use of the of the Implicit function theorem.
+It is also known by definition, that for the point $w=0$ and $z = z_0$ the equation $ |R(w,z_0)|^2 -1 = 0 $ is satisfied.
+For simplicity only the formulas for the derivative at the value $b = b_{orig}$ are given. This corresponds to $w_1 = \cdots = w_m = 0$.
+The wanted derivative  is given by
+
+\begin{equation}
+ \frac{\partial a}{\partial w_j} (w) =
+ - \left[ J_{f,a}(w,a(w))  \right] ^{-1} 
+   \left[ \frac{\partial f}{\partial w_j}(w,a(w)) \right]
+\end{equation}
+
+as long as the inverse of the Jacobian exists. This is true in most cases as it has been discussed above.
+
+The Jacobian $J_{f,a}(w,a(w))$ evaluated at $w=0$ can be calculated using the directional derivative 
+$$ J_{f,a}(b_{orig},a(w)) \Big|_{w=0} = 
+ \frac{\partial f}{\partial a} (w,a(w)) \Big|_{w=0} = 
+ \nabla |R(w,a(w)|^2 \vec{n} \Big|_{w=0}
+= \left| \nabla|R(0,z_0)|^2 \right| = \left| \nabla|R_{b_{orig}}(z_0)|^2 \right|$$
+
+The derivative in respect to $w_j$ is 
+$ \frac{\partial f}{\partial w_j}(w,a(w)) \Big|_{w=0}$
+this simplifies to 
+\begin{equation}\label{eq:derivative_to_b}
+- 2 \underbrace{R_{b_{orig}}(z_0)R^*_{b_{orig}}(z)}_{=|R_{b_{orig}}(z_0)|^2=1} + R_{b_{orig}}(z_0)R^*_{b_v}(z) + R_{b_v}(z_0)R^*_{b_{orig}}(z_0) 
+= -2 + R_{b_{orig}}(z_0)R^*_{b_v}(z_0) + R_{b_v}(z_0)R^*_{b_{orig}}(z_0)
+\end{equation}
+using the fact that $z_0$ is on the border of the stability region.
+The derivative $-2 + R_{b_{orig}}(z_0)R^*_{b_v}(z_0) + R_{b_v}(z_0)R^*_{b_{orig}}(z_0)$ is known to be finite for explicit methods, because $R_{b_v}$ and $R_{b_{orig}}$ are Polynomials and $|z_0| < \infty$
+The same process can be done for all points on the border of the stability region. 
+As long as $\left[ J_{f,a}(w,a(w))  \right] ^{-1} > 0$ the derivative $\frac{\partial a}{\partial w_j} (w)$ is finite.
+This means that a small change the $b$ only leads to a small change of the stability region.
+
 \end{document}

### 2020-01-27 latex/Image_b_convex.tex
Update overview.tex
Cleaned up the text, added some graphics and tables.

TODOs: 
     Clean up section 'Implementation aspects of the Algorithm'
     Add the experimental results

\begin{tikzpicture}
    % Draw axes
	\coordinate (borig) at (1,2.3);
    \coordinate (b_a) at (0.8,0.8);
	\coordinate (b_b) at (2.3,1);
	\coordinate (b) at (1.1,1.5);

    \draw [<->,thick] (0,2.5) node (yaxis) [above] {$b_2$}
        |- (3,0) node (xaxis) [right] {$b_1$};

    \draw[thick,red,fill = red!20] (borig) -- (b_a) -- (b_b) -- cycle;

	\draw[thick,->] (0,0) -- node[anchor=south west,pos = 0.95] {$b_{orig}$} (borig);
    \draw[thick,->] (0,0) -- node[anchor=south west,pos = 0.95] {$b_b$} (b_a);
    \draw[thick,->] (0,0) -- node[anchor=south west,pos = 0.95] {$b$} (b);
	\draw[thick,->] (0,0) -- node[anchor=north west,pos = 0.8,fill=white] {$b_b$} (b_b);

\end{tikzpicture}

### 2020-01-27 latex/Image_b_direct.tex
Update overview.tex
Cleaned up the text, added some graphics and tables.

TODOs: 
     Clean up section 'Implementation aspects of the Algorithm'
     Add the experimental results

\begin{tikzpicture}
    % Draw axes
	\coordinate (borig) at (1.2,1.8);
    \coordinate (b) at (2,1);
	\coordinate (db) at ($(b)-(borig)$);

    \draw [<->,thick] (0,2.5) node (yaxis) [above] {$b_2$}
        |- (3,0) node (xaxis) [right] {$b_1$};

%Add subspace 
	\draw[thick,red,dotted] ($-0.9*(db)+(borig)$) -- ($0.9*(db)+(b)$);
	\draw[thick,red] ($-0.6*(db)+(borig)$) -- ($0.6*(db)+(b)$);

	\draw[thick,->] (borig) -- node[anchor=south west ,pos = 0.5] {$\Delta b$} (b);
	\draw[thick,->] (0,0) -- node[anchor=south east,pos = 0.8] {$b_{orig}$} (borig);
    \draw[thick,->] (0,0) -- node[anchor=north west,pos = 0.8] {$b$} (b);

\end{tikzpicture}

### 2020-01-27 latex/overview.tex
Update overview.tex
Cleaned up the text, added some graphics and tables.

TODOs: 
     Clean up section 'Implementation aspects of the Algorithm'
     Add the experimental results

@@
 \usepackage{amsfonts}
+
+\usepackage{graphicx}
+\usepackage{subcaption}
+
+\usepackage{tikz}
+\usetikzlibrary{shapes.geometric, arrows}
+\usetikzlibrary{calc}
+\usepackage{color}
+
+\usepackage[backend=biber]{biblatex}
+\addbibresource{pos_rk_stephan.bib}
 
@@
 
+A big number of problems that preserve positifity reaction discretiesed PDEs...
+
 With increasing complexity of the right hand side a step requect has a high cost. 
@@
 
-Important things to point out:
-Approach preserves linear invariants
-Not cost anything if there are no negative components (exept for the test)
-
 (Note: The problem is connected to the theory about embedded methods, in the sense that one have different RKM with the same set of stage values.
@@
 
+Other Approaches: 
+ssp104 methods: The idea is to construct a rkm Method out of explicit euler steps, that lead to a positive result for small enough $\Delta t$.
+(Todo: Read some literature regarding this and give a short explaination)
+
+Implicit methods: Modified Patanka-Runge-Kutta (MPRK) \cite{kopecz_comparison_2019}
+
+The idea of adapting the $b$ after calculation the stage values is used in \cite{ketcheson_spatially_2013}
+In this case it is used to adapt the properties of the time integrator for a method of lines simulation on a PDE.
+With this approach it is possible to have different properties of the RKM at different parts of the domain.
+
 \section{Main Idea}
 
-When computing the solution of an ODE $u ' = f(t,u) $ using a RK Method with $s$ Stages we first compute the stage values
+When computing the solution of an ODE $u ' = f(t,u) $ using a RK-Method with $s$ Stages the stage values $k^{(1)},\cdots,k^{(s)}$ are computed according to
 
@@
 
-Afterwards we calculate the solution
+Based on these values the next solution $u^{n+1}$ is computed by
 
@@
 
-One reason why steps are requected is that the new solution does not comply with positifity. 
+One reason why steps are requected is that the new solution $u^{n+1}$ does not comply with positifity. 
 So it may be adventagious to adapt Equation\,\ref{eq:Combination} in such a way that $u^{n+1}$ is positive. This is done by adapting the weights $b$.
-As we can see in Equation\,\ref{eq:Combination} $u^{n+1}$ depends linearly on the stage values.
-When altering the weights the Order Conditions should be still meet. 
-Because $A$ is already fixed the Order Conditions reduce to linear equations.
+As it can be see in Equation\,\ref{eq:Combination} $u^{n+1}$ is a linear combination of the stage values and the previous solution $u^n$.
+When altering the weights the properties of the RKM should be preserved. Especially the Order of the RKM.  
+To ensure this the Order Conditions should be still meet. 
+Because the coeffeicents $a_{jl}$ are already fixed the Order Conditions reduce to linear equations.
 This means that finding the $b$ can be done by solving a linear Programming problem.
+
+This approach to ensure positifity preserves linear invariants like mass conservation, because it is a regular RKM. 
+If there are no negative values in $u^{n+1}$ with the original $b$, then the method does not cause additional effort, because it only has to solve the LP-Problem if the regular RKM does not lead to a positive $u^{n+1}$.
+(exept for the test)
+No additional evaluations of the rhs.
 
@@
 
-For problems where only the combination of $f(t,u)$ and $u_0$ leads to positifity tests did not show promising results, additional it is not sure if the computed solution is reasonable.
+For problems where only the combination of $f(t,u)$ and $u_0$ leads to the positifity of the exact solution, tests did not show promising results.
+Additional it is not certain if the computed solutions would be reasonable.
 
@@
 \subsection{Positifity}
-For solving an PDE with the method of lines an spartial grid  with $m$ points is introduced. 
-From this we get an ODE with an $u \in R^m$.
+
+For the further discussion we are concerned with a System of $m$ coupled ODEs. 
+In this setup  $u \in \mathbb{R}^m$.
+Systems of ODEs with a high number of dimensions can appear when solving a PDE with the method of lines.
 
@@
 has to be fulfilled.
-This infers $p$ positivity constraints to the optimisation problem. These can be written as
+This infers $m$ positivity constraints to the optimisation problem. These can be written as
 
@@
 
-This causes unnecessary many inequality constraints on the LP-Problem. To simplify the problem we want to reduce the positifity constraints at places where there is no issue with positifity.
+This causes unnecessary many inequality constraints on the LP-Problem.
+To simplify the problem one want to reduce the number of positifity constraints by ignoring it for ODEs where there is no issue with positifity.
 
-For this we write
+The new positifity constraints can be formulated as
 \begin{equation}
@@
 
-Now we have to choose a $h \subseteq \{1,\cdots,m \}$ that include all indecies that are needed for determining the $b$.
+For this a set $h \subseteq \{1,\cdots,m \}$ that include all indecies that are needed for determining the $b$ has to be chosen.
 
 A reasonable approach is to set $h_0 = \{ i \in \{1,\cdots,m \} |  u_i^{n+1}  < 0 \}$. 
-After the solution with the reduced set is computed, the algorithm has to check if the positifity of the other values $i \notin h \subseteq \{1,\cdot,m \}$ is still meet. 
-If the conditions is not meet a new set $h_{a+1} = \{ i \in \{1,\cdots,m \}|  u_i^{n+1}  < 0 \} \cup h_{a}$ using the new $u_i^{n+1}$ is generated and the LP-Problem is solved again.
-This process id repeated until a solution is found that leads to a $u_{n+1} \geq 0$.
+After the new $b$ and $u^{n+1}$ is computed using $h_0$, the algorithm has to check if the positifity of the other values $i \notin h_0 \subseteq \{1,\cdots,m \}$ is still meet. 
+If the conditions is not meet a new set $h_{a+1} = \{ i \in \{1,\cdots,m \}|  u_i^{n+1}  < 0 \} \cup h_{a}$ using the new $u^{n+1}$ is generated and the LP-Problem is solved again.
+This process id repeated until a solution is found that leads to a $u^{n+1} \geq 0$.
 The choice of $h_{a+1}$ based on the $h_{a}$ ensures that the algorithm cannot get stuck in a loop.
@@
 
-When enforcing a maximum value we can do the same.
-When both enforcing maximum and minimum values we have two separate sets of active constraints $k_{min}$ and $k_{max}$ where $k_{max}$ denotes the set of active maximum constraints and $k_{min}$ denotes the set of active minimum constraints.
+When enforcing a maximum value the number of constraints can be reduced using the same technique.
+When both enforcing maximum and minimum values two separate sets of active constraints $k_{min}$ and $k_{max}$ are used.
+The set $k_{max}$ denotes the active maximum constraints and the set $k_{min}$ denotes the active minimum constraints.
 Here it is important to update these sets simultaneously.
 
-\subsection{Order Conditions}
+\subsection{Order Conditions}\label{sec:OrderCond}
 
-The Order Conditions are an linear equations system and can be written as $Q b = r$.
+The Order Conditions for a RKM are a set of Equations depending on $a_{jl}$, $c_1,\cdots,c_s$ and $b_1,\cdots,b_s$.
+In these Equations $b_1,\cdots,b_s$ only appear linearly \cite{hairer_runge-kutta_1993}.
+If $a_{jl}$ and $c_1,\cdots,c_s$ are already known, the Order Conditions for the Order $p$ are an linear equations system and can be written as $Q b = r$.
 
-The vector $b \in R^s$ contains the b coefficients.
-$s$ is the number of stages.
-$Q$ is an $n \times s$ matrix, where $g$ is the number of Order Conditions.
-$r$ is an vector containing the right hand side of the Order Conditions
+The vector $b \in \mathbb{R}^s$ contains the weights and $Q$ is an $n \times s$ matrix, where $g$ is the number of Order Conditions.
+The vector $r$ contains the right hand side of the Order Conditions.
 
@@
 In order to get an optimization problem the equation system has to be underdetermined.
-This implies that $rank{Q} < s$. 
-Because the Quadrature Conditions are linear independent, there are at least $q$ independent vectors in $Q$. This leads to the $p < s$.
-This implies that the number of stages has to be higher as the wanted required order.
+This implies that $\mathrm{rank}(Q) < s$. 
+Because the Quadrature Conditions are linear independent, there are at least $p$ independent vectors in $Q$. This leads to the condition $p < s$.
+This implies that the number of stages has to be higher as the wanted order.
 
@@
 
-There are two possible ways to write down the adaptation of the $b$. The first way is to directly adapt the $b$ by adding a correction  $\Delta b$ that ensures positivity. 
+There are two possible ways to write down the adaptation of the $b$. The first way is to directly adapt the original weights $b_{orig}$ by adding a correction $\Delta b$ that ensures positivity.
 This is the easiest way to adapt the $b$. An advantageous property is that the step size can be reduced after calculation the stage values by using the dense output order conditions.  
 (Note: Recommendation for explicit methods)
+The direct adaption approach is explained in more detail in section\,\ref{sec:direct}.
 
 The second approach is to choose $b$ as a convex combination of a set of $b$s. This approach is more useful if some embedded methods with certain properties are know. 
-This approach is also shows more  predictable behavior because the resulting $u^{n+1}_b$ is also a convex combination of the solutions $u^{n+1}_{b_1},u^{n+1}_{b_2},\cdots$ 
+This approach also shows more  predictable behavior because the resulting $u^{n+1}_b$ is also a convex combination of the solutions $u^{n+1}_{b_a},u^{n+1}_{b_b},\cdots$ 
 (Note: Recommendation for implicit methods)
+The direct adaption approach is explained in more detail in section\,\ref{sec:convex}
 
-Both approaches are explained in more detail in the following sections.
+Both approaches are shown in figure\,\ref{fig:b_space}.
 
-\subsection{Direct Adaptation}
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.45\textwidth}
+        \input{Image_b_direct}
+        \caption{Direct Adaptation}
+        \label{fig:b_direct}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.45\textwidth}
+        \input{Image_b_convex}
+        \caption{Convex Combiantion}
+        \label{fig:b_convex}
+    \end{subfigure}
+    \caption{Graphical representation of the two different approaches to adapt the weights}\label{fig:b_space}
+\end{figure}
 
+\subsection{Direct Adaptation}\label{sec:direct}
+
+When directly adapting the $b$ a new $b$ that complies with the Order Conditions is chosen.   
+The weights $b = b_{orig} + \Delta b$ consist of the weights of the original method $b_{orig}$ and an adaption $\Delta b$. This $\Delta b \in \mathrm{ker}(Q)$. 
+In Figure\,\ref{fig:b_direct} this is is drawn. The red line represents the subspace of $b$ that compile with the Order Conditions. Both $b_{orig}$ and $b$ are in this subspace.
+
+To run it as a LP-Problem a set of constraints is constructed.  
+Also a objective function is needed to determine the optimal $b$.
+
+\subsubsection{Constraints}
 The Direct Adaptation approach leads to two set of constraints on the $b$.
@@
 
-The weights $b = b_{orig} + \Delta b$ consist of the weights $b_{orig}$ of the original method and an adaption $\Delta b$. This $\Delta b \in ker \{ Q \}$.
-
-To determine the optimal $b$ a objective function is needed.
-Different types of objective functions have been tried. (Notebook: objectives.ipynb)
+\subsubsection{Objective Function}
+Different types of objective functions have been tried.
 
@@
 
-\item[Quadrature]   The optimisation Problem opimises for $min(|bTq-r|)$
+\item[Quadrature]   The optimisation Problem opimises for $min(|b^Tq-r|)$
                             where q is the quadrature condition of the next higher order 
@@
 
-A reasonable property of the objective function would be that for a set of $b$ only constrained by the order condition the LP-Problem gives $b = b_{orig}$ as optimal solution.
-$argmin(f_{optim}(b)) = b_{orig}$.
+A reasonable property of the objective function would be that for the LP-Problem only constrained by the order condition the optimal solution is $b = b_{orig}$. 
+This can be achieved by $argmin(f_{optim}(b)) = b_{orig}$. 
 In practice this also means that there is no need to solve solve the optimization Problem if no constraints are violated because the solution of this problem is already known.
 This is adventagious for performance reasons.
-The different approaches have been tested for a test problem from Kopecz and Meister 2018 (Two substances reacting). The optimization problem was used at every step.
+The different approaches have been tested on a linear test problem from \cite{kopecz_unconditionally_2018}. The optimization problem was used at every step.
 The magnitude of $b$ can get large if the Order Conditions are used as objective function.
-By changing the $b$ the stability function is altered. This can lead to oscillations if the poles of the rhs are no longer in the stability region of the resulting method.
-For this reason $||b-b_{orig}||$ seems like a good choice.
-The Norm $||x||_1$ can be implement the with Linear Programming using slack variables.
-We choose $||b_{optim}-b_{orig}||_1$ as objective function.
-
 Large values of b can lead to numerical instability.
-Because $\sum_{i  = 1}^s b_i = 1$, a increase of $|b|$ leads to a combination of negative and positive values.
+Because $\sum_{i  = 1}^s b_i = 1$, a increase of $||b||$ leads to a combination of negative and positive values.
 Because the computation only is done with a finite precision (and usually in floating point) large $||b||$ can lead to absorption effects.
+By changing the $b$ the stability function is altered. This can lead to oscillations if the poles of the RHS are no longer in the stability region of the resulting method.
+For these reasons $||b-b_{orig}||$ seems like a good choice. The 1-Norm can be implement the with Linear Programming using slack variables. We choose $||b_{optim}-b_{orig}||_1$ as objective function.
 
@@
 
-If the original step $\Delta t$ was to large and there is no suitable $b$, we would like to calculate a $u^(n+1)$ using the stage values instead of rejecting the step.
+If the original step $\Delta t$ was to large and there is no suitable $b$, we would like to calculate a $u^{n+1}$ using the stage values instead of rejecting the step.
 For this the time step is reduced after calculation the stage values.
-This can be done using the Order Conditions for dense Output (todo: Source for this).
-We introduce $\theta \in [0,1]$. The new step taken has the length $\theta \Delta t$ whereas $\Delta t$ is the time step used when calculating the Stagevalues.
+This can be done using the Order Conditions for dense Output.
+These can be found in \cite{hairer_runge-kutta_1993}.
+We introduce $\theta \in (0,1]$. The new step taken has the length $\theta \Delta t$ whereas $\Delta t$ is the time step used when calculating the Stagevalues.
 To calculate a reduced step the Order Conditions have to be adapted and a new $b_{orig}$ for the objective function (and more important the Error detection afterwards) has to be generated. 
@@
 The dependency of $r(\theta)$ is nonlinear.
-This means that the maximum $\theta$ with a suitable $b$ can not be calculated by using a LP-Problem.
+This means that the maximum $\theta$ with a suitable $b$ can not be calculated by solving a LP-Problem.
 This means that different $\theta$ have to be tested independently.
@@
 
-\subsection{Convex combination of existing methods}
+\subsection{Convex combination of existing methods}\label{sec:convex}
 For some test problems (e.g. the Heat Equation solved with an implicit method), large distortions occur, even though a norm of the distortion of the b is used as a objective function. 
@@
 
-todo: write this down in more detail with proofs
-$b$ is in a convex hull of al b
-If all the used b complie with an order conditions the new method also has this order
+\subsubsection{Constraints}
+At first a set of embedded methods of the RKM $b_1,\cdots,b_d$ are needed. 
+If all the used b complies with an order conditions the new method also has this order.
+This can is straightforward using
 
@@
 
-The optimisation problem is changed to:
+The new $b$ is defined as 
+\begin{equation}\label{eq:b_convex_def}
+\vec{b} = \sum_i^d a_i \vec{b_i} = Ba
+\end{equation}
 
-$$u_{n+1}=u_n+\Delta K \vec{b} \geq 0$$
+where $B=\left[b_1,\cdots,b_d \right]$ is a matrix containing the embedded methods.  
+To ensure the that Equation\,\ref{eq:b_convex_def} yields a convex combination following constraints are enforced for the  the parameters $a,\cdots,a_d$.
 
-with
+\begin{equation}
+ 0 \leq a_1 \leq 1  \forall_{i \in \{1, \cdots d \}}
+\end{equation}
+\begin{equation}
+ \sum_{i=1}^d a = 1
+\end{equation}
 
-$$ \vec{b} = \sum_i^d a_i \vec{b_i}$$
+The positifity constraint changes to
 
-$$ \forall_{i \in \{1, \cdots d \}} 0 \leq a_1 \leq 1$$
+$$u_{n+1}=u_n+\Delta K B a \geq 0$$
 
-and the objective
+\subsubsection{Objective Function}
+
+As an Objective function the function
 
@@
 
-with some weighting factors $w_1,\cdots,w_d$. The weighting factors determine which is the preferred b to use.
+is used.  The $w_1,\cdots,w_d$ are weighting factors that determine which is the preferred $b_i$ to use.
 
@@
 This weight is simply set to $w_{b_{orig}} = 0$.
-For the other weights $w_i = \frac{1}{Order\{b_i\}}$ is used.
+For the other weights $w_i = \frac{1}{\mathrm{Order} \{b_i\}}$ is used.
 
-For methods with $s \geq p$ we would still like to have some degrees of freedom for the highest order.
-For this we have to introduce at least one $b$ that complies with the order Condition of the highest order and $b \neq b_{orig}$.
-These additional $b$ can be constructed by calculating a Span $S$ of the Kernel $ker Q$ of the Order Condition Matrix $Q$.
+\subsubsection{Construction of Embedded Methods}
+For methods with $s \geq p$ it is useful to have some degrees of freedom for the highest order.
+For this at least one $b$ that complies with the order Condition of the highest order and $b \neq b_{orig}$ has been introduced.
+These additional $b$ can be constructed by calculating a Span $S$ of the Kernel $\mathrm{ker} (Q)$ of the Order Condition Matrix $Q$.
 For every vector $s \in S$ one set of weights with $b = b_{orig} + a s$ and one with $b = b_{orig} - a s$ is added.
@@
 
-A Reduction of the step size suing the order Conditions for dense output would also be possible. For this 
-Doing dense Output here is a bit complicated, but it would not help ensure positifity only (maybe) increase exactness.
-(Note: is it worth exploring this or is it better to just leave this open for further research it may be also interesting regardless the study of positifity)
+A Reduction of the step size suing the order Conditions for dense output would also be possible.
+For this a set of embedded methods $b_1(\Theta),\cdots,b_d(\Theta)$ has to be constructed.
+This was not done in this research.
+
+(Note: Doing dense Output here is a bit complicated, but it would not help ensure positifity only (maybe) increase exactness.
+is it worth exploring this or is it better to just leave this open for further research it may be also interesting regardless the study of positifity)
 Everything has downsides...
 
-\subsection{Choice of Baseline method}
+\subsection{Choice of Baseline Method}
+An important property for the used baseline method is the Existence of embedded methods and the degrees of freedom for the b.
+As noted in Section\,\ref{sec:OrderCond} the number of Stages has to be higher than the Order. 
+Another important property is the number of degrees of freedom for the choice of the new weights. 
+These can be calculated using $s-\mathrm{rank}(Q)$%$\mathrm{dim}(\mathrm{ker}(Q))$.
+The number of degrees of freedom for different methods are shown in table\,\ref{table:DOF_exp} for explicit methods and in table\,\ref{table:DOF_imp} for implic methods.
 
-Existence of embedded methods and corresponding degrees of freedom for the b
+\begin{table}[h!]
+\centering    %Generated below============ 
+\begin{tabular}{|l |c c c c c c |} 
+ \hline 
+Order &1&2&3&4&5&6 \\ 
+ \hline Classical RK4&3&2&0&0& - & -  \\ 
+ SSPRK(10,4)&9&8&6&4& - & -  \\ 
+ Cash-Karp RK5(4)6&5&4&2&1&0& -  \\ 
+ Dormand-Prince RK5(4)7&6&5&3&1&0& -  \\ 
+ \hline 
+ \end{tabular}
+ \caption{Degrees of Freedom for Choice of $b$} %Generated above============ 
+ \label{table:DOF_exp}
+ \end{table}
 
-\# TODO We could introduce some table here with the dimensions of the embedded methods for different methods and orders.
+ \begin{table}[h!]
+\centering   %Generated below============ 
+ \begin{tabular}{|l |c c c c c c |} 
+ \hline 
+Order &1&2&3&4&5&6 \\ 
+ \hline Implicit Euler&0& - & - & - & - & -  \\ 
+ Im-Euler 2&2&1& - & - & - & -  \\ 
+ Im-Euler 3&5&4&2& - & - & -  \\ 
+ Im-Euler 4&9&8&6&3& - & -  \\ 
+ \hline 
+ \end{tabular}
+ \caption{Degrees of Freedom for Choice of $b$} %Generated above============ 
+ \label{table:DOF_imp}
+ \end{table}
 
-Stability region in comparison with the maximum timestep that ensures positifity.
+For all methods with the number of stages equal to the order of the method, there are no degrees of freedom without reducing the order. 
+If the Classical RK4 method is used the Order has to reduced more because the RK4 method does not have embedded methods of order 3.
+Contrary to this methods with $s > p$ have degrees of freedom even without reducing the order.
 
-Existence of a positive solution.
-This boils down to the question if there is a embedded first order method that ensures positivity.
+Another important property is the stability region in comparison with the maximum $\Delta t$ that ensures positifity.
+For this approach to be useful the RK-method has to yield negative results for $\Delta t$ smaller than the biggest stable $?Delta t$.
+
+(Note: up to this point i do not understand why this is happening. We only see it occur depending on the Proble-RKM combination)
+
+A third property is the existence of a positive solution.
+This corresponds to the question if there is a embedded first order method that ensures positivity.
 This is particularly interesting for Implicit methods, because the $\Delta t$ is not limited by the stability.
-If we have a embedded first Order Method we can ensure that there is always a positive solution, which might be of first Order.
+If we have a embedded first Order Method we can ensure that there is always a positive solution, which might be of first Order. (Todo: Source)
 
 \subsection{Error detection/Approximation}
+(Todo: edit this whole section, Add some mathematically more rigorous statement)
+
 The main question is: We see that for certain $b$ that comply with the Order Conditions we get results that do not resemble the solution. (some type of glitches, artefacts)
@@
 
-\subsection{Stability region}
+\subsection{Stability Region}
 
-By adapting the $b$ we change the used RK Method. This also alters the stability function and therefore changes the region of absolute stability. 
-It depends on the solved problem and the used $\Delta t$ on how many steps the stability function is altered.
+By adapting the $b$ the used RK Method is changed. This also alters the stability function and therefore changes the region of absolute stability. 
+It depends on the solved problem, the RKM and the used $\Delta t$ on how many steps the stability function is altered.
 For Problems where only a small number of step are affected a bigger change in the stability function would be acceptable.
@@
 
-(Note:They do not behave like a Vector space because there is no scalar Product, and all the vector have to add up to 1... Is there something simmilar so we don not have to explain everything here?)
+\subsubsection{Convex Combination}
+If the new $b$ is a chosen as a convex combination of $b$ it is easy to prove that the resulting  the Region of absolute stability is at least the Intersection of the region of absolute stability of the methods that define the convex combination.
+(Compare Section\,\ref{proof:convex_comb}).
+This also means that if all the used embedded methods used to construct the new $b$ are A-Stable, the resulting method is also A-Stable.
 
-If the new $b$ is a chosen as a convex combination of $b$ it is easy to prove that the resulting  the Region of absolute stability is at least the Intersection of the region of absolute stability of the methods that define the convex combination.
-
+\subsection{Direct Adaptation}
 If the $b$ is adapted directly there is no such proof.
@@
 
-Maybe place here some results on the stability region in dependence of the $b$
-(Note: I have some proof handy that states that for certain $A$s the boundary of the stability region does only change in regard to a changing b in a limited way.
-But the proof is a bit convoluted.
-I have to look if there is a more compact way to prove it)
+In Section\,\ref{sec:deriv_sr_b} a proof is given that for most methods a small change of $b$ does not alter the stability function dramatically.
 
@@
 
-(Note: It would also be possible to use another strategy for traversing all the combinations of $(p,\theta)$. Is it worth exploring these?)
+(Note: It would also be possible to use another strategy for traversing all the combinations of $(p,\theta)$. Is it worth exploring these? Eg, a visect algorithm for $\Theta$ or linear approximations)
 
@@
 
-We want to show that $|R_{b_a}(z)|  < 1 ^ |R_{b_b}(z)| < 1\Rightarrow |R_{\chi b_a +(1- \chi) b_b}(z)| < 1$ with $\chi \in [0,1]$.
+We want to show that $(|R_{b_a}(z)|  < 1) (|R_{b_b}(z)| < 1) \Rightarrow |R_{\chi b_a +(1- \chi) b_b}(z)| < 1$ with $\chi \in [0,1]$.
 We write $\alpha = \chi$ and $\beta = 1-\chi$ for convenience.
@@
 
-\subsection{Stability of direct adaptation}
+\subsection{Stability of direct adaptation}\label{sec:deriv_sr_b}
 
@@
 The stability function depending on the $b$ is expressed (Note: it is a bit complicated her because we cannot let it depend on the b directly, but only give b after a coordinate transformation, which also makes the expression easier because we can already introduce the order conditions here)
-$m$ is the number of dimensions of the Kernel of the Order Condition $m = dim\{Ker Q \}$
-The $b_1,\cdots,b_m$  are the Basis vectors of $Ker Q$. By this we can get the vector $w$ by $w = \left[b_1,\cdots,b_m\right]^{-1}(b-b_{orig})$.
+$m$ is the number of dimensions of the Kernel of the Order Condition $m = \mathrm{dim}\{\mathrm{Ker} (Q) \}$
+The $b_1,\cdots,b_m$  are the Basis vectors of $\mathrm{Ker} (Q)$. By this we can get the vector $w$ by $w = \left[b_1,\cdots,b_m\right]^{-1}(b-b_{orig})$.
 
@@
 \end{equation}
-This function can be vied either as $\mathbb{R}^m \times  \mathbb{C} \rightarrow \mathbb{R}$ or $\mathbb{R}^m \times  \mathbb{R}^2 \rightarrow \mathbb{R}$.
+This function can be viewed either as $\mathbb{R}^m \times  \mathbb{C} \rightarrow \mathbb{R}$ or $\mathbb{R}^m \times  \mathbb{R}^2 \rightarrow \mathbb{R}$.
 We are only concerned with the expansion or contraction. Because of this we are only interested in the change perpendicular to the Border of the stability region.
@@
  $|R(w,u,v)|^2$ is calculated. 
-If the gradient does not vanish at $z = z_0$ Equation\,\ref{eq:border} Can be transformed in a function $\mathbb{R} \rightarrow \mathbb{R}$ depending on a single variable $a$ by setting $z = z_0 + n a$.
+If the gradient does not vanish at $z = z_0$ Equation\,\ref{eq:border} Can be transformed in a function $\mathbb{R}^m \times \mathbb{R} \rightarrow \mathbb{R}$ depending on $w$ and a single variable $a$ by setting $z = z_0 + n a$.
 The complex number $n$ denotes the normal vector $\vec{n} = \frac{\nabla |R(w,u,v))|^2}{\left| \nabla |R(w,u,v))|^2 \right|}$ of the border of the stability region written in complex notation.
 This defines the function $a(w)$ with $|R(w,z_0 + n a(w))|^2 -1 = 0$.
-This gives the condition that this approach is only applicable if the gradient does not vanish on the border of the stability function. This is not true in general but for the most common RKM.
-The derivative $\frac{\mathrm d}{\mathrm d w} (a(w))$ can be calculated using the implicit function theorem.
-The defining function is $f: \mathbb{R} \rightarrow \mathbb{R}$ $f(a) = |R(b,z_0 + n a(w))|^2 -1 $
+This implies that this approach is only applicable if the gradient does not vanish on the border of the stability function. This is not true in general but for the most common RKM.
+The derivative $\frac{\mathrm d}{\mathrm d w} (a(w))$ can be calculated using the implicit function theorem. The defining function is 
+\begin{equation}\label{eq:f(w,a)}
+f: \mathbb{R}^m \times \mathbb{R} \rightarrow \mathbb{R} f(a) = |R(w,z_0 + n a)|^2 -1 
+\end{equation}
+
 If explicit methods are used this function is known to be continuously differentiable because it can be written as a Polynomial of $w_1,\cdots,w_m,u$ and $v$.
@@
 
-The Jacobian $J_{f,a}(w,a(w))$ evaluated at $w=0$ can be calculated using the directional derivative 
-$$ J_{f,a}(b_{orig},a(w)) \Big|_{w=0} = 
+The Jacobian of equation\,\ref{eq:f(w,a)} $J_{f,a}(w,a(w))$ evaluated at $w=0$ can be calculated using the directional derivative
+\begin{align*}
+ J_{f,a}(b_{orig},a(w)) \Big|_{w=0} =& 
  \frac{\partial f}{\partial a} (w,a(w)) \Big|_{w=0} = 
- \nabla |R(w,a(w)|^2 \vec{n} \Big|_{w=0}
-= \left| \nabla|R(0,z_0)|^2 \right| = \left| \nabla|R_{b_{orig}}(z_0)|^2 \right|$$
+ \nabla |R(w,a(w)|^2 \vec{n} \Big|_{w=0} \\
+=& \left| \nabla|R(0,z_0)|^2 \right| = \left| \nabla|R_{b_{orig}}(z_0)|^2 \right|
+\end{align*}
 
@@
 this simplifies to 
-\begin{equation}\label{eq:derivative_to_b}
-- 2 \underbrace{R_{b_{orig}}(z_0)R^*_{b_{orig}}(z)}_{=|R_{b_{orig}}(z_0)|^2=1} + R_{b_{orig}}(z_0)R^*_{b_v}(z) + R_{b_v}(z_0)R^*_{b_{orig}}(z_0) 
-= -2 + R_{b_{orig}}(z_0)R^*_{b_v}(z_0) + R_{b_v}(z_0)R^*_{b_{orig}}(z_0)
-\end{equation}
+\begin{align*}\label{eq:derivative_to_b}
+ \frac{\partial f}{\partial w_j}(w,a(w)) \Big|_{w=0} =&
+- 2 \underbrace{R_{b_{orig}}(z_0)R^*_{b_{orig}}(z_0)}_{=|R_{b_{orig}}(z_0)|^2=1} + R_{b_{orig}}(z_0)R^*_{b_v}(z) + R_{b_v}(z_0)R^*_{b_{orig}}(z_0) \\
+=& -2 + R_{b_{orig}}(z_0)R^*_{b_v}(z_0) + R_{b_v}(z_0)R^*_{b_{orig}}(z_0)
+\end{align*}
 using the fact that $z_0$ is on the border of the stability region.
@@
 The same process can be done for all points on the border of the stability region. 
-As long as $\left[ J_{f,a}(w,a(w))  \right] ^{-1} > 0$ the derivative $\frac{\partial a}{\partial w_j} (w)$ is finite.
+As long as $\left[ J_{f,a}(w,a(w))  \right] > 0$ the derivative $\frac{\partial a}{\partial w_j} (w)$ is finite.
 This means that a small change the $b$ only leads to a small change of the stability region.
 
+\printbibliography
+
 \end{document}

### 2020-01-29 RKimple.py
Tests with stepsize control
Implementation and some experiments on PI-Stepsize control. 
Used with explicit RKM and different problems

RKimple.py 
Added code for stepsize controll, also included some change for a better use of the dense output formulas

Stepsize_control.ipynb
Nodebook to test different stepsize controls on different Problems

@@ -10,7 +10,7 @@ from scipy.optimize import linprog
 
 
 
-def solve_LP(solver,O,rhs,rkm,u,K,dt,reduce = False,verbose_LP = False,minval = 0,maxval=np.infty, **options): 
+def solve_LP(solver,O,rhs,rkm,b_orig,u,K,dt,reduce = False,verbose_LP = False,minval = 0,maxval=np.infty, **options): 
     """
     This method solves the LP Problem
     
@@ -19,7 +19,8 @@ def solve_LP(solver,O,rhs,rkm,u,K,dt,reduce = False,verbose_LP = False,minval = 
     solver:     a string with the solver that should be used
     O:          Order condition matrix
     rhs:        Order condition right hand side
-    
+    rkm:        RKM
+    b_orig:     Original b
     reduce:     If set, the LP problem is first solved with a reduced set of constriants
     verbose_LP: prints additional messages
     minval:     Vector or scalar, is the minimum to enforce for solution. If not needed set to -inf
@@ -41,7 +42,7 @@ def solve_LP(solver,O,rhs,rkm,u,K,dt,reduce = False,verbose_LP = False,minval = 
         5 : solver crashed
         6 : Trivial Problem
     l = Array with number of constraints
-    b: found b, if solver failed rkm.b
+    b: found b, if solver failed b_origg
     
     """
     s = len(rkm.b)
@@ -62,7 +63,7 @@ def solve_LP(solver,O,rhs,rkm,u,K,dt,reduce = False,verbose_LP = False,minval = 
     i_min = np.zeros(k,bool)
     i_max = np.zeros(k,bool)
     if reduce: #We do not need it otherwise
-        u_ = u +dt*K@rkm.b
+        u_ = u +dt*K@b_orig
 
 
     while True: #We check at the end for break conditions
@@ -78,7 +79,7 @@ def solve_LP(solver,O,rhs,rkm,u,K,dt,reduce = False,verbose_LP = False,minval = 
 
         if not np.any(i_min) and not np.any(i_max):
             print('trivial Problem')
-            return (0,0,rkm.b)
+            return (0,0,b_orig)
         
         
         l.append(np.sum(i_min)+np.sum(i_max))
@@ -99,7 +100,7 @@ def solve_LP(solver,O,rhs,rkm,u,K,dt,reduce = False,verbose_LP = False,minval = 
                 method = 'revised simplex'
 
             A_eq = np.concatenate((O,-O),axis = 1)
[…]

### 2020-02-02 latex/overview.tex
Update overview.tex
Cleaned up text and corrected some minor mistakes. 
Added an example to motivate the need for a test on the change.

@@
 
-\usepackage[backend=biber]{biblatex}
+\usepackage[backend=biber,url  = false]{biblatex}
 \addbibresource{pos_rk_stephan.bib}
@@
 \maketitle
+
+\section{Introduction}
 
@@
 Other Approaches:
+
 ssp104 methods: The idea is to construct a rkm Method out of explicit euler steps, that lead to a positive result for small enough $\Delta t$.
@@
 
-\section{Main Idea}
+(TODO: Outline of paper)
+
+\section{Main Idea}\label{sec:main_idea}
 
@@
 
-This approach to ensure positifity preserves linear invariants like mass conservation, because it is a regular RKM. 
+This approach preserves linear invariants like mass conservation, because it is a regular RKM. 
 If there are no negative values in $u^{n+1}$ with the original $b$, then the method does not cause additional effort, because it only has to solve the LP-Problem if the regular RKM does not lead to a positive $u^{n+1}$.
@@
 
-\section{Applicable Problems}
+\section{Applicable Problems}\label{sec:app_problem}
 The approach can be used with ODE systems $u' = f(t,u)$ that ensure  $u(t) \geq 0 \forall_t \forall_{  u_0 \geq 0}$
@@
 
-For problems where only the combination of $f(t,u)$ and $u_0$ leads to the positifity of the exact solution, tests did not show promising results.
+For problems where the exact solution is positive for certain $u_0$ but positifity is not preserved for all $u_0 \geq 0$ tests did not show promising results.
 Additional it is not certain if the computed solutions would be reasonable.
 
-\section{Details}
+\section{Details}\label{sec:Deta	ils}
 
 \subsection{Positifity}
+
+(Note: Probably it is better to move this section further to the back)
 
@@
  u_i^{n+1} &\geq 0   \;   \forall_{i \in \{1, \cdots,m \}}  \\
- u_i^n + h \sum_{j=0}^{s-1} b_j h_i^j  &\geq 0   \;   \forall_{i \in \{1,\cdots,m \}}  
+ u_i^n + h \sum_{j=0}^{s-1} b_j k_i^j  &\geq 0   \;   \forall_{i \in \{1,\cdots,m \}}  
 \end{align}
@@
 \begin{equation}
-u_i^n + \Delta b \sum_{j=0}^{s-1} b_j h_i^j  \geq 0   \;   \forall_{i \in h \subseteq \{1,\cdot,m \}} 
+u_i^n + \Delta b \sum_{j=0}^{s-1} b_j k_i^j  \geq 0   \;   \forall_{i \in h \subseteq \{1,\cdot,m \}} 
 \end{equation}
@@
 A reasonable approach is to set $h_0 = \{ i \in \{1,\cdots,m \} |  u_i^{n+1}  < 0 \}$. 
-After the new $b$ and $u^{n+1}$ is computed using $h_0$, the algorithm has to check if the positifity of the other values $i \notin h_0 \subseteq \{1,\cdots,m \}$ is still meet. 
+After the new $b$ and $u^{n+1}$ is computed using $h_0$, the algorithm has to check if the positifity of the other values $u_i | i \notin h_0 \subseteq \{1,\cdots,m \}$ is still meet. 
 If the conditions is not meet a new set $h_{a+1} = \{ i \in \{1,\cdots,m \}|  u_i^{n+1}  < 0 \} \cup h_{a}$ using the new $u^{n+1}$ is generated and the LP-Problem is solved again.
@@
 This effect did not occur so far. 
-A approach on this would be to make a educated guess which indecies could get a problem with positifity.
-A possible estimate would be $\frac{u_i^{n+1}}{max(K_{(i,0)}, \cdots ,K_{(i,0)})} $
+(Note: A approach on this would be to make a educated guess which indecies could get a problem with positifity. A possible estimate would be $\frac{u_i^{n+1}}{max(K_{(i,0)}, \cdots ,K_{(i,0)})} $ )
 
@@
 This is the easiest way to adapt the $b$. An advantageous property is that the step size can be reduced after calculation the stage values by using the dense output order conditions.  
-(Note: Recommendation for explicit methods)
+This approach suits best for explicit methods.
 The direct adaption approach is explained in more detail in section\,\ref{sec:direct}.
@@
 This approach also shows more  predictable behavior because the resulting $u^{n+1}_b$ is also a convex combination of the solutions $u^{n+1}_{b_a},u^{n+1}_{b_b},\cdots$ 
-(Note: Recommendation for implicit methods)
+This approach works best with implicit methods.
 The direct adaption approach is explained in more detail in section\,\ref{sec:convex}
@@
 The weights $b = b_{orig} + \Delta b$ consist of the weights of the original method $b_{orig}$ and an adaption $\Delta b$. This $\Delta b \in \mathrm{ker}(Q)$. 
-In Figure\,\ref{fig:b_direct} this is is drawn. The red line represents the subspace of $b$ that compile with the Order Conditions. Both $b_{orig}$ and $b$ are in this subspace.
+In Figure\,\ref{fig:b_direct} this is drawn. The red line represents the subspace of $b$ that satisfy the Order Conditions. Both $b_{orig}$ and $b$ are in this subspace.
 
@@
 The magnitude of $b$ can get large if the Order Conditions are used as objective function.
-Large values of b can lead to numerical instability.
+Large values of b can lead to numerical problems.
 Because $\sum_{i  = 1}^s b_i = 1$, a increase of $||b||$ leads to a combination of negative and positive values.
@@
 By changing the $b$ the stability function is altered. This can lead to oscillations if the poles of the RHS are no longer in the stability region of the resulting method.
-For these reasons $||b-b_{orig}||$ seems like a good choice. The 1-Norm can be implement the with Linear Programming using slack variables. We choose $||b_{optim}-b_{orig}||_1$ as objective function.
+For these reasons $||b-b_{orig}||$ seems like a good choice. The 1-Norm can be implement the with Linear Programming using slack variables. We choose
+
+\begin{equation}
+f_{obj} = ||b_{optim}-b_{orig}||_1
+\end{equation}
+ as objective function.
 
@@
 \subsection{Convex combination of existing methods}\label{sec:convex}
-For some test problems (e.g. the Heat Equation solved with an implicit method), large distortions occur, even though a norm of the distortion of the b is used as a objective function. 
-For implicit methods it is known that implicit euler steps lead to a positive solution for any $\Delta t$.
-The backward Euler is still only first Order but shows more exact results in certain cases than a method of higher Order that shows distortions of the solution caused by the larger error coefficients ($\Delta t \gg 0$, so they do not disappear because of the nature of the polynomial).
+For some test problems (e.g. the Heat Equation solved with an implicit method of higher order), large distortions occur, even though a norm of the distortion of the b is used as a objective function. 
+For implicit methods it is known that the backward euler steps lead to a positive solution for any $\Delta t$ \cite{hundsdorfer_numerical_2003}.
+The backward Euler is still only first Order but shows more exact results than a method of higher Order that is distorted by the adaption of the $b$, at least at some steps.
 A possible method to fix this is to generate the new b based on a convex combination of embedded methods of RKM.
@@
 At first a set of embedded methods of the RKM $b_1,\cdots,b_d$ are needed. 
-If all the used b complies with an order conditions the new method also has this order.
-This can is straightforward using
+If all the used $b$ compile with the order conditions the new method is also of the same Order.
+This is straightforward using
 
-$$ q^T b = q^T (\chi b_{\alpha} + (1-\chi) b_{\beta}) = \chi q^T  b_{\alpha} + (1-\chi) q^T  b_{\beta}) = \chi rhs + (1-\chi) rhs = rhs$$
+$$ Q b = Q (\chi b_{\alpha} + (1-\chi) b_{\beta}) = \chi Q  b_{\alpha} + (1-\chi) Q b_{\beta}) = \chi r + (1-\chi) r = r$$
+
+$b_{\alpha}$ and $b_{\beta}$ are the methods that are combined to $b$ using the parameter $\chi$. The Matrix $Q$ and the vector $r$ are the Order Conditions as defined in Section\,\ref{sec:OrderCond}.
 
@@
 \begin{equation}
- \sum_{i=1}^d a = 1
+ \sum_{i=1}^d a_i = 1
 \end{equation}
@@
 
-$$u_{n+1}=u_n+\Delta K B a \geq 0$$
+$$u_{n+1}=u_n+\Delta t K B a \geq 0$$
+
+where $a = [a_1,\cdots,a_d]^T$
 
@@
 
-$$ min \left(\sum_i^d w_i a_i \right) $$
+\begin{equation}
+f_{obj} = min \left(\sum_i^d w_i a_i \right)
+\end{equation}
 
@@
 
-As a first order method a $b$ is added that corresponds to a series of backward euler steps. These are know to be yield to a positive result (Todo: source).
+As a first order method a $b$ is added that corresponds to a series of backward euler steps. These are know to be yield to a positive result.
 
@@
 Another important property is the stability region in comparison with the maximum $\Delta t$ that ensures positifity.
-For this approach to be useful the RK-method has to yield negative results for $\Delta t$ smaller than the biggest stable $?Delta t$.
+For this approach to be useful the RK-method has to yield negative results for $\Delta t$ smaller than the biggest stable $\Delta t$.
 
@@
 
-\subsection{Error detection/Approximation}
-(Todo: edit this whole section, Add some mathematically more rigorous statement)
+\subsection{Error detection and Approximation}
+When changing the weights the solution $u^{n+1}$ is changed. 
+By incorporating the Order Conditions for a certain Order the space of $b$ that can be choose is limited to a linear Subspace. 
+Now it is important to know if the new solution $u^{n+1}$ still resembles the solution of the ODE $u(t_{n+1})$. 
+We can motivate this with the following example:
 
-The main question is: We see that for certain $b$ that comply with the Order Conditions we get results that do not resemble the solution. (some type of glitches, artefacts)
+Take the linear, positifity preserving ODE from \cite{kopecz_unconditionally_2018} with $u = [u_1,u_2]^T$ with the initial conditions $u_0 = [1,0]^T$ and
 
-Note: in general it is difficult to give some mathematically rigorous statement for this statement because -by definition- it only applies for $\delta t \gg 0$ where the proofs using the Taylor Series do not give valid results any more.
-We could say that the main idea of the adaption of $b$ is to alter the error in a certain way that we know that the numerical solution gets 'better' in the sense that it only leads to positive solutions.
-When taking the Taylor Series of the numeric solution the coefficients up to the Order of the method are fixed by the Order Conditions.
-The only degrees of freedom are the coefficents above the Order Conditions.
-This requires that $\Delta t > 0$.
-(More like $\Delta t^n \approx \Delta t^(n+1) $)
-The main way is to ensure that for $\Delta t \to 0$ : $b \to b_{orig}$. For this case we already know that the numeric solution converges to the exact solution.  
-This is made sure by property of the Objective function that $argmin(f_{optim}(b)) = b_{orig}$.
+\begin{equation}
+u'(t) = A u(t) \qquad A = \left[\begin{matrix}- a & 1\\a & -1\end{matrix}\right] \quad a =5
+\end{equation}
 
-To make sure that the new adapted solution is still reasonable we measure the deviation form the Original method.
+and the first order RKM with two stages
+
+\begin{align}
+\begin{array}{c|cc}
+0 &  & \\
+\frac{1}{2} & \frac{1}{2} & \\
+\hline
+ & \frac{1}{2} & \frac{1}{2}\\
+\end{array}
+\end{align}
+
+If simulated with $\Delta t = 1$ the stage values are
+
+\begin{equation}
+k^1 = \left(\begin{matrix}-5\\5\end{matrix}\right) \qquad k^2 = \left(\begin{matrix}10\\-10\end{matrix}\right)
+\end{equation}
+
+All $b$ that comply with the order Condition for the first order can be expressed as
+
+\begin{equation}
+b= \left(\begin{matrix}\frac{1}{2}\\ \frac{1}{2}\end{matrix}\right) + \alpha \left(\begin{matrix}1\\ -1\end{matrix}\right)
+\end{equation}
+
+Therefor the general solution is
+
+\begin{equation}
+u^{1} = u^0 + \Delta t  \left[k^1,k^2\right] b = \left(\begin{matrix}\frac{5}{2}\\- \frac{5}{2}\end{matrix}\right) \alpha  \left(\begin{matrix}-15\\15\end{matrix}\right)
+\end{equation}
+
+By a suitable choice of $\alpha \in [\frac{1}{6},\frac{7}{30}]$ any $u$ that complies with mass conservation and positifity can be reached.
+
+This gives raise to the question weather the new solution is reasonable.
+This also poses the problem on developing a system that can decide if the new solution is has to be rejected.
+
+\subsubsection{Error Detection}
+For this approach is difficult to give some mathematically rigorous statement regarding the convergence of the method, because it only applies if the error of the original method causes the solution to get negative.
+This can only happen if the Truncation error is large enough.
+This requires $\Delta t \gg 0$ where lower coefficients of the Taylor Series are no longer sufficient to approximate the behavior.
+One could say that the main idea of the adaption of $b$ is to alter the error in a certain way, so that the numerical solution gets 'better' in the sense that it leads to positive solutions. 
+The local error of a RKM $e(\Delta t) =u(t_0 + \Delta t) - u^1$ can be expressed using the Taylor series of the RKM and the Taylor series of the exact solution. % Hairer I P134
+
+\begin{align}\label{eq:Taylor_sol_ref}
+u(t_0 + \Delta t) &= u(t_0) + u'(t_0) \Delta t + \frac{u''(t_0)}{2} \Delta t^2 + \cdots + \frac{u^{(p)}(t_0)}{p!} \Delta t^p + \frac{u^{(p+1)}(t_0)}{(p+1)!} \Delta t^{p+1} + \cdots \\
+u^{n+1} &= u(t_0)  + v_1 \Delta t + \frac{v_2}{2} \Delta t^2 + \cdots + \frac{v_p}{p!} \Delta t^p + \frac{v_{p+1}}{(p+1)!} \Delta t^{p+1} + \cdots \\
+\end{align}
+
+The derivatives of $u(t)$ can be solely expressed with the partial derivatives of the RHS $f(t,u)$, whereas the coefficients $v_1,v_2,\cdots$ also depend on the RKM. 
+In this case the RKM is fixed apart from the weights $b$. 
+Because of this $v_1,v_2,\cdots$ are functions of $b$.
+Because the $b$ are chosen to ensure the Order Conditions the coefficients $v_1,\cdots,v_p$ are fixed and equal the according derivatives. 
+The only remaining degrees of freedom are the coefficients $v_{p+1},v_{p+2},\cdots$.
+These can be adapted by changing the $b$.
+
+The main way is to ensure that the RKM with adapted b i to ensure that the $b$ approach $b_{orig}$ for $\Delta t \to 0$.
+For this $b$ it is already know that the numeric solution converges to the exact solution.
+This is made sure by property of the Objective function that
+
+\begin{equation}
+\mathrm{argmin}(f_{optim}(b)) = b_{orig}
+\end{equation}
+
+This approximation is not very useful for real applications. 
+To approximate the error of a new step we propose following approxiamtion of the local error:
+
+\begin{align}
+err = |u(t^n)-u^n| &= |u(t^n) - (u^n_{b_{orig}+\Delta t K(b-b_{orig})})| \\
+ &\leq \underbrace{|u(t^n)-u^n_{b_{orig}}|}_{\approx err_T}+\underbrace{|\Delta t K(b-b_{orig})|}_{= err_{adapt}}
+\end{align}
+
+After adapting the $b$ the Approximation Error is checked. If $|\Delta t K(b-b_{orig})|$ is larger than the tolerance the $b$ is requected.
+
+\subsubsection{Steppsize Control}
+An important part of an RKM method is the ability to approximate the error of the solution to use it to control the step size. 
+The truncation error $|u(t^n)-u^n_{b_{orig}}|$ is usually approximated with an standard error estimator $err_T = | u^{n}_{b_{orig}} - u^{n}_{\hat{b}} |$.
+
+Both errors are added to get an approximation of the total error $err = err_T +w_a err_{adapt}$.
+The parameter $w_a \leq 1$ is used to account for the fact that the real error is smaller than $err = err_T +err_{adapt}$.
+The factor also improves the stability of the step size control because $err_{adapt}$ is not as smooth as $err_T$, especially if negative values only occur for a small number of steps.
+If the LP-Problem was infeasible and no $b$ was found the $err_{approx}$ is set to a custom value.
+This value should be chosen sufficiently large to ensure that the next step take is smaller, but still not to large to keep the stepsize control stable.
+
+This type of error estimation is easy to implement because it can be easily incorporated in an existing step size control and takes advantage of the standard error approximation.
+
+%To make sure that the new adapted solution is still reasonable we measure the deviation form the Original method.
 In the worst case the errors add up.
@@
 
-(Todo:It is not straightforward if there are no pole zero cancellations in the stability function.
+%(Todo:It is not straightforward if there are no pole zero cancellations in the stability function.
 For explicit methods the stability function is a polynomial o degree $\leq s$ (Hundsdorfer).
@@
 \subsubsection{Convex Combination}
-If the new $b$ is a chosen as a convex combination of $b$ it is easy to prove that the resulting  the Region of absolute stability is at least the Intersection of the region of absolute stability of the methods that define the convex combination.
+If the new $b$ is chosen as a convex combination of $b$ it is easy to prove that the resulting  the Region of absolute stability is at least the Intersection of the region of absolute stability of the methods that define the convex combination.
 (Compare Section\,\ref{proof:convex_comb}).
@@
 
-\subsection{Direct Adaptation}
+\subsubsection{Direct Adaptation}
 If the $b$ is adapted directly there is no such proof.
@@
 
+\section{Results of experiments}\label{sec:Numeric_Results}
+
 (Note: Is here a good place to give a note on where supplementary code can be found?)
@@
 
-\section{Results of experiments}
 \subsection{Explicit methods}
 For most of the test problems problems with stability occur before getting negative values. This means that we need methods with large stability region.
-
-\subsubsection{ADP}
-A test problem that works good is the advection‐diffusion‐production‐destruction systems. 
-For the production‐destruction systems negative values at the intermediate steps occurred.
-These results of these could be suppressed by the optimization problem  except for some timesteps where these yield to glitches in the solution.
-
-Conclusion: it is important to control for negative Values in the intermediate stages.
-Possible solutions: Solve an positifity ensuring optimization Problem at every intermediate step.
-Can be formulated as choosing a new $b$ for every intermediate stage which then gets part of $A$ in the next step.
-
-Another idea to explore (Is kind of obvious but could also be stupid): Replace $f(t,u)$ with an $f_{pos}(t,u)$ with 
-$f(t,u)=f_{pos}(t,u) \forall u \geq 0 $
-Whereas $f_{pos}(t,u)$ is welldefined for $u \leq 0$
-
-We are going to test the solvers for multiple $\Delta t$. For the explicit method we want to show the important times.
-
-$$ \Delta t \; \underset{0}{+}---------\underset{\Delta t_{pos}}{+}\underbrace{--------}_{\gamma}\underset{\Delta t_{feasible}}{+}---\underset{\Delta t_{stable}}{+}--->  $$
-
-\begin{itemize}
-\item
-$\Delta t_{pos}$ is the largest stepsize that yields to positive results without adapting the $b$
-
-\item
-$\Delta t_{feasible}$ is the largest stepsize that can be used with the adaptive $b$
-
-\item
-$\Delta t_{stable}$ is the largest stepsize for that the method is stable
-\end{itemize}
-
-The timesteps where the adaption of the b is usefull is denoted with $\gamma$ idealy this interval is as large as possible.
-This means that $\Delta t_{pos} \ll \Delta t_{feasible} \approx \Delta t_{stable}$
-
-With stepsizes $\Delta t > \Delta t_{stable} $ we cann not expect the solution to be reasonable
-
-$\Delta t$ for the ADP Problem
-
-\begin{tabular}{|c|c|c|c|c|}
-\hline 
-Method & $\Delta t_{pos}$ &$\Delta t_{stable}$& $\Delta t_{feasible}$  \\ 
-\hline
-fe & 0.0015 & --- &  0.0051 \\
-\hline 
-ssp104 fallback =1 & 0.0134 & 0.0309 & 0.0311 \\ 
-\hline 
-dp5 &0.0031 &0.0031 & 0.0031\\ 
-\hline 
-ck5 & 0.0044&  0.0044 & 0.0044\\ 
-\hline 
-merson4 fallback= 1& 0.0038& 0.0055 & 0.0084\\
-\hline
-\end{tabular}
-
-We can see that for the ssp104 method is a good choice for this approach when solving this testproblem.  The merson4 method is also suitable. We can not get benefits form dp5 and ck5.
-
-\subsubsection{Quantum Simulations}
-Shows following shortcomings:
-
-To ensure positifity we do not only need $\rho_{i,i} \geq 0$ but also the additional constraint $|\rho_{j,k}|^2 \leq \rho_{j,j} \rho_{k,k}$ at least for the elements with $j = k +1$ and $j = k-1$.
-This conditions can not be formulated as LP-Problem.
 
@@
 
-\begin{itemize}
-\item The computational effort of solving the optimisation Problem is relatively small compared to the overall cost of the method.
-\item Implicit methods are stable for large step sizes.
-\item Implicit methods ensure positifity for the backward euler step, even for large step sizes.
-\item There are no methods for ensuring positifity for higher order methods (Which we also can not get around for the general case)
-\end{itemize}
+When tested on the diffusion equation with an spike as initial data the $b$ is only changed for the first step. The change of the $b$ can lead to reasonable changes of the hight of the maximum.
 
-Used for the ADP model the model yields good results, the main limit for the stepsize is the accuracy of the solver for the stageequations.
+\subsection{Stepsize control}
 
-When tested on the diffusion equation with an spike as initial data the $b$ is only changed for the first step. The change of the $b$ can lead to reasonable changes of the hight of the maximum.
+\subsection{Dense Output}
 

### 2020-02-09 latex/overview.tex
New structure of Paper
Changed the order of certain sections in the paper to make it more read- and understandable

@@
 
-\section{Applicable Problems}\label{sec:app_problem}
-The approach can be used with ODE systems $u' = f(t,u)$ that ensure  $u(t) \geq 0 \forall_t \forall_{  u_0 \geq 0}$
+\subsection{Example I}\label{sec:example_reac}
 
-This can be tested with the property 
+TODO: Add here the reaction equation, show that it behaves better with the new approach. (still keep it simple with fixed stepsize...)
+
+\subsection{Example II}\label{sec:example_lin}
+
+(Note: The primary goal here is to put the idea of changing the b into perspective.
+We have shown the general idea and how it works, now we want to discuss on what it does in practice.
+At first we state it is good to preserve the order...
+Get to the Example ...
+and show that there is more to it)
+
+By incorporating the Order Conditions for a certain Order the space of $b$ that can be choose is limited to a linear Subspace. 
+Now it is important to know if the new solution $u^{n+1}$ still resembles the solution of the ODE $u(t_{n+1})$. 
+We can motivate this with the following example:
+
+Take the linear, positifity preserving ODE from \cite{kopecz_unconditionally_2018} with $u = [u_1,u_2]^T$ with the initial conditions $u_0 = [1,0]^T$ and
+
 \begin{equation}
-u_i=0 \Rightarrow f_i(t,u_1,\cdots,u_i,\cdots,u_n) \geq 0  \forall_{u_c \geq 0} \forall_{t}
+u'(t) = A u(t) \qquad A = \left[\begin{matrix}- a & 1\\a & -1\end{matrix}\right] \quad a =5
 \end{equation}
 
-For problems where the exact solution is positive for certain $u_0$ but positifity is not preserved for all $u_0 \geq 0$ tests did not show promising results.
-Additional it is not certain if the computed solutions would be reasonable.
+and the first order RKM with two stages
+
+\begin{align}
+\begin{array}{c|cc}
+0 &  & \\
+\frac{1}{2} & \frac{1}{2} & \\
+\hline
+ & \frac{1}{2} & \frac{1}{2}\\
+\end{array}
+\end{align}
+
+If simulated with $\Delta t = 1$ the stage values are
+
+\begin{equation}
+k^1 = \left(\begin{matrix}-5\\5\end{matrix}\right) \qquad k^2 = \left(\begin{matrix}10\\-10\end{matrix}\right)
+\end{equation}
+
+All $b$ that comply with the order Condition for the first order can be expressed as
+
+\begin{equation}
+b= \left(\begin{matrix}\frac{1}{2}\\ \frac{1}{2}\end{matrix}\right) + \alpha \left(\begin{matrix}1\\ -1\end{matrix}\right)
+\end{equation}
+
+Therefor the general solution is
+
+\begin{equation}
+u^{1} = u^0 + \Delta t  \left[k^1,k^2\right] b = \left(\begin{matrix}\frac{5}{2}\\- \frac{5}{2}\end{matrix}\right) \alpha  \left(\begin{matrix}-15\\15\end{matrix}\right)
+\end{equation}
+
+By a suitable choice of $\alpha \in [\frac{1}{6},\frac{7}{30}]$ any $u$ that complies with mass conservation and positifity can be reached.
+
+This gives raise to the question weather the new solution is reasonable.
+This also poses the problem on developing a system that can decide if the new solution is has to be rejected.
 
 \section{Details}\label{sec:Deta	ils}
-
-\subsection{Positifity}
-
-(Note: Probably it is better to move this section further to the back)
-
-For the further discussion we are concerned with a System of $m$ coupled ODEs. 
-In this setup  $u \in \mathbb{R}^m$.
-Systems of ODEs with a high number of dimensions can appear when solving a PDE with the method of lines.
-
-To enshure that the solution is positive
-
-\begin{align}
- u_i^{n+1} &\geq 0   \;   \forall_{i \in \{1, \cdots,m \}}  \\
- u_i^n + h \sum_{j=0}^{s-1} b_j k_i^j  &\geq 0   \;   \forall_{i \in \{1,\cdots,m \}}  
-\end{align}
-
-has to be fulfilled.
-This infers $m$ positivity constraints to the optimisation problem. These can be written as
-
-\begin{equation}
-u_i + \Delta t K  b \geq 0     
-\end{equation}
-
-where $K = \big[k^1 , \cdots k^{s-1}\big]$.
-
-This causes unnecessary many inequality constraints on the LP-Problem.
-To simplify the problem one want to reduce the number of positifity constraints by ignoring it for ODEs where there is no issue with positifity.
-
-The new positifity constraints can be formulated as
-\begin{equation}
-u_i^n + \Delta b \sum_{j=0}^{s-1} b_j k_i^j  \geq 0   \;   \forall_{i \in h \subseteq \{1,\cdot,m \}} 
-\end{equation}
-
-For this a set $h \subseteq \{1,\cdots,m \}$ that include all indecies that are needed for determining the $b$ has to be chosen.
-
-A reasonable approach is to set $h_0 = \{ i \in \{1,\cdots,m \} |  u_i^{n+1}  < 0 \}$. 
-After the new $b$ and $u^{n+1}$ is computed using $h_0$, the algorithm has to check if the positifity of the other values $u_i | i \notin h_0 \subseteq \{1,\cdots,m \}$ is still meet. 
-If the conditions is not meet a new set $h_{a+1} = \{ i \in \{1,\cdots,m \}|  u_i^{n+1}  < 0 \} \cup h_{a}$ using the new $u^{n+1}$ is generated and the LP-Problem is solved again.
-This process id repeated until a solution is found that leads to a $u^{n+1} \geq 0$.
-The choice of $h_{a+1}$ based on the $h_{a}$ ensures that the algorithm cannot get stuck in a loop.
-In the worst case $h_a$ is growing very slowly until $h_a = \{1,\cdots,m \}$.
-
-This effect did not occur so far. 
-(Note: A approach on this would be to make a educated guess which indecies could get a problem with positifity. A possible estimate would be $\frac{u_i^{n+1}}{max(K_{(i,0)}, \cdots ,K_{(i,0)})} $ )
-
-When enforcing a maximum value the number of constraints can be reduced using the same technique.
-When both enforcing maximum and minimum values two separate sets of active constraints $k_{min}$ and $k_{max}$ are used.
-The set $k_{max}$ denotes the active maximum constraints and the set $k_{min}$ denotes the active minimum constraints.
-Here it is important to update these sets simultaneously.
 
@@
 Now it is important to know if the new solution $u^{n+1}$ still resembles the solution of the ODE $u(t_{n+1})$. 
-We can motivate this with the following example:
-
-Take the linear, positifity preserving ODE from \cite{kopecz_unconditionally_2018} with $u = [u_1,u_2]^T$ with the initial conditions $u_0 = [1,0]^T$ and
-
-\begin{equation}
-u'(t) = A u(t) \qquad A = \left[\begin{matrix}- a & 1\\a & -1\end{matrix}\right] \quad a =5
-\end{equation}
-
-and the first order RKM with two stages
-
-\begin{align}
-\begin{array}{c|cc}
-0 &  & \\
-\frac{1}{2} & \frac{1}{2} & \\
-\hline
- & \frac{1}{2} & \frac{1}{2}\\
-\end{array}
-\end{align}
-
-If simulated with $\Delta t = 1$ the stage values are
-
-\begin{equation}
-k^1 = \left(\begin{matrix}-5\\5\end{matrix}\right) \qquad k^2 = \left(\begin{matrix}10\\-10\end{matrix}\right)
-\end{equation}
-
-All $b$ that comply with the order Condition for the first order can be expressed as
-
-\begin{equation}
-b= \left(\begin{matrix}\frac{1}{2}\\ \frac{1}{2}\end{matrix}\right) + \alpha \left(\begin{matrix}1\\ -1\end{matrix}\right)
-\end{equation}
-
-Therefor the general solution is
-
-\begin{equation}
-u^{1} = u^0 + \Delta t  \left[k^1,k^2\right] b = \left(\begin{matrix}\frac{5}{2}\\- \frac{5}{2}\end{matrix}\right) \alpha  \left(\begin{matrix}-15\\15\end{matrix}\right)
-\end{equation}
-
+This can be seen in example in section\,\ref{sec:example_lin}.
 By a suitable choice of $\alpha \in [\frac{1}{6},\frac{7}{30}]$ any $u$ that complies with mass conservation and positifity can be reached.
@@
 
-In \ref{proof:combiningb} it is stated that one can construct the stability function of a RKM  $R_b$  with the weights $b = \alpha b_1 + \beta b_2$ by adding the weighted stability functions $R_{\alpha b_a+\beta b_b}(z) = \alpha R_{\alpha b_a}(z) + \beta R_{\beta b_b}(z) $ where all stability functions have the same $A$ and $\alpha + \beta = 1$.
+The stability function of a RKM  $R_b$  with the weights $b = \alpha b_1 + \beta b_2$ cn be calculated by adding the weighted stability functions $R_{\alpha b_a+\beta b_b}(z) = \alpha R_{\alpha b_a}(z) + \beta R_{\beta b_b}(z) $ where all stability functions have the same $A$ and $\alpha + \beta = 1$.
 
-%(Todo:It is not straightforward if there are no pole zero cancellations in the stability function.
-For explicit methods the stability function is a polynomial o degree $\leq s$ (Hundsdorfer).
-Pole zero cancellations can not happen for these.
-For implicit functions that are A-stable the poles have to be in the right halve plane.
-Here is starts to get interesting.
-Because the A-Stability of one method with a certain $A$ is not a sufficient  condition that there are no Poles for all RKM with this A-Matrix.
-But it is possible (or easy) to prove that if $a_{ii} > 0 \forall i = 1,\cdots , s$ then there are only poles in the right halve plane where they do not matter for the further considerations.)
-
-\subsubsection{Convex Combination}
-If the new $b$ is chosen as a convex combination of $b$ it is easy to prove that the resulting  the Region of absolute stability is at least the Intersection of the region of absolute stability of the methods that define the convex combination.
-(Compare Section\,\ref{proof:convex_comb}).
-This also means that if all the used embedded methods used to construct the new $b$ are A-Stable, the resulting method is also A-Stable.
-
-\subsubsection{Direct Adaptation}
-If the $b$ is adapted directly there is no such proof.
-A useful property of the resulting stability region is that it is similar to the stability region of the baseline method. 
-(Note: is it possible to give some statement like: If the stability regions of the rkm used for multiple steps are similar, then the overall method is stable for the intersection of the stability regions)
-
-In Section\,\ref{sec:deriv_sr_b} a proof is given that for most methods a small change of $b$ does not alter the stability function dramatically.
-
-\section{Implementation aspects of the Algorithm}
-
-(Todo: Maybe a flowchart and a description on how we stitch everything together.)
-
-The algorithm on adapting the b consists of two main loops.
-The outer loop loops through all the $\theta$, staring with $\theta = 1$ down to the lowest $\theta$.
-The inner loop loops through the orders, starting with the highest Order down to the lowest Order.
-Until a acceptable $b$ the algorithm tests combiantions of $\theta$ and $p$.
-At first the Order Conditions are constructed.
-On these a LP-Solver is invoiced.
-If the LP-Problem is infeasible the algorithm immediately tries the next combination.
-If the LP-Problem is feasible we test if the $b$ is acceptable.
-At first it is tested weather the new $u^{n+1}$ indeed complies with the constraints.
-This is necessary because sometimes the used LP-solvers incorrectly identify the problem as feasible.
-(Note: because of the chosen objective function it can not be unbounded) As second test the deviation from the old solution (Compare Error estimation) is calculated.
-If the error is bigger than a set maximum error the $b$ is also rejected and the algorithm also switch to the next combination.
-As soon as the algorithm finds a suitable $b$ the loops are aborted and the $b$ is used for the update step.
-If all the combinations of $\theta$ and $p$ did not lead to a acceptable $b$ the step has to be rejected and a new set of stage values has to bee computed.
-
-(Note: It would also be possible to use another strategy for traversing all the combinations of $(p,\theta)$. Is it worth exploring these? Eg, a visect algorithm for $\Theta$ or linear approximations)
-
-\section{Results of experiments}\label{sec:Numeric_Results}
-
-(Note: Is here a good place to give a note on where supplementary code can be found?)
-
-(Note: Should we discuss which solver is used. Probably it is not worth using space for some plot etc. on this... Maybe a short note?)
-
-\subsection{Explicit methods}
-For most of the test problems problems with stability occur before getting negative values. This means that we need methods with large stability region.
-
-\subsection{Implicit methods}
-Implicit methods seem like an advantageous choice for a couple of reasons.
-
-When tested on the diffusion equation with an spike as initial data the $b$ is only changed for the first step. The change of the $b$ can lead to reasonable changes of the hight of the maximum.
-
-\subsection{Stepsize control}
-
-\subsection{Dense Output}
-
-\section{Useful proofs}
-
-\subsection{Existence of b}
-Ensuring a feasible $b$ exists for a certain order condition is in general very difficult.
-If we drop the order conditions it is known that there is a positive solution as long as there is a $k_c$ with $k_c \geq 0$.
-This is no problem when using a implicit method because a backward euler step always gives a positive solution.
-
-\subsection{Stability function of linear combinations}\label{proof:combiningb}
 Let $R_b(z)$ denote the stability function of a RKM method with the butcher tableau
@@
 
-\subsection{Stability of convex combination}\label{proof:convex_comb}
+%(Todo:It is not straightforward if there are no pole zero cancellations in the stability function.
+For explicit methods the stability function is a polynomial o degree $\leq s$ (Hundsdorfer).
+Pole zero cancellations can not happen for these.
+For implicit functions that are A-stable the poles have to be in the right halve plane.
+Here is starts to get interesting.
+Because the A-Stability of one method with a certain $A$ is not a sufficient  condition that there are no Poles for all RKM with this A-Matrix.
+But it is possible (or easy) to prove that if $a_{ii} > 0 \forall i = 1,\cdots , s$ then there are only poles in the right halve plane where they do not matter for the further considerations.)
+
+\subsubsection{Stability of convex combination}
+If the new $b$ is chosen as a convex combination of $b$ it is easy to prove that the resulting  the Region of absolute stability is at least the Intersection of the region of absolute stability of the methods that define the convex combination.
+(Compare Section\,\ref{proof:convex_comb}).
+This also means that if all the used embedded methods used to construct the new $b$ are A-Stable, the resulting method is also A-Stable.
 
@@
 
-\subsection{Stability of direct adaptation}\label{sec:deriv_sr_b}
+\subsubsection{Stability of  Direct Adaptation}
+If the $b$ is adapted directly there is no such proof.
+A useful property of the resulting stability region is that it is similar to the stability region of the baseline method. 
+(Note: is it possible to give some statement like: If the stability regions of the rkm used for multiple steps are similar, then the overall method is stable for the intersection of the stability regions)
+
+For most methods a small change of $b$ does not alter the stability function dramatically.
 
@@
 
+\section{Implementation aspects of the Algorithm}
+
+TODO: Discuss in this section aspects that are needed for implementation (and improve the usability) but do not concern the mathematical idea of the approach.
+
+\subsection{Positifity}
+
+For the further discussion we are concerned with a System of $m$ coupled ODEs. 
+In this setup  $u \in \mathbb{R}^m$.
+Systems of ODEs with a high number of dimensions can appear when solving a PDE with the method of lines.
+
+To enshure that the solution is positive
+
+\begin{align}
+ u_i^{n+1} &\geq 0   \;   \forall_{i \in \{1, \cdots,m \}}  \\
+ u_i^n + h \sum_{j=0}^{s-1} b_j k_i^j  &\geq 0   \;   \forall_{i \in \{1,\cdots,m \}}  
+\end{align}
+
+has to be fulfilled.
+This infers $m$ positivity constraints to the optimisation problem. These can be written as
+
+\begin{equation}
+u_i + \Delta t K  b \geq 0     
+\end{equation}
+
+where $K = \big[k^1 , \cdots k^{s-1}\big]$.
+
+This causes unnecessary many inequality constraints on the LP-Problem.
+To simplify the problem one want to reduce the number of positifity constraints by ignoring it for ODEs where there is no issue with positifity.
+
+The new positifity constraints can be formulated as
+\begin{equation}
+u_i^n + \Delta b \sum_{j=0}^{s-1} b_j k_i^j  \geq 0   \;   \forall_{i \in h \subseteq \{1,\cdot,m \}} 
+\end{equation}
+
+For this a set $h \subseteq \{1,\cdots,m \}$ that include all indecies that are needed for determining the $b$ has to be chosen.
+
+A reasonable approach is to set $h_0 = \{ i \in \{1,\cdots,m \} |  u_i^{n+1}  < 0 \}$. 
+After the new $b$ and $u^{n+1}$ is computed using $h_0$, the algorithm has to check if the positifity of the other values $u_i | i \notin h_0 \subseteq \{1,\cdots,m \}$ is still meet. 
+If the conditions is not meet a new set $h_{a+1} = \{ i \in \{1,\cdots,m \}|  u_i^{n+1}  < 0 \} \cup h_{a}$ using the new $u^{n+1}$ is generated and the LP-Problem is solved again.
+This process id repeated until a solution is found that leads to a $u^{n+1} \geq 0$.
+The choice of $h_{a+1}$ based on the $h_{a}$ ensures that the algorithm cannot get stuck in a loop.
+In the worst case $h_a$ is growing very slowly until $h_a = \{1,\cdots,m \}$.
+
+This effect did not occur so far. 
+(Note: A approach on this would be to make a educated guess which indecies could get a problem with positifity. A possible estimate would be $\frac{u_i^{n+1}}{max(K_{(i,0)}, \cdots ,K_{(i,0)})} $ )
+
+When enforcing a maximum value the number of constraints can be reduced using the same technique.
+When both enforcing maximum and minimum values two separate sets of active constraints $k_{min}$ and $k_{max}$ are used.
+The set $k_{max}$ denotes the active maximum constraints and the set $k_{min}$ denotes the active minimum constraints.
+Here it is important to update these sets simultaneously.
+
+\subsection{•}
+(Todo: Maybe a flowchart and a description on how we stitch everything together.)
+
+The algorithm on adapting the b consists of two main loops.
+The outer loop loops through all the $\theta$, staring with $\theta = 1$ down to the lowest $\theta$.
+The inner loop loops through the orders, starting with the highest Order down to the lowest Order.
+Until a acceptable $b$ the algorithm tests combiantions of $\theta$ and $p$.
+At first the Order Conditions are constructed.
+On these a LP-Solver is invoiced.
+If the LP-Problem is infeasible the algorithm immediately tries the next combination.
+If the LP-Problem is feasible we test if the $b$ is acceptable.
+At first it is tested weather the new $u^{n+1}$ indeed complies with the constraints.
+This is necessary because sometimes the used LP-solvers incorrectly identify the problem as feasible.
+(Note: because of the chosen objective function it can not be unbounded) As second test the deviation from the old solution (Compare Error estimation) is calculated.
+If the error is bigger than a set maximum error the $b$ is also rejected and the algorithm also switch to the next combination.
+As soon as the algorithm finds a suitable $b$ the loops are aborted and the $b$ is used for the update step.
+If all the combinations of $\theta$ and $p$ did not lead to a acceptable $b$ the step has to be rejected and a new set of stage values has to bee computed.
+
+(Note: It would also be possible to use another strategy for traversing all the combinations of $(p,\theta)$. Is it worth exploring these? Eg, a visect algorithm for $\Theta$ or linear approximations)
+
+\section{Results of experiments}\label{sec:Numeric_Results}
+
+(Note: Is here a good place to give a note on where supplementary code can be found?)
+
+(Note: Should we discuss which solver is used. Probably it is not worth using space for some plot etc. on this... Maybe a short note?)
+
+\subsection{Applicable Problems}\label{sec:app_problem}
+The approach can be used with ODE systems $u' = f(t,u)$ that ensure  $u(t) \geq 0 \forall_t \forall_{  u_0 \geq 0}$
+
+This can be tested with the property 
+\begin{equation}
+u_i=0 \Rightarrow f_i(t,u_1,\cdots,u_i,\cdots,u_n) \geq 0  \forall_{u_c \geq 0} \forall_{t}
+\end{equation}
+
+For problems where the exact solution is positive for certain $u_0$ but positifity is not preserved for all $u_0 \geq 0$ tests did not show promising results.
+Additional it is not certain if the computed solutions would be reasonable.
+
+\subsection{Explicit methods}
+For most of the test problems problems with stability occur before getting negative values. This means that we need methods with large stability region.
+
+\subsection{Implicit methods}
+Implicit methods seem like an advantageous choice for a couple of reasons.
+
+When tested on the diffusion equation with an spike as initial data the $b$ is only changed for the first step. The change of the $b$ can lead to reasonable changes of the hight of the maximum.
+
+\subsection{Stepsize control}
+
+\subsection{Dense Output}
+
 \printbibliography

### 2020-02-10 latex/overview.tex
Update Paper
overview.tex:
Added an introduction, 
Added first example with description

plots\ 
Added folder for Plots

@@
 \usepackage[backend=biber,url  = false]{biblatex}
-\addbibresource{pos_rk_stephan.bib}
+\addbibresource{Bib_RK_optim.bib}
 
@@
 
-A big number of problems that preserve positifity reaction discretiesed PDEs...
+Many physical processes can be described with differential equations. 
+The physical property that are involved in these processes often only make sense if they are positive or non-negative. 
+Concentrations or number of particles have to be $\geq 0$.
+Other values may only make sense if they comply with other restrictions. An Example of these would-be probabilities or fractions which have to be $\geq 0$ and $\leq 1$.
+The ordinary differential equations (ODE) or a partial differential equations (PDEs) are often too complex to solve them analytically and therefore require numerical methods to get an approximate solution.
+Numerical methods generally do not satisfy these restrictions on the solutions.
 
-With increasing complexity of the right hand side a step requect has a high cost. 
-The goal of the project is to develop a positivity preserving method that uses adaptive b-Coefficients for every step.
+The most common methods to solve ODEs are Runge–Kutta methods (RKMs). 
+With the method of lines (MOL) it is also possible to solve PDEs with RKMs.
+At first the PDE is discretized in space. The resulting semi-discrete problem is an ODE that can be solved with a RKM. 
+It is possible to semi-discretize PDEs in a ways that preserve positivity\cite{kopecz_comparison_2019,kopecz_unconditionally_2018}.
+
+Therefore, it is interesting to know how to use RKMs that preserve positifity.
+For RKMs some theoretical limitations are known.
+One is that any unconditionally positive RKM has an order $\leq 1$.
+\cite{hundsdorfer_numerical_2003,bolley_conservation_1978,horvath_positivity_1998}.
+The only well know RKM that is unconditionally positive is the backward euler method, which is obliviously of first order.
+This causes restrictions on the step size for any RKM with a higher order.
+
+There are multiple approaches to get positivity preserving numerical schemes.
+One method is the Strong Stability Preserving Runge Kutta (SSPRK). 
+The idea is to construct a RKM that can be written as convex combination of euler steps. 
+By this the stability properties of the forward euler step are carried over to the SSPRK.
+The same happens to the positivity properties.
+By this the positivity of the method is ensured for time steps $\Delta t$ with $\Delta t \leq R \Delta t_{FE}$, where $R$ depends on the used SSPRK and $\Delta t_{FE}$ is the forward euler time step \cite{gottlieb_strong_2011}.
+The Modified Patanka Runge Kutta (MPRK) are 
+based on the Patanka RKM.
+In these destruction Terms are scaled by the new solution.
+This preserves positivity but violates the mass conservation.
+The MPRK overcome this disadvantage by also scaling the production terms.
+This is done by solving a linear equation system \cite{kopecz_comparison_2019}.
+An approach to ensure positivity in a practical application is shown in \cite{shampine_non-negative_2005}.
+Ensuring the positivity is done by redefining the ODE outside the domain.
+Another technique used are event finding methods.
+This approach is implemented in the Matlab ODE Suite.
+Unconditionally positive method of order higher than 1 are the diagonally split Runge-Kutta (DSRK) methods \cite{horvath_positivity_1998}.
+These are no longer part of the general linear methods.
+This allows them to be positive and higher order.
+(TODO: two sentences here)
+
+In the following paper an approach is taken that does not overcome the order imitation but is based on the idea that in practical problems for many time steps positivity can be achieved with a higher order methods.
+The main idea is to adapt the RKM after computation of the stage values in a way that ensures positivity.  
 This means that the computed stage values can be used and don't have to be recomputed.
+With increasing complexity of the right-hand side, a step reject has a high cost.
+This makes the additional cost of adapting the RKM economically.
 The resulting b should comply with the order conditions as far as possible.
 
-(Note: The problem is connected to the theory about embedded methods, in the sense that one have different RKM with the same set of stage values.
-It may be interesting to have some literature regarding this)
+A common way to calculate two different solutions with one set of stage values is the use of embedded methods for error approximation. 
+The idea of adapting the weights $b$ after calculating the stage values is used in \cite{ketcheson_spatially_2013}
+In this case it is used to adapt the properties of the time integrator for a method of lines simulation on a PDE.
+With this approach, it is possible to have different properties of the RKM at different parts of the domain.
 
-Other Approaches:
+In this paper, an adaption of the $b$ is done by solving a Linear programming (LP)-Problem.
+(TODO: Description LP-Problem)
 
-ssp104 methods: The idea is to construct a rkm Method out of explicit euler steps, that lead to a positive result for small enough $\Delta t$.
-(Todo: Read some literature regarding this and give a short explaination)
+The paper unfolds as follows. In section\,\ref{sec:main_idea} the main idea is explained. How the method can be written as LP-Problem is explained in section\,\ref{sec:LP-Problem}.
+In section\,\ref{sec:integration} it is described how the new approach can be used with different RKM, how it can be integrated in a step size control and how the Region of absolute stability can be approximated.
+In section\,\ref{sec:imple} further details that were used when implementing the algorithm are described.
+In section\,\ref{sec:Numeric_Results} Numerical results are given for multiple testproblems.
+A conclusion is given in section\,\ref{sec:conclusion}
 
-Implicit methods: Modified Patanka-Runge-Kutta (MPRK) \cite{kopecz_comparison_2019}
-
-The idea of adapting the $b$ after calculation the stage values is used in \cite{ketcheson_spatially_2013}
-In this case it is used to adapt the properties of the time integrator for a method of lines simulation on a PDE.
-With this approach it is possible to have different properties of the RKM at different parts of the domain.
-
-(TODO: Outline of paper)
+%(TODO: We know that we cannot work around the limitation from \cite{hundsdorfer_numerical_2003}, Where would be a good place to put this?)
 
@@
 
-When computing the solution of an ODE $u ' = f(t,u) $ using a RK-Method with $s$ Stages the stage values $k^{(1)},\cdots,k^{(s)}$ are computed according to
-
+When computing the solution of an ODE $u ' = f(t,u) $ using a RK-Method with $s$ Stages and the Butcher tableau
+\begin{align}
+\begin{array}{c|c}
+c &  A \\
+\hline
+ & b\\
+\end{array}
+\end{align}
+the stage values $k^{(1)},\cdots,k^{(s)}$ are computed according to
 \begin{equation}
@@
 \end{equation}
-
 Based on these values the next solution $u^{n+1}$ is computed by
-
 \begin{equation}
@@
 \end{equation}
-
 This equation can be formulated to
-
 \begin{equation}\label{eq:Combination}
@@
 \end{equation}
-
 where $K$ is a matrix containing the stage values $k_1,\cdots,k_s$.
 
-One reason why steps are requected is that the new solution $u^{n+1}$ does not comply with positifity. 
-So it may be adventagious to adapt Equation\,\ref{eq:Combination} in such a way that $u^{n+1}$ is positive. This is done by adapting the weights $b$.
+The new solution $u^{n+1}$ does not comply with positifity for some cases. 
+So it may be adventagious to adapt Equation\,\ref{eq:Combination} in such a way that $u^{n+1}$ is positive, instead of rejecting the Solution and computing a new set of stagevalues.
+This is done by adapting the weights $b$.
+
+For the computation of the stage-values only the $A$ has to be known. The $b$ is not needed for this computation. 
+Based on this Observation we can define $RKM_{A}$ as the set of RKMs that share the same $A$ and only differ by $b$.
+Calculating the stage-values can be considered as solving the ODE for all RKM in $RKM_a$ at the same time.
+The idea now is to choose one $RKM_{(A,b)} \in RKM_{A}$ by fixing the $b$. 
+Because the stage-values are already known we can choose a $RKM_{(A,b)}$ that leads to a positive $u^{n+1}$, if $RKM_{(A,b)} \in RKM_{A} | u^{n+1} \geq 0$ exists.
+When altering the $b$ the properties of an RKM are preserved. 
+Especially this approach preserves linear invariants like mass conservation, because it is a regular RKM.
+
 As it can be see in Equation\,\ref{eq:Combination} $u^{n+1}$ is a linear combination of the stage values and the previous solution $u^n$.
-When altering the weights the properties of the RKM should be preserved. Especially the Order of the RKM.  
-To ensure this the Order Conditions should be still meet. 
 Because the coeffeicents $a_{jl}$ are already fixed the Order Conditions reduce to linear equations.
@@
 
-This approach preserves linear invariants like mass conservation, because it is a regular RKM. 
-If there are no negative values in $u^{n+1}$ with the original $b$, then the method does not cause additional effort, because it only has to solve the LP-Problem if the regular RKM does not lead to a positive $u^{n+1}$.
-(exept for the test)
-No additional evaluations of the rhs.
-
-Flowchart to evaluate a step:
-
-\begin{itemize}
-\item Calculate the stage values $k_1,\cdots,k_s$
-\item Calculate the $u_{n+1} = u_n + \Delta t K b_{d}$ with a default b
-\item Test $u_{n+1} \geq 0$, if not:
-\item find a new b with $u_n + \Delta t K b \geq 0$
-\item Calculate the $u_{n+1} = u_n + \Delta t K b$ with the new b
-\end{itemize}
+If there are no negative values in $u^{n+1}$ with the original $b$, then the method does not cause additional computational effort, because it only has to solve the LP-Problem if the regular RKM does not lead to a positive $u^{n+1}$.
+%(exept for the test)
 
@@
 
-TODO: Add here the reaction equation, show that it behaves better with the new approach. (still keep it simple with fixed stepsize...)
+To illustrate the use of the method we consider the Following reaction equation from\,\cite{kopecz_comparison_2019}.
+
+\begin{subequations}
+\label{eq:Reaction}
+\begin{align}
+u_1' &= 0.01u_2 + 0.01 u_3 +0.003u_4 - \frac{u_1 u_2}{0.01+u_1} \\
+u_2' &= \frac{u_1u_2}{0.01+u_1}-0.01 u_2-0.5(1-\exp(-1.21 u_2^2)) u_3 -0.05 u_2 \\
+u_3' &= 0.5(1-\exp(-1.21u_2^2)) u_3 - 0.01 u_3 -0.02 u_3 \\
+u_4' &=0.05 u_2 + 0.02 u_3 0.003u_4
+\end{align}
+\end{subequations}
+and the initial conditions
+\begin{equation}
+u(0) = (8,2,1,4)^T
+\end{equation}
+When solved with the Cash-Karp (CK5) method and $\Delta t = 0.005$ the approximated solution contains negative values. This causes qualitatively wrong solutions to the problem. 
+In Figure\,\ref{fig:exampleI} the obtained results are plotted with a dashed line. 
+At $t=1.905$ $u_1$ gets negative. This leads to a diverging solution.
+
+Now the $b$ are adapted. The results are also plotted in Figure\,\ref{fig:exampleI}. 
+All $u\geq0$ is now fulfilled. Because of this a qualitatively correct solution is obtained.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.75\textwidth]{plots/exampleI.pdf}
+    \caption{Numerical approxiamtion of Reaction Problem from \cite{kopecz_comparison_2019}.
+Computed with ck5 and $\Delta t = 0.005$.
+The dashed line are the reference solution without the adaption of the $b$.
+The lower plot are the weights $b$ }
+    \label{fig:exampleI}
+\end{figure}
+
+The used weights $b_1,\cdots,b_5$ are also plotted in Figure\,\ref{fig:exampleI}. 
+The used weights are equal for $t<1.905$.
+At $t=1.905$ the weights are first adapted to ensure the positivity of the solution.
+For $t>2.63$ the original set of weights lead to a positive solution.
+Because of this the used weights equal the original weights again.
+
+\section{Details (TODO: better title)}\label{sec:LP-Problem}
+
+In the following section it is explained how the approach can be written as a LP-Problem. 
+For this we have to consider which $b$ to choose. 
+Generally any $b \in \mathbb{R}$ can be used. 
+A first requirement would be that the new solution is positive.
+We can further reduce this space by enforcing the Order Conditions. 
+By incorporating the Order Conditions the space of $b$ is limited to a subspace of $\mathbb{R}$. %(Note: not linear because 0 is not insubspace)
+Finally we have to give a objective function to get the optimal $b$.
 
@@
 
-By incorporating the Order Conditions for a certain Order the space of $b$ that can be choose is limited to a linear Subspace. 
-Now it is important to know if the new solution $u^{n+1}$ still resembles the solution of the ODE $u(t_{n+1})$. 
-We can motivate this with the following example:
+The main goal is to choose a $RKM_{(A,b)}$ that for which $u^{n+1}$ resembles the solution of the ODE $u(t_{n+1})$. 
+A obvious objective is to use a $RKM_{(A,b)}$ with a high order, but this is not enough.
+To get a better understanding for the method we consider the behavior of a simple problem.
 
-Take the linear, positifity preserving ODE from \cite{kopecz_unconditionally_2018} with $u = [u_1,u_2]^T$ with the initial conditions $u_0 = [1,0]^T$ and
+We take the linear, positifity preserving ODE from \cite{kopecz_unconditionally_2018} with $u = [u_1,u_2]^T$ with the initial conditions $u_0 = [1,0]^T$ and
 
 \begin{equation}
-u'(t) = A u(t) \qquad A = \left[\begin{matrix}- a & 1\\a & -1\end{matrix}\right] \quad a =5
+u'(t) = L u(t) \qquad L = \left[\begin{matrix}- a & 1\\a & -1\end{matrix}\right] \quad a =5
 \end{equation}
@@
 By a suitable choice of $\alpha \in [\frac{1}{6},\frac{7}{30}]$ any $u$ that complies with mass conservation and positifity can be reached. 
+By adding additional constraints on the $b$ the choice of $\alpha$ can be narrowed down. 
+An objective function is also needed to make the choice unique. This should be designed in a way to prefer $b$ that are close to the original weights $b_{orig}$
 
-This gives raise to the question weather the new solution is reasonable.
-This also poses the problem on developing a system that can decide if the new solution is has to be rejected.
-
-\section{Details}\label{sec:Deta	ils}
+%This gives raise to the question weather the new solution is reasonable.
+%This also poses the problem on developing a system that can decide if the new solution is has to be rejected.
 
@@
 
+\section{TODO: Title} \label{sec:integration}
+
+In the previous sections a algorithm for choosing a $b$ has been presented. 
+In the following section it is explained how this Algorithm can be integrated into the framework of existing solvers.
+
 \subsection{Choice of Baseline Method}
@@
 
-The main way is to ensure that the RKM with adapted b i to ensure that the $b$ approach $b_{orig}$ for $\Delta t \to 0$.
+The main way is to ensure that the RKM with adapted b converges to the correct solution is to ensure that the $b$ approaches $b_{orig}$ for $\Delta t \to 0$.
 For this $b$ it is already know that the numeric solution converges to the exact solution.
@@
 
-\section{Implementation aspects of the Algorithm}
+\section{Implementation aspects of the Algorithm}\label{sec:imple}
 
@@
 
+\section{Conclusion} \label{sec:conclusion}
+
 \printbibliography

### 2020-02-11 RKimple.py
Added Convergence Plots for Heat Equation and ADR-Problem
RK-Imple.py
Minor bug in Error estimator and minor changes to make it run better

utils.py
Added canges to make it display more information

Convex_Implicit.ipynb:
Added Convergence Plots,
New Embeded methods. This field requires more research

ADP.ipynb
Added Convergence Plots

@@ -42,7 +42,7 @@ def solve_LP(solver,O,rhs,rkm,b_orig,u,K,dt,reduce = False,verbose_LP = False,mi
         5 : solver crashed
         6 : Trivial Problem
     l = Array with number of constraints
-    b: found b, if solver failed b_origg
+    b: found b, if solver failed b_orig
     
     """
     s = len(rkm.b)
@@ -435,6 +435,7 @@ def adapt_b(rkm,K,dt,u,minval,maxval,tol_neg,tol_change,p,theta,solver,solveropt
     """
     message = ''
     status = {}
+    change = None
 
     for i,the in enumerate(theta): #loop through all the sub-timesteps
         for p_new in p:     #loop through orders
@@ -460,7 +461,7 @@ def adapt_b(rkm,K,dt,u,minval,maxval,tol_neg,tol_change,p,theta,solver,solveropt
                     if verbose:    print('LP-Solve returned a b that leads to a false solution')
                     if verbose >= 2:    print(min(u_n-minval)); print(max(u_n-maxval)); print(u_n)
                 else:
-                    change = np.linalg.norm(K@(b-rkm.b))
+                    change = dt*np.linalg.norm(K@(b-rkm.b))
                     if change > tol_change: # to big adaption...
                         #do some error handling here
                         if verbose:    print('a to big adaptation to the solution by changing the b')
@@ -469,7 +470,7 @@ def adapt_b(rkm,K,dt,u,minval,maxval,tol_neg,tol_change,p,theta,solver,solveropt
                         if verbose: print('found new b')
                         return True, u_n,b, dt*the, message, change,p_new,the,status
 
-    return False, None,np.zeros_like(rkm.b)*np.nan, 0, message, None, 0, 0,status
+    return False, None,np.zeros_like(rkm.b)*np.nan, 0, message, change, 0, 0,status
 
 def adapt_b_convex(rkm,K,dt,u,minval,maxval,tol_neg,tol_change,p,theta,solver,solveropts,verbose = False):
     """
@@ -500,6 +501,7 @@ def adapt_b_convex(rkm,K,dt,u,minval,maxval,tol_neg,tol_change,p,theta,solver,so
     """
     message = ''
     status = {}
+    change = None
     print('changed')
 
     for i,the in enumerate(theta): #loop through all the sub-timesteps
@@ -515,7 +517,7 @@ def adapt_b_convex(rkm,K,dt,u,minval,maxval,tol_neg,tol_change,p,theta,solver,so
[…]

### 2020-02-11 utils.py

@@ -27,11 +27,12 @@ def plot_convergence(problem,solver,dt,refernce,step=1,error='abs',dx='1',Norm =
     Return:
     sol:                Array with the solutions used for calculationg the errors
     err:                Array with errors
+    change:             Array, True if b was changed False otherwise
 
     """
     
     err = np.zeros_like(dt)
-    
+    change = np.zeros_like(dt,dtype=bool)
     sol = []
     
     
@@ -40,12 +41,13 @@ def plot_convergence(problem,solver,dt,refernce,step=1,error='abs',dx='1',Norm =
         solver.dt = dt[i]
         status,t,u,b = RK_integrate(solver=solver,problem=problem,**Params)
         dif = refernce[:,i]-u[step]
+        change[i] = 'c' in status['b']
         if error == 'abs':
             err[i] = np.linalg.norm(dif,ord=Norm)
         elif error == 'rel':
             err[i] = np.linalg.norm(dif,ord=Norm)/np.linalg.norm(refernce[:,i],ord=Norm)
         elif error == 'grid': #Grid function Norm (LeVeque Appendix A.5)
-            error[i] = dx**(1/Norm)*np.linalg.norm(dif,ord=Norm)
+            err[i] = dx**(1/Norm)*np.linalg.norm(dif,ord=Norm)
         else:
             print('Error not defined')
             print(error)
@@ -59,7 +61,7 @@ def plot_convergence(problem,solver,dt,refernce,step=1,error='abs',dx='1',Norm =
     plt.ylabel('Error')
     plt.xlabel('dt')
     
-    return sol,err
+    return sol,err,change
     
         
 
@@ -277,7 +279,7 @@ def show_status(status):
     #Number of step reqects
     print('Number of step reqects:');print(status['b'].count('r'))
 
-    fig, axs = plt.subplots(3,1)
+    fig, axs = plt.subplots(5,1)
 
     #plot on which timesteps the b was adabpted
 
@@ -286,6 +288,10 @@ def show_status(status):
 
     #plot the step reqects
     axs[0].eventplot(np.nonzero(np.array(status['b'])=='r'), colors='red', lineoffsets=-0.5,
+                    linelengths=0.5)
+
+    #plot step reqect caused by tol
+    axs[0].eventplot(np.nonzero(np.array(status['sc'])=='r'), colors='red', lineoffsets=-1.5,
                     linelengths=0.5)
 
[…]

### 2020-02-11 latex/overview.tex
Spelling mistakes and minor corrections in Paper

@@
 
-When computing the solution of an ODE $u ' = f(t,u) $ using a RK-Method with $s$ Stages and the Butcher tableau
+When computing the solution of an ODE $u ' = f(t,u) $ using a RKM with $s$ Stages and the Butcher tableau
 \begin{align}
@@
 \end{align}
-the stage values $k^{(1)},\cdots,k^{(s)}$ are computed according to
+the stage values $y^{(1)},\cdots,y^{(s)}$ are computed according to
 \begin{equation}
-k^{(j)} =  f(t^n + \Delta t c_j, u^n + \Delta t \sum_{l = 1}^{s} a_{jl} k^{(l)}  \quad j = 1,\cdots,s)
+y^{(j)} =  u^n + \Delta t \sum_{k = 1}^{s} a_{jk} f(t^n + \Delta t c_k,y^{(k)})  \quad j = 1,\cdots,s
 \end{equation}
@@
 \begin{equation}
-u^{n+1} = u^n + \Delta t \sum_{j  = 1}^s k^{(j)} b_j .
+u^{n+1} = u^n + \Delta t \sum_{j  = 1}^s f(t^n + \Delta t c_j,y^{(j)}) b_j .
 \end{equation}
-This equation can be formulated to
+The stage derivatives $f^{(1)},\cdots,f^{(s)}$ can be calculated with
+\begin{equation}
+f^{(j)} = f(t^n + \Delta t c_j,y^{(j)})
+\end{equation}
+The new solution can now be calculated with
 \begin{equation}\label{eq:Combination}
-u^{n+1} = u^n + \Delta t K b
+u^{n+1} = u^n + \Delta t F b
 \end{equation}
-where $K$ is a matrix containing the stage values $k_1,\cdots,k_s$.
+where $F$ is a matrix containing the stage derivatives $f^{(1)},\cdots,f^{(s)}$.
 
-The new solution $u^{n+1}$ does not comply with positifity for some cases. 
-So it may be adventagious to adapt Equation\,\ref{eq:Combination} in such a way that $u^{n+1}$ is positive, instead of rejecting the Solution and computing a new set of stagevalues.
+The new solution $u^{n+1}$ does not comply with positivity for some cases. 
+So it may be advantageous to adapt \eqref{eq:Combination} in such a way that $u^{n+1}$ is positive, instead of rejecting the Solution and computing a new set of stage values.
 This is done by adapting the weights $b$.
 
-For the computation of the stage-values only the $A$ has to be known. The $b$ is not needed for this computation. 
-Based on this Observation we can define $RKM_{A}$ as the set of RKMs that share the same $A$ and only differ by $b$.
-Calculating the stage-values can be considered as solving the ODE for all RKM in $RKM_a$ at the same time.
+For the computation of the stage derivatives only the $A$ has to be known. The $b$ is not needed for this computation. 
+Based on this Observation we can define $RKM_{A}$ as the set of RKMs that share the same $A$ and only differ in $b$.
+Calculating the stage derivatives can be considered as solving the ODE for all RKM in $RKM_A$ at the same time.
 The idea now is to choose one $RKM_{(A,b)} \in RKM_{A}$ by fixing the $b$. 
-Because the stage-values are already known we can choose a $RKM_{(A,b)}$ that leads to a positive $u^{n+1}$, if $RKM_{(A,b)} \in RKM_{A} | u^{n+1} \geq 0$ exists.
+Because the stage derivatives are already known we can choose a $RKM_{(A,b)}$ that leads to a positive $u^{n+1}$, if $RKM_{(A,b)} \in RKM_{A} | u^{n+1} \geq 0$ exists.
 When altering the $b$ the properties of an RKM are preserved. 
@@
 
-As it can be see in Equation\,\ref{eq:Combination} $u^{n+1}$ is a linear combination of the stage values and the previous solution $u^n$.
-Because the coeffeicents $a_{jl}$ are already fixed the Order Conditions reduce to linear equations.
+As it can be see in \eqref{eq:Combination} $u^{n+1}$ is a linear combination of the stage values and the previous solution $u^n$.
+Because the coefficients $a_{jl}$ are already fixed the Order Conditions reduce to linear equations.
 This means that finding the $b$ can be done by solving a linear Programming problem.
@@
 If there are no negative values in $u^{n+1}$ with the original $b$, then the method does not cause additional computational effort, because it only has to solve the LP-Problem if the regular RKM does not lead to a positive $u^{n+1}$.
-%(exept for the test)
+%(except for the test)
 
@@
 
-To illustrate the use of the method we consider the Following reaction equation from\,\cite{kopecz_comparison_2019}.
+To illustrate the use of the method we consider the following reaction equation from\,\cite{kopecz_comparison_2019}.
 
@@
 The used weights $b_1,\cdots,b_5$ are also plotted in Figure\,\ref{fig:exampleI}. 
-The used weights are equal for $t<1.905$.
+The used weights are equal the original weights for $t<1.905$.
 At $t=1.905$ the weights are first adapted to ensure the positivity of the solution.
@@
 
-In the following section it is explained how the approach can be written as a LP-Problem. 
+In the following section, it is explained how the approach can be written as a LP-Problem. 
 For this we have to consider which $b$ to choose. 
-Generally any $b \in \mathbb{R}$ can be used. 
+Generally, any $b \in \mathbb{R}$ can be used. 
 A first requirement would be that the new solution is positive.
 We can further reduce this space by enforcing the Order Conditions. 
-By incorporating the Order Conditions the space of $b$ is limited to a subspace of $\mathbb{R}$. %(Note: not linear because 0 is not insubspace)
-Finally we have to give a objective function to get the optimal $b$.
+By incorporating the Order Conditions, the space of $b$ is limited to a subspace of $\mathbb{R}$. %(Note: not linear because 0 is not in subspace)
+Finally, we have to give a objective function to get the optimal $b$.
 
@@
 We have shown the general idea and how it works, now we want to discuss on what it does in practice.
-At first we state it is good to preserve the order...
+At first, we state it is good to preserve the order...
 Get to the Example ...
@@
 
-We take the linear, positifity preserving ODE from \cite{kopecz_unconditionally_2018} with $u = [u_1,u_2]^T$ with the initial conditions $u_0 = [1,0]^T$ and
+We take the linear, positivity preserving ODE from \cite{kopecz_unconditionally_2018} with $u = [u_1,u_2]^T$ with the initial conditions $u_0 = [1,0]^T$ and
 
@@
 \end{equation}
-
 and the first order RKM with two stages
-
 \begin{align}
@@
 \end{align}
-
 If simulated with $\Delta t = 1$ the stage values are 
-
 \begin{equation}
-k^1 = \left(\begin{matrix}-5\\5\end{matrix}\right) \qquad k^2 = \left(\begin{matrix}10\\-10\end{matrix}\right)
+f^1 = \left(\begin{matrix}-5\\5\end{matrix}\right) \qquad f^2 = \left(\begin{matrix}10\\-10\end{matrix}\right)
 \end{equation}
-
 All $b$ that comply with the order Condition for the first order can be expressed as
-
 \begin{equation}
@@
 \end{equation}
-
-Therefor the general solution is
-
+With this we have one degree of freedom for the choice of $b$. This is done by the parameter $\alpha$.
+If the general expression for $b$ is inserted in \eqref{eq:Combination} the general solution is 
 \begin{equation}
-u^{1} = u^0 + \Delta t  \left[k^1,k^2\right] b = \left(\begin{matrix}\frac{5}{2}\\- \frac{5}{2}\end{matrix}\right) \alpha  \left(\begin{matrix}-15\\15\end{matrix}\right)
+u^{1} = u^0 + \Delta t  \left[f^1,f^2\right] b = \left(\begin{matrix}\frac{5}{2}\\- \frac{5}{2}\end{matrix}\right) \alpha  \left(\begin{matrix}-15\\15\end{matrix}\right)
 \end{equation}
-
-By a suitable choice of $\alpha \in [\frac{1}{6},\frac{7}{30}]$ any $u$ that complies with mass conservation and positifity can be reached. 
+The solution for the original RKM would be $u^{1} = \left(\frac{5}{2} ,- \frac{5}{2}\right)^T$ which contains a negative value.
+But by changing the parameter $\alpha$ and therefore changing the $b$ a positive solution can be reached.
+By a suitable choice of $\alpha \in \left[\frac{1}{6},\frac{7}{30}\right]$ any $u$ that complies with mass conservation and positivity can be reached. 
 By adding additional constraints on the $b$ the choice of $\alpha$ can be narrowed down. 
@@
 
-%This gives raise to the question weather the new solution is reasonable.
+%This gives raise to the question whether the new solution is reasonable.
 %This also poses the problem on developing a system that can decide if the new solution is has to be rejected.
@@
 If the $A$ of a preexisting RK-method is used, there is at least one solution for the linear equation system.
-In order to get an optimization problem the equation system has to be underdetermined.
+In order to get an optimization problem, the equation system has to be underdetermined.
 This implies that $\mathrm{rank}(Q) < s$. 
@@
 
-The second approach is to choose $b$ as a convex combination of a set of $b$s. This approach is more useful if some embedded methods with certain properties are know. 
+The second approach is to choose $b$ as a convex combination of a set of $b$s. This approach is more useful if some embedded methods with certain properties are known. 
 This approach also shows more predictable behavior because the resulting $u^{n+1}_b$ is also a convex combination of the solutions $u^{n+1}_{b_a},u^{n+1}_{b_b},\cdots$ 
@@
 When directly adapting the $b$ a new $b$ that complies with the Order Conditions is chosen.   
-The weights $b = b_{orig} + \Delta b$ consist of the weights of the original method $b_{orig}$ and an adaption $\Delta b$. This $\Delta b \in \mathrm{ker}(Q)$. 
+The weights $b = b_{orig} + \Delta b$ consist of the weights of the original method $b_{orig}$ and an adaption $\Delta b$. The adaption is $\Delta b \in \mathrm{ker}(Q)$. 
 In Figure\,\ref{fig:b_direct} this is drawn. The red line represents the subspace of $b$ that satisfy the Order Conditions. Both $b_{orig}$ and $b$ are in this subspace.
@@
 To run it as a LP-Problem a set of constraints is constructed.  
-Also a objective function is needed to determine the optimal $b$.
+Also, an objective function is needed to determine the optimal $b$.
 
@@
 The Direct Adaptation approach leads to two set of constraints on the $b$.
-
-The first set of constraints is the positifity and can be written as 
-$$u_{n+1}=u_n+\Delta K \vec{b} \geq 0$$
-
+The first set of constraints is the positivity and can be written as 
+\begin{equation}
+u_{n+1}=u_n+\Delta K \vec{b} \geq 0
+\end{equation}
 The second set of constraints are the Order Conditions
-$$Qb=r$$
+\begin{equation}
+Qb=r
+\end{equation}
 
@@
 
-\item[L1 norm] The optimisation Problem opimises for $min(|b-rkm.b|_1)$
+\item[L1 norm] The optimization Problem optimizes for $min(|b-rkm.b|_1)$
 
-\item[Quadrature]   The optimisation Problem opimises for $min(|b^Tq-r|)$
+\item[Quadrature]   The optimization Problem optimizes for $min(|b^Tq-r|)$
                             where q is the quadrature condition of the next higher order 
@@
 
-\item[Hom. Order]    The optimisation Problem opimises for $min(|bTO|)$ where O are the homogenus Order Conditions of the next higher orders
+\item[Hom. Order]    The optimization Problem optimizes for $min(|bTO|)$ where O are the homogenous Order Conditions of the next higher orders
 
@@
 This can be achieved by $argmin(f_{optim}(b)) = b_{orig}$. 
-In practice this also means that there is no need to solve solve the optimization Problem if no constraints are violated because the solution of this problem is already known.
-This is adventagious for performance reasons.
+In practice, this also means that there is no need to solve solve the optimization Problem if no constraints are violated because the solution of this problem is already known.
+This is advantageous for performance reasons.
 The different approaches have been tested on a linear test problem from \cite{kopecz_unconditionally_2018}. The optimization problem was used at every step.
@@
 Large values of b can lead to numerical problems.
-Because $\sum_{i  = 1}^s b_i = 1$, a increase of $||b||$ leads to a combination of negative and positive values.
-Because the computation only is done with a finite precision (and usually in floating point) large $||b||$ can lead to absorption effects.
+Because $\sum_{i  = 1}^s b_i = 1$, an increase of $||b||$ leads to a combination of negative and positive values.
+Because the computation is done with a finite precision (and usually in floating point) large $||b||$ can lead to absorption effects.
 By changing the $b$ the stability function is altered. This can lead to oscillations if the poles of the RHS are no longer in the stability region of the resulting method.
@@
 \begin{equation}
-f_{obj} = ||b_{optim}-b_{orig}||_1
+f_{obj} = ||b-b_{orig}||_1
 \end{equation}
@@
 
-If the original step $\Delta t$ was to large and there is no suitable $b$, we would like to calculate a $u^{n+1}$ using the stage values instead of rejecting the step.
+If the original step $\Delta t$ was too large and there is no suitable $b$, we would like to calculate a $u^{n+1}$ using the stage values instead of rejecting the step.
 For this the time step is reduced after calculation the stage values.
@@
 These can be found in \cite{hairer_runge-kutta_1993}.
-We introduce $\theta \in (0,1]$. The new step taken has the length $\theta \Delta t$ whereas $\Delta t$ is the time step used when calculating the Stagevalues.
+We introduce $\theta \in (0,1]$. The new step taken has the length $\theta \Delta t$ whereas $\Delta t$ is the time step used when calculating the stage values.
 To calculate a reduced step the Order Conditions have to be adapted and a new $b_{orig}$ for the objective function (and more important the Error detection afterwards) has to be generated. 
@@
 The dependency of $r(\theta)$ is nonlinear.
-This means that the maximum $\theta$ with a suitable $b$ can not be calculated by solving a LP-Problem.
-This means that different $\theta$ have to be tested independently.
+This means that the maximum $\theta$ with a suitable $b$ cannot be calculated by solving a LP-Problem.
+This means that different $\theta$ have to be tested.
 (Note: we could include some algorithm to search for it)
@@
 \subsection{Convex combination of existing methods}\label{sec:convex}
-For some test problems (e.g. the Heat Equation solved with an implicit method of higher order), large distortions occur, even though a norm of the distortion of the b is used as a objective function. 
-For implicit methods it is known that the backward euler steps lead to a positive solution for any $\Delta t$ \cite{hundsdorfer_numerical_2003}.
+For some test problems (e.g.
+the Heat Equation solved with an implicit method of higher order), large distortions occur, even though a norm of the distortion of the b is used as an objective function.
+For implicit methods, it is known that the backward euler steps lead to a positive solution for any $\Delta t$ \cite{hundsdorfer_numerical_2003}.
 The backward Euler is still only first Order but shows more exact results than a method of higher Order that is distorted by the adaption of the $b$, at least at some steps.
@@
 This is straightforward using
-
 $$ Q b = Q (\chi b_{\alpha} + (1-\chi) b_{\beta}) = \chi Q  b_{\alpha} + (1-\chi) Q b_{\beta}) = \chi r + (1-\chi) r = r$$
-
 $b_{\alpha}$ and $b_{\beta}$ are the methods that are combined to $b$ using the parameter $\chi$. The Matrix $Q$ and the vector $r$ are the Order Conditions as defined in Section\,\ref{sec:OrderCond}.
@@
 \begin{equation}\label{eq:b_convex_def}
-\vec{b} = \sum_i^d a_i \vec{b_i} = Ba
+b = \sum_i^d a_i b_i = Ba
 \end{equation}
-
 where $B=\left[b_1,\cdots,b_d \right]$ is a matrix containing the embedded methods.  
-To ensure the that Equation\,\ref{eq:b_convex_def} yields a convex combination following constraints are enforced for the  the parameters $a,\cdots,a_d$.
-
+To ensure the that \eqref{eq:b_convex_def} yields a convex combination following constraints are enforced for the  the parameters $a,\cdots,a_d$.
 \begin{equation}
- 0 \leq a_1 \leq 1  \forall_{i \in \{1, \cdots d \}}
+ 0 \leq a_1 \leq 1  \forall {i \in \{1, \cdots d \}}
 \end{equation}
@@
 \end{equation}
-
-The positifity constraint changes to
-
+The positivity constraint changes to 
 $$u_{n+1}=u_n+\Delta t K B a \geq 0$$
-
 where $a = [a_1,\cdots,a_d]^T$
@@
 
-As a first order method a $b$ is added that corresponds to a series of backward euler steps. These are know to be yield to a positive result.
+As a first order method, a $b$ is added that corresponds to a series of backward euler steps. These are known to be yield to a positive result.
 
@@
 
-(Note: Doing dense Output here is a bit complicated, but it would not help ensure positifity only (maybe) increase exactness.
-is it worth exploring this or is it better to just leave this open for further research it may be also interesting regardless the study of positifity)
+(Note: Doing dense Output here is a bit complicated, but it would not help ensure positivity only (maybe) increase exactness.
+is it worth exploring this or is it better to just leave this open for further research it may be also interesting regardless the study of positivity)
 Everything has downsides...
@@
 
-In the previous sections a algorithm for choosing a $b$ has been presented. 
-In the following section it is explained how this Algorithm can be integrated into the framework of existing solvers.
+In the previous sections an algorithm for choosing a $b$ has been presented. 
+In the following section, it is explained how this Algorithm can be integrated into the framework of existing solvers.
 
 \subsection{Choice of Baseline Method}
-An important property for the used baseline method is the Existence of embedded methods and the degrees of freedom for the b.
-As noted in Section\,\ref{sec:OrderCond} the number of Stages has to be higher than the Order. 
+An important property for the used baseline method is the existence of embedded methods and the degrees of freedom for the b.
+As noted in Section\,\ref{sec:OrderCond} the number of Stages has to be higher than the order. 
 Another important property is the number of degrees of freedom for the choice of the new weights. 
 These can be calculated using $s-\mathrm{rank}(Q)$%$\mathrm{dim}(\mathrm{ker}(Q))$.
-The number of degrees of freedom for different methods are shown in table\,\ref{table:DOF_exp} for explicit methods and in table\,\ref{table:DOF_imp} for implic methods.
+The number of degrees of freedom for different methods are shown in table\,\ref{table:DOF_exp} for explicit methods and in table\,\ref{table:DOF_imp} for implicit methods.
 
@@
 
-Another important property is the stability region in comparison with the maximum $\Delta t$ that ensures positifity.
+Another important property is the stability region in comparison with the maximum $\Delta t$ that ensures positivity.
 For this approach to be useful the RK-method has to yield negative results for $\Delta t$ smaller than the biggest stable $\Delta t$.
 
-(Note: up to this point i do not understand why this is happening. We only see it occur depending on the Proble-RKM combination)
+(Note: up to this point I do not understand why this is happening. We only see it occur depending on the Problem-RKM combination)
 
 A third property is the existence of a positive solution.
-This corresponds to the question if there is a embedded first order method that ensures positivity.
+This corresponds to the question if there is an embedded first order method that ensures positivity.
 This is particularly interesting for Implicit methods, because the $\Delta t$ is not limited by the stability.
-If we have a embedded first Order Method we can ensure that there is always a positive solution, which might be of first Order. (Todo: Source)
+If we have an embedded first Order Method we can ensure that there is always a positive solution, which might be of first Order.
 
@@
 
-This gives raise to the question weather the new solution is reasonable.
+This gives raise to the question whether the new solution is reasonable.
 This also poses the problem on developing a system that can decide if the new solution is has to be rejected.
@@
 The main way is to ensure that the RKM with adapted b converges to the correct solution is to ensure that the $b$ approaches $b_{orig}$ for $\Delta t \to 0$.
-For this $b$ it is already know that the numeric solution converges to the exact solution.
+For this $b$ it is already known that the numeric solution converges to the exact solution.
 This is made sure by property of the Objective function that 
-
 \begin{equation}
@@
 \end{equation}
-
 This approximation is not very useful for real applications. 
-To approximate the error of a new step we propose following approxiamtion of the local error:
+To approximate the error of a new step we propose following approximation of the local error:
 
@@
 
-After adapting the $b$ the Approximation Error is checked. If $|\Delta t K(b-b_{orig})|$ is larger than the tolerance the $b$ is requected.
+After adapting the $b$ the Approximation Error is checked. If $|\Delta t K(b-b_{orig})|$ is larger than the tolerance the $b$ is rejected.
 
@@
 If the LP-Problem was infeasible and no $b$ was found the $err_{approx}$ is set to a custom value.
-This value should be chosen sufficiently large to ensure that the next step take is smaller, but still not to large to keep the stepsize control stable.
+This value should be chosen sufficiently large to ensure that the next step take is smaller, but still not too large to keep the step size control stable.
 
@@
 
-%To make sure that the new adapted solution is still reasonable we measure the deviation form the Original method.
+%To make sure that the new adapted solution is still reasonable we measure the deviation from the Original method.
 In the worst case the errors add up.
 (todo: here some math formula...
-(We know that can not be the case because we already know that we will reduce the error for the quantity we will make positive, still we can use it as a upper bound.))
+(We know that cannot be the case because we already know that we will reduce the error for the quantity we will make positive, still we can use it as a upper bound.))
 
@@
 For Problems where only a small number of step are affected a bigger change in the stability function would be acceptable.
-For Problems with require a adaptation of the weights for every step one needs to make sure that the resulting method is stable.
+For Problems with require an adaptation of the weights for every step one needs to make sure that the resulting method is stable.
 
@@
 We start with the definition of the RKM
-
 \begin{align}
@@
 \end{align}
-
 Now we insert the Stability functions for \ref{eq:u_n+1} and \ref{eq:u_n+1_sum} which we already know to be equal. We get 
-
 \begin{align}
@@
 \end{align}
-
 because this has to be true for all $u_n$ we get
@@
 
-%(Todo:It is not straightforward if there are no pole zero cancellations in the stability function.
-For explicit methods the stability function is a polynomial o degree $\leq s$ (Hundsdorfer).
-Pole zero cancellations can not happen for these.
+%(Todo: It is not straightforward if there are no pole zero cancellations in the stability function.
+For explicit methods, the stability function is a polynomial o degree $\leq s$ (Hundsdorfer).
+Pole zero cancellations cannot happen for these.
 For implicit functions that are A-stable the poles have to be in the right halve plane.
@@
 \subsubsection{Stability of convex combination}
-If the new $b$ is chosen as a convex combination of $b$ it is easy to prove that the resulting  the Region of absolute stability is at least the Intersection of the region of absolute stability of the methods that define the convex combination.
-(Compare Section\,\ref{proof:convex_comb}).
+If the new $b$ is chosen as a convex combination of $b$ it is easy to prove that the resulting region of absolute stability is at least the Intersection of the regions of absolute stability of the methods that define the convex combination.
 This also means that if all the used embedded methods used to construct the new $b$ are A-Stable, the resulting method is also A-Stable.
 
-With this we can proof that the convex combination of two methods is absolutely stable at the points where both original methods were absolutely stable.
-
-We want to show that $(|R_{b_a}(z)|  < 1) (|R_{b_b}(z)| < 1) \Rightarrow |R_{\chi b_a +(1- \chi) b_b}(z)| < 1$ with $\chi \in [0,1]$.
+We want to proof that the convex combination of two methods is absolutely stable at the points where both original methods were absolutely stable.
+This can be formulated as $(|R_{b_a}(z)|  < 1) (|R_{b_b}(z)| < 1) \Rightarrow |R_{\chi b_a +(1- \chi) b_b}(z)| < 1$ with $\chi \in [0,1]$.
 We write $\alpha = \chi$ and $\beta = 1-\chi$ for convenience.
@@
 A useful property of the resulting stability region is that it is similar to the stability region of the baseline method. 
-(Note: is it possible to give some statement like: If the stability regions of the rkm used for multiple steps are similar, then the overall method is stable for the intersection of the stability regions)
+(Note: is it possible to give some statement like: If the stability regions of the RKM used for multiple steps are similar, then the overall method is stable for the intersection of the stability regions)
 
-For most methods a small change of $b$ does not alter the stability function dramatically.
+For most methods, a small change of $b$ does not alter the stability function dramatically.
+(TODO: I am missing a piece here from ‘the derivative is finite’ to ‘is simmilar’)
 
@@
 \end{equation}
-
 The border is defined by the implicit function 
@@
  $|R(w,u,v)|^2$ is calculated. 
-If the gradient does not vanish at $z = z_0$ Equation\,\ref{eq:border} Can be transformed in a function $\mathbb{R}^m \times \mathbb{R} \rightarrow \mathbb{R}$ depending on $w$ and a single variable $a$ by setting $z = z_0 + n a$.
+If the gradient does not vanish at $z = z_0$ \eqref{eq:border} Can be transformed in a function $\mathbb{R}^m \times \mathbb{R} \rightarrow \mathbb{R}$ depending on $w$ and a single variable $a$ by setting $z = z_0 + n a$.
 The complex number $n$ denotes the normal vector $\vec{n} = \frac{\nabla |R(w,u,v))|^2}{\left| \nabla |R(w,u,v))|^2 \right|}$ of the border of the stability region written in complex notation.
@@
 \end{equation}
-
 If explicit methods are used this function is known to be continuously differentiable because it can be written as a Polynomial of $w_1,\cdots,w_m,u$ and $v$.
@@
 \end{equation}
-
 as long as the inverse of the Jacobian exists. This is true in most cases as it has been discussed above.
 
-The Jacobian of equation\,\ref{eq:f(w,a)} $J_{f,a}(w,a(w))$ evaluated at $w=0$ can be calculated using the directional derivative
+The Jacobian of \eqref{eq:f(w,a)} $J_{f,a}(w,a(w))$ evaluated at $w=0$ can be calculated using the directional derivative
 \begin{align*}
@@
 
-For the further discussion we are concerned with a System of $m$ coupled ODEs. 
+For the further discussion, we are concerned with a System of $m$ coupled ODEs. 
 In this setup $u \in \mathbb{R}^m$.
@@
 \begin{align}
- u_i^{n+1} &\geq 0   \;   \forall_{i \in \{1, \cdots,m \}}  \\
- u_i^n + h \sum_{j=0}^{s-1} b_j k_i^j  &\geq 0   \;   \forall_{i \in \{1,\cdots,m \}}  
+ u_i^{n+1} &\geq 0   \;   \forall {i \in \{1, \cdots,m \}}  \\
+ u_i^n + h \sum_{j=0}^{s-1} b_j k_i^j  &\geq 0   \;   \forall {i \in \{1,\cdots,m \}}  
 \end{align}
-
 has to be fulfilled.
-This infers $m$ positivity constraints to the optimisation problem. These can be written as
+This infers $m$ positivity constraints to the optimization problem. These can be written as
 
@@
 \end{equation}
-
 where $K = \big[k^1 , \cdots k^{s-1}\big]$.
@@
 This causes unnecessary many inequality constraints on the LP-Problem.
-To simplify the problem one want to reduce the number of positifity constraints by ignoring it for ODEs where there is no issue with positifity.
-
-The new positifity constraints can be formulated as
+To simplify the problem, one want to reduce the number of positivity constraints by ignoring it for ODEs where there is no issue with positivity.
+The new positivity constraints can be formulated as
 \begin{equation}
-u_i^n + \Delta b \sum_{j=0}^{s-1} b_j k_i^j  \geq 0   \;   \forall_{i \in h \subseteq \{1,\cdot,m \}} 
+u_i^n + \Delta b \sum_{j=0}^{s-1} b_j k_i^j  \geq 0   \;   \forall {i \in h \subseteq \{1,\cdot,m \}} 
 \end{equation}
-
 For this a set $h \subseteq \{1,\cdots,m \}$ that include all indecies that are needed for determining the $b$ has to be chosen.
@@
 A reasonable approach is to set $h_0 = \{ i \in \{1,\cdots,m \} |  u_i^{n+1}  < 0 \}$. 
-After the new $b$ and $u^{n+1}$ is computed using $h_0$, the algorithm has to check if the positifity of the other values $u_i | i \notin h_0 \subseteq \{1,\cdots,m \}$ is still meet. 
+After the new $b$ and $u^{n+1}$ is computed using $h_0$, the algorithm has to check if the positivity of the other values $u_i | i \notin h_0 \subseteq \{1,\cdots,m \}$ is still meet. 
 If the conditions is not meet a new set $h_{a+1} = \{ i \in \{1,\cdots,m \}|  u_i^{n+1}  < 0 \} \cup h_{a}$ using the new $u^{n+1}$ is generated and the LP-Problem is solved again.
@@
 This effect did not occur so far. 
-(Note: A approach on this would be to make a educated guess which indecies could get a problem with positifity. A possible estimate would be $\frac{u_i^{n+1}}{max(K_{(i,0)}, \cdots ,K_{(i,0)})} $ )
+(Note: A approach on this would be to make a educated guess which indices could get a problem with positivity. A possible estimate would be $\frac{u_i^{n+1}}{max(K_{(i,0)}, \cdots ,K_{(i,0)})} $ )
 
@@
 The inner loop loops through the orders, starting with the highest Order down to the lowest Order.
-Until a acceptable $b$ the algorithm tests combiantions of $\theta$ and $p$.
+Until an acceptable $b$ the algorithm tests combinations of $\theta$ and $p$.
 At first the Order Conditions are constructed.
@@
 This is necessary because sometimes the used LP-solvers incorrectly identify the problem as feasible.
-(Note: because of the chosen objective function it can not be unbounded) As second test the deviation from the old solution (Compare Error estimation) is calculated.
+(Note: because of the chosen objective function it cannot be unbounded) As second test the deviation from the old solution (Compare Error estimation) is calculated.
 If the error is bigger than a set maximum error the $b$ is also rejected and the algorithm also switch to the next combination.
 As soon as the algorithm finds a suitable $b$ the loops are aborted and the $b$ is used for the update step.
-If all the combinations of $\theta$ and $p$ did not lead to a acceptable $b$ the step has to be rejected and a new set of stage values has to bee computed.
+If all the combinations of $\theta$ and $p$ did not lead to an acceptable $b$ the step has to be rejected and a new set of stage values has to be computed.
 
-(Note: It would also be possible to use another strategy for traversing all the combinations of $(p,\theta)$. Is it worth exploring these? Eg, a visect algorithm for $\Theta$ or linear approximations)
+(Note: It would also be possible to use another strategy for traversing all the combinations of $(p,\theta)$. Is it worth exploring these? Eg, a bisect algorithm for $\Theta$ or linear approximations)
 
@@
 \subsection{Applicable Problems}\label{sec:app_problem}
-The approach can be used with ODE systems $u' = f(t,u)$ that ensure  $u(t) \geq 0 \forall_t \forall_{  u_0 \geq 0}$
-
-This can be tested with the property 
+The approach can be used with ODE systems $u' = f(t,u)$ that ensure  $u(t) \geq 0 \forall t \forall {  u_0 \geq 0}$ 
+A sufficient condition is  
 \begin{equation}
-u_i=0 \Rightarrow f_i(t,u_1,\cdots,u_i,\cdots,u_n) \geq 0  \forall_{u_c \geq 0} \forall_{t}
+u_i=0 \Rightarrow f_i(t,u_1,\cdots,u_i,\cdots,u_n) \geq 0  \forall {u_c \geq 0} \forall {t}
 \end{equation}
-
-For problems where the exact solution is positive for certain $u_0$ but positifity is not preserved for all $u_0 \geq 0$ tests did not show promising results.
+For problems where the exact solution is positive for certain $u_0$ but positivity is not preserved for all $u_0 \geq 0$ tests did not show promising results.
 Additional it is not certain if the computed solutions would be reasonable.

### 2020-02-16 latex/overview.tex
Update overview.tex
Updated Header, minor corrections

@@
-\documentclass{article}
+\documentclass[a4paper]{scrartcl}
+
 \usepackage[utf8]{inputenc}
+%\usepackage[USenglish]{babel}
+\usepackage{csquotes}
+
+%More margins for annotations
+%\usepackage[a4paper, top=2.5cm, bottom=2.5cm, left=2.5cm, right=2.5cm]{geometry}
+\usepackage[a4paper, top=3.5cm, bottom=3.5cm, left=3.5cm, right=3.5cm]{geometry}
+
+\usepackage[plainpages=false,pdfpagelabels,hidelinks,unicode]{hyperref}
+
 \usepackage{amsmath}
@@
 
+% math packages
+\usepackage{amsmath}
+\numberwithin{equation}{section}
+\allowdisplaybreaks
+\usepackage{amssymb}
+\usepackage{commath}
+\usepackage{mathtools}
+\usepackage{bbm}
+\usepackage{nicefrac}
+
+\usepackage{siunitx}
+
+\usepackage{amsthm}
+\theoremstyle{plain}
+  \newtheorem{theorem}{Theorem}
+  \newtheorem{lemma}[theorem]{Lemma}
+  \newtheorem{corollary}[theorem]{Corollary}
+  \newtheorem{proposition}[theorem]{Proposition}
+\theoremstyle{definition}
+  \newtheorem{definition}[theorem]{Definition}
+  \newtheorem{remark}[theorem]{Remark}
+  \newtheorem{example}[theorem]{Example}
+\numberwithin{theorem}{section}
+
+% definitions
+\newcommand{\N}{\mathbb{N}}
+\newcommand{\R}{\mathbb{R}}
+\newcommand{\CN}{\mathbb{C}}
+\renewcommand{\H}{\mathcal{H}}
+\renewcommand{\O}{\mathcal{O}}
+\newcommand{\dt}{{\Delta t}}
+\newcommand{\1}{\mathbbm{1}}
+\newcommand{\e}{\mathrm{e}}
+\newcommand{\scp}[2]{\left\langle{#1,\, #2}\right\rangle}
+\newcommand{\I}{\operatorname{I}}
+\newcommand{\lot}{\ell_{\mathrm{ot}}}
+
+%tikz
 \usepackage{tikz}
@@
 
-\usepackage[backend=biber,url  = false]{biblatex}
+% bibliography
+\usepackage[%
+%   backend=bibtex,bibencoding=ascii,
+  backend=biber,
+%   style=authoryear-comp, dashed=false,
+  style=numeric-comp,
+%   firstinits=true, uniquename=init, %abbreviate first names
+  giveninits=true, uniquename=init, %abbreviate first names
+  natbib=true,
+  url=false,
+  doi=true,
+  isbn=false,
+  backref=false,
+  maxnames=99,
+  ]{biblatex}
 \addbibresource{Bib_RK_optim.bib}
 
-\title{Positivity-preserving adaptive Runge-Kutta methods: results}
-\author{Stephan Nuesslein}
+\title{Positivity-Preserving Adaptive Runge-Kutta Methods}
+\author{Stephan Nüßlein}
+
+\makeatletter
+\hypersetup{pdfauthor={\@author}}
+\hypersetup{pdftitle={\@title}}
+\makeatother
 
@@
 The same happens to the positivity properties.
-By this the positivity of the method is ensured for time steps $\Delta t$ with $\Delta t \leq R \Delta t_{FE}$, where $R$ depends on the used SSPRK and $\Delta t_{FE}$ is the forward euler time step \cite{gottlieb_strong_2011}.
+By this the positivity of the method is ensured for time steps $\dt$ with $\dt \leq R \dt_{FE}$, where $R$ depends on the used SSPRK and $\dt_{FE}$ is the forward euler time step \cite{gottlieb_strong_2011}.
 The Modified Patanka Runge Kutta (MPRK) are 
@@
 \begin{equation}
-y^{(j)} =  u^n + \Delta t \sum_{k = 1}^{s} a_{jk} f(t^n + \Delta t c_k,y^{(k)})  \quad j = 1,\cdots,s
+y^{(j)} =  u^n + \dt \sum_{k = 1}^{s} a_{jk} f(t^n + \dt c_k,y^{(k)})  \quad j = 1,\cdots,s
 \end{equation}
@@
 \begin{equation}
-u^{n+1} = u^n + \Delta t \sum_{j  = 1}^s f(t^n + \Delta t c_j,y^{(j)}) b_j .
+u^{n+1} = u^n + \dt \sum_{j  = 1}^s f(t^n + \dt c_j,y^{(j)}) b_j .
 \end{equation}
@@
 \begin{equation}
-f^{(j)} = f(t^n + \Delta t c_j,y^{(j)})
+f^{(j)} = f(t^n + \dt c_j,y^{(j)})
 \end{equation}
@@
 \begin{equation}\label{eq:Combination}
-u^{n+1} = u^n + \Delta t F b
+u^{n+1} = u^n + \dt F b
 \end{equation}
@@
 \end{equation}
-When solved with the Cash-Karp (CK5) method and $\Delta t = 0.005$ the approximated solution contains negative values. This causes qualitatively wrong solutions to the problem. 
+When solved with the Cash-Karp (CK5) method and $\dt = 0.005$ the approximated solution contains negative values. This causes qualitatively wrong solutions to the problem. 
 In Figure\,\ref{fig:exampleI} the obtained results are plotted with a dashed line. 
@@
 Now the $b$ are adapted. The results are also plotted in Figure\,\ref{fig:exampleI}. 
-All $u\geq0$ is now fulfilled. Because of this a qualitatively correct solution is obtained.
+All $u\geq0$ is now fulfilled. A qualitatively correct solution is obtained.
 
@@
     \caption{Numerical approxiamtion of Reaction Problem from \cite{kopecz_comparison_2019}.
-Computed with ck5 and $\Delta t = 0.005$.
+Computed with ck5 and $\dt = 0.005$.
 The dashed line are the reference solution without the adaption of the $b$.
@@
 For $t>2.63$ the original set of weights lead to a positive solution.
-Because of this the used weights equal the original weights again.
+Because of this, the used weights equal the original weights again.
 
@@
 For this we have to consider which $b$ to choose. 
-Generally, any $b \in \mathbb{R}$ can be used. 
+Generally, any $b \in \R$ can be used. 
 A first requirement would be that the new solution is positive.
 We can further reduce this space by enforcing the Order Conditions. 
-By incorporating the Order Conditions, the space of $b$ is limited to a subspace of $\mathbb{R}$. %(Note: not linear because 0 is not in subspace)
+By incorporating the Order Conditions, the space of $b$ is limited to a subspace of $\R$. %(Note: not linear because 0 is not in subspace)
 Finally, we have to give a objective function to get the optimal $b$.
@@
 \end{align}
-If simulated with $\Delta t = 1$ the stage values are 
+If simulated with $\dt = 1$ the stage values are 
 \begin{equation}
@@
 \begin{equation}
-u^{1} = u^0 + \Delta t  \left[f^1,f^2\right] b = \left(\begin{matrix}\frac{5}{2}\\- \frac{5}{2}\end{matrix}\right) \alpha  \left(\begin{matrix}-15\\15\end{matrix}\right)
+u^{1} = u^0 + \dt  \left[f^1,f^2\right] b = \left(\begin{matrix}\frac{5}{2}\\- \frac{5}{2}\end{matrix}\right) \alpha  \left(\begin{matrix}-15\\15\end{matrix}\right)
 \end{equation}
@@
 
-The vector $b \in \mathbb{R}^s$ contains the weights and $Q$ is an $n \times s$ matrix, where $g$ is the number of Order Conditions.
-The vector $r$ contains the right hand side of the Order Conditions.
+The vector $b \in \R^s$ contains the weights and $Q$ is an $n \times s$ matrix, where $g$ is the number of Order Conditions. The vector $r$ contains the right hand side of the Order Conditions.
 
@@
 \begin{equation}
-u_{n+1}=u_n+\Delta K \vec{b} \geq 0
+u_{n+1}=u_n+\dt K \vec{b} \geq 0
 \end{equation}
@@
 
-If the original step $\Delta t$ was too large and there is no suitable $b$, we would like to calculate a $u^{n+1}$ using the stage values instead of rejecting the step.
+If the original step $\dt$ was too large and there is no suitable $b$, we would like to calculate a $u^{n+1}$ using the stage values instead of rejecting the step.
 For this the time step is reduced after calculation the stage values.
@@
 These can be found in \cite{hairer_runge-kutta_1993}.
-We introduce $\theta \in (0,1]$. The new step taken has the length $\theta \Delta t$ whereas $\Delta t$ is the time step used when calculating the stage values.
+We introduce $\theta \in (0,1]$. The new step taken has the length $\theta \dt$ whereas $\dt$ is the time step used when calculating the stage values.
 To calculate a reduced step the Order Conditions have to be adapted and a new $b_{orig}$ for the objective function (and more important the Error detection afterwards) has to be generated. 
@@
 the Heat Equation solved with an implicit method of higher order), large distortions occur, even though a norm of the distortion of the b is used as an objective function.
-For implicit methods, it is known that the backward euler steps lead to a positive solution for any $\Delta t$ \cite{hundsdorfer_numerical_2003}.
+For implicit methods, it is known that the backward euler steps lead to a positive solution for any $\dt$ \cite{hundsdorfer_numerical_2003}.
 The backward Euler is still only first Order but shows more exact results than a method of higher Order that is distorted by the adaption of the $b$, at least at some steps.
@@
 The positivity constraint changes to 
-$$u_{n+1}=u_n+\Delta t K B a \geq 0$$
+$$u_{n+1}=u_n+\dt K B a \geq 0$$
 where $a = [a_1,\cdots,a_d]^T$
@@
 \subsubsection{Construction of Embedded Methods}
+(TODO: have a look on which of the ideas are still valid, maybe rewrite section completely)
+
 For methods with $s \geq p$ it is useful to have some degrees of freedom for the highest order.
@@
 
-Another important property is the stability region in comparison with the maximum $\Delta t$ that ensures positivity.
-For this approach to be useful the RK-method has to yield negative results for $\Delta t$ smaller than the biggest stable $\Delta t$.
+Another important property is the stability region in comparison with the maximum $\dt$ that ensures positivity.
+For this approach to be useful the RK-method has to yield negative results for $\dt$ smaller than the biggest stable $\dt$.
 
@@
 This corresponds to the question if there is an embedded first order method that ensures positivity.
-This is particularly interesting for Implicit methods, because the $\Delta t$ is not limited by the stability.
+This is particularly interesting for Implicit methods, because the $\dt$ is not limited by the stability.
 If we have an embedded first Order Method we can ensure that there is always a positive solution, which might be of first Order.
@@
 This can only happen if the Truncation error is large enough.
-This requires $\Delta t \gg 0$ where lower coefficients of the Taylor Series are no longer sufficient to approximate the behavior.
+This requires $\dt \gg 0$ where lower coefficients of the Taylor Series are no longer sufficient to approximate the behavior.
 One could say that the main idea of the adaption of $b$ is to alter the error in a certain way, so that the numerical solution gets 'better' in the sense that it leads to positive solutions. 
-The local error of a RKM $e(\Delta t) =u(t_0 + \Delta t) - u^1$ can be expressed using the Taylor series of the RKM and the Taylor series of the exact solution. % Hairer I P134
+The local error of a RKM $e(\dt) =u(t_0 + \dt) - u^1$ can be expressed using the Taylor series of the RKM and the Taylor series of the exact solution. % Hairer I P134
 
 \begin{align}\label{eq:Taylor_sol_ref}
-u(t_0 + \Delta t) &= u(t_0) + u'(t_0) \Delta t + \frac{u''(t_0)}{2} \Delta t^2 + \cdots + \frac{u^{(p)}(t_0)}{p!} \Delta t^p + \frac{u^{(p+1)}(t_0)}{(p+1)!} \Delta t^{p+1} + \cdots \\
-u^{n+1} &= u(t_0)  + v_1 \Delta t + \frac{v_2}{2} \Delta t^2 + \cdots + \frac{v_p}{p!} \Delta t^p + \frac{v_{p+1}}{(p+1)!} \Delta t^{p+1} + \cdots \\
+u(t_0 + \dt) &= u(t_0) + u'(t_0) \dt + \frac{u''(t_0)}{2} \dt^2 + \cdots + \frac{u^{(p)}(t_0)}{p!} \dt^p + \frac{u^{(p+1)}(t_0)}{(p+1)!} \dt^{p+1} + \cdots \\
+u^{n+1} &= u(t_0)  + v_1 \dt + \frac{v_2}{2} \dt^2 + \cdots + \frac{v_p}{p!} \dt^p + \frac{v_{p+1}}{(p+1)!} \dt^{p+1} + \cdots \\
 \end{align}
@@
 
-The main way is to ensure that the RKM with adapted b converges to the correct solution is to ensure that the $b$ approaches $b_{orig}$ for $\Delta t \to 0$.
+The main way is to ensure that the RKM with adapted b converges to the correct solution is to ensure that the $b$ approaches $b_{orig}$ for $\dt \to 0$.
 For this $b$ it is already known that the numeric solution converges to the exact solution.
@@
 \begin{align}
-err = |u(t^n)-u^n| &= |u(t^n) - (u^n_{b_{orig}+\Delta t K(b-b_{orig})})| \\
- &\leq \underbrace{|u(t^n)-u^n_{b_{orig}}|}_{\approx err_T}+\underbrace{|\Delta t K(b-b_{orig})|}_{= err_{adapt}}
+err = |u(t^n)-u^n| &= |u(t^n) - (u^n_{b_{orig}+\dt K(b-b_{orig})})| \\
+ &\leq \underbrace{|u(t^n)-u^n_{b_{orig}}|}_{\approx err_T}+\underbrace{|\dt K(b-b_{orig})|}_{= err_{adapt}}
 \end{align}
 
-After adapting the $b$ the Approximation Error is checked. If $|\Delta t K(b-b_{orig})|$ is larger than the tolerance the $b$ is rejected.
+After adapting the $b$ the Approximation Error is checked. If $|\dt K(b-b_{orig})|$ is larger than the tolerance the $b$ is rejected.
 
@@
 By adapting the $b$ the used RK Method is changed. This also alters the stability function and therefore changes the region of absolute stability. 
-It depends on the solved problem, the RKM and the used $\Delta t$ on how many steps the stability function is altered.
+It depends on the solved problem, the RKM and the used $\dt$ on how many steps the stability function is altered.
 For Problems where only a small number of step are affected a bigger change in the stability function would be acceptable.
@@
 \begin{align}
-u_{n+1} &= u_n + \Delta t K (\alpha b_a+\beta b_b) \label{eq:u_n+1}  \\
-u_{n+1}^{sum} &= \alpha (u_n + \Delta t K b_a) + \beta (u_n + \Delta t K b_b) \label{eq:u_n+1_sum} \\
-& = (\alpha + \beta) u_n + \Delta t \alpha K b_a + \Delta t \beta K b_b \\
-&= u_n + \Delta t K (\alpha b_a+\beta b_b) \\
+u_{n+1} &= u_n + \dt K (\alpha b_a+\beta b_b) \label{eq:u_n+1}  \\
+u_{n+1}^{sum} &= \alpha (u_n + \dt K b_a) + \beta (u_n + \dt K b_b) \label{eq:u_n+1_sum} \\
+& = (\alpha + \beta) u_n + \dt \alpha K b_a + \dt \beta K b_b \\
+&= u_n + \dt K (\alpha b_a+\beta b_b) \\
 &=u_{n+1} 
@@
 \end{equation}
-This function can be viewed either as $\mathbb{R}^m \times  \mathbb{C} \rightarrow \mathbb{R}$ or $\mathbb{R}^m \times  \mathbb{R}^2 \rightarrow \mathbb{R}$.
+This function can be viewed either as $\R^m \times  \CN \rightarrow \R$ or $\R^m \times  \R^2 \rightarrow \R$.
 We are only concerned with the expansion or contraction. Because of this we are only interested in the change perpendicular to the Border of the stability region.
@@
  $|R(w,u,v)|^2$ is calculated. 
-If the gradient does not vanish at $z = z_0$ \eqref{eq:border} Can be transformed in a function $\mathbb{R}^m \times \mathbb{R} \rightarrow \mathbb{R}$ depending on $w$ and a single variable $a$ by setting $z = z_0 + n a$.
+If the gradient does not vanish at $z = z_0$ \eqref{eq:border} Can be transformed in a function $\R^m \times \R \rightarrow \R$ depending on $w$ and a single variable $a$ by setting $z = z_0 + n a$.
 The complex number $n$ denotes the normal vector $\vec{n} = \frac{\nabla |R(w,u,v))|^2}{\left| \nabla |R(w,u,v))|^2 \right|}$ of the border of the stability region written in complex notation.
@@
 \begin{equation}\label{eq:f(w,a)}
-f: \mathbb{R}^m \times \mathbb{R} \rightarrow \mathbb{R} f(a) = |R(w,z_0 + n a)|^2 -1 
+f: \R^m \times \R \rightarrow \R f(a) = |R(w,z_0 + n a)|^2 -1 
 \end{equation}
@@
 For the further discussion, we are concerned with a System of $m$ coupled ODEs. 
-In this setup $u \in \mathbb{R}^m$.
+In this setup $u \in \R^m$.
 Systems of ODEs with a high number of dimensions can appear when solving a PDE with the method of lines.
@@
 \begin{equation}
-u_i + \Delta t K  b \geq 0     
+u_i + \dt K  b \geq 0     
 \end{equation}
@@
 
+Reaction Problem with step-size control
+
 \subsection{Dense Output}

### 2020-02-17 latex/overview.tex
Updated the section regarding stability, Added some plots
overview.tex:
Rewrote the section regarding the region of absolut stability

plots/Stability.ipynb:
Notebook to produce Plots of stabiliy regions of altered methods

@@
 
-The stability function of a RKM $R_b$ with the weights $b = \alpha b_1 + \beta b_2$ cn be calculated by adding the weighted stability functions $R_{\alpha b_a+\beta b_b}(z) = \alpha R_{\alpha b_a}(z) + \beta R_{\beta b_b}(z) $ where all stability functions have the same $A$ and $\alpha + \beta = 1$.
+\begin{figure}
+     \centering
+     \begin{subfigure}[b]{0.45\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{plots/stab_ck5.pdf}
+         \caption{Cash-Karp 5}
+         \label{fig:stab_ck5}
+     \end{subfigure}
+     \hfill
+     \begin{subfigure}[b]{0.45\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{plots/stab_ex3.pdf}
+         \caption{Implicit Extrapolation 3}
+         \label{fig:stab_ex3}
+     \end{subfigure}
+        \caption{Change of stability region for direct adaptation(a) and for convex combination(b) (TODO: legend, axis, description in text)}
+        \label{fig:stab}
+\end{figure}
 
-Let $R_b(z)$ denote the stability function of a RKM method with the butcher tableau
+\paragraph*{Version 1} $ $\\
 
-$$
-\begin{array}
-{c|c}
-c & A\\
-\hline
-& b^T
-\end{array}
-$$
+The stability function is defined as 
+\begin{equation}\label{eq:stab_func}
+R(z) = 1 + zb^T(I - zA)^{-1}e
+\end{equation}
+We can define a function $P(z)$
+\begin{equation}
+P(z) = R(z)-1 = z b^T (I - zA)^{-1}
+\end{equation}
+These Polynomials are $\in \mathcal{L}^\infty$.
+(Note: for all explicit methods this is always true, for implicit methods we have to exclude the poles $\{z | \det(I - zA) = 0\}$ or  $\{z | \det(I - zA) > \epsilon \}$ )
+We define $\mathcal{P}$ as all $P(z)$ for the class of $RKM_A$. 
+This is a vector space.
+Proof:
+\begin{multline}
+\alpha P_a(z) + \gamma P_c(z) =  \alpha z a^T (I - zA)^{-1} + \gamma z c^T (I - zA)^{-1} =\\
+z (\alpha a+ \gamma c)^T (I - zA)^{-1} = P_{\alpha a + \gamma c}(z) \in \mathcal{P}
+\end{multline}
 
-We want to prove that $R_{\alpha b_a+\beta b_b}(z)$ = $\alpha R_{ b_a}(z)+\beta R_{b_b}(z)$ when $\alpha + \beta = 1$ and $A$ is the same for all methods.
+The $P(z)$ is a linear combination of the Polynomials
+\begin{subequations}
+\begin{align}
+p_1(z) &= z e_1^T (I - zA)^{-1}\\
+\vdots \\
+p_s(z) &= z e_s^T (I - zA)^{-1}
+\end{align}
+\end{subequations}
+whereas $e_1,\cdots,e_s$ is the standard basis of $\R^s$.
+These span $\mathcal{P}$. If $p_{e_1}(z),\cdots,p_{e_s}(z)$ are linearly independent $p_{e_1}(z),\cdots,p_{e_s}(z)$ are a basis of $\mathcal{P}$.
 
-We start with the definition of the RKM
-\begin{align}
-u_{n+1} &= u_n + \dt K (\alpha b_a+\beta b_b) \label{eq:u_n+1}  \\
-u_{n+1}^{sum} &= \alpha (u_n + \dt K b_a) + \beta (u_n + \dt K b_b) \label{eq:u_n+1_sum} \\
-& = (\alpha + \beta) u_n + \dt \alpha K b_a + \dt \beta K b_b \\
-&= u_n + \dt K (\alpha b_a+\beta b_b) \\
-&=u_{n+1} 
-\end{align}
-Now we insert the Stability functions for \ref{eq:u_n+1} and \ref{eq:u_n+1_sum} which we already know to be equal. We get 
-\begin{align}
-u_{n+1} &= u_{n+1}^{sum} \\
-R_{\alpha b_a+\beta b_b}(z) u_n &= \alpha R_{\alpha b_a}(z) u_n + \beta R_{\beta b_b}(z) u_n \\
-&= (\alpha R_{\alpha b_a}(z) + \beta R_{\beta b_b}(z)) u_n
-\end{align}
-because this has to be true for all $u_n$ we get
+With this results the stability function $R_b(z)$ of any RKM in $RKM_{(A)}$ can be calculated with 
+\begin{equation}
+R_b(z) = 1 + P_b(z) = 1 + \sum_{j = 1}^s b_s p_{e_1}
+\end{equation}
 
-\begin{equation}
-R_{\alpha b_a+\beta b_b}(z) = \alpha R_{\alpha b_a}(z) + \beta R_{\beta b_b}(z) 
+\paragraph*{Version 2} $ $\\
+The stability function is defined as a function $\CN \to \CN$ with 
+\begin{equation}\label{eq:stab_func}
+R(z) = 1 + zb^T(I - zA)^{-1}e
 \end{equation}
+We can add $b$ as a second parameter and get a function $\CN \times \R^s \to \CN$ with
+\begin{equation}\label{eq:stab_func}
+R(z,b) = 1 + zb^T(I - zA)^{-1}e
+\end{equation}
+The stability function is a linear function in $b$.
+
+\paragraph*{Version 3} $ $\\
+(Ignore this topic, because the following sections also can work without it)
 
@@
 \subsubsection{Stability of convex combination}
-If the new $b$ is chosen as a convex combination of $b$ it is easy to prove that the resulting region of absolute stability is at least the Intersection of the regions of absolute stability of the methods that define the convex combination.
+If the new $b$ is chosen as a convex combination of $b$ it is easy to prove properties of the stability region.
+The stability function $R_{\chi a + (1-\chi) c}(z)$ of a convex combination can be calculated by the weighted sum of the Stability functions $R_a(z)$ and $R_c(z)$
+
+Proof:
+\begin{multline}
+R_{\chi a + (1-\chi) c}(z) = R(z) = 1 + (\chi a + (1-\chi) c)^Tz(I - zA)^{-1}e \\
+= \chi + (1-\chi) + \chi a ^Tz(I - zA)^{-1}e + (1-\chi) c^Tz(I - zA)^{-1}e \\
+= \chi R_a(z) + (1-\chi) R_c(z)
+\end{multline}
+
+The resulting region of absolute stability is at least the Intersection of the regions of absolute stability of the methods that define the convex combination.
 This also means that if all the used embedded methods used to construct the new $b$ are A-Stable, the resulting method is also A-Stable.
@@
 This can be formulated as $(|R_{b_a}(z)|  < 1) (|R_{b_b}(z)| < 1) \Rightarrow |R_{\chi b_a +(1- \chi) b_b}(z)| < 1$ with $\chi \in [0,1]$.
-We write $\alpha = \chi$ and $\beta = 1-\chi$ for convenience.
 
-\begin{align}
-|R_{\alpha b_a +\beta b_b}(z)| &= |\alpha R_{b_a}(z) + \beta R_{b_a}(z)| \leq |\alpha R_{b_a}(z)| + |\beta R_{b_a}(z)|\\
- &=| \alpha| \underbrace{|R_{b_a}(z)|}_{<1} + |\beta| \underbrace{|R_{b_a}(z)|}_{<1} < \alpha + \beta = 1
-\end{align}
+\begin{multline}
+|R_{\chi b_a +(1-\chi) b_b}(z)| = |\chi R_{b_a}(z) + (1-\chi) R_{b_a}(z)| \leq |\chi R_{b_a}(z)| + |(1-\chi) R_{b_a}(z)|\\
+ =| \chi| \underbrace{|R_{b_a}(z)|}_{<1} + |(1-\chi)| \underbrace{|R_{b_a}(z)|}_{<1} < \chi + (1-\chi)= 1
+\end{multline}
 
@@
 A useful property of the resulting stability region is that it is similar to the stability region of the baseline method. 
-(Note: is it possible to give some statement like: If the stability regions of the RKM used for multiple steps are similar, then the overall method is stable for the intersection of the stability regions)
+(Note: is it possible (or needed at all) to give some statement like: If the stability regions of the RKM used for multiple steps are similar, then the overall method is stable for the intersection of the stability regions.)
 
@@
 
-The stability function depending on the $b$ is expressed (Note: it is a bit complicated her because we cannot let it depend on the b directly, but only give b after a coordinate transformation, which also makes the expression easier because we can already introduce the order conditions here)
+The stability function depending on the $b$ is expressed 
+When adapting the $b$ we cannot change them independently but only in ways that comply with the order conditions
 $m$ is the number of dimensions of the Kernel of the Order Condition $m = \mathrm{dim}\{\mathrm{Ker} (Q) \}$
-The $b_1,\cdots,b_m$  are the Basis vectors of $\mathrm{Ker} (Q)$. By this we can get the vector $w$ by $w = \left[b_1,\cdots,b_m\right]^{-1}(b-b_{orig})$.
+The $b_1,\cdots,b_m$  are the Basis vectors of $\mathrm{Ker} (Q)$. With an coordinate Transformation we get the coordinate vector $\hat{b}$ with the Matrix $B = [b_1,\cdots,b_m]$ and $b = B \hat{b}$.
 
@@
 \begin{equation}\label{eq:gen_stabilityf}
-R(w,z) = (1-(w_1,\cdots,w_m))R_{b_{orig}}(z) + w_1 R_{b_1}(z) + \cdots + w_m R_{b_m}(z)
+R(\hat{b},z) = 1 +  (b_{orig} +B \hat{b})^T z(I - zA)^{-1}
 \end{equation}
@@
 \begin{equation}\label{eq:border}
-|R(w,u+iv)|^2 -1 = 0
+|R(\hat{b},u+iv)|^2 -1 = 0
 \end{equation}
@@
  $|R(w,u,v)|^2$ is calculated. 
-If the gradient does not vanish at $z = z_0$ \eqref{eq:border} Can be transformed in a function $\R^m \times \R \rightarrow \R$ depending on $w$ and a single variable $a$ by setting $z = z_0 + n a$.
-The complex number $n$ denotes the normal vector $\vec{n} = \frac{\nabla |R(w,u,v))|^2}{\left| \nabla |R(w,u,v))|^2 \right|}$ of the border of the stability region written in complex notation.
-This defines the function $a(w)$ with $|R(w,z_0 + n a(w))|^2 -1 = 0$.
+If the gradient does not vanish at $z = z_0$ \eqref{eq:border} Can be transformed in a function $\R^m \times \R \rightarrow \R$ depending on $\hat{b}$ and a single variable $a$ by setting $z = z_0 + n a$.
+The complex number $n$ denotes the normal vector $\vec{n} = \frac{\nabla |R(\hat{b},u,v))|^2}{\left| \nabla |R(\hat{b},u,v))|^2 \right|}$ of the border of the stability region written in complex notation.
+This defines the function $a(\hat{b})$ with $|R(\hat{b},z_0 + n a(\hat{b}))|^2 -1 = 0$.
 This implies that this approach is only applicable if the gradient does not vanish on the border of the stability function. This is not true in general but for the most common RKM. 
-The derivative $\frac{\mathrm d}{\mathrm d w} (a(w))$ can be calculated using the implicit function theorem. The defining function is 
+The derivative $\frac{\mathrm d}{\mathrm d \hat{b}} (a(\hat{b}))$ can be calculated using the implicit function theorem. The defining function is 
 \begin{equation}\label{eq:f(w,a)}
-f: \R^m \times \R \rightarrow \R f(a) = |R(w,z_0 + n a)|^2 -1 
+f: \R^m \times \R \rightarrow \R f(\hat{b},a) = |R(\hat{b},z_0 + n a)|^2 -1 
 \end{equation}
-If explicit methods are used this function is known to be continuously differentiable because it can be written as a Polynomial of $w_1,\cdots,w_m,u$ and $v$.
+If explicit methods are used this function is known to be continuously differentiable because it can be written as a Polynomial of $\hat{b},u$ and $v$.
 This is a necessary condition for the use of the of the Implicit function theorem.
-It is also known by definition, that for the point $w=0$ and $z = z_0$ the equation $ |R(w,z_0)|^2 -1 = 0 $ is satisfied.
-For simplicity only the formulas for the derivative at the value $b = b_{orig}$ are given. This corresponds to $w_1 = \cdots = w_m = 0$.
+It is also known by definition, that for the point $b = b_{orig}$ and $z = z_0$ the equation $ |R(\hat{b},z_0)|^2 -1 = 0 $ is satisfied.
+For simplicity only the formulas for the derivative at the value $b = b_{orig}$ are given. This corresponds to $\hat{b} = 0$.
 The wanted derivative  is given by
@@
 \begin{equation}
- \frac{\partial a}{\partial w_j} (w) =
- - \left[ J_{f,a}(w,a(w))  \right] ^{-1} 
-   \left[ \frac{\partial f}{\partial w_j}(w,a(w)) \right]
+ \frac{\partial a}{\partial \hat{b}_j} (w) =
+ - \left[ \frac{\partial f}{\partial a} (\hat{b},a(\hat{b}))  \right] ^{-1} 
+   \left[ \frac{\partial f}{\partial \hat{b}_j}(\hat{b},a(\hat{b})) \right]
 \end{equation}
-as long as the inverse of the Jacobian exists. This is true in most cases as it has been discussed above.
+as long as $\frac{\partial f}{\partial a} \neq 0$ on the border of the region of absolute stability. This is true in most cases as it has been discussed above.
 
-The Jacobian of \eqref{eq:f(w,a)} $J_{f,a}(w,a(w))$ evaluated at $w=0$ can be calculated using the directional derivative
+The Derivative of \eqref{eq:f(w,a)} in respect to $a$ evaluated at $\hat{b}=0$ can be calculated using the directional derivative
 \begin{align*}
- J_{f,a}(b_{orig},a(w)) \Big|_{w=0} =& 
- \frac{\partial f}{\partial a} (w,a(w)) \Big|_{w=0} = 
- \nabla |R(w,a(w)|^2 \vec{n} \Big|_{w=0} \\
+ \frac{\partial f}{\partial a} \Big|_{\hat{b}=0} =& 
+ \frac{\partial f}{\partial a} (\hat{b},a(\hat{b})) \Big|_{\hat{b}=0} = 
+ \nabla |R(\hat{b},a(\hat{b})|^2 \vec{n} \Big|_{\hat{b}=0} \\
 =& \left| \nabla|R(0,z_0)|^2 \right| = \left| \nabla|R_{b_{orig}}(z_0)|^2 \right|
@@
 
-The derivative in respect to $w_j$ is 
-$ \frac{\partial f}{\partial w_j}(w,a(w)) \Big|_{w=0}$
+The derivative in respect to $\hat{b}_j$ is 
+$ \frac{\partial f}{\partial \hat{b}_j}(\hat{b},a(\hat{b})) \Big|_{\hat{b}=0}$
 this simplifies to 
-\begin{align*}\label{eq:derivative_to_b}
- \frac{\partial f}{\partial w_j}(w,a(w)) \Big|_{w=0} =&
-- 2 \underbrace{R_{b_{orig}}(z_0)R^*_{b_{orig}}(z_0)}_{=|R_{b_{orig}}(z_0)|^2=1} + R_{b_{orig}}(z_0)R^*_{b_v}(z) + R_{b_v}(z_0)R^*_{b_{orig}}(z_0) \\
-=& -2 + R_{b_{orig}}(z_0)R^*_{b_v}(z_0) + R_{b_v}(z_0)R^*_{b_{orig}}(z_0)
-\end{align*}
+\begin{multline}\label{eq:derivative_to_b}
+ \frac{\partial f}{\partial \hat{b}_j}(w,a(w)) \Big|_{\hat{b}=0} = 
+ \frac{\partial }{\partial \hat{b}}(R(\hat{b},z)R(\hat{b},z^*)) \\
+ = (B\hat{b})^T (I-zA)^{-1} R_{b_{orig}}(z^*) + (B\hat{b})^T (I-z^*A)^{-1} R_{b_{orig}}(z)
+\end{multline}
 using the fact that $z_0$ is on the border of the stability region.
-The derivative $-2 + R_{b_{orig}}(z_0)R^*_{b_v}(z_0) + R_{b_v}(z_0)R^*_{b_{orig}}(z_0)$ is known to be finite for explicit methods, because $R_{b_v}$ and $R_{b_{orig}}$ are Polynomials and $|z_0| < \infty$
+The derivative $(B\hat{b})^T (I-zA)^{-1} R_{b_{orig}}(z^*) (B\hat{b})^T (I-z^*A)^{-1} R_{b_{orig}}(z)$ is known to be finite for explicit methods, because it is a polynomial and $|z_0| < \infty$
 The same process can be done for all points on the border of the stability region. 
-As long as $\left[ J_{f,a}(w,a(w))  \right] > 0$ the derivative $\frac{\partial a}{\partial w_j} (w)$ is finite.
+As long as $\nabla|R(\hat{b},u+iv)|^2 \neq 0 \forall \hat{b} < \epsilon  \forall z$ the derivative $\frac{\partial a}{\partial \hat{b}_j} (\hat{b})$ is finite.
 This means that a small change the $b$ only leads to a small change of the stability region.

### 2020-02-18 latex/overview.tex
Changes in Introduction, Corrections
overview.tex
Minor corrections

Bib_RK_optim.bib:
New sources

@@
 
-Therefore, it is interesting to know how to use RKMs that preserve positifity.
+Therefore, it is interesting to know how to use RKMs that preserve positivity.
 For RKMs some theoretical limitations are known.
@@
 This allows them to be positive and higher order.
-(TODO: two sentences here)
 
@@
 With this approach, it is possible to have different properties of the RKM at different parts of the domain.
+Another method that adapts the $b$ are the relaxation Runge–Kutta (RRK) methods. 
+In these the weights are scaled by a scalar relaxation parameter to conserve convex quantities.
+This can be used to conserve the energy or entropy \cite{ranocha_relaxation_2019,ketcheson_relaxation_2019}
 
 In this paper, an adaption of the $b$ is done by solving a Linear programming (LP)-Problem.
-(TODO: Description LP-Problem)
+These are convex optimization problems where all constraints and the objective function are all affine. These can incorporate both equality and inequality constraints. 
+LP-Problems can be quite reliably solved with the Simplex method or with interior point methods. This can be done in reasonable time even for large problems. 
+\cite{boyd_convex_2004}
 
@@
 In section\,\ref{sec:imple} further details that were used when implementing the algorithm are described.
-In section\,\ref{sec:Numeric_Results} Numerical results are given for multiple testproblems.
+In section\,\ref{sec:Numeric_Results} Numerical results are given for multiple test problems.
 A conclusion is given in section\,\ref{sec:conclusion}
@@
 A first requirement would be that the new solution is positive.
-We can further reduce this space by enforcing the Order Conditions. 
-By incorporating the Order Conditions, the space of $b$ is limited to a subspace of $\R$. %(Note: not linear because 0 is not in subspace)
+We can further reduce this space by enforcing the order conditions. 
+By incorporating the order conditions, the space of $b$ is limited to a subspace of $\R$. %(Note: not linear because 0 is not in subspace)
 Finally, we have to give a objective function to get the optimal $b$.
@@
 \end{align}
-If simulated with $\dt = 1$ the stage values are 
+If simulated with $\dt = 1$ the stage derivatives are 
 \begin{equation}
@@
 The solution for the original RKM would be $u^{1} = \left(\frac{5}{2} ,- \frac{5}{2}\right)^T$ which contains a negative value.
-But by changing the parameter $\alpha$ and therefore changing the $b$ a positive solution can be reached.
+But by changing the parameter $\alpha$ the weights $b$ and the solution is altered.
 By a suitable choice of $\alpha \in \left[\frac{1}{6},\frac{7}{30}\right]$ any $u$ that complies with mass conservation and positivity can be reached. 
@@
 
-The Order Conditions for a RKM are a set of Equations depending on $a_{jl}$, $c_1,\cdots,c_s$ and $b_1,\cdots,b_s$.
+The order conditions for a RKM are a set of Equations depending on $a_{jl}$, $c_1,\cdots,c_s$ and $b_1,\cdots,b_s$.
 In these Equations $b_1,\cdots,b_s$ only appear linearly \cite{hairer_runge-kutta_1993}.
-If $a_{jl}$ and $c_1,\cdots,c_s$ are already known, the Order Conditions for the Order $p$ are an linear equations system and can be written as $Q b = r$.
+If $a_{jl}$ and $c_1,\cdots,c_s$ are already known, the order conditions for the Order $p$ are an linear equations system and can be written as $Q b = r$.
 
-The vector $b \in \R^s$ contains the weights and $Q$ is an $n \times s$ matrix, where $g$ is the number of Order Conditions. The vector $r$ contains the right hand side of the Order Conditions.
+The vector $b \in \R^s$ contains the weights and $Q$ is an $g \times s$ matrix, where $g$ is the number of order conditions. The vector $r$ contains the right hand side of the order conditions.
 
@@
 This implies that $\mathrm{rank}(Q) < s$. 
-Because the Quadrature Conditions are linear independent, there are at least $p$ independent vectors in $Q$. This leads to the condition $p < s$.
-This implies that the number of stages has to be higher as the wanted order.
+Because the Quadrature conditions are linear independent, there are at least $p$ independent vectors in $Q$. This leads to the condition $p < s$.
+This implies that the number of stages has to be higher than the wanted order.
 
@@
 
-Both approaches are shown in figure\,\ref{fig:b_space}.
+Both approaches are illustrated in figure\,\ref{fig:b_space}.
 
@@
 
-When directly adapting the $b$ a new $b$ that complies with the Order Conditions is chosen.   
+When directly adapting the $b$ a new $b$ that complies with the order conditions is chosen.   
 The weights $b = b_{orig} + \Delta b$ consist of the weights of the original method $b_{orig}$ and an adaption $\Delta b$. The adaption is $\Delta b \in \mathrm{ker}(Q)$. 
-In Figure\,\ref{fig:b_direct} this is drawn. The red line represents the subspace of $b$ that satisfy the Order Conditions. Both $b_{orig}$ and $b$ are in this subspace.
+In Figure\,\ref{fig:b_direct} this is drawn. The red line represents the subspace of $b$ that satisfy the order conditions. Both $b_{orig}$ and $b$ are in this subspace.
 
@@
 \end{equation}
-The second set of constraints are the Order Conditions
+The second set of constraints are the order conditions
 \begin{equation}
@@
 The different approaches have been tested on a linear test problem from \cite{kopecz_unconditionally_2018}. The optimization problem was used at every step.
-The magnitude of $b$ can get large if the Order Conditions are used as objective function.
+The magnitude of $b$ can get large if the order conditions are used as objective function.
 Large values of b can lead to numerical problems.
@@
 For this the time step is reduced after calculation the stage values.
-This can be done using the Order Conditions for dense Output.
+This can be done using the order conditions for dense Output.
 These can be found in \cite{hairer_runge-kutta_1993}.
 We introduce $\theta \in (0,1]$. The new step taken has the length $\theta \dt$ whereas $\dt$ is the time step used when calculating the stage values.
-To calculate a reduced step the Order Conditions have to be adapted and a new $b_{orig}$ for the objective function (and more important the Error detection afterwards) has to be generated. 
+To calculate a reduced step the order conditions have to be adapted and a new $b_{orig}$ for the objective function (and more important the Error detection afterwards) has to be generated. 
 (TODO: the $b_{orig}$ is still a open question)
@@
 $$ Q b = Q (\chi b_{\alpha} + (1-\chi) b_{\beta}) = \chi Q  b_{\alpha} + (1-\chi) Q b_{\beta}) = \chi r + (1-\chi) r = r$$
-$b_{\alpha}$ and $b_{\beta}$ are the methods that are combined to $b$ using the parameter $\chi$. The Matrix $Q$ and the vector $r$ are the Order Conditions as defined in Section\,\ref{sec:OrderCond}.
+$b_{\alpha}$ and $b_{\beta}$ are the methods that are combined to $b$ using the parameter $\chi$. The Matrix $Q$ and the vector $r$ are the order conditions as defined in Section\,\ref{sec:OrderCond}.
 
@@
 \subsubsection{Construction of Embedded Methods}
-(TODO: have a look on which of the ideas are still valid, maybe rewrite section completely)
 
-For methods with $s \geq p$ it is useful to have some degrees of freedom for the highest order.
-For this at least one $b$ that complies with the order Condition of the highest order and $b \neq b_{orig}$ has been introduced.
-These additional $b$ can be constructed by calculating a Span $S$ of the Kernel $\mathrm{ker} (Q)$ of the Order Condition Matrix $Q$.
-For every vector $s \in S$ one set of weights with $b = b_{orig} + a s$ and one with $b = b_{orig} - a s$ is added.
-The $a$ is a parameter that determines how big a change is allowed.
-
-As a first order method, a $b$ is added that corresponds to a series of backward euler steps. These are known to be yield to a positive result.
-
+As a first order method, a $b$ is added that corresponds to a series of backward euler steps. These are known to be yield to a positive result for any step-size.
+It is also possible to add embedded methods of higher order.
 An important property of the embedded methods is that they should show the desired stability characteristics.
 
-A Reduction of the step size suing the order Conditions for dense output would also be possible.
+A Reduction of the step size suing the order conditions for dense output would also be possible.
 For this a set of embedded methods $b_1(\Theta),\cdots,b_d(\Theta)$ has to be constructed.
@@
 is it worth exploring this or is it better to just leave this open for further research it may be also interesting regardless the study of positivity)
-Everything has downsides...
 
@@
 When changing the weights the solution $u^{n+1}$ is changed. 
-By incorporating the Order Conditions for a certain Order the space of $b$ that can be choose is limited to a linear Subspace. 
+By incorporating the order conditions for a certain order the space of $b$ that can be choose is limited to a linear Subspace. 
 Now it is important to know if the new solution $u^{n+1}$ still resembles the solution of the ODE $u(t_{n+1})$. 
@@
 In this case the RKM is fixed apart from the weights $b$. 
-Because of this $v_1,v_2,\cdots$ are functions of $b$.
+Therefore $v_1,v_2,\cdots$ are functions of $b$.
 Because the $b$ are chosen to ensure the Order Conditions the coefficients $v_1,\cdots,v_p$ are fixed and equal the according derivatives. 
@@
 \end{equation}
-These Polynomials are $\in \mathcal{L}^\infty$.
+These rational functions are $\in \mathcal{L}^\infty$.
 (Note: for all explicit methods this is always true, for implicit methods we have to exclude the poles $\{z | \det(I - zA) = 0\}$ or  $\{z | \det(I - zA) > \epsilon \}$ )
@@
 
-The $P(z)$ is a linear combination of the Polynomials
+The $P(z)$ is a linear combination of the rational functions
 \begin{subequations}
@@
 If the new $b$ is chosen as a convex combination of $b$ it is easy to prove properties of the stability region.
-The stability function $R_{\chi a + (1-\chi) c}(z)$ of a convex combination can be calculated by the weighted sum of the Stability functions $R_a(z)$ and $R_c(z)$
+The stability function $R_{\chi b_a + (1-\chi) b_b}(z)$ of a convex combination can be calculated by the weighted sum of the Stability functions $R_{b_a}(z)$ and $R_{b_b}(z)$
 
@@
 \begin{multline}
-R_{\chi a + (1-\chi) c}(z) = R(z) = 1 + (\chi a + (1-\chi) c)^Tz(I - zA)^{-1}e \\
-= \chi + (1-\chi) + \chi a ^Tz(I - zA)^{-1}e + (1-\chi) c^Tz(I - zA)^{-1}e \\
-= \chi R_a(z) + (1-\chi) R_c(z)
+R_{\chi b_a + (1-\chi) b_b}(z) = R(z) = 1 + (\chi b_a + (1-\chi) b_b)^Tz(I - zA)^{-1}e \\
+= \chi + (1-\chi) + \chi b_a ^Tz(I - zA)^{-1}e + (1-\chi) b_b^Tz(I - zA)^{-1}e \\
+= \chi R_{b_a}(z) + (1-\chi) R_{b_b}(z)
 \end{multline}
@@
 For most methods, a small change of $b$ does not alter the stability function dramatically.
-(TODO: I am missing a piece here from ‘the derivative is finite’ to ‘is simmilar’)
+(TODO: I am missing a piece here from ‘the derivative is finite’ to ‘is similar’)
 
@@
 The stability function depending on the $b$ is expressed 
-When adapting the $b$ we cannot change them independently but only in ways that comply with the order conditions
+When adapting the $b$ we cannot change them independently but only in ways that comply with the order conditions.
 $m$ is the number of dimensions of the Kernel of the Order Condition $m = \mathrm{dim}\{\mathrm{Ker} (Q) \}$
@@
 This function can be viewed either as $\R^m \times  \CN \rightarrow \R$ or $\R^m \times  \R^2 \rightarrow \R$.
-We are only concerned with the expansion or contraction. Because of this we are only interested in the change perpendicular to the Border of the stability region.
+We are only concerned with the expansion or contraction. Because of this we are only interested in the change perpendicular to the border of the stability region.
 
@@
 \begin{multline}\label{eq:derivative_to_b}
- \frac{\partial f}{\partial \hat{b}_j}(w,a(w)) \Big|_{\hat{b}=0} = 
+ \frac{\partial f}{\partial \hat{b}_j}(\hat{b},a(\hat{b})) \Big|_{\hat{b}=0} = 
  \frac{\partial }{\partial \hat{b}}(R(\hat{b},z)R(\hat{b},z^*)) \\
- = (B\hat{b})^T (I-zA)^{-1} R_{b_{orig}}(z^*) + (B\hat{b})^T (I-z^*A)^{-1} R_{b_{orig}}(z)
+ = (B\hat{b}_j)^T (I-zA)^{-1} R_{b_{orig}}(z^*) + (B\hat{b}_j)^T (I-z^*A)^{-1} R_{b_{orig}}(z)
 \end{multline}
@@
  u_i^{n+1} &\geq 0   \;   \forall {i \in \{1, \cdots,m \}}  \\
- u_i^n + h \sum_{j=0}^{s-1} b_j k_i^j  &\geq 0   \;   \forall {i \in \{1,\cdots,m \}}  
+ u_i^n + h \sum_{j=0}^{s-1} b_j f_i^j  &\geq 0   \;   \forall {i \in \{1,\cdots,m \}}  
 \end{align}
@@
 \begin{equation}
-u_i + \dt K  b \geq 0     
+u_i + \dt F  b \geq 0     
 \end{equation}
-where $K = \big[k^1 , \cdots k^{s-1}\big]$.
+where $F = \big[f^1 , \cdots f^{s-1}\big]$.
 
@@
 \begin{equation}
-u_i^n + \Delta b \sum_{j=0}^{s-1} b_j k_i^j  \geq 0   \;   \forall {i \in h \subseteq \{1,\cdot,m \}} 
+u_i^n + \Delta b \sum_{j=0}^{s-1} b_j f_i^j  \geq 0   \;   \forall {i \in h \subseteq \{1,\cdot,m \}} 
 \end{equation}

### 2020-02-18 latex/overview.tex
Added convergence plots for Diffusion problem and Advection-Decay problem
overview.ipynb:
Added Plots and descriptions 

utils.py:
Made output more clear if solver failed

plots/Convergence.ipynb:
Notebook to compute and plot convergence

@@
 
+As test problem the PDE
+\begin{equation}
+\frac{\partial u}{\partial t} = -a \frac{\partial u}{\partial z} - K (u)
+\end{equation}
+is simulated with the method of lines on the domain $x = [0,1]$ with $N =100$ points.
+The Problem consists of a Advection part and an Exponential decay.
+At $x =0$ a inflow boundary with $u(0,t)=1$ is used.
+The initial condition is $u(x,0)=0$
+As semidiscretization an upwinding scheme is used. This leads to positivity preserving ODE
+\begin{equation}
+\frac{\mathrm d}{\mathrm d t} u_i = \frac{a}{\Delta x} \left( u_{i-1} - u_i \right) - K u_{i-1}
+\end{equation}
+The parameters are set to $a=1$ and $K=1$.
+For $\dt \leq 0.0082$ the unaltered method lead to positive solutions. For larger $\dt$ the original method leads to negative values and $b$ is altered. For $\dt >0.016$ the dp5 is no longer stable.
+In Figure\,\ref{fig:conv_expl} the convergence is plotted for the altered and unaltered method.
+
+\begin{figure}[h]
+\centering
+\includegraphics[scale=0.75]{plots/conv_adde.pdf}
+\caption{Convergence adapted dp5}
+\label{fig:conv_expl}
+\end{figure}
+
 \subsection{Implicit methods}
@@
 
-When tested on the diffusion equation with an spike as initial data the $b$ is only changed for the first step. The change of the $b$ can lead to reasonable changes of the hight of the maximum.
+The method is tested on the diffusion equation 
+\begin{equation}
+\frac{\partial }{\partial t} = D \frac{\partial^2}{\partial x^2}
+\end{equation}
+with homogeneous Dirichlet boundary conditions on the domain $x = [-0.5,0.5]$ with $N=100$ points. The equation is semidiscreticed using the 3-point-scheme
+\begin{equation}
+\frac{\mathrm d}{\mathrm d t} u_i = \frac{d}{\Delta x^2} \left( u_{i-1} - 2u_i + u_{i+1} \right)
+\end{equation}
+As initial condition $u^0 = (0,\cdots,0,1,0,\cdots,0)^T$ is used. The diffusion coefficient is$D=1$
+The ODE is solved using the implicit EX3 method. 
+For large (Note: better word for 'not small enough') $\dt$ the method computes negative values for $u^1$.
+These can be corrected by adapting the $b$. 
+If the direct adaptation approach is used the $u^1$ can show gliches/is nonphysical. 
+Alternatively the convex combination approach is used. A first order embedded method is added. The $u^1$ computed with the convex combination are physical.
+For both approaches the remaining steps can be computed with the standard $b$.
+
+In figure\,\ref{fig:conv_impl} the convergence is shown. For $\dt < \num{3e-5} $ the standard $b$ yields to a positive result. For larger $\dt$ the $b$ have to be adapted to ensure positivity.
+The EX3a shows similar convergence properties than the original method. 
+The EX3c has larger errors than the EX3a but leads to physical solutions for all time-steps. 
+\begin{figure}[h]
+\centering
+\includegraphics[scale=0.75]{plots/conv_heat.pdf}
+\caption{Convergence adapted ex3}
+\label{fig:conv_impl}
+\end{figure}
 

### 2020-02-18 utils.py

@@ -42,16 +42,19 @@ def plot_convergence(problem,solver,dt,refernce,step=1,error='abs',dx='1',Norm =
         status,t,u,b = RK_integrate(solver=solver,problem=problem,**Params)
         dif = refernce[:,i]-u[step]
         change[i] = 'c' in status['b']
-        if error == 'abs':
-            err[i] = np.linalg.norm(dif,ord=Norm)
-        elif error == 'rel':
-            err[i] = np.linalg.norm(dif,ord=Norm)/np.linalg.norm(refernce[:,i],ord=Norm)
-        elif error == 'grid': #Grid function Norm (LeVeque Appendix A.5)
-            err[i] = dx**(1/Norm)*np.linalg.norm(dif,ord=Norm)
+        if status['success']:
+            if error == 'abs':
+                err[i] = np.linalg.norm(dif,ord=Norm)
+            elif error == 'rel':
+                err[i] = np.linalg.norm(dif,ord=Norm)/np.linalg.norm(refernce[:,i],ord=Norm)
+            elif error == 'grid': #Grid function Norm (LeVeque Appendix A.5)
+                err[i] = dx**(1/Norm)*np.linalg.norm(dif,ord=Norm)
+            else:
+                print('Error not defined')
+                print(error)
+                raise ValueError
         else:
-            print('Error not defined')
-            print(error)
-            raise ValueError
+            err[i] = np.nan
         sol.append(u[step])
         
     plt.plot(dt,err,'o-')

### 2020-03-09 latex/Image_b_convex.tex
Multiple small corrections
overview.tex:
spelling mistakes and minor notation changes
Changed the second example to make it more reasonable
TODOs: Abstract, Introduction, Formulation LPs, Numerical results, Conclussion

Bib_RK_optim.bib:
Hairer as complete book 

Image_b_direct.tex:
Notation change

Image_b_convex.tex:
Notation change, 
Make it fit into context of direct adaptation

@@
     % Draw axes
-	\coordinate (borig) at (1,2.3);
-    \coordinate (b_a) at (0.8,0.8);
-	\coordinate (b_b) at (2.3,1);
-	\coordinate (b) at (1.1,1.5);
+
+	\coordinate (bo) at (1.2,1.8);
+    \coordinate (b) at (2,1);
+    \coordinate (db) at ($(b)-(bo)$);
+    \coordinate (b1) at ($0.4*(db)+(b)$);
+    \coordinate (b0) at ($-0.2*(db)+(bo)$);;
 
@@
 
-    \draw[thick,red,fill = red!20] (borig) -- (b_a) -- (b_b) -- cycle;
+%\draw[thick,red,*-*] (b0) -- (b1);
+   \node[circle,inner sep=1.5pt,draw,fill,red] (A) at (b0) {};
+   \node[circle,inner sep=1.5pt,draw,fill,red] (B) at (b1) {};
+   \draw[thick,red,-] (b0) -- (b1);
 
-	\draw[thick,->] (0,0) -- node[anchor=south west,pos = 0.95] {$b_{orig}$} (borig);
-    \draw[thick,->] (0,0) -- node[anchor=south west,pos = 0.95] {$b_b$} (b_a);
-    \draw[thick,->] (0,0) -- node[anchor=south west,pos = 0.95] {$b$} (b);
-	\draw[thick,->] (0,0) -- node[anchor=north west,pos = 0.8,fill=white] {$b_b$} (b_b);
+	\draw[thick,->] (0,0) -- node[anchor=south east,pos = 0.8] {$b_{a}$} (b0);
+    \draw[thick,->] (0,0) -- node[anchor=south east,pos = 0.8] {$\tilde{b}$} (b);
+    \draw[thick,->] (0,0) -- node[anchor=north west,pos = 0.8] {$b_{b}$} (b1);
 

### 2020-03-09 latex/overview.tex
Multiple small corrections
overview.tex:
spelling mistakes and minor notation changes
Changed the second example to make it more reasonable
TODOs: Abstract, Introduction, Formulation LPs, Numerical results, Conclussion

Bib_RK_optim.bib:
Hairer as complete book 

Image_b_direct.tex:
Notation change

Image_b_convex.tex:
Notation change, 
Make it fit into context of direct adaptation

@@
 \title{Positivity-Preserving Adaptive Runge-Kutta Methods}
-\author{Stephan Nüßlein}
+\author{Nüßlein, Stephan and Ketcheson, David I.}
 
@@
 Many physical processes can be described with differential equations. 
-The physical property that are involved in these processes often only make sense if they are positive or non-negative. 
-Concentrations or number of particles have to be $\geq 0$.
-Other values may only make sense if they comply with other restrictions. An Example of these would-be probabilities or fractions which have to be $\geq 0$ and $\leq 1$.
-The ordinary differential equations (ODE) or a partial differential equations (PDEs) are often too complex to solve them analytically and therefore require numerical methods to get an approximate solution.
+The physical quantities that are involved in these processes often only make sense if they are positive or non-negative. 
+Concentrations or the number of particles have to be $\geq 0$.
+Other values may only make sense if they comply with other restrictions. An example of these would be probabilities or fractions which have to be $\geq 0$ and $\leq 1$.
+The ordinary differential equations (ODE) or partial differential equations (PDEs) are often too complex to solve them analytically and therefore require numerical methods to get an approximate solution.
 Numerical methods generally do not satisfy these restrictions on the solutions.
@@
 At first the PDE is discretized in space. The resulting semi-discrete problem is an ODE that can be solved with a RKM. 
-It is possible to semi-discretize PDEs in a ways that preserve positivity\cite{kopecz_comparison_2019,kopecz_unconditionally_2018}.
+It is possible to semi-discretize PDEs in ways that preserve positivity\cite{kopecz_comparison_2019,kopecz_unconditionally_2018}.
 
@@
 \cite{hundsdorfer_numerical_2003,bolley_conservation_1978,horvath_positivity_1998}.
-The only well know RKM that is unconditionally positive is the backward euler method, which is obliviously of first order.
+The only well-known RKM that is unconditionally positive is the backward euler method, which is obviously of first order.
 This causes restrictions on the step size for any RKM with a higher order.
@@
 The same happens to the positivity properties.
-By this the positivity of the method is ensured for time steps $\dt$ with $\dt \leq R \dt_{FE}$, where $R$ depends on the used SSPRK and $\dt_{FE}$ is the forward euler time step \cite{gottlieb_strong_2011}.
-The Modified Patanka Runge Kutta (MPRK) are 
-based on the Patanka RKM.
+By this the positivity of the method is ensured for time steps $\dt$ with $\dt \leq R \dt_{FE}$, where $R$ depends on the used SSPRK and $\dt_{FE}$ is the forward Euler time step \cite{gottlieb_strong_2011}.
+The Modified Patankar Runge Kutta (MPRK) are 
+based on the Patankar RKM.
 In these destruction Terms are scaled by the new solution.
@@
 Ensuring the positivity is done by redefining the ODE outside the domain.
-Another technique used are event finding methods.
+Another technique used is event finding methods.
 This approach is implemented in the Matlab ODE Suite.
@@
 
-In the following paper an approach is taken that does not overcome the order imitation but is based on the idea that in practical problems for many time steps positivity can be achieved with a higher order methods.
-The main idea is to adapt the RKM after computation of the stage values in a way that ensures positivity.  
+In the following paper an approach is taken that does not overcome the order limitation but is based on the idea that in practical problems for many time steps positivity can be achieved with higher order methods.
+The main idea is to adapt the RKM after the computation of the stage values in a way that ensures positivity.  
 This means that the computed stage values can be used and don't have to be recomputed.
 With increasing complexity of the right-hand side, a step reject has a high cost.
-This makes the additional cost of adapting the RKM economically.
-The resulting b should comply with the order conditions as far as possible.
+This makes the additional cost of adapting the RKM economical.
+The resulting weights should comply with the order conditions as far as possible.
 
-A common way to calculate two different solutions with one set of stage values is the use of embedded methods for error approximation. 
-The idea of adapting the weights $b$ after calculating the stage values is used in \cite{ketcheson_spatially_2013}
-In this case it is used to adapt the properties of the time integrator for a method of lines simulation on a PDE.
+A common way to calculate two different solutions with one set of stage values is the usage of embedded methods for error approximation. 
+The idea of adapting the weights after calculating the stage values is used in \cite{ketcheson_spatially_2013}.
+In this case it is used to adapt the properties of the time integrator for a method of lines solution of a PDE.
 With this approach, it is possible to have different properties of the RKM at different parts of the domain.
-Another method that adapts the $b$ are the relaxation Runge–Kutta (RRK) methods. 
-In these the weights are scaled by a scalar relaxation parameter to conserve convex quantities.
+Other methods that adapts the weights are the relaxation Runge–Kutta (RRK) methods. 
+In these, the weights are scaled by a scalar relaxation parameter to conserve convex quantities.
 This can be used to conserve the energy or entropy \cite{ranocha_relaxation_2019,ketcheson_relaxation_2019}
 
-In this paper, an adaption of the $b$ is done by solving a Linear programming (LP)-Problem.
-These are convex optimization problems where all constraints and the objective function are all affine. These can incorporate both equality and inequality constraints. 
-LP-Problems can be quite reliably solved with the Simplex method or with interior point methods. This can be done in reasonable time even for large problems. 
+In this paper, an adaptation of the weights is done by solving a Linear program (LP).
+These are convex optimization problems where all constraints and the objective function are affine. These can incorporate both equality and inequality constraints. 
+LPs can be quite reliably solved with the Simplex method or with interior point methods. This can be done in reasonable time even for large problems. 
 \cite{boyd_convex_2004}
 
-The paper unfolds as follows. In section\,\ref{sec:main_idea} the main idea is explained. How the method can be written as LP-Problem is explained in section\,\ref{sec:LP-Problem}.
-In section\,\ref{sec:integration} it is described how the new approach can be used with different RKM, how it can be integrated in a step size control and how the Region of absolute stability can be approximated.
+The paper unfolds as follows. In section\,\ref{sec:main_idea} the main idea is explained. How the method can be written as LP is explained in section\,\ref{sec:LP}.
+Section\,\ref{sec:integration} describes how the new approach can be used with different RKM, how it can be integrated in a step size control and how the Region of absolute stability can be approximated.
 In section\,\ref{sec:imple} further details that were used when implementing the algorithm are described.
 In section\,\ref{sec:Numeric_Results} Numerical results are given for multiple test problems.
-A conclusion is given in section\,\ref{sec:conclusion}
+A conclusion is given in section\,\ref{sec:conclusion}.
 
@@
 \end{align}
-the stage values $y^{(1)},\cdots,y^{(s)}$ are computed according to
+the stage values $y_1,\cdots,y_s$ are computed according to
 \begin{equation}
-y^{(j)} =  u^n + \dt \sum_{k = 1}^{s} a_{jk} f(t^n + \dt c_k,y^{(k)})  \quad j = 1,\cdots,s
+y_j =  u^n + \dt \sum_{k = 1}^{s} a_{jk} f(t^n + \dt c_k,y_k)  \quad j = 1,\cdots,s
 \end{equation}
@@
 \begin{equation}
-u^{n+1} = u^n + \dt \sum_{j  = 1}^s f(t^n + \dt c_j,y^{(j)}) b_j .
+u^{n+1} = u^n + \dt \sum_{j  = 1}^s f(t^n + \dt c_j,y_j) b_j .
 \end{equation}
-The stage derivatives $f^{(1)},\cdots,f^{(s)}$ can be calculated with
+The stage derivatives $f_1,\cdots,f_s$ can be calculated with
 \begin{equation}
-f^{(j)} = f(t^n + \dt c_j,y^{(j)})
+f_j = f(t^n + \dt c_j,y_j)
 \end{equation}
@@
 \end{equation}
-where $F$ is a matrix containing the stage derivatives $f^{(1)},\cdots,f^{(s)}$.
+where $F$ is a matrix containing the stage derivatives $F = \big[f(y_1) , \cdots f(y_s)\big]$.
 
 The new solution $u^{n+1}$ does not comply with positivity for some cases. 
-So it may be advantageous to adapt \eqref{eq:Combination} in such a way that $u^{n+1}$ is positive, instead of rejecting the Solution and computing a new set of stage values.
-This is done by adapting the weights $b$.
+So it might be advantageous to adapt \eqref{eq:Combination} in such a way that $u^{n+1}$ is positive, instead of rejecting the solution and computing a new set of stage values.
+This is done by adapting the weights.
 
-For the computation of the stage derivatives only the $A$ has to be known. The $b$ is not needed for this computation. 
-Based on this Observation we can define $RKM_{A}$ as the set of RKMs that share the same $A$ and only differ in $b$.
-Calculating the stage derivatives can be considered as solving the ODE for all RKM in $RKM_A$ at the same time.
-The idea now is to choose one $RKM_{(A,b)} \in RKM_{A}$ by fixing the $b$. 
-Because the stage derivatives are already known we can choose a $RKM_{(A,b)}$ that leads to a positive $u^{n+1}$, if $RKM_{(A,b)} \in RKM_{A} | u^{n+1} \geq 0$ exists.
-When altering the $b$ the properties of an RKM are preserved. 
+For the computation of the stage derivatives only the $A$ has to be known. The weights is not needed for this computation. 
+Based on this observation we can define $(A,\bullet)$ as the set of RKMs that share the same $A$ and only differ in the weights.
+Calculating the stage derivatives can be considered as solving the ODE for all RKM in $(A,\bullet)$ at the same time.
+The idea is to choose one $(A,b) \in (A,\bullet)$ by fixing the $b$. 
+Because the stage derivatives are already known we can choose a $(A,b)$ that leads to a positive $u^{n+1}$, if $(A,b) \in (A,\bullet) | u^{n+1} \geq 0$ exists.
+When altering the weights the properties of a RKM are preserved. 
 Especially this approach preserves linear invariants like mass conservation, because it is a regular RKM.
 
-As it can be see in \eqref{eq:Combination} $u^{n+1}$ is a linear combination of the stage values and the previous solution $u^n$.
-Because the coefficients $a_{jl}$ are already fixed the Order Conditions reduce to linear equations.
-This means that finding the $b$ can be done by solving a linear Programming problem.
+As it can be seen in \eqref{eq:Combination} $u^{n+1}$ is a linear combination of the stage values and the previous solution $u^n$.
+Because the coefficients $A$ are already fixed the Order Conditions reduce to linear equations.
+This means that finding the weights can be done by solving a linear Programming problem.
 
-If there are no negative values in $u^{n+1}$ with the original $b$, then the method does not cause additional computational effort, because it only has to solve the LP-Problem if the regular RKM does not lead to a positive $u^{n+1}$.
+If there are no negative values in $u^{n+1}$ with the original weights, the method does not cause additional computational effort, because it only has to solve the LP if the regular RKM does not lead to a positive $u^{n+1}$.
 %(except for the test)
@@
 
-To illustrate the use of the method we consider the following reaction equation from\,\cite{kopecz_comparison_2019}.
+To illustrate the usage of the method we consider the following reaction equation from\,\cite{kopecz_comparison_2019}.
 
@@
 u_3' &= 0.5(1-\exp(-1.21u_2^2)) u_3 - 0.01 u_3 -0.02 u_3 \\ 
-u_4' &=0.05 u_2 + 0.02 u_3 0.003u_4 
+u_4' &=0.05 u_2 + 0.02 u_3 + 0.003u_4 
 \end{align}
@@
 
-Now the $b$ are adapted. The results are also plotted in Figure\,\ref{fig:exampleI}. 
+Now the weights are adapted. The results are also plotted in Figure\,\ref{fig:exampleI}. 
 All $u\geq0$ is now fulfilled. A qualitatively correct solution is obtained.
@@
     \includegraphics[width=0.75\textwidth]{plots/exampleI.pdf}
-    \caption{Numerical approxiamtion of Reaction Problem from \cite{kopecz_comparison_2019}.
+    \caption{Numerical approximation of the Reaction Problem from \cite{kopecz_comparison_2019}.
 Computed with ck5 and $\dt = 0.005$.
-The dashed line are the reference solution without the adaption of the $b$.
-The lower plot are the weights $b$ }
+The dashed lines are the reference (Note: Reference might be misleading here because real solution is different) solution without the adaption of the weights.
+The lower plot are the weights }
     \label{fig:exampleI}
@@
 The used weights $b_1,\cdots,b_5$ are also plotted in Figure\,\ref{fig:exampleI}. 
-The used weights are equal the original weights for $t<1.905$.
+The used weights are equal to the original weights for $t<1.905$.
 At $t=1.905$ the weights are first adapted to ensure the positivity of the solution.
@@
 
-\section{Details (TODO: better title)}\label{sec:LP-Problem}
+\section{Details (TODO: better title)}\label{sec:LP}
 
-In the following section, it is explained how the approach can be written as a LP-Problem. 
-For this we have to consider which $b$ to choose. 
+In the following section, it is explained how the approach can be written as a LP. 
+For this we have to consider which weights to choose. 
 Generally, any $b \in \R$ can be used. 
@@
 We can further reduce this space by enforcing the order conditions. 
-By incorporating the order conditions, the space of $b$ is limited to a subspace of $\R$. %(Note: not linear because 0 is not in subspace)
-Finally, we have to give a objective function to get the optimal $b$.
+By incorporating the order conditions, the space of possible weights is limited to a subspace of $\R$. %(Note: not linear because 0 is not in subspace)
+Finally, we have to give an objective function to get the optimal weights.
 
@@
 
-The main goal is to choose a $RKM_{(A,b)}$ that for which $u^{n+1}$ resembles the solution of the ODE $u(t_{n+1})$. 
-A obvious objective is to use a $RKM_{(A,b)}$ with a high order, but this is not enough.
+The main goal is to choose a $(A,b)$ in such a way that $u^{n+1}$ resembles the solution of the ODE $u(t_{n+1})$. 
+A obvious objective is to use a $(A,b)$ with a high order, but this is not enough.
 To get a better understanding for the method we consider the behavior of a simple problem.
@@
 \begin{equation}
-u'(t) = L u(t) \qquad L = \left[\begin{matrix}- a & 1\\a & -1\end{matrix}\right] \quad a =5
+u'(t) = L u(t) \qquad L = \left[\begin{matrix}- 5 & 1\\5 & -1\end{matrix}\right] 
 \end{equation}
-and the first order RKM with two stages
+and and use the three Stage and third order SSP method (SSP33)
 \begin{align}
-\begin{array}{c|cc}
-0 &  & \\
-\frac{1}{2} & \frac{1}{2} & \\
+\begin{array}{c|ccc}
+0 &  &  & \\
+1 & 1 &  & \\
+\frac{1}{2} & \frac{1}{4} & \frac{1}{4} & \\
 \hline
- & \frac{1}{2} & \frac{1}{2}\\
+ & \frac{1}{6} & \frac{1}{6} & \frac{2}{3}\\
 \end{array}
 \end{align}
-If simulated with $\dt = 1$ the stage derivatives are 
+The Matrix $L$ has the eigenvalues $\lambda_1 = -6,\lambda_2=0$.
+The real stability interval of the SSP33 method includes $[-2.5,0]$.
+The method is absolut stable for this problem with $\dt = \frac{1}{3}$
+The stage derivatives are
 \begin{equation}
-f^1 = \left(\begin{matrix}-5\\5\end{matrix}\right) \qquad f^2 = \left(\begin{matrix}10\\-10\end{matrix}\right)
+f(y_1)=\left[\begin{matrix}-5\\5\end{matrix}\right]\qquad f(y_2)=\left[\begin{matrix}5\\-5\end{matrix}\right]\qquad f(y_3)=\left[\begin{matrix}-5\\5\end{matrix}\right]\end{equation}
+The values of the next step using the standard weights is
+\begin{equation}
+u_n=\left[\begin{matrix}- \frac{1}{9}\\\frac{10}{9}\end{matrix}\right]
 \end{equation}
-All $b$ that comply with the order Condition for the first order can be expressed as
+The new solution contains negative values. 
+Therefor we want to adapt the weights to ensure positifity. 
+All weights that comply with the order Condition for the first and second order can be expressed as
 \begin{equation}
-b= \left(\begin{matrix}\frac{1}{2}\\ \frac{1}{2}\end{matrix}\right) + \alpha \left(\begin{matrix}1\\ -1\end{matrix}\right)
+\tilde b=\left[\begin{matrix}\frac{1}{6}\\\frac{1}{6}\\\frac{2}{3}\end{matrix}\right]+\alpha \left[\begin{matrix}\frac{1}{2}\\\frac{1}{2}\\-1\end{matrix}\right]
 \end{equation}
-With this we have one degree of freedom for the choice of $b$. This is done by the parameter $\alpha$.
-If the general expression for $b$ is inserted in \eqref{eq:Combination} the general solution is 
+With this we have one degree of freedom for the choice of the weights. This is done by the parameter $\alpha$.
+If the general expression for the weights is inserted in \eqref{eq:Combination} the general solution is 
 \begin{equation}
-u^{1} = u^0 + \dt  \left[f^1,f^2\right] b = \left(\begin{matrix}\frac{5}{2}\\- \frac{5}{2}\end{matrix}\right) \alpha  \left(\begin{matrix}-15\\15\end{matrix}\right)
+u^{1} = u^0 + \dt  \left[f(y_1),f(y_2),f(y_3)\right] b =
+\left[\begin{matrix}- \frac{1}{9}\\\frac{10}{9}\end{matrix}\right] +\alpha \left[\begin{matrix}\frac{5}{3}\\- \frac{5}{3}\end{matrix}\right]
 \end{equation}
-The solution for the original RKM would be $u^{1} = \left(\frac{5}{2} ,- \frac{5}{2}\right)^T$ which contains a negative value.
-But by changing the parameter $\alpha$ the weights $b$ and the solution is altered.
-By a suitable choice of $\alpha \in \left[\frac{1}{6},\frac{7}{30}\right]$ any $u$ that complies with mass conservation and positivity can be reached. 
-By adding additional constraints on the $b$ the choice of $\alpha$ can be narrowed down. 
-An objective function is also needed to make the choice unique. This should be designed in a way to prefer $b$ that are close to the original weights $b_{orig}$
+By changing the parameter $\alpha$ the weights $b$ and the solution is altered.
+With a suitable choice of $\alpha \in \left[\frac{1}{15},\frac{2}{3}\right]$ any $u$ that complies with mass conservation and positivity can be reached. 
+By adding additional constraints on the weights the choice of $\alpha$ can be narrowed down. 
+An objective function is also needed to make the choice unique. This should be designed in a way to prefer weights that are close to the original weights.
 
@@
 
-The order conditions for a RKM are a set of Equations depending on $a_{jl}$, $c_1,\cdots,c_s$ and $b_1,\cdots,b_s$.
-In these Equations $b_1,\cdots,b_s$ only appear linearly \cite{hairer_runge-kutta_1993}.
-If $a_{jl}$ and $c_1,\cdots,c_s$ are already known, the order conditions for the Order $p$ are an linear equations system and can be written as $Q b = r$.
+The order conditions for a RKM are a set of Equations depending on $A$, $c$ and $b$. In these Equations $b$ only appears linearly \cite{hairer_solving_1993}.
+If $A$ and $c$ are already known, the order conditions for the Order $p$ are a linear equations system and can be written as $Q b = r$.
 
@@
 This implies that $\mathrm{rank}(Q) < s$. 
-Because the Quadrature conditions are linear independent, there are at least $p$ independent vectors in $Q$. This leads to the condition $p < s$.
+Because the quadrature conditions are linearly independent, there are at least $p$ independent vectors in $Q$. This leads to the condition $p < s$.
 This implies that the number of stages has to be higher than the wanted order.
@@
 
-There are two possible ways to write down the adaptation of the $b$. The first way is to directly adapt the original weights $b_{orig}$ by adding a correction $\Delta b$ that ensures positivity.
-This is the easiest way to adapt the $b$. An advantageous property is that the step size can be reduced after calculation the stage values by using the dense output order conditions.  
-This approach suits best for explicit methods.
-The direct adaption approach is explained in more detail in section\,\ref{sec:direct}.
+There are two possible ways to write down the adaptation of the weights. The first way is to directly adapt the original weights $b$ by adding a correction $\Delta b$ that ensures positivity.
+This is the easiest way to adapt the weights. An advantageous property is that the step size can be reduced after calculating the stage values by using the dense output order conditions.  
+The direct adaption is explained in more detail in section\,\ref{sec:direct}.
 
-The second approach is to choose $b$ as a convex combination of a set of $b$s. This approach is more useful if some embedded methods with certain properties are known. 
-This approach also shows more predictable behavior because the resulting $u^{n+1}_b$ is also a convex combination of the solutions $u^{n+1}_{b_a},u^{n+1}_{b_b},\cdots$ 
-This approach works best with implicit methods.
-The direct adaption approach is explained in more detail in section\,\ref{sec:convex}
+The second way to formulate a LP is to choose weights as a convex combination of a set of weights. This is more useful if some embedded methods with certain properties are known. 
+The convex combination also shows more predictable behavior because the resulting $u^{n+1}_b$ is also a convex combination of the solutions $u^{n+1}_{b_a},u^{n+1}_{b_b},\cdots$ 
+This is explained in more detail in section\,\ref{sec:convex}
 
@@
         \input{Image_b_convex}
-        \caption{Convex Combiantion}
+        \caption{Convex Combination}
         \label{fig:b_convex}
@@
 
-When directly adapting the $b$ a new $b$ that complies with the order conditions is chosen.   
-The weights $b = b_{orig} + \Delta b$ consist of the weights of the original method $b_{orig}$ and an adaption $\Delta b$. The adaption is $\Delta b \in \mathrm{ker}(Q)$. 
-In Figure\,\ref{fig:b_direct} this is drawn. The red line represents the subspace of $b$ that satisfy the order conditions. Both $b_{orig}$ and $b$ are in this subspace.
+When directly adapting the weights new weights that complies with the order conditions is chosen.   
+The weights $\tilde{b} = b + \Delta b$ consist of the weights of the original method $b$ and an adaption $\Delta b$. The adaption is $\Delta b \in \mathrm{ker}(Q)$. 
+In Figure\,\ref{fig:b_direct} this is drawn. The red line represents the subspace of weights that satisfies the order conditions. Both $b$ and $\tilde{b}$ are in this subspace.
 
-To run it as a LP-Problem a set of constraints is constructed.  
-Also, an objective function is needed to determine the optimal $b$.
+To run it as a LP a set of constraints is constructed.  
+Also, an objective function is needed to determine the optimal weights.
 
 \subsubsection{Constraints}
-The Direct Adaptation approach leads to two set of constraints on the $b$.
+The Direct Adaptation approach leads to two sets of constraints on $b$.
 The first set of constraints is the positivity and can be written as 
 \begin{equation}
-u_{n+1}=u_n+\dt K \vec{b} \geq 0
+u_{n+1}=u_n+\dt F \tilde{b} \geq 0
 \end{equation}
@@
 
-A reasonable property of the objective function would be that for the LP-Problem only constrained by the order condition the optimal solution is $b = b_{orig}$. 
+A reasonable property of the objective function would be that for the LP only constrained by the order condition the optimal solution is $b = b_{orig}$. 
 This can be achieved by $argmin(f_{optim}(b)) = b_{orig}$. 
-In practice, this also means that there is no need to solve solve the optimization Problem if no constraints are violated because the solution of this problem is already known.
+In practice, this also means that there is no need to solve the optimization problem if no constraints are violated because the solution of this problem is already known.
 This is advantageous for performance reasons.
 The different approaches have been tested on a linear test problem from \cite{kopecz_unconditionally_2018}. The optimization problem was used at every step.
-The magnitude of $b$ can get large if the order conditions are used as objective function.
+The magnitude of the weights can get large if the order conditions are used as objective function.
 Large values of b can lead to numerical problems.
@@
 Because the computation is done with a finite precision (and usually in floating point) large $||b||$ can lead to absorption effects.
-By changing the $b$ the stability function is altered. This can lead to oscillations if the poles of the RHS are no longer in the stability region of the resulting method.
-For these reasons $||b-b_{orig}||$ seems like a good choice. The 1-Norm can be implement the with Linear Programming using slack variables. We choose
+By changing the weights the stability function is altered. This can lead to oscillations if the poles of the RHS are no longer in the stability region of the resulting method.
+For these reasons $||b-b_{orig}||$ seems like a good choice. The 1-Norm can be implemented the with Linear Programming using slack variables. We choose
 
@@
 
-If the original step $\dt$ was too large and there is no suitable $b$, we would like to calculate a $u^{n+1}$ using the stage values instead of rejecting the step.
+If the original step $\dt$ was too large and there is no suitable set of weights, the LP is infeasible.
+In this case we would like to calculate $u^{n+\theta}$ using the stage values instead of rejecting the step.
 For this the time step is reduced after calculation the stage values.
-This can be done using the order conditions for dense Output.
-These can be found in \cite{hairer_runge-kutta_1993}.
+This can be done using the order conditions for dense output.
+These can be found in \cite{hairer_solving_1993}.
 We introduce $\theta \in (0,1]$. The new step taken has the length $\theta \dt$ whereas $\dt$ is the time step used when calculating the stage values.
-To calculate a reduced step the order conditions have to be adapted and a new $b_{orig}$ for the objective function (and more important the Error detection afterwards) has to be generated. 
+To calculate a reduced step the order conditions have to be adapted and a new $b_{orig}$ for the objective function (and more important the error detection afterwards) has to be generated. 
 (TODO: the $b_{orig}$ is still a open question)
 
-When adapting the Order Condition to a new $\theta$ the Order Condition Matrix $Q$ stays unchanged whereas the right hand side depends on $\theta$.
+When adapting the Order Condition to a new $\theta$ the Order Condition Matrix $Q$ stays unchanged whereas the right hand side $r(\theta)$ depends on $\theta$.
 
-The dependency of $r(\theta)$ is nonlinear.
-This means that the maximum $\theta$ with a suitable $b$ cannot be calculated by solving a LP-Problem.
-This means that different $\theta$ have to be tested.
+The dependency of $r(\theta)$ is nonlinear. This means that the maximum $\theta$ with suitable weights cannot be calculated by solving a LP. This means that different $\theta$ have to be tested. 
 (Note: we could include some algorithm to search for it)
@@
 For some test problems (e.g.
-the Heat Equation solved with an implicit method of higher order), large distortions occur, even though a norm of the distortion of the b is used as an objective function.
-For implicit methods, it is known that the backward euler steps lead to a positive solution for any $\dt$ \cite{hundsdorfer_numerical_2003}.
-The backward Euler is still only first Order but shows more exact results than a method of higher Order that is distorted by the adaption of the $b$, at least at some steps.
-A possible method to fix this is to generate the new b based on a convex combination of embedded methods of RKM.
+the Heat Equation solved with an implicit method of higher order), large changes occur, even though a norm of $\Delta b$ is used as an objective function.
+A numerical example is given in Section\,\ref{sec:Numeric_Results}
+It is known that the backward Euler method lead to a positive solution for any $\dt$ \cite{hundsdorfer_numerical_2003}.
+The backward Euler is still only first Order but shows more exact results than a method of higher Order that is distorted by the adaption of the weights, at least at some steps.
+A possible method to fix this is to generate the new b based on a convex combination of embedded methods.
 
 \subsubsection{Constraints}
-At first a set of embedded methods of the RKM $b_1,\cdots,b_d$ are needed. 
-If all the used $b$ compile with the order conditions the new method is also of the same Order.
+At first a set of embedded methods of the RKM $b_1,\cdots,b_d$ is needed. 
+If all the used weights complie with the order conditions the new method is also of the same Order.
 This is straightforward using
-$$ Q b = Q (\chi b_{\alpha} + (1-\chi) b_{\beta}) = \chi Q  b_{\alpha} + (1-\chi) Q b_{\beta}) = \chi r + (1-\chi) r = r$$
+$$ Q b = Q (\chi b_{\alpha} + (1-\chi) b_{\beta}) = \chi Q  b_{\alpha} + (1-\chi) Q b_{\beta} = \chi r + (1-\chi) r = r$$
 $b_{\alpha}$ and $b_{\beta}$ are the methods that are combined to $b$ using the parameter $\chi$. The Matrix $Q$ and the vector $r$ are the order conditions as defined in Section\,\ref{sec:OrderCond}.
 
-The new $b$ is defined as 
+The new weights are defined as 
 \begin{equation}\label{eq:b_convex_def}
-b = \sum_i^d a_i b_i = Ba
+b = \sum_i^d g_i b_i = Bg
 \end{equation}
 where $B=\left[b_1,\cdots,b_d \right]$ is a matrix containing the embedded methods.  
-To ensure the that \eqref{eq:b_convex_def} yields a convex combination following constraints are enforced for the  the parameters $a,\cdots,a_d$.
+To ensure that \eqref{eq:b_convex_def} yields to a convex combination the following constraints are enforced for the parameters $g_1,\cdots,g_d$.
 \begin{equation}
- 0 \leq a_1 \leq 1  \forall {i \in \{1, \cdots d \}}
+ 0 \leq g_i \leq 1  \forall {i \in \{1, \cdots d \}}
 \end{equation}
 \begin{equation}
- \sum_{i=1}^d a_i = 1
+ \sum_{i=1}^d g_i = 1
 \end{equation}
 The positivity constraint changes to 
-$$u_{n+1}=u_n+\dt K B a \geq 0$$
-where $a = [a_1,\cdots,a_d]^T$
+$$u_{n+1}=u_n+\dt K B g \geq 0$$
+where $g = [g_1,\cdots,g_d]^T$
 
@@
 
-As an Objective function the function
-
+As an objective function,
 \begin{equation}
@@
 \end{equation}
-
 is used.  The $w_1,\cdots,w_d$ are weighting factors that determine which is the preferred $b_i$ to use.
 
-Because the optimal solution of the unconstrained problem has to be the original $b$ the weight $w_0$ corresponding to $b_{orig}$ should be the smallest $w$.
+Because the optimal solution of the unconstrained problem has to be the original weights the weight $w_0$ corresponding to $b_{orig}$ should be the smallest $w$.
 This weight is simply set to $w_{b_{orig}} = 0$.
@@
 
-As a first order method, a $b$ is added that corresponds to a series of backward euler steps. These are known to be yield to a positive result for any step-size.
+As a first order method, the weights that corresponds to a series of backward Euler steps are added. These are known to yield a positive result for any step size.
 It is also possible to add embedded methods of higher order.
@@
 
-A Reduction of the step size suing the order conditions for dense output would also be possible.
-For this a set of embedded methods $b_1(\Theta),\cdots,b_d(\Theta)$ has to be constructed.
-This was not done in this research.
+A reduction of the step size using the order conditions for dense output would also be possible.
+For this a set of embedded methods $b_1(\theta),\cdots,b_d(\theta)$ has to be constructed.
+This was not the aim of this research.
 
@@
 
-In the previous sections an algorithm for choosing a $b$ has been presented. 
+In the previous sections an algorithm for choosing positifity preserving weights $\tilde{b}$ has been presented. 
 In the following section, it is explained how this Algorithm can be integrated into the framework of existing solvers.
@@
 An important property for the used baseline method is the existence of embedded methods and the degrees of freedom for the b.
-As noted in Section\,\ref{sec:OrderCond} the number of Stages has to be higher than the order. 
+As noted in Section\,\ref{sec:OrderCond} the number of stages has to be higher than the order. 
 Another important property is the number of degrees of freedom for the choice of the new weights. 
-These can be calculated using $s-\mathrm{rank}(Q)$%$\mathrm{dim}(\mathrm{ker}(Q))$.
+These can be calculated using $s-\mathrm{rank}(Q)$.%$\mathrm{dim}(\mathrm{ker}(Q))$.
 The number of degrees of freedom for different methods are shown in table\,\ref{table:DOF_exp} for explicit methods and in table\,\ref{table:DOF_imp} for implicit methods.
@@
  \end{tabular}
- \caption{Degrees of Freedom for Choice of $b$} %Generated above============ 
+ \caption{Degrees of Freedom for Choice of the weights} %Generated above============ 
  \label{table:DOF_exp}
@@
  \end{tabular}
- \caption{Degrees of Freedom for Choice of $b$} %Generated above============ 
+ \caption{Degrees of Freedom for Choice of the weights} %Generated above============ 
  \label{table:DOF_imp}
@@
 For all methods with the number of stages equal to the order of the method, there are no degrees of freedom without reducing the order. 
-If the Classical RK4 method is used the Order has to reduced more because the RK4 method does not have embedded methods of order 3.
+If the classical RK4 method is used the Order has to be reduced more because the RK4 method does not have embedded methods of order 3.
 Contrary to this methods with $s > p$ have degrees of freedom even without reducing the order.
 
-Another important property is the stability region in comparison with the maximum $\dt$ that ensures positivity.
-For this approach to be useful the RK-method has to yield negative results for $\dt$ smaller than the biggest stable $\dt$.
+Another important property is the stability region.
+The adaption of the weights does solve issues with stabilitz when used with a too big $\dt$.
+Therefore the RK-method has to be stable for large $\dt$.
 
-(Note: up to this point I do not understand why this is happening. We only see it occur depending on the Problem-RKM combination)
+%(Note: up to this point I do not understand why some methods perform better than other. We only see it occur depending on the Problem-RKM combination)
 
@@
 This corresponds to the question if there is an embedded first order method that ensures positivity.
-This is particularly interesting for Implicit methods, because the $\dt$ is not limited by the stability.
+This is particularly interesting for implicit methods, because the $\dt$ is not limited by the stability.
 If we have an embedded first Order Method we can ensure that there is always a positive solution, which might be of first Order.
@@
 When changing the weights the solution $u^{n+1}$ is changed. 
-By incorporating the order conditions for a certain order the space of $b$ that can be choose is limited to a linear Subspace. 
+By incorporating the order conditions for a certain order the space of possible weights that can be chosen is limited to a linear Subspace. 
 Now it is important to know if the new solution $u^{n+1}$ still resembles the solution of the ODE $u(t_{n+1})$. 
@@
 
-This gives raise to the question whether the new solution is reasonable.
+This gives rise to the question whether the new solution is reasonable.
 This also poses the problem on developing a system that can decide if the new solution is has to be rejected.
 
-\subsubsection{Error Detection}
-For this approach is difficult to give some mathematically rigorous statement regarding the convergence of the method, because it only applies if the error of the original method causes the solution to get negative.
-This can only happen if the Truncation error is large enough.
-This requires $\dt \gg 0$ where lower coefficients of the Taylor Series are no longer sufficient to approximate the behavior.
-One could say that the main idea of the adaption of $b$ is to alter the error in a certain way, so that the numerical solution gets 'better' in the sense that it leads to positive solutions. 
+\subsubsection{Error Detection and Appproximation}
+For this approach it is difficult to give some mathematically rigorous statement regarding the convergence of the method, because it only applies if the error of the original method causes the solution to get negative.
+This can only happen if the truncation error is large enough.
+This requires $\dt \gg 0$ where lower coefficients of the Taylor series are no longer sufficient to approximate the behavior.
+One could say that the main idea of the adaption of the weights alters the error in a certain way, so that the numerical solution gets 'better' in the sense that it leads to positive solutions. 
 The local error of a RKM $e(\dt) =u(t_0 + \dt) - u^1$ can be expressed using the Taylor series of the RKM and the Taylor series of the exact solution. % Hairer I P134
@@
 u(t_0 + \dt) &= u(t_0) + u'(t_0) \dt + \frac{u''(t_0)}{2} \dt^2 + \cdots + \frac{u^{(p)}(t_0)}{p!} \dt^p + \frac{u^{(p+1)}(t_0)}{(p+1)!} \dt^{p+1} + \cdots \\
-u^{n+1} &= u(t_0)  + v_1 \dt + \frac{v_2}{2} \dt^2 + \cdots + \frac{v_p}{p!} \dt^p + \frac{v_{p+1}}{(p+1)!} \dt^{p+1} + \cdots \\
+u^{n+1} &= u(t_0)  + v_1 \dt + \frac{v_2}{2} \dt^2 + \cdots + \frac{v_p}{p!} \dt^p + \frac{v_{p+1}}{(p+1)!} \dt^{p+1} + \cdots 
 \end{align}
@@
 The derivatives of $u(t)$ can be solely expressed with the partial derivatives of the RHS $f(t,u)$, whereas the coefficients $v_1,v_2,\cdots$ also depend on the RKM. 
-In this case the RKM is fixed apart from the weights $b$. 
+In this case the RKM is fixed apart from the weights. 
 Therefore $v_1,v_2,\cdots$ are functions of $b$.
-Because the $b$ are chosen to ensure the Order Conditions the coefficients $v_1,\cdots,v_p$ are fixed and equal the according derivatives. 
+Because the the weights are chosen to ensure the Order Conditions the coefficients $v_1,\cdots,v_p$ are fixed and equal the according derivatives. 
 The only remaining degrees of freedom are the coefficients $v_{p+1},v_{p+2},\cdots$.
-These can be adapted by changing the $b$.
+These can be adapted by changing the weights.
 
-The main way is to ensure that the RKM with adapted b converges to the correct solution is to ensure that the $b$ approaches $b_{orig}$ for $\dt \to 0$.
-For this $b$ it is already known that the numeric solution converges to the exact solution.
-This is made sure by property of the Objective function that 
-\begin{equation}
-\mathrm{argmin}(f_{optim}(b)) = b_{orig}
-\end{equation}
+%The main way to ensure that the RKM with adapted b converges to the correct solution is to ensure that the weights approaches $b_{orig}$ for $\dt \to 0$.
+For this the weights it is already known that the numeric solution converges to the exact solution.
+%This is made sure by the property of the Objective function that 
+%\begin{equation}
+%\mathrm{argmin}(f_{optim}(b)) = b_{orig}
+%\end{equation}
 This approximation is not very useful for real applications. 
@@
 
-After adapting the $b$ the Approximation Error is checked. If $|\dt K(b-b_{orig})|$ is larger than the tolerance the $b$ is rejected.
+After adapting the weights the approximation error is checked. If $|\dt K(b-b_{orig})|$ is larger than the tolerance the weights are rejected.
 
-\subsubsection{Steppsize Control}
-An important part of an RKM method is the ability to approximate the error of the solution to use it to control the step size. 
+%\subsubsection{Steppsize Control}
+%An important part of an RKM method is the ability to approximate the error of the solution to use it to control the step size. 
 The truncation error $|u(t^n)-u^n_{b_{orig}}|$ is usually approximated with an standard error estimator $err_T = | u^{n}_{b_{orig}} - u^{n}_{\hat{b}} |$.
-
-Both errors are added to get an approximation of the total error $err = err_T +w_a err_{adapt}$.
-The parameter $w_a \leq 1$ is used to account for the fact that the real error is smaller than $err = err_T +err_{adapt}$.
+Both errors are added to get an approximation of the total error $err = err_T + err_{adapt}$.
+%The parameter $w_a \leq 1$ is used to account for the fact that the real error is smaller than $err = err_T +err_{adapt}$.
 The factor also improves the stability of the step size control because $err_{adapt}$ is not as smooth as $err_T$, especially if negative values only occur for a small number of steps.
-If the LP-Problem was infeasible and no $b$ was found the $err_{approx}$ is set to a custom value.
+%If the LP was infeasible and no weights were found the $err_{approx}$ is set to a custom value.
 This value should be chosen sufficiently large to ensure that the next step take is smaller, but still not too large to keep the step size control stable.
-
 This type of error estimation is easy to implement because it can be easily incorporated in an existing step size control and takes advantage of the standard error approximation.
@@
 By adapting the $b$ the used RK Method is changed. This also alters the stability function and therefore changes the region of absolute stability. 
-It depends on the solved problem, the RKM and the used $\dt$ on how many steps the stability function is altered.
-For Problems where only a small number of step are affected a bigger change in the stability function would be acceptable.
-For Problems with require an adaptation of the weights for every step one needs to make sure that the resulting method is stable.
+It depends on the solved problem, the RKM and the used $\dt$ by how many steps the stability function is altered.
+For problems where only a small number of steps are affected a bigger change in the stability function would be acceptable.
+For problems which require an adaptation of the weights for every step one needs to make sure that the resulting method is stable.
 
@@
 \begin{equation}
-P(z) = R(z)-1 = z b^T (I - zA)^{-1}
+P(z) = R(z)-1 = z b^T (I - zA)^{-1} e
 \end{equation}
@@
 \end{equation}
-The stability function is a linear function in $b$.
+The stability function is a linear function in the weights.
 
@@
 \subsubsection{Stability of convex combination}
-If the new $b$ is chosen as a convex combination of $b$ it is easy to prove properties of the stability region.
-The stability function $R_{\chi b_a + (1-\chi) b_b}(z)$ of a convex combination can be calculated by the weighted sum of the Stability functions $R_{b_a}(z)$ and $R_{b_b}(z)$
+If the new weights are chosen as a convex combination of weights it is easy to prove properties of the stability region.
+The stability function $R_{\chi b_a + (1-\chi) b_b}(z)$ of a convex combination can be calculated by the weighted sum of the stability functions $R_{b_a}(z)$ and $R_{b_b}(z)$
 
@@
 
-The resulting region of absolute stability is at least the Intersection of the regions of absolute stability of the methods that define the convex combination.
-This also means that if all the used embedded methods used to construct the new $b$ are A-Stable, the resulting method is also A-Stable.
+The resulting region of absolute stability is at least the intersection of the regions of absolute stability of the methods that define the convex combination.
+This also means that if all the used embedded methods used to construct the new weights are A-Stable, the resulting method is also A-Stable.
 
-We want to proof that the convex combination of two methods is absolutely stable at the points where both original methods were absolutely stable.
+We want to show that the convex combination of two methods is absolutely stable at the points where both original methods were absolutely stable.
 This can be formulated as $(|R_{b_a}(z)|  < 1) (|R_{b_b}(z)| < 1) \Rightarrow |R_{\chi b_a +(1- \chi) b_b}(z)| < 1$ with $\chi \in [0,1]$.
@@
 \subsubsection{Stability of  Direct Adaptation}
-If the $b$ is adapted directly there is no such proof.
+If the weights is adapted directly there is no such proof.
 A useful property of the resulting stability region is that it is similar to the stability region of the baseline method. 
@@
 
-For most methods, a small change of $b$ does not alter the stability function dramatically.
+For most methods, a small change of the weights does not alter the stability function dramatically.
 (TODO: I am missing a piece here from ‘the derivative is finite’ to ‘is similar’)
 
-We would like to calculate the derivative of the border in respect to a change of $b$.
+We would like to calculate the derivative of the border in respect to a change of the weights.
 
-The stability function depending on the $b$ is expressed 
-When adapting the $b$ we cannot change them independently but only in ways that comply with the order conditions.
+The stability function depending on the weights is expressed. 
+When adapting the weights we cannot change them independently but only in ways that comply with the order conditions.
 $m$ is the number of dimensions of the Kernel of the Order Condition $m = \mathrm{dim}\{\mathrm{Ker} (Q) \}$
-The $b_1,\cdots,b_m$  are the Basis vectors of $\mathrm{Ker} (Q)$. With an coordinate Transformation we get the coordinate vector $\hat{b}$ with the Matrix $B = [b_1,\cdots,b_m]$ and $b = B \hat{b}$.
+The $b_1,\cdots,b_m$  are the Basis vectors of $\mathrm{Ker} (Q)$. With a coordinate transformation we get the coordinate vector $\hat{b}$ with the Matrix $B = [b_1,\cdots,b_m]$ and $b = B \hat{b}$.
 
-The General Stability function of all embedded RKM that satisfy the Order Conditions $Q$ is given by
+The general stability function of all embedded RKM that satisfy the Order Conditions $Q$ is given by
 
@@
  $|R(w,u,v)|^2$ is calculated. 
-If the gradient does not vanish at $z = z_0$ \eqref{eq:border} Can be transformed in a function $\R^m \times \R \rightarrow \R$ depending on $\hat{b}$ and a single variable $a$ by setting $z = z_0 + n a$.
+If the gradient does not vanish at $z = z_0$ \eqref{eq:border} can be transformed in a function $\R^m \times \R \rightarrow \R$ depending on $\hat{b}$ and a single variable $a$ by setting $z = z_0 + n a$.
 The complex number $n$ denotes the normal vector $\vec{n} = \frac{\nabla |R(\hat{b},u,v))|^2}{\left| \nabla |R(\hat{b},u,v))|^2 \right|}$ of the border of the stability region written in complex notation.
@@
 \end{equation}
-If explicit methods are used this function is known to be continuously differentiable because it can be written as a Polynomial of $\hat{b},u$ and $v$.
-This is a necessary condition for the use of the of the Implicit function theorem.
+If explicit methods are used this function is known to be continuously differentiable because it can be written as a polynomial of $\hat{b},u$ and $v$.
+This is a necessary condition for the use of the Implicit function theorem.
 It is also known by definition, that for the point $b = b_{orig}$ and $z = z_0$ the equation $ |R(\hat{b},z_0)|^2 -1 = 0 $ is satisfied.
@@
 
-The Derivative of \eqref{eq:f(w,a)} in respect to $a$ evaluated at $\hat{b}=0$ can be calculated using the directional derivative
+The derivative of \eqref{eq:f(w,a)} in respect to $a$ evaluated at $\hat{b}=0$ can be calculated using the directional derivative
 \begin{align*}
@@
 The derivative in respect to $\hat{b}_j$ is 
-$ \frac{\partial f}{\partial \hat{b}_j}(\hat{b},a(\hat{b})) \Big|_{\hat{b}=0}$
-this simplifies to 
+$ \frac{\partial f}{\partial \hat{b}_j}(\hat{b},a(\hat{b})) \Big|_{\hat{b}=0}$.
+This simplifies to 
 \begin{multline}\label{eq:derivative_to_b}
@@
 As long as $\nabla|R(\hat{b},u+iv)|^2 \neq 0 \forall \hat{b} < \epsilon  \forall z$ the derivative $\frac{\partial a}{\partial \hat{b}_j} (\hat{b})$ is finite.
-This means that a small change the $b$ only leads to a small change of the stability region.
+This means that a small change of the weights only leads to a small change of the stability region.
 
@@
 
-\subsection{Positifity}
+\subsection{Positivity}
 
-For the further discussion, we are concerned with a System of $m$ coupled ODEs. 
+For the further discussion, we are concerned with a system of $m$ coupled ODEs. 
 In this setup $u \in \R^m$.
@@
 
-To enshure that the solution is positive
+To ensure that the solution is positive
 
-\begin{align}
- u_i^{n+1} &\geq 0   \;   \forall {i \in \{1, \cdots,m \}}  \\
- u_i^n + h \sum_{j=0}^{s-1} b_j f_i^j  &\geq 0   \;   \forall {i \in \{1,\cdots,m \}}  
-\end{align}
+\begin{equation}
+ u_i^{n+1} \geq 0   \;   \forall {i \in \{1, \cdots,m \}}  
+\end{equation}
 has to be fulfilled.
@@
 \begin{equation}
-u_i + \dt F  b \geq 0     
+u_i + \dt \sum_{j=0}^s F_{i,j}  b_j \geq 0 \forall {i \in \{1, \cdots,m \}}  
 \end{equation}
-where $F = \big[f^1 , \cdots f^{s-1}\big]$.
 
-This causes unnecessary many inequality constraints on the LP-Problem.
-To simplify the problem, one want to reduce the number of positivity constraints by ignoring it for ODEs where there is no issue with positivity.
+This causes unnecessarily many inequality constraints on the LP.
+To simplify the problem, one wants to reduce the number of positivity constraints by ignoring it for ODEs where there is no issue with positivity.
 The new positivity constraints can be formulated as
 \begin{equation}
-u_i^n + \Delta b \sum_{j=0}^{s-1} b_j f_i^j  \geq 0   \;   \forall {i \in h \subseteq \{1,\cdot,m \}} 
+u_i + \dt \sum_{j=0}^s F_{i,j}  b_j  \geq 0   \;   \forall {i \in h \subseteq \{1,\cdots,m \}} 
 \end{equation}
-For this a set $h \subseteq \{1,\cdots,m \}$ that include all indecies that are needed for determining the $b$ has to be chosen.
+For this a set $h$ that includes all indecies $i \in \{1,\cdots,m \}$ that are needed for determining the weights has to be chosen.
 
 A reasonable approach is to set $h_0 = \{ i \in \{1,\cdots,m \} |  u_i^{n+1}  < 0 \}$. 
-After the new $b$ and $u^{n+1}$ is computed using $h_0$, the algorithm has to check if the positivity of the other values $u_i | i \notin h_0 \subseteq \{1,\cdots,m \}$ is still meet. 
-If the conditions is not meet a new set $h_{a+1} = \{ i \in \{1,\cdots,m \}|  u_i^{n+1}  < 0 \} \cup h_{a}$ using the new $u^{n+1}$ is generated and the LP-Problem is solved again.
-This process id repeated until a solution is found that leads to a $u^{n+1} \geq 0$.
+After the new weights and $u^{n+1}$ is computed using $h_0$, the algorithm has to check if the positivity of the other values $u_i | i \notin h_0 \subseteq \{1,\cdots,m \}$ is still met. 
+If the condition is not met a new set $h_{a+1} = \{ i \in \{1,\cdots,m \}|  u_i^{n+1}  < 0 \} \cup h_{a}$ using the new $u^{n+1}$ is generated and the LP is solved again.
+This process is repeated until a solution is found that leads to a $u^{n+1} \geq 0$.
 The choice of $h_{a+1}$ based on the $h_{a}$ ensures that the algorithm cannot get stuck in a loop.
 In the worst case $h_a$ is growing very slowly until $h_a = \{1,\cdots,m \}$.
-
-This effect did not occur so far. 
-(Note: A approach on this would be to make a educated guess which indices could get a problem with positivity. A possible estimate would be $\frac{u_i^{n+1}}{max(K_{(i,0)}, \cdots ,K_{(i,0)})} $ )
+This worst case did not occur when runing the testproblems. 
+%(Note: A approach on this would be to make a educated guess which indices could get a problem with positivity. A possible estimate would be $\frac{u_i^{n+1}}{max(K_{(i,0)}, \cdots ,K_{(i,0)})} $ )
 
 When enforcing a maximum value the number of constraints can be reduced using the same technique.
-When both enforcing maximum and minimum values two separate sets of active constraints $k_{min}$ and $k_{max}$ are used.
+When nforcing both maximum and minimum values two separate sets of active constraints $k_{min}$ and $k_{max}$ are used.
 The set $k_{max}$ denotes the active maximum constraints and the set $k_{min}$ denotes the active minimum constraints.
-Here it is important to update these sets simultaneously.
+In this case it is important to update these sets simultaneously.
 
@@
 The inner loop loops through the orders, starting with the highest Order down to the lowest Order.
-Until an acceptable $b$ the algorithm tests combinations of $\theta$ and $p$.
+Until acceptable weights are found the algorithm tests combinations of $\theta$ and $p$.
 At first the Order Conditions are constructed.
 On these a LP-Solver is invoiced.
-If the LP-Problem is infeasible the algorithm immediately tries the next combination.
-If the LP-Problem is feasible we test if the $b$ is acceptable.
+If the LP is infeasible the algorithm immediately tries the next combination.
+If the LP is feasible we test if the $b$ is acceptable.
 At first it is tested weather the new $u^{n+1}$ indeed complies with the constraints.
@@
 (Note: because of the chosen objective function it cannot be unbounded) As second test the deviation from the old solution (Compare Error estimation) is calculated.
-If the error is bigger than a set maximum error the $b$ is also rejected and the algorithm also switch to the next combination.
-As soon as the algorithm finds a suitable $b$ the loops are aborted and the $b$ is used for the update step.
-If all the combinations of $\theta$ and $p$ did not lead to an acceptable $b$ the step has to be rejected and a new set of stage values has to be computed.
+If the error is bigger than a set maximum error the $b$ is also rejected and the algorithm also switches to the next combination.
+As soon as the algorithm finds suitable weights the loops are aborted and the weights are used for the update step.
+If all the combinations of $\theta$ and $p$ did not lead to acceptable weights the step has to be rejected and a new set of stage values has to be computed.
 
@@
 For problems where the exact solution is positive for certain $u_0$ but positivity is not preserved for all $u_0 \geq 0$ tests did not show promising results.
-Additional it is not certain if the computed solutions would be reasonable.
+Additionaly it is not certain if the computed solutions would be reasonable.
 
 \subsection{Explicit methods}
-For most of the test problems problems with stability occur before getting negative values. This means that we need methods with large stability region.
+For most of the test problems with stability occur before getting negative values. This means that we need methods with large stability region.
 
@@
 The parameters are set to $a=1$ and $K=1$.
-For $\dt \leq 0.0082$ the unaltered method lead to positive solutions. For larger $\dt$ the original method leads to negative values and $b$ is altered. For $\dt >0.016$ the dp5 is no longer stable.
+For $\dt \leq 0.0082$ the unaltered method leads to positive solutions.
+For a larger $\dt$ the original method leads to negative values and the weights are altered.
+For $\dt >0.016$ the dp5 is no longer stable.
 In Figure\,\ref{fig:conv_expl} the convergence is plotted for the altered and unaltered method.
@@
 For large (Note: better word for 'not small enough') $\dt$ the method computes negative values for $u^1$.
-These can be corrected by adapting the $b$. 
+These can be corrected by adapting the weights. 
 If the direct adaptation approach is used the $u^1$ can show gliches/is nonphysical. 
 Alternatively the convex combination approach is used. A first order embedded method is added. The $u^1$ computed with the convex combination are physical.
-For both approaches the remaining steps can be computed with the standard $b$.
+For both approaches the remaining steps can be computed with the standard weights.
 
-In figure\,\ref{fig:conv_impl} the convergence is shown. For $\dt < \num{3e-5} $ the standard $b$ yields to a positive result. For larger $\dt$ the $b$ have to be adapted to ensure positivity.
+In figure\,\ref{fig:conv_impl} the convergence is shown. For $\dt < \num{3e-5} $ the standard weights yield to a positive result. For larger $\dt$ the weights have to be adapted to ensure positivity.
 The EX3a shows similar convergence properties than the original method. 

### 2020-03-11 latex/overview.tex
Abstract and Conclusion, Additional changes
overview.tex
Abstract, Conclusion
Formulated LP
Updated section about algorithm to reduce importance of stepsize reduction

TODOs: Introduction (FE,BE), Numerical results, Section about existence of b

@@
 
+% LP-Box
+\newcommand{\LPblock}[1]{
+\tikzstyle{textbox} = [draw=black, fill=white, very thick,
+    rectangle, rounded corners, inner sep=10pt, inner ysep=20pt]
+\tikzstyle{titlebox} =[fill=white,draw=black, text=black]
+\begin{center}
+\begin{tikzpicture}
+\node [textbox] (box){%
+\begin{minipage}{0.8\textwidth}
+\vspace{-0.3cm}
+{#1}
+\vspace{-0.3cm}
+\end{minipage}
+};
+\node[titlebox, right=10pt] at (box.north west) {LP};
+\end{tikzpicture}
+\end{center}
+}
+
 % bibliography
@@
 \maketitle
+
+\section*{Abstract}
+%(i) General motivation 
+Positivity preserving numerical methods are a ongoing topic in research. 
+(Note: Important?)
+%(ii) Specific problem 
+Many differential equations are known to preserve positivity. 
+This property is not preserved when calculating numerical approximations using most Runge Kutta Methods.
+%(iii) how you propose to solve the problem
+The main idea is to adapt the weights after calculating the stage derivatives. 
+Basically choosing the Runge Kutta Method after evaluating the right hand side. 
+The weights are chosen by solving a Linear Program. 
+%(iv) how you verified your idea and  important results and possible/important limitations
+In this research different ways to formulate the Linear Program were stated. 
+Properties of the adaptive Runge Kutta Methods were investigated analytically.
+It is possible to adapt explicit and diagonally implicit methods. 
+The adaptive Runge-Kutta Methods can be embedded in existing solver frameworks. 
+The behavior was tested numerical on test-cases for explicit and implicit Methods.
 
@@
 The weights $\tilde{b} = b + \Delta b$ consist of the weights of the original method $b$ and an adaption $\Delta b$. The adaption is $\Delta b \in \mathrm{ker}(Q)$. 
-In Figure\,\ref{fig:b_direct} this is drawn. The red line represents the subspace of weights that satisfies the order conditions. Both $b$ and $\tilde{b}$ are in this subspace.
+This is drawn in Figure\,\ref{fig:b_direct}. The red line represents the subspace of weights that satisfie the order conditions. Both $b$ and $\tilde{b}$ are in this subspace. 
+The LP can be expressed as
 
-To run it as a LP a set of constraints is constructed.  
-Also, an objective function is needed to determine the optimal weights.
+\LPblock{
+\begin{subequations}
+\begin{align}
+u^{n+1}&=u^n+\dt F \tilde{b} \geq 0 \label{eq:direct_pos}\\
+Q\tilde{b} &=r \label{eq:direct_Order} \\
+\min_{\tilde b}&( ||\tilde{b} -b ||_1) \label{eq:direct_obj} 
+\end{align}
+\end{subequations}
+}
 
@@
 The Direct Adaptation approach leads to two sets of constraints on $b$.
-The first set of constraints is the positivity and can be written as 
-\begin{equation}
-u_{n+1}=u_n+\dt F \tilde{b} \geq 0
-\end{equation}
-The second set of constraints are the order conditions
-\begin{equation}
-Qb=r
-\end{equation}
+The first set of constraints is the positivity \eqref{eq:direct_pos}.
+The second set of constraints are the order conditions in \eqref{eq:direct_Order}.
 
 \subsubsection{Objective Function}
-Different types of objective functions have been tried.
-
-\begin{description}
-
-\item[L1 norm] The optimization Problem optimizes for $min(|b-rkm.b|_1)$
-
-\item[Quadrature]   The optimization Problem optimizes for $min(|b^Tq-r|)$
-                            where q is the quadrature condition of the next higher order 
-                            and r is the expected right hand side
-
-\item[Hom. Order]    The optimization Problem optimizes for $min(|bTO|)$ where O are the homogenous Order Conditions of the next higher orders
-
-\end{description}
-
-A reasonable property of the objective function would be that for the LP only constrained by the order condition the optimal solution is $b = b_{orig}$. 
-This can be achieved by $argmin(f_{optim}(b)) = b_{orig}$. 
+A reasonable property of the objective function would be that for the LP only constrained by the order condition the optimal solution is $\tilde{b} = b$. 
+This can be achieved by $argmin(f_{optim}(\tilde{b})) = b$. 
 In practice, this also means that there is no need to solve the optimization problem if no constraints are violated because the solution of this problem is already known.
 This is advantageous for performance reasons.
-The different approaches have been tested on a linear test problem from \cite{kopecz_unconditionally_2018}. The optimization problem was used at every step.
-The magnitude of the weights can get large if the order conditions are used as objective function.
 Large values of b can lead to numerical problems.
 Because $\sum_{i  = 1}^s b_i = 1$, an increase of $||b||$ leads to a combination of negative and positive values.
-Because the computation is done with a finite precision (and usually in floating point) large $||b||$ can lead to absorption effects.
-By changing the weights the stability function is altered. This can lead to oscillations if the poles of the RHS are no longer in the stability region of the resulting method.
-For these reasons $||b-b_{orig}||$ seems like a good choice. The 1-Norm can be implemented the with Linear Programming using slack variables. We choose
-
-\begin{equation}
-f_{obj} = ||b-b_{orig}||_1
-\end{equation}
+Because the computation is done with a finite precision (and usually in floating point) large $||b||$ can lead to numerical difficulties.
+By changing the weights the stability function is altered. 
+For these reasons $||\tilde{b}-b||$ seems like a good choice. The 1-Norm can be implemented in a  LP using slack variables. We choose \eqref{eq:direct_obj}
 as objective function.
@@
 We introduce $\theta \in (0,1]$. The new step taken has the length $\theta \dt$ whereas $\dt$ is the time step used when calculating the stage values.
-To calculate a reduced step the order conditions have to be adapted and a new $b_{orig}$ for the objective function (and more important the error detection afterwards) has to be generated. 
-(TODO: the $b_{orig}$ is still a open question)
-
+To calculate a reduced step the order conditions have to be adapted and a new $b(\theta)$ for the objective function has to be generated. 
 When adapting the Order Condition to a new $\theta$ the Order Condition Matrix $Q$ stays unchanged whereas the right hand side $r(\theta)$ depends on $\theta$.
@@
 The dependency of $r(\theta)$ is nonlinear. This means that the maximum $\theta$ with suitable weights cannot be calculated by solving a LP. This means that different $\theta$ have to be tested. 
-(Note: we could include some algorithm to search for it)
+%(Note: we could include some algorithm to search for it)
 
@@
 the Heat Equation solved with an implicit method of higher order), large changes occur, even though a norm of $\Delta b$ is used as an objective function.
-A numerical example is given in Section\,\ref{sec:Numeric_Results}
+A numerical example is given in Section\,\ref{sec:Numeric_Results}.
 It is known that the backward Euler method lead to a positive solution for any $\dt$ \cite{hundsdorfer_numerical_2003}.
@@
 A possible method to fix this is to generate the new b based on a convex combination of embedded methods. 
+The new weights are defined as 
+\begin{equation}\label{eq:b_convex_def}
+\tilde{b} = \sum_i^d g_i b_i = Bg
+\end{equation}
+where $B=\left[b_1,\cdots,b_d \right]$ is a matrix containing the embedded methods.  
+The LP is defined as
+\LPblock{
+\begin{subequations}
+\begin{align}
+ u^{n+1}&=u^n+\dt K B g \geq 0 \label{eq:Convex_pos}  \\
+ 0 \leq g_i &\leq 1  \forall {i \in \{1, \cdots d \}} \quad \sum_{i=1}^d g_i = 1 \label{eq:Convex_cvx} \\
+ \min_{g} &\left(\sum_i^d w_i g_i \right) \label{eq:Convex_obj}
+\end{align}
+\end{subequations}
+}
 
@@
 $b_{\alpha}$ and $b_{\beta}$ are the methods that are combined to $b$ using the parameter $\chi$. The Matrix $Q$ and the vector $r$ are the order conditions as defined in Section\,\ref{sec:OrderCond}.
-
-The new weights are defined as 
-\begin{equation}\label{eq:b_convex_def}
-b = \sum_i^d g_i b_i = Bg
-\end{equation}
-where $B=\left[b_1,\cdots,b_d \right]$ is a matrix containing the embedded methods.  
-To ensure that \eqref{eq:b_convex_def} yields to a convex combination the following constraints are enforced for the parameters $g_1,\cdots,g_d$.
-\begin{equation}
- 0 \leq g_i \leq 1  \forall {i \in \{1, \cdots d \}}
-\end{equation}
-\begin{equation}
- \sum_{i=1}^d g_i = 1
-\end{equation}
-The positivity constraint changes to 
-$$u_{n+1}=u_n+\dt K B g \geq 0$$
-where $g = [g_1,\cdots,g_d]^T$
+To ensure that \eqref{eq:b_convex_def} yields to a convex combination the constraints \eqref{eq:Convex_cvx} are enforced for the parameters $g_1,\cdots,g_d$.
+The positivity constraint changes to \eqref{eq:Convex_pos}
 
@@
 
-As an objective function,
-\begin{equation}
-f_{obj} = min \left(\sum_i^d w_i a_i \right)
-\end{equation}
+As an objective function \eqref{eq:Convex_obj}
 is used.  The $w_1,\cdots,w_d$ are weighting factors that determine which is the preferred $b_i$ to use.
 
-Because the optimal solution of the unconstrained problem has to be the original weights the weight $w_0$ corresponding to $b_{orig}$ should be the smallest $w$.
-This weight is simply set to $w_{b_{orig}} = 0$.
+Because the optimal solution of the unconstrained problem has to be the original weights the weight corresponding to $b$ should be the smallest $w$.
+This weight is simply set to $w_{b} = 0$.
 For the other weights $w_i = \frac{1}{\mathrm{Order} \{b_i\}}$ is used.
+This prefers embeded methods of higher order.
 
@@
 err = |u(t^n)-u^n| &= |u(t^n) - (u^n_{b_{orig}+\dt K(b-b_{orig})})| \\
- &\leq \underbrace{|u(t^n)-u^n_{b_{orig}}|}_{\approx err_T}+\underbrace{|\dt K(b-b_{orig})|}_{= err_{adapt}}
+ &\leq \underbrace{|u(t^n)-u^n_{b_{orig}}|}_{\approx err_T}+\underbrace{|\dt K(b-b_{orig})|}_{= err_{adapt}} \label{eq:Err}
 \end{align}
@@
 \begin{equation}
- u_i^{n+1} \geq 0   \;   \forall {i \in \{1, \cdots,m \}}  
+ u_i^{n+1} \geq 0   \quad   \forall {i \in \{1, \cdots,m \}}  
 \end{equation}
@@
 \begin{equation}
-u_i + \dt \sum_{j=0}^s F_{i,j}  b_j \geq 0 \forall {i \in \{1, \cdots,m \}}  
+u_i + \dt \sum_{j=0}^s F_{i,j}  b_j \geq 0 \quad \forall {i \in \{1, \cdots,m \}}  
 \end{equation}
@@
 \begin{equation}
-u_i + \dt \sum_{j=0}^s F_{i,j}  b_j  \geq 0   \;   \forall {i \in h \subseteq \{1,\cdots,m \}} 
+u_i + \dt \sum_{j=0}^s F_{i,j}  b_j  \geq 0 \quad   \forall {i \in h \subseteq \{1,\cdots,m \}} 
 \end{equation}
@@
 
-\subsection{•}
+\subsection{Implemention}
 (Todo: Maybe a flowchart and a description on how we stitch everything together.)
 
-The algorithm on adapting the b consists of two main loops.
-The outer loop loops through all the $\theta$, staring with $\theta = 1$ down to the lowest $\theta$.
-The inner loop loops through the orders, starting with the highest Order down to the lowest Order.
-Until acceptable weights are found the algorithm tests combinations of $\theta$ and $p$.
+At fist the stagevaules and stagedeerivatives are calculated.
+After this $u^{n+1}$ is computed using the original weights. If the positivity is not violated, this solution is used to update the solution. 
+If $u^{n+1}$ contains negative values, the weights are adapted using the following algorithm.
+For this different orders are tried, stariting with the highest order that has degrees of Freedom for $\tilde{b}$. 
 At first the Order Conditions are constructed.
 On these a LP-Solver is invoiced.
-If the LP is infeasible the algorithm immediately tries the next combination.
-If the LP is feasible we test if the $b$ is acceptable.
+If the LP is infeasible the algorithm immediately reduces the order.
+If the LP is feasible we test if the solution is acceptable.
 At first it is tested weather the new $u^{n+1}$ indeed complies with the constraints.
 This is necessary because sometimes the used LP-solvers incorrectly identify the problem as feasible.
-(Note: because of the chosen objective function it cannot be unbounded) As second test the deviation from the old solution (Compare Error estimation) is calculated.
-If the error is bigger than a set maximum error the $b$ is also rejected and the algorithm also switches to the next combination.
-As soon as the algorithm finds suitable weights the loops are aborted and the weights are used for the update step.
-If all the combinations of $\theta$ and $p$ did not lead to acceptable weights the step has to be rejected and a new set of stage values has to be computed.
-
-(Note: It would also be possible to use another strategy for traversing all the combinations of $(p,\theta)$. Is it worth exploring these? Eg, a bisect algorithm for $\Theta$ or linear approximations)
+%(Note: because of the chosen objective function it cannot be unbounded)
+Afterwards the adaption error is calculated using \eqref{eq:Err}. If the error is bigger than a set maximum error the $\tilde{b}$ is also rejected.
+If the new set of weights is requected in one of the tests the order is reduced.
+As soon as the algorithm finds suitable weights these are used to update the solution.
+If all acceptable orders are tried and no weights were found the step hast to be requected.
+If the reduction of stepsize is used this process is done first for $\theta = 1$. If no weights were found, the process is repeated for a smaler $\theta$.
 
@@
 
-Reaction Problem with step-size control
+\section{Conclusion} \label{sec:conclusion}
+It is possible to adapt the weights to enforce positivity for RKMs that are not positifity preserving.
+A main limitation is that the resulting Order has to be lower than the number of stages. 
+An error approximation for this method was given. 
+The region of absolute stability is altered by changing the weights. This effect can be predicted or controlled.
+Used with explicit methods the positivity for some test problmes could be recovered. 
+Because the time steps is limited by the stability it is only useful for a small interval of time steps.
+The adaptive method mainly interesting for diagonally implicit methods. 
+The times teps here are not limited by stability.
+Also the cost of solving the LP is not crucial factor. 
+If the negative values occurring are not to large, which can be expected for most computations, adapting the weights is a potential way to fix these.
 
-\subsection{Dense Output}
-
-\section{Conclusion} \label{sec:conclusion}
+%also if exactnes is not the limiting factor but positivity
 

### 2020-03-15 latex/overview.tex
Added Plots for simple examples
overview.tex:
Added Plots, description

Convergence.ipynb
Added indicators for Order
More Points for dp5 convergence plot

and multiple pdf-Files

@@
 \subsection{Explicit methods}
+When used with explicit methods
+
+Cost of solving the LP significant (Depending on complexity of RHS)
+
 For most of the test problems with stability occur before getting negative values. This means that we need methods with large stability region.
 
-As test problem the PDE
+(Positifity of stagevaulues not esured, therefore the right had side has to be defined for negative values. If this is not ensured, the right hand side has to be defined for negative values.
+%All the following testproblems did not nead this modification.
+
+As first test problem we consider the PDE
 \begin{equation}
@@
 The Problem consists of a Advection part and an Exponential decay.
-At $x =0$ a inflow boundary with $u(0,t)=1$ is used.
+At $x=0$ a inflow boundary with $u(0,t)=1$ is used.
 The initial condition is $u(x,0)=0$
@@
 The parameters are set to $a=1$ and $K=1$.
+We use the Dorman-Prince 5 RKM and adapt the weights using the direct adaptation.
+In Figure\,\ref{fig:sol_AdDe} the results for $\Delta t = 0.02$ are plottet for different timesteps. We can see that the solution approaches a exponential function with $t \rightarrow \infty$.
+In Figure\,\ref{fig:weights_AdDe} the used weights are plotted. For $t\leq 0.6$ the weights are altered. For $t > 0.6$ the original weights lead to a positive solution.
+\begin{figure}
+\centering
+\begin{subfigure}[b]{0.45\textwidth}
+\includegraphics[width=1\textwidth]{plots/Advection_Decay.pdf}
+\label{fig:sol_AdDe}
+\caption{Solution of Advection Decay Problem}
+\end{subfigure}
+\\
+\begin{subfigure}[b]{0.45\textwidth}
+\includegraphics[width=1\textwidth]{plots/b_Advection_Decay.pdf}\\
+\label{fig:weights_AdDe}
+\caption{Weights used for Advecion Decay Problem}
+\end{subfigure}
+\end{figure}
+
+Next different timesteps are used.
 For $\dt \leq 0.0082$ the unaltered method leads to positive solutions.
@@
 \end{figure}
+TODO: Add converence bars and give exolaination
 
@@
 Implicit methods seem like an advantageous choice for a couple of reasons.
+At first the cost of solving the LP is relativeley small compared to the cost of solving the stageequations.
+Second the timestep is not limited by the stability of the method.
+Therefore it is possible to use larger timesteps, that are more likely to lead to negative values.
+A very interesting calss of methods are the implicit extraploation methods.
+These are Diagonaly implict and $p < s$.
+This means that they allow for changes of the weights without a reduction of the order.
+Also all stegevalues are computed using a the BE method. This means that all intermedaite stages are nonnegative. 
+Also a embeded BE euler step is included. This ensures that an positive solution always exists, even if it is of first order.
 
-The method is tested on the diffusion equation 
+We test the adaptive RKM on the diffusion equation 
 \begin{equation}
@@
 These can be corrected by adapting the weights. 
-If the direct adaptation approach is used the $u^1$ can show gliches/is nonphysical. 
-Alternatively the convex combination approach is used. A first order embedded method is added. The $u^1$ computed with the convex combination are physical.
+The siolutions are computed using the direct adaptation and convex combination for $\dt = \num{1e-3}$
+The results for the direct adaptation are plotted in Figure\,\ref{fig:sol_Diff_a}. The used weights are plotted in Figure\,\ref{fig:weights_Diff_a}.
+The original solution for the first step are negative. Therefore the weights have to be changed.
+If we take a look at the solution after the first timestep at $t=0.001$ we can see that $x=0$ the solution is samler than the solution at the surrounding points. 
+This is not physical. 
+The next timesteps lead to physical solutions again.
+To prevent this glitch from happening we choose the weights based on a convex combiantion.
+A first order embedded method is added. The solution is shown in Figure\,\ref{fig:sol_Diff_c} and the weights are plotted in Figure\,\ref{fig:weights_Diff_c}.
+The weights for the first step are altered again.
+The weights obtained by the convex combination are different from the weighst obtained
+by taking the direct adaptation.
+The solution for $t=0.001$ computed with the convex combination is physical.
 For both approaches the remaining steps can be computed with the standard weights.
 
-In figure\,\ref{fig:conv_impl} the convergence is shown. For $\dt < \num{3e-5} $ the standard weights yield to a positive result. For larger $\dt$ the weights have to be adapted to ensure positivity.
+\begin{figure}
+\centering
+\begin{subfigure}[b]{0.45\textwidth}
+\includegraphics[width=1\textwidth]{plots/Diff_Direct.pdf}\\
+\label{fig:sol_Diff_a}
+\caption{Solution of Diffusion for direct adaptation}
+\includegraphics[width=1\textwidth]{plots/b_Diff_Direct.pdf}\\
+\caption{Weights used for direct adaptation}
+\label{fig:weights_Diff_a}
+\end{subfigure}
+\begin{subfigure}[b]{0.45\textwidth}
+\includegraphics[width=1\textwidth]{plots/Diff_Convex.pdf}\\
+\label{fig:sol_Diff_c}
+\caption{Solution of Diffusion for convex combiantion}
+\includegraphics[width=1\textwidth]{plots/b_Diff_Convex.pdf}\\
+\label{fig:weights_Diff_c}
+\caption{Weights used for convex combination}
+\end{subfigure}
+\end{figure}
+
+In figure\,\ref{fig:conv_impl} the convergence is shown for the unalteres EX3 method, that potentially leads to negative vaules, the adaptive method with direct adaptation (EX3a) and the adapted method using convex combiantions (EX3c).
+Also the BE method is plotted. It is only of 1st order but preserves positivity for all $\dt$.
+For $\dt < \num{3e-5} $ the standard weights yield to a positive result. For larger $\dt$ the weights have to be adapted to ensure positivity.
 The EX3a shows similar convergence properties than the original method.
+This can be expected, because the adapted method is still of 3rd order.
 The EX3c has larger errors than the EX3a but leads to physical solutions for all time-steps. 
+This is no supprise because the adapted RKM used for the first step is only of 1st order.
+But the method still outperformes the BE
+
 \begin{figure}[h]

### 2020-03-16 latex/overview.tex
Added ADR-Problem to paper
plots/AddvectionDiffusionReaction.ipynb:
Notebook to solve ADR-Problem and generate Plots
Multiple pdf-Files

overview.tex:
Changed descriptions for examples,
Added plots and description for ADR 

Bib_RK_optim.bib:
Added Solving ODEII

@@
 \subsection{Explicit methods}
-When used with explicit methods
-
-Cost of solving the LP significant (Depending on complexity of RHS)
-
-For most of the test problems with stability occur before getting negative values. This means that we need methods with large stability region.
-
-(Positifity of stagevaulues not esured, therefore the right had side has to be defined for negative values. If this is not ensured, the right hand side has to be defined for negative values.
+First we want to test the adaptive RKM used with explicit methods. 
+When used with explicit methods the cost of solving the LP significant because the computation of the stagevalues only requires $s$ evaluations of the RHS.
+For most of the linear test problems tried the explicit methods jield to positive results. 
+When increasing the stepsize issues with stability occur before getting negative values. 
+An example therefore the ODE in in Section\,\ref{sec:example_lin}.
+Some nonlinear RHS may require very small timesteps to preserve positifity. 
+For these adapting the weights could be a possible way to solve them.
+An Exmample is the Reaction equation solved in Section\,\ref{sec:example_reac}.
+The positivity of stagevaulues is not esured, therefore the right had side has to be defined for negative values. If this is not the case, the right hand side has to be defined for negative values.
 %All the following testproblems did not nead this modification.
 
-As first test problem we consider the PDE
+As test problem the PDE
 \begin{equation}
@@
 \includegraphics[width=1\textwidth]{plots/Advection_Decay.pdf}
+\caption{Solution of Advection Decay Problem}
 \label{fig:sol_AdDe}
-\caption{Solution of Advection Decay Problem}
 \end{subfigure}
@@
 \includegraphics[width=1\textwidth]{plots/b_Advection_Decay.pdf}\\
+\caption{Weights used for Advecion Decay Problem}
 \label{fig:weights_AdDe}
-\caption{Weights used for Advecion Decay Problem}
 \end{subfigure}
@@
 For $\dt >0.016$ the dp5 is no longer stable.
-In Figure\,\ref{fig:conv_expl} the convergence is plotted for the altered and unaltered method.
+For the ODE the refernce solution can be computed using the matrix exponential. 
+In Figure\,\ref{fig:conv_expl} the convergence for $t=0.5$ is plotted for the altered and unaltered method.
 
@@
 \end{figure}
-TODO: Add converence bars and give exolaination
+
+The unaltered method has the order $p=5$. 
+The values marked with an cross denote the numerical experiments that required an adaption of the weights. 
+Even though the order is reduced most errors are still close to the error of the unaltered method.
 
@@
 This means that they allow for changes of the weights without a reduction of the order.
-Also all stegevalues are computed using a the BE method. This means that all intermedaite stages are nonnegative. 
+Also all stegevalues are computed using the BE method. This means that all intermedaite stages are nonnegative. 
 Also a embeded BE euler step is included. This ensures that an positive solution always exists, even if it is of first order.
@@
 These can be corrected by adapting the weights. 
-The siolutions are computed using the direct adaptation and convex combination for $\dt = \num{1e-3}$
+The solutions are computed using the direct adaptation and convex combination for $\dt = \num{1e-3}$
 The results for the direct adaptation are plotted in Figure\,\ref{fig:sol_Diff_a}. The used weights are plotted in Figure\,\ref{fig:weights_Diff_a}.
@@
 \includegraphics[width=1\textwidth]{plots/Diff_Direct.pdf}\\
+\caption{Solution of Diffusion for direct adaptation}
 \label{fig:sol_Diff_a}
-\caption{Solution of Diffusion for direct adaptation}
 \includegraphics[width=1\textwidth]{plots/b_Diff_Direct.pdf}\\
@@
 \includegraphics[width=1\textwidth]{plots/Diff_Convex.pdf}\\
+\caption{Solution of Diffusion for convex combiantion}
 \label{fig:sol_Diff_c}
-\caption{Solution of Diffusion for convex combiantion}
 \includegraphics[width=1\textwidth]{plots/b_Diff_Convex.pdf}\\
+\caption{Weights used for convex combination}
 \label{fig:weights_Diff_c}
-\caption{Weights used for convex combination}
 \end{subfigure}
@@
 This is no supprise because the adapted RKM used for the first step is only of 1st order.
-But the method still outperformes the BE
+But the adaptive method still outperformes the BE, even when acocunting for the higher cost per step.
 
@@
 \subsection{Stepsize control}
+
+Next we test the adaptive RKM on a more compelx problem. 
+For this we consider the advection-diffusion-production-destruction systems form \citep{kopecz_comparison_2019}. The system is defined by
+\begin{subequations}
+\label{eq:ADR}
+\begin{align}
+\frac{\partial u_1}{\partial t} &=-a \frac{\partial u_1}{\partial x} + d\frac{\partial^2 u_1}{\partial x ^2} + 0.01u_2 + 0.01 u_3 +0.003u_4 - \frac{u_1 u_2}{0.01+u_1} \\ 
+\frac{\partial u_2}{\partial t} &=-a \frac{\partial u_2}{\partial x} + d\frac{\partial^2 u_2}{\partial x ^2}\frac{u_1u_2}{0.01+u_1} -0.01 u_2-0.5(1-\exp(-1.21 u_2^2)) u_3 -0.05 u_2 \\ 
+\frac{\partial u_3}{\partial t} &=-a \frac{\partial u_3}{\partial x} + d\frac{\partial^2 u_3}{\partial x ^2} + 0.5(1-\exp(-1.21u_2^2)) u_3 - 0.01 u_3 -0.02 u_3 \\ 
+\frac{\partial u_4}{\partial t} &=-a \frac{\partial u_4}{\partial x} + d\frac{\partial^2 u_4}{\partial x ^2} + 0.05 u_2 + 0.02 u_3 + 0.003u_4 
+\end{align}
+\end{subequations}
+
+With the parameters $    a=\num{1e-2} $ and $ d=\num{1e-6}$.
+The PDE is simulated on the doamin $x = [0,1]$ with $N=100$ points. Periodic boundarie conditions are used.
+The advecion part is semidiscreticed using a first order upwind sheme and the diffusion part is semidiscreticed using a 3-poin-scheme. This leads to a positivity preserving system of ODEs. 
+The computation is done using the EX3 method with direct adaptation.
+As stepsize control a PI-Control from\,\cite{hairer_solving_1996} is used. The Error was estimatet using \eqref{eq:Err}. The Tolerance was set to $Tol = 0.01$. 
+The final time is $t_{end} = 50$.
+The simulation required 264 steps. Of these 72 required a adaptation of the weights.
+The solutions for $t=9,t=18,t=27$ and $t=50$ are plotted in Figure\,\ref{fig:Sol_ADP}.
+In Figure\,\ref{fig:sol_ADP18} it can be seen that the reaction occurs in a small interfaces.
+Outside of this regions quantities are close to $0$. Therefore it is very likely that negative values orrur in the numerical approxiamtion.
+In Figure\,\ref{fig:sol_ADP27} it can be seen that at $T=27$ the two reaction interfaces merged.
+Afterwards the reaction stops and the behavoir is mainly controlled by the advection and diffusion part.
+
+In Figure\,\ref{fig:Stats_ADP} different vaules are plottet. 
+In the first subplot the stepsize is plotted. For $t<25$ the timesteps are small. After $t = 30$ the the stepsize increases, because the solution only evovlves slowly afterwards.
+In the second subplot the $min(u_1,u_2,u_3,u_4)$ is plotted for all timesteps that initaily lead to negative values. This value is computet before and after adapting the weights. 
+We can see that relativley large negative values occured at some timesteps.
+After the adaption of the weights all values are cloes to $0$. Therfor the adaption of weights succesfully preserved positifity.
+In the third subplot the approxiamted truncation error $err_T$ and the adaption errro $err_{adapt}$ are plotted.
+We can see that the adaption error is of a similar magnitude as the truncation error. Therefore the totoal error of the method is not increased drastically. 
+In the last subplots the weights are ploted. 
+We can see that the cahnges to the weights are only very small. The adapted RKM is still very close to the original RKM.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.49\textwidth}
+        \includegraphics[width=\textwidth]{plots/ADP_sol_00.pdf}
+        \caption{$t=0$}
+        \label{fig:sol_ADP00}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.49\textwidth}
+        \includegraphics[width=\textwidth]{plots/ADP_sol_18.pdf}
+        \caption{$t=$}
+        \label{fig:sol_ADP18}
+    \end{subfigure}
+
+    \begin{subfigure}[b]{0.49\textwidth}
+        \includegraphics[width=\textwidth]{plots/ADP_sol_27.pdf}
+        \caption{$t=$}
+        \label{fig:sol_ADP27}
+    \end{subfigure}
+	\begin{subfigure}[b]{0.49\textwidth}
+        \includegraphics[width=\textwidth]{plots/ADP_sol_50.pdf}
+        \caption{$t=$}
+        \label{fig:sol_ADP50}
+    \end{subfigure}
+
+    \caption{Solution of Advection-Diffusion-Reaction Problems at different times}\label{fig:Sol_ADP}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.85\textwidth]{plots/ADP_stepsize,b.pdf}
+\caption{Different statistics of computation of ADPD-Problem}
+\label{fig:Stats_ADP}
+\end{figure}
 

### 2020-03-17 latex/overview.tex
Revise introduction.

@@
-\documentclass[a4paper]{scrartcl}
+\documentclass[a4paper]{article}
 
@@
 
-% bibliography
-\usepackage[%
-%   backend=bibtex,bibencoding=ascii,
-  backend=biber,
-%   style=authoryear-comp, dashed=false,
-  style=numeric-comp,
-%   firstinits=true, uniquename=init, %abbreviate first names
-  giveninits=true, uniquename=init, %abbreviate first names
-  natbib=true,
-  url=false,
-  doi=true,
-  isbn=false,
-  backref=false,
-  maxnames=99,
-  ]{biblatex}
-\addbibresource{Bib_RK_optim.bib}
-
 \title{Positivity-Preserving Adaptive Runge-Kutta Methods}
-\author{Nüßlein, Stephan and Ketcheson, David I.}
+\author{Stephan Nüßlein \and David I. Ketcheson}
 
 \makeatletter
-\hypersetup{pdfauthor={\@author}}
-\hypersetup{pdftitle={\@title}}
 \makeatother
@@
 
-\section*{Abstract}
+\begin{abstract}
 %(i) General motivation 
-Positivity preserving numerical methods are a ongoing topic in research. 
-(Note: Important?)
+Many important differential equations model quantities whose value
+must remain positive or stay in some bounded interval.
 %(ii) Specific problem 
-Many differential equations are known to preserve positivity. 
-This property is not preserved when calculating numerical approximations using most Runge Kutta Methods.
+These bounds may not be preserved when the model is solved numerically.
 %(iii) how you propose to solve the problem
-The main idea is to adapt the weights after calculating the stage derivatives. 
-Basically choosing the Runge Kutta Method after evaluating the right hand side. 
-The weights are chosen by solving a Linear Program. 
+We propose to ensure positivity or other bounds by applying Runge--Kutta
+integration in which the method weights are adapted in order to
+enforce the bounds.  The weights are chosen at each step after calculating the
+stage derivatives, in a way that also preserves (when possible) the order of
+accuracy of the method.  The choice of weights is given by the solution
+of a linear program (LP).
 %(iv) how you verified your idea and  important results and possible/important limitations
-In this research different ways to formulate the Linear Program were stated. 
-Properties of the adaptive Runge Kutta Methods were investigated analytically.
-It is possible to adapt explicit and diagonally implicit methods. 
-The adaptive Runge-Kutta Methods can be embedded in existing solver frameworks. 
-The behavior was tested numerical on test-cases for explicit and implicit Methods.
+We investigate different approaches to choosing the weights by considering
+different objective functions.  We also provide some analysis of the properties
+of Runge--Kutta methods with perturbed weights.  Numerical examples demonstrate
+the effectiveness of the approach, including application to both explicit and
+implicit methods.
+\end{abstract}
 
@@
 Many physical processes can be described with differential equations. 
-The physical quantities that are involved in these processes often only make sense if they are positive or non-negative. 
-Concentrations or the number of particles have to be $\geq 0$.
-Other values may only make sense if they comply with other restrictions. An example of these would be probabilities or fractions which have to be $\geq 0$ and $\leq 1$.
-The ordinary differential equations (ODE) or partial differential equations (PDEs) are often too complex to solve them analytically and therefore require numerical methods to get an approximate solution.
-Numerical methods generally do not satisfy these restrictions on the solutions.
-
-The most common methods to solve ODEs are Runge–Kutta methods (RKMs). 
-With the method of lines (MOL) it is also possible to solve PDEs with RKMs.
-At first the PDE is discretized in space. The resulting semi-discrete problem is an ODE that can be solved with a RKM. 
-It is possible to semi-discretize PDEs in ways that preserve positivity\cite{kopecz_comparison_2019,kopecz_unconditionally_2018}.
-
-Therefore, it is interesting to know how to use RKMs that preserve positivity.
-For RKMs some theoretical limitations are known.
-One is that any unconditionally positive RKM has an order $\leq 1$.
-\cite{hundsdorfer_numerical_2003,bolley_conservation_1978,horvath_positivity_1998}.
-The only well-known RKM that is unconditionally positive is the backward euler method, which is obviously of first order.
-This causes restrictions on the step size for any RKM with a higher order.
-
-There are multiple approaches to get positivity preserving numerical schemes.
-One method is the Strong Stability Preserving Runge Kutta (SSPRK). 
-The idea is to construct a RKM that can be written as convex combination of euler steps. 
-By this the stability properties of the forward euler step are carried over to the SSPRK.
-The same happens to the positivity properties.
-By this the positivity of the method is ensured for time steps $\dt$ with $\dt \leq R \dt_{FE}$, where $R$ depends on the used SSPRK and $\dt_{FE}$ is the forward Euler time step \cite{gottlieb_strong_2011}.
-The Modified Patankar Runge Kutta (MPRK) are 
-based on the Patankar RKM.
-In these destruction Terms are scaled by the new solution.
-This preserves positivity but violates the mass conservation.
-The MPRK overcome this disadvantage by also scaling the production terms.
-This is done by solving a linear equation system \cite{kopecz_comparison_2019}.
-An approach to ensure positivity in a practical application is shown in \cite{shampine_non-negative_2005}.
-Ensuring the positivity is done by redefining the ODE outside the domain.
-Another technique used is event finding methods.
-This approach is implemented in the Matlab ODE Suite.
-Unconditionally positive method of order higher than 1 are the diagonally split Runge-Kutta (DSRK) methods \cite{horvath_positivity_1998}.
-These are no longer part of the general linear methods.
-This allows them to be positive and higher order.
-
-In the following paper an approach is taken that does not overcome the order limitation but is based on the idea that in practical problems for many time steps positivity can be achieved with higher order methods.
-The main idea is to adapt the RKM after the computation of the stage values in a way that ensures positivity.  
-This means that the computed stage values can be used and don't have to be recomputed.
-With increasing complexity of the right-hand side, a step reject has a high cost.
-This makes the additional cost of adapting the RKM economical.
-The resulting weights should comply with the order conditions as far as possible.
-
-A common way to calculate two different solutions with one set of stage values is the usage of embedded methods for error approximation. 
-The idea of adapting the weights after calculating the stage values is used in \cite{ketcheson_spatially_2013}.
+The physical quantities that are involved in these processes often only make sense if they remain within certain bounds.
+For instance, concentrations must be non-negative (we will often say simply {\em positive} for short), while
+probabilities or mass fractions must remain in $[0,1]$.
+The ordinary differential equations (ODEs) or partial differential equations
+(PDEs) that model these quantities are often too complex to be solved
+analytically and therefore require numerical approximation.
+Numerical methods generally may not satisfy these bound constraints.
+In the present work, we develop an approach to ensuring positivity
+or other bound constraints using Runge--Kutta methods (RKMs) for the
+solution of ODEs or semi-discretized PDEs.
+
+%It is possible to semi-discretize PDEs in ways that preserve positivity\cite{kopecz_comparison_2019,kopecz_unconditionally_2018}.
+We say an ODE
+\begin{align} \label{ode}
+    u'(t) & = f(t,u)
+\end{align}
+is positive if 
+\begin{align}
+    u(0)\ge 0 \implies u(t) \ge 0 \text{ for all } t>0.
+\end{align}
+Any RKM (or in fact any genearal linear method) that is unconditionally positivity preserving for all positive ODEs
+must have order $\le 1$ \cite{bolley_conservation_1978}. Notably, the (first-order)
+backward Euler method is unconditionally positivity preserving.
+For any higher order method, we expect positivity only under some restriction
+on the step size.
+
+Several approaches to ensuring numerical positivity exist in the literature.
+One technique is to use event finding methods in order to stop when some quantity
+reaches zero and then proceed in some special way. This approach is implemented in the MATLAB ODE Suite.
+If positivity is preserved under a forward Euler step (with
+some step size restriction$\dt \le \dt_\text{FE}$), then any strong stability preserving Runge--Kutta (SSPRK)
+method will also preserve positivity (with a modified step size restriction)~\cite{gottlieb_strong_2011}.
+Specifically, the positivity of the method is ensured for time steps $\dt$ with
+$\dt \leq {\mathcal C} \dt_\text{FE}$, where ${\mathcal C}$ depends on the SSPRK method.
+The modified Patankar Runge Kutta (MPRK) methods represent another approach to ensuring
+positivity, for specific classes of ODEs.  MPRK methods introduce multiplicative
+factors within the Runge--Kutta stages to ensure positivity, but require the solution
+of a linear algebraic system; see e.g. \cite{kopecz_comparison_2019} and references therein.
+Yet another approach to ensure positivity in a practical application is shown in \cite{shampine_non-negative_2005},
+where the ODE is redefined outside the physical domain. 
+Finally, we mention diagonally split Runge-Kutta (DSRK) methods, which can be unconditionally
+positive and have order higher than one (they avoid the restriction mentioned above because
+they are not general linear methods \cite{horvath_positivity_1998}.  However, in practice such
+methods are less accurate than backward Euler for large step sizes \cite{macdonald2007}.
+
+The rather discouraging theoretical result of \cite{bolley_conservation_1978} shows that one
+should not hope to preserve positivity with a single method for every problem and every
+initial condition.  In the present work we take an approach based on the idea that for
+a particular problem and initial condition, there often exists a method of high order
+that is positivity preserving, at least for a single step.
+The main idea is to adaptively choose the weights $b$ of the RKM, after the
+the stage values are known, in a way that ensures positivity.  
+The selection of the weights requires the solution of a linear program (LP) at every
+step for which the numerical solution would otherwise be non-positive.
+This is a significant cost, but may in fact be an economical alternative to rejecting
+a step or using excessively small step sizes.
+
+The idea of using different weights within a RKM is not new; for instance it is
+the basis of error approximation using embedded RK pairs \cite{hairer_solving_1993}.
+The idea of adapting the weights after calculating the stage values has also been used,
+for instance in \cite{ketcheson_spatially_2013}.
 In this case it is used to adapt the properties of the time integrator for a method of lines solution of a PDE.
 With this approach, it is possible to have different properties of the RKM at different parts of the domain.
-Other methods that adapts the weights are the relaxation Runge–Kutta (RRK) methods. 
-In these, the weights are scaled by a scalar relaxation parameter to conserve convex quantities.
-This can be used to conserve the energy or entropy \cite{ranocha_relaxation_2019,ketcheson_relaxation_2019}
-
-In this paper, an adaptation of the weights is done by solving a Linear program (LP).
-These are convex optimization problems where all constraints and the objective function are affine. These can incorporate both equality and inequality constraints. 
-LPs can be quite reliably solved with the Simplex method or with interior point methods. This can be done in reasonable time even for large problems. 
-\cite{boyd_convex_2004}
-
-The paper unfolds as follows. In section\,\ref{sec:main_idea} the main idea is explained. How the method can be written as LP is explained in section\,\ref{sec:LP}.
+Another class of methods that adapt the weights at the end of a RK step
+are the relaxation Runge–Kutta (RRK) methods. 
+In these, the weights are scaled by a scalar relaxation parameter in order to guarantee
+conservation or monotonicity of a desired functional; e.g. to conserve or
+dissipate energy or entropy
+\cite{ketcheson_relaxation_2019,ranocha_relaxation_2019}.
+
+The paper unfolds as follows. In Section\,\ref{sec:main_idea} the main idea is explained. How the method can be written as LP is explained in Section\,\ref{sec:LP}.
 Section\,\ref{sec:integration} describes how the new approach can be used with different RKM, how it can be integrated in a step size control and how the Region of absolute stability can be approximated.
-In section\,\ref{sec:imple} further details that were used when implementing the algorithm are described.
-In section\,\ref{sec:Numeric_Results} Numerical results are given for multiple test problems.
-A conclusion is given in section\,\ref{sec:conclusion}.
+In Section\,\ref{sec:imple} further details that were used when implementing the algorithm are described.
+In Section\,\ref{sec:Numeric_Results} Numerical results are given for multiple test problems.
+A conclusion is given in Section\,\ref{sec:conclusion}.
 
@@
 Next we test the adaptive RKM on a more compelx problem. 
-For this we consider the advection-diffusion-production-destruction systems form \citep{kopecz_comparison_2019}. The system is defined by
+For this we consider the advection-diffusion-production-destruction systems form \cite{kopecz_comparison_2019}. The system is defined by
 \begin{subequations}
@@
 
-\printbibliography
+\bibliographystyle{apalike}
+\bibliography{Bib_RK_optim}
 

### 2020-03-17 latex/overview.tex
Multiple changes to paper: Spelling mistakes Extended section regarding the suitable RKM with short part about existence of b Minor Changes in multiple places
Overview.tex:
Spelling Mistakes
Rewrote some short parts
Changed names to free adaptation and convex adaptation
...

Bib_RK_optim.bib
Added one source,
Removed filenames

Plots:
Multiple corrections, mainly adding legends, changing names, fixing bounding boxes of pdf


Bib_RK_optim.bib, Advection_Decay.pdf und 14 weitere dateien aktualisiert...

@@
 
-When computing the solution of an ODE $u ' = f(t,u) $ using a RKM with $s$ Stages and the Butcher tableau
+When computing the solution of an ODE $u ' = f(t,u) $ using a RKM with $s$ stages and the Butcher tableau
 \begin{align}
@@
 This is done by adapting the weights.
-
-For the computation of the stage derivatives only the $A$ has to be known. The weights is not needed for this computation. 
+The original weights $b$ are replaced by the adapted weights $\tilde{b}$.
+With this adaptation we get
+\begin{equation}\label{eq:def_b_tilde}
+u^{n+1} = u^n + \dt F \tilde{b}
+\end{equation}
+form \ref{eq:Combination}. 
+This equation is contains $\tilde{b}$ as second unknown alongside $u^{n+1}$. 
+First we have to determine the adapted weights to compute the new solution afterwards.
+For the computation of the stage derivatives only the $A$ has to be known. The weights are not needed for this computation. 
 Based on this observation we can define $(A,\bullet)$ as the set of RKMs that share the same $A$ and only differ in the weights.
 Calculating the stage derivatives can be considered as solving the ODE for all RKM in $(A,\bullet)$ at the same time.
-The idea is to choose one $(A,b) \in (A,\bullet)$ by fixing the $b$. 
-Because the stage derivatives are already known we can choose a $(A,b)$ that leads to a positive $u^{n+1}$, if $(A,b) \in (A,\bullet) | u^{n+1} \geq 0$ exists.
-When altering the weights the properties of a RKM are preserved. 
+The idea is to choose one $(A,\tilde{b}) \in (A,\bullet)$ by fixing the weights. 
+Because the stage derivatives are already known, we can choose a $(A,\tilde{b})$, that leads to a positive $u^{n+1}$, if $(A,\tilde{b}) \in (A,\bullet) | u^{n+1} \geq 0$ exists.
+When altering the weights, the properties of a RKM are preserved. 
 Especially this approach preserves linear invariants like mass conservation, because it is a regular RKM.
@@
 As it can be seen in \eqref{eq:Combination} $u^{n+1}$ is a linear combination of the stage values and the previous solution $u^n$.
-Because the coefficients $A$ are already fixed the Order Conditions reduce to linear equations.
-This means that finding the weights can be done by solving a linear Programming problem.
+Because the coefficients in $A$ are already fixed, the Order Conditions reduce to linear equations.
+This means that finding the weights can be done by solving a LP.
 
@@
 In Figure\,\ref{fig:exampleI} the obtained results are plotted with a dashed line. 
-At $t=1.905$ $u_1$ gets negative. This leads to a diverging solution.
+At $t=1.905$ the value of $u_1$ gets negative. This leads to a diverging solution.
 
 Now the weights are adapted. The results are also plotted in Figure\,\ref{fig:exampleI}. 
-All $u\geq0$ is now fulfilled. A qualitatively correct solution is obtained.
+The positivity constraint is now fulfilled. A qualitatively correct solution is obtained.
 
@@
 Computed with ck5 and $\dt = 0.005$.
-The dashed lines are the reference (Note: Reference might be misleading here because real solution is different) solution without the adaption of the weights.
-The lower plot are the weights }
+The dashed lines are the approximation obtained without the adaption of the weights.
+The lower plot are the weights.
+}
     \label{fig:exampleI}
@@
 
-The used weights $b_1,\cdots,b_5$ are also plotted in Figure\,\ref{fig:exampleI}. 
+The used weights are also plotted in Figure\,\ref{fig:exampleI}. 
 The used weights are equal to the original weights for $t<1.905$.
@@
 
-\section{Details (TODO: better title)}\label{sec:LP}
-
-In the following section, it is explained how the approach can be written as a LP. 
+\section{Adaptive RKM}\label{sec:LP}
+
+In the following Section,  how the approach can be written as a LP. 
 For this we have to consider which weights to choose. 
-Generally, any $b \in \R$ can be used. 
-A first requirement would be that the new solution is positive.
+Generally, any $b \in \R^s$ can be used. 
+One requirement would be that the new solution is positive.
 We can further reduce this space by enforcing the order conditions. 
-By incorporating the order conditions, the space of possible weights is limited to a subspace of $\R$. %(Note: not linear because 0 is not in subspace)
-Finally, we have to give an objective function to get the optimal weights.
+By incorporating the order conditions, the space of possible weights is limited to a subset of $\R^s$. %(Note: not linear because 0 is not in subspace)
+Finally, we have to state an objective function to get the optimal weights.
 
@@
 The main goal is to choose a $(A,b)$ in such a way that $u^{n+1}$ resembles the solution of the ODE $u(t_{n+1})$. 
-A obvious objective is to use a $(A,b)$ with a high order, but this is not enough.
+An obvious objective is to use a $(A,b)$ with a high order, but this is not enough.
 To get a better understanding for the method we consider the behavior of a simple problem.
@@
 \end{equation}
-and and use the three Stage and third order SSP method (SSP33)
+and use the three stage and third order SSP method (SSP33)
 \begin{align}
@@
 \begin{equation}
-f(y_1)=\left[\begin{matrix}-5\\5\end{matrix}\right]\qquad f(y_2)=\left[\begin{matrix}5\\-5\end{matrix}\right]\qquad f(y_3)=\left[\begin{matrix}-5\\5\end{matrix}\right]\end{equation}
+f(y_1)=\left(\begin{matrix}-5\\5\end{matrix}\right)\qquad f(y_2)=\left(\begin{matrix}5\\-5\end{matrix}\right)\qquad f(y_3)=\left(\begin{matrix}-5\\5\end{matrix}\right)\end{equation}
 The values of the next step using the standard weights is
 \begin{equation}
-u_n=\left[\begin{matrix}- \frac{1}{9}\\\frac{10}{9}\end{matrix}\right]
+u_n=\left(\begin{matrix}- \frac{1}{9}\\\frac{10}{9}\end{matrix}\right)
 \end{equation}
 The new solution contains negative values. 
-Therefor we want to adapt the weights to ensure positifity. 
-All weights that comply with the order Condition for the first and second order can be expressed as
-\begin{equation}
-\tilde b=\left[\begin{matrix}\frac{1}{6}\\\frac{1}{6}\\\frac{2}{3}\end{matrix}\right]+\alpha \left[\begin{matrix}\frac{1}{2}\\\frac{1}{2}\\-1\end{matrix}\right]
-\end{equation}
-With this we have one degree of freedom for the choice of the weights. This is done by the parameter $\alpha$.
+Therefore, we want to adapt the weights to ensure positivity. 
+All weights, that comply with the order Condition for the first and second order can be expressed as
+\begin{equation}
+\tilde b=\left(\begin{matrix}\frac{1}{6}\\\frac{1}{6}\\\frac{2}{3}\end{matrix}\right)+\alpha \left(\begin{matrix}\frac{1}{2}\\\frac{1}{2}\\-1\end{matrix}\right).
+\end{equation}
+We have one degree of freedom for the choice of the weights. This is done by the parameter $\alpha$.
 If the general expression for the weights is inserted in \eqref{eq:Combination} the general solution is 
 \begin{equation}
-u^{1} = u^0 + \dt  \left[f(y_1),f(y_2),f(y_3)\right] b =
-\left[\begin{matrix}- \frac{1}{9}\\\frac{10}{9}\end{matrix}\right] +\alpha \left[\begin{matrix}\frac{5}{3}\\- \frac{5}{3}\end{matrix}\right]
-\end{equation}
-By changing the parameter $\alpha$ the weights $b$ and the solution is altered.
+u^{1} = u^0 + \dt  \left(f(y_1),f(y_2),f(y_3)\right) \tilde{b} =
+\left(\begin{matrix}- \frac{1}{9}\\\frac{10}{9}\end{matrix}\right) +\alpha \left(\begin{matrix}\frac{5}{3}\\- \frac{5}{3}\end{matrix}\right)
+\end{equation}
+By changing the parameter $\alpha$, the weights and the solution are altered.
 With a suitable choice of $\alpha \in \left[\frac{1}{15},\frac{2}{3}\right]$ any $u$ that complies with mass conservation and positivity can be reached. 
-By adding additional constraints on the weights the choice of $\alpha$ can be narrowed down. 
+By adding additional constraints on the weights, the choice of $\alpha$ can be narrowed down. 
 An objective function is also needed to make the choice unique. This should be designed in a way to prefer weights that are close to the original weights.
@@
 The order conditions for a RKM are a set of Equations depending on $A$, $c$ and $b$. In these Equations $b$ only appears linearly \cite{hairer_solving_1993}.
-If $A$ and $c$ are already known, the order conditions for the Order $p$ are a linear equations system and can be written as $Q b = r$.
+If $A$ and $c$ are already known, the order conditions for the order $p$ are a linear equations system and can be written as $Q b = r$.
 
@@
 
-\subsection{Ways to formulate the Optimization Problem}
-
-There are two possible ways to write down the adaptation of the weights. The first way is to directly adapt the original weights $b$ by adding a correction $\Delta b$ that ensures positivity.
+\subsection{Formulation of the Optimization Problem}
+
+There are two possible ways to write down the adaptation of the weights. The first way is to freely adapt the original weights $b$ by adding a correction $\Delta b$ that ensures positivity.
 This is the easiest way to adapt the weights. An advantageous property is that the step size can be reduced after calculating the stage values by using the dense output order conditions.  
-The direct adaption is explained in more detail in section\,\ref{sec:direct}.
-
-The second way to formulate a LP is to choose weights as a convex combination of a set of weights. This is more useful if some embedded methods with certain properties are known. 
-The convex combination also shows more predictable behavior because the resulting $u^{n+1}_b$ is also a convex combination of the solutions $u^{n+1}_{b_a},u^{n+1}_{b_b},\cdots$ 
-This is explained in more detail in section\,\ref{sec:convex}
+The free adaption is explained in more detail in Section\,\ref{sec:direct}.
+
+The second way to formulate a LP is to choose weights as a convex adaptation of a set of weights. This is more useful if some embedded methods with certain properties are known. 
+The convex adaptation also shows more predictable behavior because the resulting $u^{n+1}_b$ is also a convex adaptation of the solutions $u^{n+1}_{b_a},u^{n+1}_{b_b},\cdots$ 
+This is explained in more detail in Section\,\ref{sec:convex}
 
@@
         \input{Image_b_direct}
-        \caption{Direct Adaptation}
+        \caption{Free adaptation}
         \label{fig:b_direct}
@@
         \input{Image_b_convex}
-        \caption{Convex Combination}
+        \caption{Convex adaptation}
         \label{fig:b_convex}
@@
 
-\subsection{Direct Adaptation}\label{sec:direct}
-
-When directly adapting the weights new weights that complies with the order conditions is chosen.   
+\subsection{Free Adaptation}\label{sec:direct}
+
+When freely adapting the weights, new weights that complie with the order conditions is chosen.   
 The weights $\tilde{b} = b + \Delta b$ consist of the weights of the original method $b$ and an adaption $\Delta b$. The adaption is $\Delta b \in \mathrm{ker}(Q)$. 
-This is drawn in Figure\,\ref{fig:b_direct}. The red line represents the subspace of weights that satisfie the order conditions. Both $b$ and $\tilde{b}$ are in this subspace. 
+This is drawn in Figure\,\ref{fig:b_direct}. The red line represents the subspace of weights that satisfy the order conditions. Both $b$ and $\tilde{b}$ are in this subspace. 
 The LP can be expressed as
@@
 \subsubsection{Constraints}
-The Direct Adaptation approach leads to two sets of constraints on $b$.
+The free adaptation approach leads to two sets of constraints on $b$.
 The first set of constraints is the positivity \eqref{eq:direct_pos}.
@@
 \subsubsection{Objective Function}
-A reasonable property of the objective function would be that for the LP only constrained by the order condition the optimal solution is $\tilde{b} = b$. 
+A reasonable property of the LP is that the optimal solution for the problem without positivity constraints is $\tilde{b} = b$. 
 This can be achieved by $argmin(f_{optim}(\tilde{b})) = b$. 
@@
 Because the computation is done with a finite precision (and usually in floating point) large $||b||$ can lead to numerical difficulties.
-By changing the weights the stability function is altered. 
+By changing the weights, the stability function is altered. 
 For these reasons $||\tilde{b}-b||$ seems like a good choice. The 1-Norm can be implemented in a  LP using slack variables. We choose \eqref{eq:direct_obj}
@@
 In this case we would like to calculate $u^{n+\theta}$ using the stage values instead of rejecting the step.
-For this the time step is reduced after calculation the stage values.
+The time step is reduced after calculation the stage values.
 This can be done using the order conditions for dense output.
@@
 
-\subsection{Convex combination of existing methods}\label{sec:convex}
+\subsection{Convex Adaptation of Existing Methods}\label{sec:convex}
 For some test problems (e.g.
@@
 A numerical example is given in Section\,\ref{sec:Numeric_Results}.
-It is known that the backward Euler method lead to a positive solution for any $\dt$ \cite{hundsdorfer_numerical_2003}.
+It is known that the backward Euler method leads to a positive solution for any $\dt$ \cite{hundsdorfer_numerical_2003}.
 The backward Euler is still only first Order but shows more exact results than a method of higher Order that is distorted by the adaption of the weights, at least at some steps.
-A possible method to fix this is to generate the new b based on a convex combination of embedded methods. 
+A possible method to fix this is to generate the new b based on a convex adaptation of embedded methods. 
 The new weights are defined as 
@@
 At first a set of embedded methods of the RKM $b_1,\cdots,b_d$ is needed. 
-If all the used weights complie with the order conditions the new method is also of the same Order.
+If all the used weights complie with the order conditions, the new method is also of the same order.
 This is straightforward using
@@
 $b_{\alpha}$ and $b_{\beta}$ are the methods that are combined to $b$ using the parameter $\chi$. The Matrix $Q$ and the vector $r$ are the order conditions as defined in Section\,\ref{sec:OrderCond}.
-To ensure that \eqref{eq:b_convex_def} yields to a convex combination the constraints \eqref{eq:Convex_cvx} are enforced for the parameters $g_1,\cdots,g_d$.
+To ensure that \eqref{eq:b_convex_def} yields to a convex adaptation the constraints \eqref{eq:Convex_cvx} are enforced for the parameters $g_1,\cdots,g_d$.
 The positivity constraint changes to \eqref{eq:Convex_pos}
@@
 
-As a first order method, the weights that corresponds to a series of backward Euler steps are added. These are known to yield a positive result for any step size.
+As a first order method, the weights that correspond to a series of backward Euler steps are added. These are known to yield a positive result for any step size.
 It is also possible to add embedded methods of higher order.
@@
 
-\section{TODO: Title} \label{sec:integration}
-
-In the previous sections an algorithm for choosing positifity preserving weights $\tilde{b}$ has been presented. 
-In the following section, it is explained how this Algorithm can be integrated into the framework of existing solvers.
+\section{Properties of Adaptive RKM} \label{sec:integration}
+
+In the previous Sections an algorithm for choosing positifity preserving weights $\tilde{b}$ has been presented. 
+In the following Section, it is explained how this Algorithm can be integrated into the framework of existing solvers.
 
@@
 As noted in Section\,\ref{sec:OrderCond} the number of stages has to be higher than the order.  
+Therefore, explicit RKM are interesting, because an explicit RKM of Order $p$ requires at least $p$ stages. 
+Fully implicit RKM can achieve higher orders.  
+To fulfill $p<s$ the order has to be reduced drastically. Therefore, these methods are not good candidates for the adaptation. 
+As stated in \cite{norsett_attainable_1977} the highest achievable order of a Diagonally implicit RKM is $s+1$. This condition is far closer to the condition required for the adaptation. 
 Another important property is the number of degrees of freedom for the choice of the new weights.  
-These can be calculated using $s-\mathrm{rank}(Q)$.%$\mathrm{dim}(\mathrm{ker}(Q))$.
-The number of degrees of freedom for different methods are shown in table\,\ref{table:DOF_exp} for explicit methods and in table\,\ref{table:DOF_imp} for implicit methods.
+These can be calculated using $s-\mathrm{rank}(Q)$. %$\mathrm{dim}(\mathrm{ker}(Q))$. 
+The number of degrees of freedom for different explicit methods are shown in Table\,\ref{table:DOF_exp}.
 
@@
 \centering    %Generated below============ 
-\begin{tabular}{|l |c c c c c c |} 
+\begin{tabular}{|l c|c c c c c c |} 
  \hline 
-Order &1&2&3&4&5&6 \\ 
- \hline Classical RK4&3&2&0&0& - & -  \\ 
- SSPRK(10,4)&9&8&6&4& - & -  \\ 
- Cash-Karp RK5(4)6&5&4&2&1&0& -  \\ 
- Dormand-Prince RK5(4)7&6&5&3&1&0& -  \\ 
+Order & &1&2&3&4&5&6 \\ 
+ \hline Classical RK4& s=4&3&2&0&0& - & -  \\ 
+ SSPRK(10,4)& s=10&9&8&6&4& - & -  \\ 
+ Cash-Karp RK5(4)6& s=6&5&4&2&1&0& -  \\ 
+ Dormand-Prince RK5(4)7& s=7&6&5&3&1&0& -  \\ 
  \hline 
@@
 \centering   %Generated below============ 
- \begin{tabular}{|l |c c c c c c |} 
+\begin{tabular}{|l c|c c c c c c |} 
  \hline 
-Order &1&2&3&4&5&6 \\ 
- \hline Implicit Euler&0& - & - & - & - & -  \\ 
- Im-Euler 2&2&1& - & - & - & -  \\ 
- Im-Euler 3&5&4&2& - & - & -  \\ 
- Im-Euler 4&9&8&6&3& - & -  \\ 
+Order & &1&2&3&4&5&6 \\ 
+ \hline Implicit Euler& s=1&0& - & - & - & - & -  \\ 
+ Lobatto IIIC4& s=4&3&2&1&0&0&0 \\ 
+ Radau IIA3& s=3&2&1&0&0&0& -  \\ 
+ SDIRK 54& s=5&4&3&1&0& - & -  \\ 
+ TR-BDF2& s=3&2&1& - & - & - & -  \\ 
+ Im-Euler 2& s=3&2&1& - & - & - & -  \\ 
+ Im-Euler 3& s=6&5&4&2& - & - & -  \\ 
+ Im-Euler 4& s=10&9&8&6&3& - & -  \\ 
  \hline 
@@
 If the classical RK4 method is used the Order has to be reduced more because the RK4 method does not have embedded methods of order 3. 
-Contrary to this methods with $s > p$ have degrees of freedom even without reducing the order.
+Contrary to this, methods with $s > p$ have degrees of freedom even without reducing the order.  
+For Cash-Karp RK5 and Dorman-Prince RK5 the Order has to be reduced to get degrees of freedom even though the number of stages is higher than the order.
+
+In Table\,\ref{table:DOF_imp} same was done for implicit methods. 
+We can see that the fully implicit methods Lobatto IIIC4 and Radua IIA3 require a drastic reduction of the order as expected. 
+The Diagonally implicit SDIRK 54 method only requires an order reduction of one to get one degree of freedom for the weights.  
+The TR-BDF2 method even allows adaptations without reducing the order.  
+The Implicit-Euler extrapolation methods also exhibit degrees of freedom without a reduction of the order.
 
 Another important property is the stability region. 
-The adaption of the weights does solve issues with stabilitz when used with a too big $\dt$.
-Therefore the RK-method has to be stable for large $\dt$.
+The adaption of the weights does not solve issues with stability when used with a too big $\dt$. Therefore, the RK-method has to be stable for large $\dt$.
 
@@
 This is particularly interesting for implicit methods, because the $\dt$ is not limited by the stability.
-If we have an embedded first Order Method we can ensure that there is always a positive solution, which might be of first Order.
+If we have an embedded first order method, we can ensure that there is always a positive solution, which might be of first Order.  
+This is the case for the Im-Euler extrapolation methods, because they contain a chain of BE steps. 
+For explicit methods the FE method is always embedded, but a reduction of the step size might be required.
 
@@
 Now it is important to know if the new solution $u^{n+1}$ still resembles the solution of the ODE $u(t_{n+1})$. 
-This can be seen in example in section\,\ref{sec:example_lin}.
-By a suitable choice of $\alpha \in [\frac{1}{6},\frac{7}{30}]$ any $u$ that complies with mass conservation and positifity can be reached.
-
+This can be seen in example in Section\,\ref{sec:example_lin}. 
 This gives rise to the question whether the new solution is reasonable.
-This also poses the problem on developing a system that can decide if the new solution is has to be rejected.
+This also poses the problem on developing a system that can decide whether the new solution is acceptable or has to be rejected.
 
@@
 \begin{align}
-err = |u(t^n)-u^n| &= |u(t^n) - (u^n_{b_{orig}+\dt K(b-b_{orig})})| \\
- &\leq \underbrace{|u(t^n)-u^n_{b_{orig}}|}_{\approx err_T}+\underbrace{|\dt K(b-b_{orig})|}_{= err_{adapt}} \label{eq:Err}
+err = |u(t^n)-u^n| &= |u(t^n) - (u^n_{unadapted}+\dt K(\tilde{b}-b))| \\
+ &\leq \underbrace{|u(t^n)-u^n_{unadapted}|}_{\approx err_T}+\underbrace{|\dt K(\tilde{b}-b)|}_{= err_{adapt}} \label{eq:Err}
 \end{align}
 
-After adapting the weights the approximation error is checked. If $|\dt K(b-b_{orig})|$ is larger than the tolerance the weights are rejected.
-
+The total error is split up in the truncation error $err_T$ and the adaption error $err_{adapt}$ using the triangle inequality. 
+The truncation error can be estimated using the standard error estimators $err_T = | u^{n}_{b_{orig}} - u^{n}_{\hat{b}} |$. 
+After adapting the weights, the adaption error is calculated. If the adaption error is larger than the tolerance the weights are rejected.  
+Both errors are added to get an approximation of the total error $err = err_T + err_{adapt}$.
+The truncation error and the adaption error never have the same direction.  
+Therefore, the triangle inequality becomes a strict inequality and the real error is always smaller than the estimate. (Note: as long as the Truncation error estimator is right) 
 %\subsubsection{Steppsize Control}
 %An important part of an RKM method is the ability to approximate the error of the solution to use it to control the step size. 
-The truncation error $|u(t^n)-u^n_{b_{orig}}|$ is usually approximated with an standard error estimator $err_T = | u^{n}_{b_{orig}} - u^{n}_{\hat{b}} |$.
-Both errors are added to get an approximation of the total error $err = err_T + err_{adapt}$.
+%The truncation error $|u(t^n)-u^n_{b_{orig}}|$ is usually approximated with an standard error estimator $err_T = | u^{n}_{b_{orig}} - u^{n}_{\hat{b}} |$.
  %The parameter $w_a \leq 1$ is used to account for the fact that the real error is smaller than $err = err_T +err_{adapt}$.
@@
 
+In Figure\,\ref{fig:stab} the stabilit regions of adapted RKM are plotted. 
+In Figure\,\ref{fig:stab_ck5} the Dorman-Prince 5 method is freely adapted. 
+The weights are taken form the example in Section\,\ref{sec:Ex_expl}.
+
+In Figure\,\ref{fig:stab_ex3} the stabilty regions of the BE-Extraplotaion 3 method and the embedded chain of three BE steps is plotted.
+Additionally the stabilty regions of convex combiantions of these two methods are plotted.
+
 \begin{figure}
@@
          \centering
-         \includegraphics[width=\textwidth]{plots/stab_ck5.pdf}
-         \caption{Cash-Karp 5}
+         \includegraphics[width=\textwidth]{plots/stab_dp5.pdf}
+         \caption{Dorman-Prince 5\\
+         with $\tilde{b}$ form \ref{fig:weights_AdDe}
+         }
          \label{fig:stab_ck5}
@@
          \includegraphics[width=\textwidth]{plots/stab_ex3.pdf}
-         \caption{Implicit Extrapolation 3}
+         \caption{Implicit Extrapolation 3 and embedded chain of three BE steps with convex combinations of both weights}
          \label{fig:stab_ex3}
      \end{subfigure}
-        \caption{Change of stability region for direct adaptation(a) and for convex combination(b) (TODO: legend, axis, description in text)}
+        \caption{Change of stability region for free adaptation(a) and for convex adaptation(b)}
         \label{fig:stab}
@@
 \begin{multline}
-\alpha P_a(z) + \gamma P_c(z) =  \alpha z a^T (I - zA)^{-1} + \gamma z c^T (I - zA)^{-1} =\\
-z (\alpha a+ \gamma c)^T (I - zA)^{-1} = P_{\alpha a + \gamma c}(z) \in \mathcal{P}
+\alpha P_a(z) + \gamma P_c(z) =  \alpha z a^T (I - zA)^{-1} e + \gamma z c^T (I - zA)^{-1} e=\\
+z (\alpha a+ \gamma c)^T (I - zA)^{-1} e= P_{\alpha a + \gamma c}(z) \in \mathcal{P}
 \end{multline}
@@
 \paragraph*{Version 3} $ $\\
-(Ignore this topic, because the following sections also can work without it)
+(Ignore this topic, because the following Sections also can work without it)
 
@@
 
-\subsubsection{Stability of convex combination}
-If the new weights are chosen as a convex combination of weights it is easy to prove properties of the stability region.
-The stability function $R_{\chi b_a + (1-\chi) b_b}(z)$ of a convex combination can be calculated by the weighted sum of the stability functions $R_{b_a}(z)$ and $R_{b_b}(z)$
+\subsubsection{Stability of Convex Adaptation}
+If the new weights are chosen as a convex adaptation of weights it is easy to prove properties of the stability region.
+The stability function $R_{\chi b_a + (1-\chi) b_b}(z)$ of a convex adaptation can be calculated by the weighted sum of the stability functions $R_{b_a}(z)$ and $R_{b_b}(z)$
 
@@
 
-The resulting region of absolute stability is at least the intersection of the regions of absolute stability of the methods that define the convex combination.
+The resulting region of absolute stability is at least the intersection of the regions of absolute stability of the methods that define the convex adaptation.
 This also means that if all the used embedded methods used to construct the new weights are A-Stable, the resulting method is also A-Stable.
 
-We want to show that the convex combination of two methods is absolutely stable at the points where both original methods were absolutely stable.
+We want to show that the convex adaptation of two methods is absolutely stable at the points where both original methods were absolutely stable.
 This can be formulated as $(|R_{b_a}(z)|  < 1) (|R_{b_b}(z)| < 1) \Rightarrow |R_{\chi b_a +(1- \chi) b_b}(z)| < 1$ with $\chi \in [0,1]$.
@@
 
-\subsubsection{Stability of  Direct Adaptation}
-If the weights is adapted directly there is no such proof.
+\subsubsection{Stability of  Free Adaptation}
+If the weights is adapted freely there is no such proof.
 A useful property of the resulting stability region is that it is similar to the stability region of the baseline method. 
@@
 
-TODO: Discuss in this section aspects that are needed for implementation (and improve the usability) but do not concern the mathematical idea of the approach.
+TODO: Discuss in this Section aspects that are needed for implementation (and improve the usability) but do not concern the mathematical idea of the approach.
 
@@
 When enforcing a maximum value the number of constraints can be reduced using the same technique.
-When nforcing both maximum and minimum values two separate sets of active constraints $k_{min}$ and $k_{max}$ are used.
+When enforcing both maximum and minimum values two separate sets of active constraints $k_{min}$ and $k_{max}$ are used.
 The set $k_{max}$ denotes the active maximum constraints and the set $k_{min}$ denotes the active minimum constraints.
@@
 
-\subsection{Explicit methods}
+\subsection{Explicit methods}\label{sec:Ex_expl}
 First we want to test the adaptive RKM used with explicit methods.  
-When used with explicit methods the cost of solving the LP significant because the computation of the stagevalues only requires $s$ evaluations of the RHS.
-For most of the linear test problems tried the explicit methods jield to positive results. 
-When increasing the stepsize issues with stability occur before getting negative values. 
+When used with explicit methods the cost of solving the LP significant because the computation of the stagederivatives only requires $s$ evaluations of the RHS. 
+For most of the linear test problems tried, the explicit methods yield to positive results.  
+When increasing the step size, issues with stability occur before getting negative values.  
 An example therefore the ODE in in Section\,\ref{sec:example_lin}. 
-Some nonlinear RHS may require very small timesteps to preserve positifity. 
-For these adapting the weights could be a possible way to solve them.
-An Exmample is the Reaction equation solved in Section\,\ref{sec:example_reac}.
-The positivity of stagevaulues is not esured, therefore the right had side has to be defined for negative values. If this is not the case, the right hand side has to be defined for negative values.
+Some nonlinear RHS may require very small timesteps to preserve positivity.  
+For these, adapting the weights could be a possible way to solve them. 
+An example is the reaction equation solved in Section\,\ref{sec:example_reac}. 
+The positivity of stage values is not ensured, therefore the right hand side has to be defined for negative values. If this is not the case, the right hand side has to be defined for negative values. 
 %All the following testproblems did not nead this modification.
@@
 is simulated with the method of lines on the domain $x = [0,1]$ with $N =100$ points.
-The Problem consists of a Advection part and an Exponential decay.
+The problem consists of an Advection part and an Exponential decay.
 At $x=0$ a inflow boundary with $u(0,t)=1$ is used.
-The initial condition is $u(x,0)=0$
+The initial condition is $u(x,0)=0$.
 As semidiscretization an upwinding scheme is used. This leads to positivity preserving ODE
@@
 The parameters are set to $a=1$ and $K=1$.
-We use the Dorman-Prince 5 RKM and adapt the weights using the direct adaptation.
-In Figure\,\ref{fig:sol_AdDe} the results for $\Delta t = 0.02$ are plottet for different timesteps. We can see that the solution approaches a exponential function with $t \rightarrow \infty$.
+We use the Dorman-Prince 5 RKM and adapt the weights using the free adaptation.
+In Figure\,\ref{fig:sol_AdDe} the results for $\Delta t = 0.02$ are plotted for different timesteps. We can see that the solution approaches an exponential function with $t \rightarrow \infty$.
 In Figure\,\ref{fig:weights_AdDe} the used weights are plotted. For $t\leq 0.6$ the weights are altered. For $t > 0.6$ the original weights lead to a positive solution.
@@
 
-Next different timesteps are used.
+Next, different timesteps are used.
 For $\dt \leq 0.0082$ the unaltered method leads to positive solutions.
@@
 For $\dt >0.016$ the dp5 is no longer stable.
-For the ODE the refernce solution can be computed using the matrix exponential. 
+For the ODE, the reference solution can be computed using the matrix exponential. 
 In Figure\,\ref{fig:conv_expl} the convergence for $t=0.5$ is plotted for the altered and unaltered method.
@@
 The unaltered method has the order $p=5$. 
-The values marked with an cross denote the numerical experiments that required an adaption of the weights. 
-Even though the order is reduced most errors are still close to the error of the unaltered method.
+The values marked with a cross denote the numerical experiments that required an adaption of the weights. 
+Even though the order is reduced, most errors are still close to the error of the unaltered method.
 
@@
 Implicit methods seem like an advantageous choice for a couple of reasons. 
-At first the cost of solving the LP is relativeley small compared to the cost of solving the stageequations.
-Second the timestep is not limited by the stability of the method.
-Therefore it is possible to use larger timesteps, that are more likely to lead to negative values.
-A very interesting calss of methods are the implicit extraploation methods.
-These are Diagonaly implict and $p < s$.
+Fistly the cost of solving the LP is relatively small compared to the cost of solving the stage equations. 
+Secondly the timestep is not limited by the stability of the method. 
+Therefore, it is possible to use larger timesteps, that are more likely to lead to negative values. 
+A very interesting class of methods are the implicit extrapolation methods.
+These are Diagonally implicit and $p < s$.
 This means that they allow for changes of the weights without a reduction of the order.
-Also all stegevalues are computed using the BE method. This means that all intermedaite stages are nonnegative. 
-Also a embeded BE euler step is included. This ensures that an positive solution always exists, even if it is of first order.
-
+Also, all stage values are computed using the BE method. This means that all intermediate stages are positive.  
+Also, an embedded BE Euler step is included. This ensures that a positive solution always exists, even if it is of first order.  
 We test the adaptive RKM on the diffusion equation 
@@
 \end{equation}
-As initial condition $u^0 = (0,\cdots,0,1,0,\cdots,0)^T$ is used. The diffusion coefficient is$D=1$
+As initial condition $u^0 = (0,\cdots,0,1,0,\cdots,0)^T$ is used. The diffusion coefficient is $D=1$ 
 The ODE is solved using the implicit EX3 method.  
@@
 These can be corrected by adapting the weights.  
-The solutions are computed using the direct adaptation and convex combination for $\dt = \num{1e-3}$
-The results for the direct adaptation are plotted in Figure\,\ref{fig:sol_Diff_a}. The used weights are plotted in Figure\,\ref{fig:weights_Diff_a}.
-The original solution for the first step are negative. Therefore the weights have to be changed.
-If we take a look at the solution after the first timestep at $t=0.001$ we can see that $x=0$ the solution is samler than the solution at the surrounding points. 
+The solutions are computed using the free adaptation and convex adaptation for $\dt = \num{1e-3}$.
+The results for the free adaptation are plotted in Figure\,\ref{fig:sol_Diff_a}. The used weights are plotted in Figure\,\ref{fig:weights_Diff_a}. 
+The original solution for the first step is negative. Therefore, the weights have to be changed. 
+If we take a look at the solution after the first timestep at $t=0.001$ we can see that at $x=0$ the solution is smaller than the solution at the surrounding points.  
 This is not physical.  
 The next timesteps lead to physical solutions again. 
-To prevent this glitch from happening we choose the weights based on a convex combiantion.
+To prevent this glitch from happening we choose the weights based on a convex adaptation. 
 A first order embedded method is added. The solution is shown in Figure\,\ref{fig:sol_Diff_c} and the weights are plotted in Figure\,\ref{fig:weights_Diff_c}. 
 The weights for the first step are altered again. 
-The weights obtained by the convex combination are different from the weighst obtained
-by taking the direct adaptation.
-The solution for $t=0.001$ computed with the convex combination is physical.
-For both approaches the remaining steps can be computed with the standard weights.
-
+The weights obtained by the convex adaptation are different from the weights obtained 
+by taking the free adaptation. 
+The solution for $t=0.001$ computed with the convex adaptation is physical. 
+For both approaches, the remaining steps can be computed with the standard weights. 
 \begin{figure}
@@
 \includegraphics[width=1\textwidth]{plots/Diff_Direct.pdf}\\
-\caption{Solution of Diffusion for direct adaptation}
+\caption{Solution of Diffusion for free adaptation \newline $ $}
 \label{fig:sol_Diff_a}
-\includegraphics[width=1\textwidth]{plots/b_Diff_Direct.pdf}\\
-\caption{Weights used for direct adaptation}
-\label{fig:weights_Diff_a}
 \end{subfigure}
@@
 \includegraphics[width=1\textwidth]{plots/Diff_Convex.pdf}\\
-\caption{Solution of Diffusion for convex combiantion}
+\caption{Solution of Diffusion for convex adaptation}
 \label{fig:sol_Diff_c}
+\end{subfigure}
+
+\begin{subfigure}[b]{0.45\textwidth}
+\includegraphics[width=1\textwidth]{plots/b_Diff_Direct.pdf}\\
+\caption{Weights used for free adaptation}
+\label{fig:weights_Diff_a}
+\end{subfigure}
+\begin{subfigure}[b]{0.45\textwidth}
 \includegraphics[width=1\textwidth]{plots/b_Diff_Convex.pdf}\\
-\caption{Weights used for convex combination}
+\caption{Weights used for convex adaptation}
 \label{fig:weights_Diff_c}
@@
 
-In figure\,\ref{fig:conv_impl} the convergence is shown for the unalteres EX3 method, that potentially leads to negative vaules, the adaptive method with direct adaptation (EX3a) and the adapted method using convex combiantions (EX3c).
-Also the BE method is plotted. It is only of 1st order but preserves positivity for all $\dt$.
+In Figure\,\ref{fig:conv_impl} the convergence is shown for the unaltered EX3 method, that potentially leads to negative values, the adaptive method with free adaptation (EX3a) and the adapted method using convex adaptations (EX3c).
+Also, the BE method is plotted. It is only of 1st order but preserves positivity for all $\dt$. 
 For $\dt < \num{3e-5} $ the standard weights yield to a positive result. For larger $\dt$ the weights have to be adapted to ensure positivity. 
-The EX3a shows similar convergence properties than the original method.
+The EX3a shows similar convergence properties as the original method. 
 This can be expected, because the adapted method is still of 3rd order. 
 The EX3c has larger errors than the EX3a but leads to physical solutions for all time-steps.  
-This is no supprise because the adapted RKM used for the first step is only of 1st order.
-But the adaptive method still outperformes the BE, even when acocunting for the higher cost per step.
+This is no surprise because the adapted RKM used for the first step is only of 1st order. 
+But the adaptive method still outperforms the BE, even when accounting for the higher cost per step.
 
@@
 
-Next we test the adaptive RKM on a more compelx problem. 
-For this we consider the advection-diffusion-production-destruction systems form \cite{kopecz_comparison_2019}. The system is defined by
+Next we test the adaptive RKM on a more complex problem. 
+For this, we consider the advection-diffusion-production-destruction systems from \cite{kopecz_comparison_2019}. The system is defined by
 \begin{subequations}
@@
 \frac{\partial u_1}{\partial t} &=-a \frac{\partial u_1}{\partial x} + d\frac{\partial^2 u_1}{\partial x ^2} + 0.01u_2 + 0.01 u_3 +0.003u_4 - \frac{u_1 u_2}{0.01+u_1} \\ 
-\frac{\partial u_2}{\partial t} &=-a \frac{\partial u_2}{\partial x} + d\frac{\partial^2 u_2}{\partial x ^2}\frac{u_1u_2}{0.01+u_1} -0.01 u_2-0.5(1-\exp(-1.21 u_2^2)) u_3 -0.05 u_2 \\ 
+\frac{\partial u_2}{\partial t} &=-a \frac{\partial u_2}{\partial x} + d\frac{\partial^2 u_2}{\partial x ^2} + \frac{u_1u_2}{0.01+u_1} -0.01 u_2-0.5(1-\exp(-1.21 u_2^2)) u_3 -0.05 u_2 \\ 
 \frac{\partial u_3}{\partial t} &=-a \frac{\partial u_3}{\partial x} + d\frac{\partial^2 u_3}{\partial x ^2} + 0.5(1-\exp(-1.21u_2^2)) u_3 - 0.01 u_3 -0.02 u_3 \\ 
@@
 
-With the parameters $    a=\num{1e-2} $ and $ d=\num{1e-6}$.
-The PDE is simulated on the doamin $x = [0,1]$ with $N=100$ points. Periodic boundarie conditions are used.
-The advecion part is semidiscreticed using a first order upwind sheme and the diffusion part is semidiscreticed using a 3-poin-scheme. This leads to a positivity preserving system of ODEs. 
-The computation is done using the EX3 method with direct adaptation.
-As stepsize control a PI-Control from\,\cite{hairer_solving_1996} is used. The Error was estimatet using \eqref{eq:Err}. The Tolerance was set to $Tol = 0.01$. 
+With the parameters $a=\num{1e-2} $ and $ d=\num{1e-6}$. 
+The PDE is simulated on the domain $x = [0,1]$ with $N=100$ points. Periodic boundary conditions are used. 
+The advection part is semidiscreticed using a first order upwind scheme and the diffusion part is semidiscreticed using a 3-point-scheme. This leads to a positivity preserving system of ODEs.  
+The computation is done using the EX3 method with free adaptation. 
+As step size control a PI-Control from\,\cite{hairer_solving_1996} is used. The Error was estimated using \eqref{eq:Err}. The Tolerance was set to $Tol = 0.01$.  
 The final time is $t_{end} = 50$. 
-The simulation required 264 steps. Of these 72 required a adaptation of the weights.
+The simulation required 264 steps. Of these, 72 required an adaptation of the weights. 
 The solutions for $t=9,t=18,t=27$ and $t=50$ are plotted in Figure\,\ref{fig:Sol_ADP}. 
 In Figure\,\ref{fig:sol_ADP18} it can be seen that the reaction occurs in a small interfaces. 
-Outside of this regions quantities are close to $0$. Therefore it is very likely that negative values orrur in the numerical approxiamtion.
+Outside of this regions quantities are close to $0$. Therefore, it is very likely that negative values occur in the numerical approximation. 
 In Figure\,\ref{fig:sol_ADP27} it can be seen that at $T=27$ the two reaction interfaces merged.
-Afterwards the reaction stops and the behavoir is mainly controlled by the advection and diffusion part.
-
-In Figure\,\ref{fig:Stats_ADP} different vaules are plottet. 
-In the first subplot the stepsize is plotted. For $t<25$ the timesteps are small. After $t = 30$ the the stepsize increases, because the solution only evovlves slowly afterwards.
-In the second subplot the $min(u_1,u_2,u_3,u_4)$ is plotted for all timesteps that initaily lead to negative values. This value is computet before and after adapting the weights. 
-We can see that relativley large negative values occured at some timesteps.
-After the adaption of the weights all values are cloes to $0$. Therfor the adaption of weights succesfully preserved positifity.
-In the third subplot the approxiamted truncation error $err_T$ and the adaption errro $err_{adapt}$ are plotted.
-We can see that the adaption error is of a similar magnitude as the truncation error. Therefore the totoal error of the method is not increased drastically. 
-In the last subplots the weights are ploted. 
-We can see that the cahnges to the weights are only very small. The adapted RKM is still very close to the original RKM.
+Afterwards the reaction stops and the behavior is mainly controlled by the advection and diffusion part.
+
+In Figure\,\ref{fig:Stats_ADP} different values are plotted.  
+In the first subplot the step size is plotted. For $t<25$ the timesteps are small. After $t = 30$ the step size increases, because the solution only evolves slowly afterwards. 
+In the second subplot the $min(u_1,u_2,u_3,u_4)$ is plotted for all timesteps that initially lead to negative values. This value is computed before and after adapting the weights.  
+We can see that relatively large negative values occurred at some timesteps. 
+After the adaption of the weights, all values are close to $0$. Therefore, the adaption of weights successfully preserved positivity. 
+In the third subplot the approximated truncation error $err_T$ and the adaption errro $err_{adapt}$ are plotted. 
+We can see that the adaption error is of a similar magnitude as the truncation error. Therefore, the total error of the method is not increased drastically.  
+In the last subplots the weights are plotted.  
+We can see that the changes to the weights are only very small. The adapted RKM is still very close to the original RKM.
 
@@
 \section{Conclusion} \label{sec:conclusion}
-It is possible to adapt the weights to enforce positivity for RKMs that are not positifity preserving.
-A main limitation is that the resulting Order has to be lower than the number of stages. 
+It is possible to adapt the weights to enforce positivity for RKMs that are not positivity preserving. 
+One main limitation is that the resulting Order has to be lower than the number of stages.  
 An error approximation for this method was given.  
 The region of absolute stability is altered by changing the weights. This effect can be predicted or controlled. 
-Used with explicit methods the positivity for some test problmes could be recovered. 
-Because the time steps is limited by the stability it is only useful for a small interval of time steps.
+Used with explicit methods the positivity for some test problems could be recovered.  
+Because the time step size is limited by the stability it is only useful for a small interval of time steps. 
 The adaptive method mainly interesting for diagonally implicit methods.  
-The times teps here are not limited by stability.
-Also the cost of solving the LP is not crucial factor. 
-If the negative values occurring are not to large, which can be expected for most computations, adapting the weights is a potential way to fix these.
+The times step size is not limited by stability. 
+Also, the cost of solving the LP is not a crucial factor.  
+If the negative values occurring are not too large, which can be expected for most computations, adapting the weights is a potential way to ensure positivity.
 

### 2020-03-23 latex/overview.tex
Revise beginning of Section 2.

@@
 \newcommand{\lot}{\ell_{\mathrm{ot}}}
+\newcommand{\ut}{\tilde{u}}
+\newcommand{\bt}{\tilde{b}}
 
@@
 is positive if 
-\begin{align}
+\begin{align} \label{continuous-positivity}
     u(0)\ge 0 \implies u(t) \ge 0 \text{ for all } t>0.
@@
 
-The paper unfolds as follows. In Section\,\ref{sec:main_idea} the main idea is explained. How the method can be written as LP is explained in Section\,\ref{sec:LP}.
+The paper unfolds as follows. In Section\,\ref{sec:main_idea} the main idea is explained. How the method can be written as a LP is explained in Section\,\ref{sec:LP}.
 Section\,\ref{sec:integration} describes how the new approach can be used with different RKM, how it can be integrated in a step size control and how the Region of absolute stability can be approximated.
@@
 
-\section{Main Idea}\label{sec:main_idea}
+\section{Bound-preserving adaptive Runge--Kutta methods}\label{sec:main_idea}
 
@@
 \begin{equation}
-y_j =  u^n + \dt \sum_{k = 1}^{s} a_{jk} f(t^n + \dt c_k,y_k)  \quad j = 1,\cdots,s
+y_j =  u^n + \dt \sum_{k = 1}^{s} a_{jk} f(t^n + \dt c_k,y_k)  \quad j = 1,\cdots,s.
 \end{equation}
 Based on these values the next solution $u^{n+1}$ is computed by
-\begin{equation}
+\begin{equation} \label{eq:rkstep}
 u^{n+1} = u^n + \dt \sum_{j  = 1}^s f(t^n + \dt c_j,y_j) b_j .
 \end{equation}
-The stage derivatives $f_1,\cdots,f_s$ can be calculated with
-\begin{equation}
-f_j = f(t^n + \dt c_j,y_j)
-\end{equation}
-The new solution can now be calculated with
+Let $f_j = f(t^n + \dt c_j,y_j)$; then we can write \eqref{eq:rkstep} as
 \begin{equation}\label{eq:Combination}
@@
 \end{equation}
-where $F$ is a matrix containing the stage derivatives $F = \big[f(y_1) , \cdots f(y_s)\big]$.
-
-The new solution $u^{n+1}$ does not comply with positivity for some cases. 
-So it might be advantageous to adapt \eqref{eq:Combination} in such a way that $u^{n+1}$ is positive, instead of rejecting the solution and computing a new set of stage values.
-This is done by adapting the weights.
-The original weights $b$ are replaced by the adapted weights $\tilde{b}$.
-With this adaptation we get
-\begin{equation}\label{eq:def_b_tilde}
-u^{n+1} = u^n + \dt F \tilde{b}
-\end{equation}
-form \ref{eq:Combination}. 
-This equation is contains $\tilde{b}$ as second unknown alongside $u^{n+1}$. 
-First we have to determine the adapted weights to compute the new solution afterwards.
-For the computation of the stage derivatives only the $A$ has to be known. The weights are not needed for this computation. 
-Based on this observation we can define $(A,\bullet)$ as the set of RKMs that share the same $A$ and only differ in the weights.
-Calculating the stage derivatives can be considered as solving the ODE for all RKM in $(A,\bullet)$ at the same time.
-The idea is to choose one $(A,\tilde{b}) \in (A,\bullet)$ by fixing the weights. 
-Because the stage derivatives are already known, we can choose a $(A,\tilde{b})$, that leads to a positive $u^{n+1}$, if $(A,\tilde{b}) \in (A,\bullet) | u^{n+1} \geq 0$ exists.
-When altering the weights, the properties of a RKM are preserved. 
-Especially this approach preserves linear invariants like mass conservation, because it is a regular RKM.
-
-As it can be seen in \eqref{eq:Combination} $u^{n+1}$ is a linear combination of the stage values and the previous solution $u^n$.
-Because the coefficients in $A$ are already fixed, the Order Conditions reduce to linear equations.
-This means that finding the weights can be done by solving a LP.
-
-If there are no negative values in $u^{n+1}$ with the original weights, the method does not cause additional computational effort, because it only has to solve the LP if the regular RKM does not lead to a positive $u^{n+1}$.
-%(except for the test)
+where the $j$th column of $F$ is $f_j$.
+
+We wish to impose the discrete analog of \eqref{continuous-positivity}; i.e.
+\begin{align} \label{positivity}
+    u^n\ge 0 \implies u{n+1} \ge 0,
+\end{align}
+or more general bound constraints
+\begin{align}
+    \alpha \le u^n\le \beta \implies \alpha \le u{n+1} \le \beta.
+\end{align}
+We will focus on the case of positivity while keeping in mind that
+the methodology extends to general bounds.
+The main idea of the present work is that if the new solution $u^{n+1}$ contains
+negative entries, we can replace the weights in \eqref{eq:Combination} with
+a set of modified weights $\bt$ such that the resulting solution is positive:
+\begin{equation}\label{eq:ut}
+\ut^{n+1} = u^n + \dt F \bt^n \ge 0.
+\end{equation}
+Indeed, we can view \eqref{eq:ut} as a linear constraint on the choice of
+modified weights $\bt^n$.  Since we have already computed the intermediate stages,
+$F$ is a known, fixed matrix.  In order to ensure that the modified solution $\ut^{n+1}$
+is accurate, we can also constrain $\bt$ to satisfy the Runge--Kutta order conditions
+up to some order (ideally, the same order as the original method).  Observe that
+all of the order conditions are linear in the weights, so that these additional constraints
+take the form
+$$
+Qb = r
+$$
+for some fixed matrix $Q$ and vector $r$.
+By applying this technique at each step, we integrate \eqref{eq:ode} with a
+sequence of Runge--Kutta methods with coefficients $(A,\bt^n)$.  At any step
+for which the solution $u^n$ (produced by method $(A,b)$ is positive, we do not
+need to modify the weights and can simply accept this unmodified solution.
+Note that linear invariants (such as mass conservation) of the solution are
+automatically preserved in this approach, since at each step we use a Runge--Kutta
+method.
+
+We see that the choice of $\bt$ is subject to linear equality and inequality
+constraints.  If we choose a linear objective function, the resulting problem
+for finding the modified weights is a linear program.  Two important questions
+arise:
+\begin{itemize}
+    \item How should the objective function be chosen?
+    \item Does the resulting LP always have a solution?
+\end{itemize}
 

### 2020-03-24 latex/overview.tex
Revisions to section 2.1.

@@
 for some fixed matrix $Q$ and vector $r$.
-By applying this technique at each step, we integrate \eqref{eq:ode} with a
+By applying this technique at each step, we integrate \eqref{ode} with a
 sequence of Runge--Kutta methods with coefficients $(A,\bt^n)$.  At any step
@@
 constraints.  If we choose a linear objective function, the resulting problem
-for finding the modified weights is a linear program.  Two important questions
-arise:
-\begin{itemize}
-    \item How should the objective function be chosen?
-    \item Does the resulting LP always have a solution?
-\end{itemize}
+for finding the modified weights is a linear program, which can be efficiently
+solved by standard algorithms.  A natural choice of objective function is
+$$
+\text{minimize } \|\bt - b\|_1.
+$$
+The resulting problem can be phrased as an LP by using slack variables.
+In general, this LP may not have a solution; we can relax the constraints
+by requiring a lower order of consistency than the design order of the
+method.  These choices and alternatives will be considered in Section \ref{sec:LP}.
 
@@
 When solved with the Cash-Karp (CK5) method and $\dt = 0.005$ the approximated solution contains negative values. This causes qualitatively wrong solutions to the problem. 
-In Figure\,\ref{fig:exampleI} the obtained results are plotted with a dashed line. 
+In Figure\,\ref{fig:exampleI} the obtained results are plotted with dashed lines. 
 At $t=1.905$ the value of $u_1$ gets negative. This leads to a diverging solution.
 
-Now the weights are adapted. The results are also plotted in Figure\,\ref{fig:exampleI}. 
+Now the weights are adapted. The results are also plotted in Figure\,\ref{fig:exampleI}, with solid lines. 
 The positivity constraint is now fulfilled. A qualitatively correct solution is obtained.
+{\bf Need to mention here what the order is when the weights are adapted.}
 

### 2020-03-24 latex/overview.tex
Revise section 3.2.

@@
 \newcommand{\bt}{\tilde{b}}
+\newcommand{\pt}{{\tilde{p}}}
 
@@
 
-\section{Adaptive RKM}\label{sec:LP}
-
-In the following Section,  how the approach can be written as a LP. 
-For this we have to consider which weights to choose. 
-Generally, any $b \in \R^s$ can be used. 
-One requirement would be that the new solution is positive.
-We can further reduce this space by enforcing the order conditions. 
-By incorporating the order conditions, the space of possible weights is limited to a subset of $\R^s$. %(Note: not linear because 0 is not in subspace)
-Finally, we have to state an objective function to get the optimal weights.
+\section{Selection of modified weights}\label{sec:LP}
+
+In this Section, we consider the formulation of the LP to choose
+the modified weights $\bt$.  In particular, we focus on the choice
+of objective function and how to relax the constraints to ensure that
+a feasible solution exists.
 
@@
 
-The order conditions for a RKM are a set of Equations depending on $A$, $c$ and $b$. In these Equations $b$ only appears linearly \cite{hairer_solving_1993}.
-If $A$ and $c$ are already known, the order conditions for the order $p$ are a linear equations system and can be written as $Q b = r$.
-
-The vector $b \in \R^s$ contains the weights and $Q$ is an $g \times s$ matrix, where $g$ is the number of order conditions. The vector $r$ contains the right hand side of the order conditions.
-
-If the $A$ of a preexisting RK-method is used, there is at least one solution for the linear equation system.
-In order to get an optimization problem, the equation system has to be underdetermined.
-This implies that $\mathrm{rank}(Q) < s$. 
-Because the quadrature conditions are linearly independent, there are at least $p$ independent vectors in $Q$. This leads to the condition $p < s$.
-This implies that the number of stages has to be higher than the wanted order.
+The order conditions for an $s$-stage, order $p$ RKM are a set of equations depending on $A$, $b$ and
+$c$.  As mentioned already, if $A$ and $c$ are given, the order conditions are
+linear in $b$ and can be written as $Q_p b = r_p$, where
+and $Q_p\in{\mathbb R}^{s\times s}, r_p\in{\mathbb R}^s$ represent the set of all conditions up to
+and including order $p$.  It may not be possible to find modified weights
+that also satisfy the conditions of order $p$ and yield positivity, so
+in general the modified weights will be a solution of
+$$
+  Q_\pt \bt = r_\pt
+$$
+for some $\pt \le p$.   Since we have $s$ degrees of freedom $\bt_j$,  we need
+at a minimum to choose $\pt$ so that $\mathrm{rank}(Q_\pt) < s$. 
+Because the quadrature conditions are linearly independent, we have $\mathrm{rank}(Q_p)\ge p$,
+so we must take $\pt \le s$.  In general we may need to take $\pt$ even smaller
+in order to achieve positivity.
 
@@
 
-\subsubsection{Adaption of Timestep using the dense output formual}
+\subsubsection{Adaption of Timestep using the dense output formula}
 

### 2020-03-24 latex/overview.tex
Revising section 3 (not finished).

@@
 
-\subsection{Formulation of the Optimization Problem}
-
-There are two possible ways to write down the adaptation of the weights. The first way is to freely adapt the original weights $b$ by adding a correction $\Delta b$ that ensures positivity.
-This is the easiest way to adapt the weights. An advantageous property is that the step size can be reduced after calculating the stage values by using the dense output order conditions.  
-The free adaption is explained in more detail in Section\,\ref{sec:direct}.
-
-The second way to formulate a LP is to choose weights as a convex adaptation of a set of weights. This is more useful if some embedded methods with certain properties are known. 
-The convex adaptation also shows more predictable behavior because the resulting $u^{n+1}_b$ is also a convex adaptation of the solutions $u^{n+1}_{b_a},u^{n+1}_{b_b},\cdots$ 
-This is explained in more detail in Section\,\ref{sec:convex}
+\subsection{Choice of objective function and additional constraints}
+In the design of Runge--Kutta methods, weights are carefully chosen
+not only to satisfy the order conditions but also
+to give desirable properties such as a good region of absolute
+stability, small error coefficients, and so forth.
+Replacing these carefully-chosen weights $b$ with arbitary weights
+$\bt$ could lead to the loss of these desirable properties.
+In order to preserve good properties of the method, we use as
+objective function $\|\bt - b\|_1$.  This has the additional benefit
+of penalizing weights with large magnitude in general, avoiding
+large truncation or cancellation errors.  This also ensures that if
+no negative solution values appear, the solution of the LP is
+simply the original method weights.  Thus we have the following LP:
+
+\LPblock{
+(Free adaptation) Given $F, p$, and $b$, find $\bt$ that minimizes $||\tilde{b} - b||_1$ subject to
+\begin{subequations}
+\begin{align}
+u^{n+1}&=u^n+\dt F \tilde{b} \geq 0 \label{eq:direct_pos}\\
+Q_\pt\tilde{b} &=r_\pt \label{eq:direct_Order}.
+\end{align}
+\end{subequations}
+}
+
+Of course, there is still no guarantee that the modified weights
+will be close to the original method weights.  In some examples
+we have observed that large modifications of the weights can lead
+to inaccurate solutions even though the order conditions are satisfied.
+In order to avoid issues that might be caused by poor weights, we can
+additionally impose either or both of the following strategies:
+
+\begin{enumerate}
+    \item Select in advance a set of desirable weight vectors $b^1, b^2, \dots, b^m$
+        corresponding to known good methods,
+        and restrict the choice of $\bt$ to convex combinations of this set.
+    \item Require that $\|bt - b\|_1$ be small, rejecting the step if it is not.
+\end{enumerate}
+
+We discuss the first strategy here; the second is deferred to a later section.
+Ideally every element of the set of potential weight vectors would correspond to
+a method of the same order as the original method.   Due to linearity of the
+order conditions, any linear combination of such weights would also yield a 
+method of the same order.  On the other hand, 
+it is natural to include a weight vector corresponding to the forward Euler method
+(for explicit methods) or backward Euler method (for implicit methods), since
+these two methods guarantee positivity (unconditionally for backward Euler
+and conditionally for forward Euler).
+We can formulate an LP using the first strategy (restriction to a convex set)
+as follows.  Let $B$ denote the matrix with columns $b^1, b^2, \dots, b^m$ and let
+$g\in{\mathbb R}^m$.  The LP is then
+
+\LPblock{
+(Convex adaptation) Given $F, p$, and $b$, find $g$ that minimizes $||\tilde{b} - b||_1$ subject to
+\begin{subequations}
+\begin{align}
+\bt & = Bg \\
+0 & \le g_i \le 1 \\
+u^{n+1}&=u^n+\dt F \tilde{b} \geq 0 \label{eq:direct_pos}\\
+Q_\pt\tilde{b} &=r_\pt \label{eq:direct_Order}.
+\end{align}
+\end{subequations}
+}
 
@@
 
-\subsection{Free Adaptation}\label{sec:direct}
-
-When freely adapting the weights, new weights that complie with the order conditions is chosen.   
-The weights $\tilde{b} = b + \Delta b$ consist of the weights of the original method $b$ and an adaption $\Delta b$. The adaption is $\Delta b \in \mathrm{ker}(Q)$. 
-This is drawn in Figure\,\ref{fig:b_direct}. The red line represents the subspace of weights that satisfy the order conditions. Both $b$ and $\tilde{b}$ are in this subspace. 
-The LP can be expressed as
-
-\LPblock{
-\begin{subequations}
-\begin{align}
-u^{n+1}&=u^n+\dt F \tilde{b} \geq 0 \label{eq:direct_pos}\\
-Q\tilde{b} &=r \label{eq:direct_Order} \\
-\min_{\tilde b}&( ||\tilde{b} -b ||_1) \label{eq:direct_obj} 
-\end{align}
-\end{subequations}
-}
-
-\subsubsection{Constraints}
-The free adaptation approach leads to two sets of constraints on $b$.
-The first set of constraints is the positivity \eqref{eq:direct_pos}.
-The second set of constraints are the order conditions in \eqref{eq:direct_Order}.
-
-\subsubsection{Objective Function}
-A reasonable property of the LP is that the optimal solution for the problem without positivity constraints is $\tilde{b} = b$. 
-This can be achieved by $argmin(f_{optim}(\tilde{b})) = b$. 
-In practice, this also means that there is no need to solve the optimization problem if no constraints are violated because the solution of this problem is already known.
-This is advantageous for performance reasons.
-Large values of b can lead to numerical problems.
-Because $\sum_{i  = 1}^s b_i = 1$, an increase of $||b||$ leads to a combination of negative and positive values.
-Because the computation is done with a finite precision (and usually in floating point) large $||b||$ can lead to numerical difficulties.
-By changing the weights, the stability function is altered. 
-For these reasons $||\tilde{b}-b||$ seems like a good choice. The 1-Norm can be implemented in a  LP using slack variables. We choose \eqref{eq:direct_obj}
-as objective function.
-
-\subsubsection{Adaption of Timestep using the dense output formula}
+\subsection{Adaption of Timestep using the dense output formula}
 

### 2020-03-26 latex/overview.tex
Order of adapted weights, Different versions of Plot
overview.tex:
Added a short text regarding the order of the adapted steps.

Plots:
Changed plots to display the objective function and not the used weights.
More revision of Section 3.

@@
 
-Now the weights are adapted. The results are also plotted in Figure\,\ref{fig:exampleI}, with solid lines. 
+Now the weights are adapted. The adapted weights are of 4th order. The results are also plotted in Figure\,\ref{fig:exampleI}, with solid lines. 
 The positivity constraint is now fulfilled. A qualitatively correct solution is obtained.
-{\bf Need to mention here what the order is when the weights are adapted.}
 
@@
         and restrict the choice of $\bt$ to convex combinations of this set.
-    \item Require that $\|bt - b\|_1$ be small, rejecting the step if it is not.
+    \item Require that $\|\bt - b\|_1$ be small, rejecting the step if it is not.
 \end{enumerate}
@@
 as follows.  Let $B$ denote the matrix with columns $b^1, b^2, \dots, b^m$ and let
-$g\in{\mathbb R}^m$.  The LP is then
+$g\in{\mathbb R}^m$.  The LP is then as follows:
 
@@
 0 & \le g_i \le 1 \\
-u^{n+1}&=u^n+\dt F \tilde{b} \geq 0 \label{eq:direct_pos}\\
-Q_\pt\tilde{b} &=r_\pt \label{eq:direct_Order}.
+u^{n+1}&=u^n+\dt F \tilde{b} \geq 0 \label{eq:direct_pos}.
 \end{align}
@@
 }
+
+Note that we do not need to impose the order conditions here, since they will be
+satisfied by each of the methods and thus (by linearity) by the modified
+method.  The order of the modified method will in general be equal to the lowest
+order among the component methods.
 
@@
     \end{subfigure}
-    \caption{Graphical representation of the two different approaches to adapt the weights}\label{fig:b_space}
+    \caption{Graphical representation of the two different approaches to adapt the weights, for
+    a two-stage method.}\label{fig:b_space}
 \end{figure}
@@
 \subsection{Adaption of Timestep using the dense output formula}
+
+{\bf Need to also discuss relaxing the order conditions}
 
@@
 %(Note: we could include some algorithm to search for it)
-
-\subsection{Convex Adaptation of Existing Methods}\label{sec:convex}
-For some test problems (e.g.
-the Heat Equation solved with an implicit method of higher order), large changes occur, even though a norm of $\Delta b$ is used as an objective function.
-A numerical example is given in Section\,\ref{sec:Numeric_Results}.
-It is known that the backward Euler method leads to a positive solution for any $\dt$ \cite{hundsdorfer_numerical_2003}.
-The backward Euler is still only first Order but shows more exact results than a method of higher Order that is distorted by the adaption of the weights, at least at some steps.
-A possible method to fix this is to generate the new b based on a convex adaptation of embedded methods. 
-The new weights are defined as 
-\begin{equation}\label{eq:b_convex_def}
-\tilde{b} = \sum_i^d g_i b_i = Bg
-\end{equation}
-where $B=\left[b_1,\cdots,b_d \right]$ is a matrix containing the embedded methods.  
-The LP is defined as
-\LPblock{
-\begin{subequations}
-\begin{align}
- u^{n+1}&=u^n+\dt K B g \geq 0 \label{eq:Convex_pos}  \\
- 0 \leq g_i &\leq 1  \forall {i \in \{1, \cdots d \}} \quad \sum_{i=1}^d g_i = 1 \label{eq:Convex_cvx} \\
- \min_{g} &\left(\sum_i^d w_i g_i \right) \label{eq:Convex_obj}
-\end{align}
-\end{subequations}
-}
-
-\subsubsection{Constraints}
-At first a set of embedded methods of the RKM $b_1,\cdots,b_d$ is needed. 
-If all the used weights complie with the order conditions, the new method is also of the same order.
-This is straightforward using
-$$ Q b = Q (\chi b_{\alpha} + (1-\chi) b_{\beta}) = \chi Q  b_{\alpha} + (1-\chi) Q b_{\beta} = \chi r + (1-\chi) r = r$$
-$b_{\alpha}$ and $b_{\beta}$ are the methods that are combined to $b$ using the parameter $\chi$. The Matrix $Q$ and the vector $r$ are the order conditions as defined in Section\,\ref{sec:OrderCond}.
-To ensure that \eqref{eq:b_convex_def} yields to a convex adaptation the constraints \eqref{eq:Convex_cvx} are enforced for the parameters $g_1,\cdots,g_d$.
-The positivity constraint changes to \eqref{eq:Convex_pos}
-
-\subsubsection{Objective Function}
-
-As an objective function \eqref{eq:Convex_obj}
-is used.  The $w_1,\cdots,w_d$ are weighting factors that determine which is the preferred $b_i$ to use.
-
-Because the optimal solution of the unconstrained problem has to be the original weights the weight corresponding to $b$ should be the smallest $w$.
-This weight is simply set to $w_{b} = 0$.
-For the other weights $w_i = \frac{1}{\mathrm{Order} \{b_i\}}$ is used.
-This prefers embeded methods of higher order.
 
@@
 The simulation required 264 steps. Of these, 72 required an adaptation of the weights. 
+All adapted weights are still of 3rd order.
 The solutions for $t=9,t=18,t=27$ and $t=50$ are plotted in Figure\,\ref{fig:Sol_ADP}. 

### 2020-03-30 latex/overview.tex
switch to a bibliographystyle that can handle arXiv and first adaptations

@@
 \usepackage[a4paper, top=3.5cm, bottom=3.5cm, left=3.5cm, right=3.5cm]{geometry}
-
-\usepackage[plainpages=false,pdfpagelabels,hidelinks,unicode]{hyperref}
 
@@
 \numberwithin{theorem}{section}
+
+\usepackage[%
+  backend=bibtex,bibencoding=ascii,
+%   backend=biber,
+%   style=authoryear-comp, dashed=false,
+  style=numeric-comp,
+%   firstinits=true, uniquename=init, %abbreviate first names
+  giveninits=true, uniquename=init, %abbreviate first names
+  natbib=true,
+  url=true,
+  doi=true,
+  isbn=false,
+  backref=false,
+  maxnames=99,
+  ]{biblatex}
+\addbibresource{Bib_RK_optim.bib}
+
+% load hyperref after amsmath to get rid of stupid ``destination with the same identifier...'' warnings
+\usepackage[plainpages=false,pdfpagelabels,hidelinks,unicode]{hyperref}
+
+% suppress "multiple pdfs with page group included in a single page"
+% http://tex.stackexchange.com/questions/198586/conditional-based-on-the-version-of-pdflatex
+\begingroup\expandafter\expandafter\expandafter\endgroup
+\expandafter\ifx\csname pdfsuppresswarningpagegroup\endcsname\relax
+\else
+  \pdfsuppresswarningpagegroup=1\relax
+\fi
 
@@
 \newcommand{\pt}{{\tilde{p}}}
+\newcommand{\todo}[1]{{\Large{\color{red}{#1}}}}
 
@@
 
-\title{Positivity-Preserving Adaptive Runge-Kutta Methods}
-\author{Stephan Nüßlein \and David I. Ketcheson}
+\title{Positivity-Preserving Adaptive Runge--Kutta Methods}
+\author{Stephan Nüßlein \and Hendrik Ranocha \and David I. Ketcheson}
 
@@
 \end{align}
-Any RKM (or in fact any genearal linear method) that is unconditionally positivity preserving for all positive ODEs
+Any RKM (or in fact any general linear method) that is unconditionally positivity preserving for all positive ODEs
 must have order $\le 1$ \cite{bolley_conservation_1978}. Notably, the (first-order)
@@
 For any higher order method, we expect positivity only under some restriction
-on the step size.
+on the time step size.
 
@@
 If positivity is preserved under a forward Euler step (with
-some step size restriction$\dt \le \dt_\text{FE}$), then any strong stability preserving Runge--Kutta (SSPRK)
+some step size restriction $\dt \le \dt_\text{FE}$), then any strong stability preserving Runge--Kutta (SSPRK)
 method will also preserve positivity (with a modified step size restriction)~\cite{gottlieb_strong_2011}.
-Specifically, the positivity of the method is ensured for time steps $\dt$ with
+Specifically, the positivity of the method is ensured for time steps
 $\dt \leq {\mathcal C} \dt_\text{FE}$, where ${\mathcal C}$ depends on the SSPRK method.
-The modified Patankar Runge Kutta (MPRK) methods represent another approach to ensuring
+The modified Patankar--Runge--Kutta (MPRK) methods represent another approach to ensuring
 positivity, for specific classes of ODEs.  MPRK methods introduce multiplicative
 factors within the Runge--Kutta stages to ensure positivity, but require the solution
-of a linear algebraic system; see e.g. \cite{kopecz_comparison_2019} and references therein.
+of a linear algebraic system; see e.g.\ \cite{kopecz_comparison_2019} and references therein.
 Yet another approach to ensure positivity in a practical application is shown in \cite{shampine_non-negative_2005},
 where the ODE is redefined outside the physical domain. 
-Finally, we mention diagonally split Runge-Kutta (DSRK) methods, which can be unconditionally
+Finally, we mention diagonally split Runge--Kutta (DSRK) methods, which can be unconditionally
 positive and have order higher than one (they avoid the restriction mentioned above because
@@
 
-The idea of using different weights within a RKM is not new; for instance it is
+The idea of using different weights within an RKM is not new; for instance it is
 the basis of error approximation using embedded RK pairs \cite{hairer_solving_1993}.
@@
 With this approach, it is possible to have different properties of the RKM at different parts of the domain.
-Another class of methods that adapt the weights at the end of a RK step
+Another class of methods that adapt the weights at the end of an RK step
 are the relaxation Runge–Kutta (RRK) methods. 
@@
 dissipate energy or entropy
-\cite{ketcheson_relaxation_2019,ranocha_relaxation_2019}.
-
-The paper unfolds as follows. In Section\,\ref{sec:main_idea} the main idea is explained. How the method can be written as a LP is explained in Section\,\ref{sec:LP}.
-Section\,\ref{sec:integration} describes how the new approach can be used with different RKM, how it can be integrated in a step size control and how the Region of absolute stability can be approximated.
-In Section\,\ref{sec:imple} further details that were used when implementing the algorithm are described.
-In Section\,\ref{sec:Numeric_Results} Numerical results are given for multiple test problems.
-A conclusion is given in Section\,\ref{sec:conclusion}.
+\cite{ketcheson_relaxation_2019,ranocha_relaxation_2019,ranocha2020general}.
+
+Our means to ensure positivity can be interpreted as a projection
+approach, where the numerical solution is adapted to satisfy the
+positivity constraint at the end of each step. In contrast to the
+usual orthogonal projection, which has also been proposed to deal
+with positivity constraints \cite{shampine1986conservation},
+our approach preserves all linear invariants of the given ODE. These
+invariants can be very important. e.g.\ the total mass for a transport
+problem or in reaction systems. The property to preserve all linear
+invariants is shared by RRK methods and has been shown to be an
+important advantage compared to orthogonal projection methods
+\cite{ranocha2020relaxationHamiltonian}.
+
+The paper unfolds as follows. In Section~\ref{sec:main_idea} the main idea is explained. How the method can be written as a LP is explained in Section~\ref{sec:LP}.
+Section~\ref{sec:integration} describes how the new approach can be used with different RKM, how it can be integrated in a step size control and how the region of absolute stability can be approximated.
+In Section~\ref{sec:imple} further details that were used when implementing the algorithm are described.
+In Section~\ref{sec:Numeric_Results} numerical results are given for multiple test problems.
+A conclusion is given in Section~\ref{sec:conclusion}.
 
@@
 
-When computing the solution of an ODE $u ' = f(t,u) $ using a RKM with $s$ stages and the Butcher tableau
+When computing the solution of an ODE $u ' = f(t,u) $ using an RKM with $s$ stages and the Butcher tableau
 \begin{align}
+\renewcommand{\arraystretch}{1.2}
 \begin{array}{c|c}
@@
 \hline
- & b\\
+ & b^T\\
 \end{array}
@@
 \begin{equation}
-y_j =  u^n + \dt \sum_{k = 1}^{s} a_{jk} f(t^n + \dt c_k,y_k)  \quad j = 1,\cdots,s.
+y_j =  u^n + \dt \sum_{k = 1}^{s} a_{jk} f(t^n + \dt c_k,y_k),  \quad j = 1,\cdots,s.
 \end{equation}
@@
 \begin{align} \label{positivity}
-    u^n\ge 0 \implies u{n+1} \ge 0,
+    u^n\ge 0 \implies u^{n+1} \ge 0,
 \end{align}
@@
 \begin{align}
-    \alpha \le u^n\le \beta \implies \alpha \le u{n+1} \le \beta.
+    \alpha \le u^n\le \beta \implies \alpha \le u^{n+1} \le \beta.
 \end{align}
@@
 sequence of Runge--Kutta methods with coefficients $(A,\bt^n)$.  At any step
-for which the solution $u^n$ (produced by method $(A,b)$ is positive, we do not
+for which the solution $u^{n+1}$ produced by method $(A,b)$ is positive, we do not
 need to modify the weights and can simply accept this unmodified solution.
@@
 by requiring a lower order of consistency than the design order of the
-method.  These choices and alternatives will be considered in Section \ref{sec:LP}.
+method.  These choices and alternatives will be considered in Section~\ref{sec:LP}.
 
@@
 \begin{equation}
-u(0) = (8,2,1,4)^T
+u(0) = (8,2,1,4)^T.
 \end{equation}
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
     \centering
     \includegraphics[width=0.75\textwidth]{plots/exampleI.pdf}
-    \caption{Numerical approximation of the Reaction Problem from \cite{kopecz_comparison_2019}.
-Computed with ck5 and $\dt = 0.005$.
-The dashed lines are the approximation obtained without the adaption of the weights.
-The lower plot are the weights.
+    \caption{Numerical approximation of the reaction problem from \cite{kopecz_comparison_2019} computed with CK5 and $\dt = 0.005$.
+The dashed lines are the approximations obtained without the adaption of the weights.
+The lower plot shows the adaptation of the weights.
 }
@@
 
-In this Section, we consider the formulation of the LP to choose
+In this section, we consider the formulation of the LP to choose
 the modified weights $\bt$.  In particular, we focus on the choice
@@
 
-The main goal is to choose a $(A,b)$ in such a way that $u^{n+1}$ resembles the solution of the ODE $u(t_{n+1})$. 
+The main goal is to choose a method $(A,b)$ in such a way that $u^{n+1}$ resembles the solution of the ODE $u(t_{n+1})$.
 An obvious objective is to use a $(A,b)$ with a high order, but this is not enough.
@@
 1 & 1 &  & \\
-\frac{1}{2} & \frac{1}{4} & \frac{1}{4} & \\
+\nicefrac{1}{2} & \nicefrac{1}{4} & \nicefrac{1}{4} & \\
 \hline
- & \frac{1}{6} & \frac{1}{6} & \frac{2}{3}\\
+ & \nicefrac{1}{6} & \nicefrac{1}{6} & \nicefrac{2}{3}\\
 \end{array}
 \end{align}
-The Matrix $L$ has the eigenvalues $\lambda_1 = -6,\lambda_2=0$.
+The matrix $L$ has the eigenvalues $\lambda_1 = -6,\lambda_2=0$.
 The real stability interval of the SSP33 method includes $[-2.5,0]$.
-The method is absolut stable for this problem with $\dt = \frac{1}{3}$
+The method is absolut stable for this problem with $\dt = \frac{1}{3}$.
 The stage derivatives are
 \begin{equation}
-f(y_1)=\left(\begin{matrix}-5\\5\end{matrix}\right)\qquad f(y_2)=\left(\begin{matrix}5\\-5\end{matrix}\right)\qquad f(y_3)=\left(\begin{matrix}-5\\5\end{matrix}\right)\end{equation}
+f(y_1)=\left(\begin{matrix}-5\\5\end{matrix}\right),\qquad f(y_2)=\left(\begin{matrix}5\\-5\end{matrix}\right),\qquad f(y_3)=\left(\begin{matrix}-5\\5\end{matrix}\right).
+\end{equation}
 The values of the next step using the standard weights is
 \begin{equation}
-u_n=\left(\begin{matrix}- \frac{1}{9}\\\frac{10}{9}\end{matrix}\right)
+u_n=\left(\begin{matrix} \nicefrac{-1}{9}\\ \nicefrac{10}{9}\end{matrix}\right)
 \end{equation}
@@
 Therefore, we want to adapt the weights to ensure positivity. 
-All weights, that comply with the order Condition for the first and second order can be expressed as
-\begin{equation}
-\tilde b=\left(\begin{matrix}\frac{1}{6}\\\frac{1}{6}\\\frac{2}{3}\end{matrix}\right)+\alpha \left(\begin{matrix}\frac{1}{2}\\\frac{1}{2}\\-1\end{matrix}\right).
-\end{equation}
-We have one degree of freedom for the choice of the weights. This is done by the parameter $\alpha$.
+All weights that comply with the order condition for the first and second order can be expressed as
+\begin{equation}
+\bt=\begin{pmatrix}\nicefrac{1}{6}\\ \nicefrac{1}{6}\\ \nicefrac{2}{3}\end{pmatrix}+\alpha \begin{pmatrix}\nicefrac{1}{2}\\\nicefrac{1}{2}\\-1\end{pmatrix}.
+\end{equation}
+We have one degree of freedom for the choice of the weights,
+parameterized by $\alpha$.
 If the general expression for the weights is inserted in \eqref{eq:Combination} the general solution is 
 \begin{equation}
-u^{1} = u^0 + \dt  \left(f(y_1),f(y_2),f(y_3)\right) \tilde{b} =
-\left(\begin{matrix}- \frac{1}{9}\\\frac{10}{9}\end{matrix}\right) +\alpha \left(\begin{matrix}\frac{5}{3}\\- \frac{5}{3}\end{matrix}\right)
+u^{1} = u^0 + \dt  \left(f_1, f_2, f_3\right) \bt =
+\begin{pmatrix} \nicefrac{-1}{9}\\\nicefrac{10}{9}\end{pmatrix}
++\alpha \begin{pmatrix}\nicefrac{5}{3}\\ \nicefrac{-5}{3}\end{pmatrix}
 \end{equation}
 By changing the parameter $\alpha$, the weights and the solution are altered.
-With a suitable choice of $\alpha \in \left[\frac{1}{15},\frac{2}{3}\right]$ any $u$ that complies with mass conservation and positivity can be reached. 
+With a suitable choice of $\alpha \in \left[\nicefrac{1}{15},\nicefrac{2}{3}\right]$ any $u$ that complies with mass conservation and positivity can be reached.
 By adding additional constraints on the weights, the choice of $\alpha$ can be narrowed down. 
@@
 
-\subsection{Order Conditions}\label{sec:OrderCond}
+\todo{Note (HR)} %TODO
+This example has to be fixed.
+Since the matrix is not normal, eigenvalue analysis
+and classical linear stability theory do not apply. By a direct
+computation, we can see that SSP33 is absolutely stable for this
+initial condition and the first step for $\dt \lesssim 0.309$
+but not for $\dt = \nicefrac{1}{3}$.
+
+\subsection{Order conditions}\label{sec:OrderCond}
 
@@
 \begin{align}
-u^{n+1}&=u^n+\dt F \tilde{b} \geq 0 \label{eq:direct_pos}\\
+u^{n+1}&=u^n+\dt F \tilde{b} \geq 0, \label{eq:direct_pos}\\
 Q_\pt\tilde{b} &=r_\pt \label{eq:direct_Order}.
@@
 \begin{align}
-\bt & = Bg \\
-0 & \le g_i \le 1 \\
-u^{n+1}&=u^n+\dt F \tilde{b} \geq 0 \label{eq:direct_pos}.
+\bt & = Bg, \\
+0 & \le g_i \le 1, \\
+u^{n+1}&=u^n+\dt F \tilde{b} \geq 0.
 \end{align}
@@
 
-Both approaches are illustrated in figure\,\ref{fig:b_space}.
+Both approaches are illustrated in Figure\,\ref{fig:b_space}.
 
@@
     \begin{subfigure}[b]{0.45\textwidth}
+        \centering
         \input{Image_b_direct}
@@
     \begin{subfigure}[b]{0.45\textwidth}
+        \centering
         \input{Image_b_convex}
@@
     \end{subfigure}
-    \caption{Graphical representation of the two different approaches to adapt the weights, for
+    \caption{Graphical representation of the two different approaches to adapt the weights for
     a two-stage method.}\label{fig:b_space}
@@
 
-\subsection{Adaption of Timestep using the dense output formula}
+\subsection{Adaption of the timestep using the dense output formula}
 
@@
 In this case we would like to calculate $u^{n+\theta}$ using the stage values instead of rejecting the step.
-The time step is reduced after calculation the stage values.
-This can be done using the order conditions for dense output.
-These can be found in \cite{hairer_solving_1993}.
-We introduce $\theta \in (0,1]$. The new step taken has the length $\theta \dt$ whereas $\dt$ is the time step used when calculating the stage values.
+The time step is reduced after computing the stage values.
+This can be done using the order conditions for dense output \cite[Section~II.6]{hairer_solving_1993}.
+We introduce $\theta \in (0,1]$. The new step taken has the length $\theta \dt$, where $\dt$ is the time step used when calculating the stage values.
 To calculate a reduced step the order conditions have to be adapted and a new $b(\theta)$ for the objective function has to be generated. 
-When adapting the Order Condition to a new $\theta$ the Order Condition Matrix $Q$ stays unchanged whereas the right hand side $r(\theta)$ depends on $\theta$.
+When adapting the order condition to a new $\theta$ the order condition matrix $Q$ stays unchanged whereas the right hand side $r(\theta)$ depends on $\theta$.
 
@@
 
-\subsubsection{Construction of Embedded Methods}
+\subsubsection{Construction of embedded methods}
 
@@
 
-\section{Properties of Adaptive RKM} \label{sec:integration}
-
-In the previous Sections an algorithm for choosing positifity preserving weights $\tilde{b}$ has been presented. 
-In the following Section, it is explained how this Algorithm can be integrated into the framework of existing solvers.
-
-\subsection{Choice of Baseline Method}
-An important property for the used baseline method is the existence of embedded methods and the degrees of freedom for the b. 
-As noted in Section\,\ref{sec:OrderCond} the number of stages has to be higher than the order.  
-Therefore, explicit RKM are interesting, because an explicit RKM of Order $p$ requires at least $p$ stages. 
-Fully implicit RKM can achieve higher orders.  
+\section{Properties of adaptive RKM} \label{sec:integration}
+
+In the previous sections an algorithm for choosing positivity preserving weights $\tilde{b}$ has been presented.
+Here, we explain how this algorithm can be integrated into the framework of existing solvers.
+
+\subsection{Choice of baseline method}
+An important property of the baseline method is the existence of embedded methods and the degrees of freedom for the weights $b$.
+As noted in Section~\ref{sec:OrderCond} the number of stages has to be higher than the order.
+Therefore, explicit RKMs are interesting, because an explicit RKM of order $p$ requires at least $p$ stages.
+Fully implicit RKMs can achieve higher orders.
 To fulfill $p<s$ the order has to be reduced drastically. Therefore, these methods are not good candidates for the adaptation. 
-As stated in \cite{norsett_attainable_1977} the highest achievable order of a Diagonally implicit RKM is $s+1$. This condition is far closer to the condition required for the adaptation. 
+As stated in \cite{norsett_attainable_1977} the highest achievable order of a diagonally implicit RKM is $s+1$. This condition is far closer to the condition required for the adaptation.
 Another important property is the number of degrees of freedom for the choice of the new weights.  
 These can be calculated using $s-\mathrm{rank}(Q)$. %$\mathrm{dim}(\mathrm{ker}(Q))$. 
-The number of degrees of freedom for different explicit methods are shown in Table\,\ref{table:DOF_exp}.
+The number of degrees of freedom for different explicit methods are shown in Tables\,\ref{table:DOF_exp} and~\ref{table:DOF_imp}.
 
@@
  \end{tabular}
- \caption{Degrees of Freedom for Choice of the weights} %Generated above============ 
+ \caption{Degrees of freedom for the choice of the weights for some explicit methods.} %Generated above============
  \label{table:DOF_exp}
@@
  \hline Implicit Euler& s=1&0& - & - & - & - & -  \\ 
- Lobatto IIIC4& s=4&3&2&1&0&0&0 \\ 
- Radau IIA3& s=3&2&1&0&0&0& -  \\ 
+ Lobatto~IIIC4& s=4&3&2&1&0&0&0 \\
+ Radau~IIA3& s=3&2&1&0&0&0& -  \\
  SDIRK 54& s=5&4&3&1&0& - & -  \\ 
@@
  \end{tabular}
- \caption{Degrees of Freedom for Choice of the weights} %Generated above============ 
+ \caption{Degrees of freedom for the choice of the weights for some implicit methods.} %Generated above============
  \label{table:DOF_imp}
@@
 For all methods with the number of stages equal to the order of the method, there are no degrees of freedom without reducing the order.  
-If the classical RK4 method is used the Order has to be reduced more because the RK4 method does not have embedded methods of order 3. 
-Contrary to this, methods with $s > p$ have degrees of freedom even without reducing the order.  
-For Cash-Karp RK5 and Dorman-Prince RK5 the Order has to be reduced to get degrees of freedom even though the number of stages is higher than the order.
+If the classical RK4 method is used the order has to be reduced more because the RK4 method does not have embedded methods of order 3.
+In contrast to this, methods with $s > p$ have degrees of freedom even without reducing the order.
+For Cash--Karp RK5 and Dormand--Prince RK5 the order has to be reduced to get degrees of freedom even though the number of stages is higher than the order.
 
 In Table\,\ref{table:DOF_imp} same was done for implicit methods. 
-We can see that the fully implicit methods Lobatto IIIC4 and Radua IIA3 require a drastic reduction of the order as expected. 
+We can see that the fully implicit methods Lobatto~IIIC4 and Radau~IIA3 require a drastic reduction of the order as expected.
 The Diagonally implicit SDIRK 54 method only requires an order reduction of one to get one degree of freedom for the weights.  
@@
 
-\subsection{Error detection and Approximation}
+\subsection{Error detection and approximation}
 When changing the weights the solution $u^{n+1}$ is changed. 
@@
 Now it is important to know if the new solution $u^{n+1}$ still resembles the solution of the ODE $u(t_{n+1})$. 
-This can be seen in example in Section\,\ref{sec:example_lin}. 
+This can be seen in example in Section~\ref{sec:example_lin}.
 This gives rise to the question whether the new solution is reasonable.
@@
 
-\subsubsection{Error Detection and Appproximation}
+\subsubsection{Error detection and appproximation}
 For this approach it is difficult to give some mathematically rigorous statement regarding the convergence of the method, because it only applies if the error of the original method causes the solution to get negative.
@@
 One could say that the main idea of the adaption of the weights alters the error in a certain way, so that the numerical solution gets 'better' in the sense that it leads to positive solutions. 
-The local error of a RKM $e(\dt) =u(t_0 + \dt) - u^1$ can be expressed using the Taylor series of the RKM and the Taylor series of the exact solution. % Hairer I P134
+The local error of an RKM $e(\dt) =u(t_0 + \dt) - u^1$ can be expressed using the Taylor series of the RKM and the Taylor series of the exact solution. % Hairer I P134
 
@@
 
-\subsection{Stability Region}
-
-By adapting the $b$ the used RK Method is changed. This also alters the stability function and therefore changes the region of absolute stability. 
+\subsection{Stability region}
+
+By adapting the $b$ the used RK method is changed. This also alters the stability function and therefore changes the region of absolute stability.
 It depends on the solved problem, the RKM and the used $\dt$ by how many steps the stability function is altered.
@@
 In Figure\,\ref{fig:stab} the stabilit regions of adapted RKM are plotted. 
-In Figure\,\ref{fig:stab_ck5} the Dorman-Prince 5 method is freely adapted. 
-The weights are taken form the example in Section\,\ref{sec:Ex_expl}.
-
-In Figure\,\ref{fig:stab_ex3} the stabilty regions of the BE-Extraplotaion 3 method and the embedded chain of three BE steps is plotted.
-Additionally the stabilty regions of convex combiantions of these two methods are plotted.
+In Figure\,\ref{fig:stab_ck5} the Dormand--Prince 5 method is freely adapted.
+The weights are taken form the example in Section~\ref{sec:Ex_expl}.
+
+In Figure\,\ref{fig:stab_ex3} the stability regions of the BE-Extraplotaion 3 method and the embedded chain of three BE steps is plotted.
+Additionally the stability regions of convex combinations of these two methods are plotted.
 
@@
          \includegraphics[width=\textwidth]{plots/stab_dp5.pdf}
-         \caption{Dorman-Prince 5\\
+         \caption{Dormand--Prince 5\\
          with $\tilde{b}$ form \ref{fig:weights_AdDe}
@@
 The stability function is defined as a function $\CN \to \CN$ with 
-\begin{equation}\label{eq:stab_func}
+\begin{equation}
 R(z) = 1 + zb^T(I - zA)^{-1}e
@@
 We can add $b$ as a second parameter and get a function $\CN \times \R^s \to \CN$ with
-\begin{equation}\label{eq:stab_func}
+\begin{equation}
 R(z,b) = 1 + zb^T(I - zA)^{-1}e
@@
 
-\subsubsection{Stability of Convex Adaptation}
+\subsubsection{Stability of convex adaptation}
 If the new weights are chosen as a convex adaptation of weights it is easy to prove properties of the stability region.
@@
 
-\subsubsection{Stability of  Free Adaptation}
+\subsubsection{Stability of free adaptation}
 If the weights is adapted freely there is no such proof.
@@
 
-\section{Implementation aspects of the Algorithm}\label{sec:imple}
+\section{Implementation aspects of the algorithm}\label{sec:imple}
 
@@
 
-\subsection{Applicable Problems}\label{sec:app_problem}
+\subsection{Applicable problems}\label{sec:app_problem}
 The approach can be used with ODE systems $u' = f(t,u)$ that ensure  $u(t) \geq 0 \forall t \forall {  u_0 \geq 0}$ 
@@
 When increasing the step size, issues with stability occur before getting negative values.  
-An example therefore the ODE in in Section\,\ref{sec:example_lin}. 
+An example therefore the ODE in in Section~\ref{sec:example_lin}.
 Some nonlinear RHS may require very small timesteps to preserve positivity.  
 For these, adapting the weights could be a possible way to solve them. 
-An example is the reaction equation solved in Section\,\ref{sec:example_reac}. 
+An example is the reaction equation solved in Section~\ref{sec:example_reac}.
 The positivity of stage values is not ensured, therefore the right hand side has to be defined for negative values. If this is not the case, the right hand side has to be defined for negative values. 
@@
 The parameters are set to $a=1$ and $K=1$.
-We use the Dorman-Prince 5 RKM and adapt the weights using the free adaptation.
+We use the Dormand--Prince 5 RKM and adapt the weights using the free adaptation.
 In Figure\,\ref{fig:sol_AdDe} the results for $\Delta t = 0.02$ are plotted for different timesteps. We can see that the solution approaches an exponential function with $t \rightarrow \infty$.
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
 \centering
@@
 
-\bibliographystyle{apalike}
-\bibliography{Bib_RK_optim}
+\printbibliography
 

### 2020-03-30 latex/overview.tex
some additional revisions while reading till section 6

@@
 accuracy of the method.  The choice of weights is given by the solution
-of a linear program (LP).
+of a linear program.
 %(iv) how you verified your idea and  important results and possible/important limitations
@@
 These can be calculated using $s-\mathrm{rank}(Q)$. %$\mathrm{dim}(\mathrm{ker}(Q))$. 
-The number of degrees of freedom for different explicit methods are shown in Tables\,\ref{table:DOF_exp} and~\ref{table:DOF_imp}.
+The number of degrees of freedom for different explicit methods are shown in Table\,\ref{table:DOF_exp}.
 
@@
 For all methods with the number of stages equal to the order of the method, there are no degrees of freedom without reducing the order.  
-If the classical RK4 method is used the order has to be reduced more because the RK4 method does not have embedded methods of order 3.
+If the classical RK4 method is used the order has to be reduced more because the RK4 method does not have embedded methods of order~3.
 In contrast to this, methods with $s > p$ have degrees of freedom even without reducing the order.
@@
 
-In Table\,\ref{table:DOF_imp} same was done for implicit methods. 
+In Table\,\ref{table:DOF_imp}, the same properties of some implicit
+methods are reported.
 We can see that the fully implicit methods Lobatto~IIIC4 and Radau~IIA3 require a drastic reduction of the order as expected.
-The Diagonally implicit SDIRK 54 method only requires an order reduction of one to get one degree of freedom for the weights.  
+The diagonally implicit SDIRK 54 method only requires an order reduction of one to get one degree of freedom for the weights.
 The TR-BDF2 method even allows adaptations without reducing the order.  
@@
 Another important property is the stability region. 
-The adaption of the weights does not solve issues with stability when used with a too big $\dt$. Therefore, the RK-method has to be stable for large $\dt$.
+The adaption of the weights does not solve issues with stability when used with a too big $\dt$. Therefore, the RK method has to be stable for large $\dt$.
 
@@
 A third property is the existence of a positive solution.
-This corresponds to the question if there is an embedded first order method that ensures positivity.
+This corresponds to the question whether there is an embedded first order method that ensures positivity.
 This is particularly interesting for implicit methods, because the $\dt$ is not limited by the stability.
-If we have an embedded first order method, we can ensure that there is always a positive solution, which might be of first Order.  
+If we have an embedded first order method, we can ensure that there is always a positive solution, which might be of first order.
 This is the case for the Im-Euler extrapolation methods, because they contain a chain of BE steps. 
@@
 When changing the weights the solution $u^{n+1}$ is changed. 
-By incorporating the order conditions for a certain order the space of possible weights that can be chosen is limited to a linear Subspace. 
+By incorporating the order conditions for a certain order the space of possible weights that can be chosen is limited to a linear subspace.
 Now it is important to know if the new solution $u^{n+1}$ still resembles the solution of the ODE $u(t_{n+1})$. 
-This can be seen in example in Section~\ref{sec:example_lin}.
+This can be seen in the example in Section~\ref{sec:example_lin}.
 This gives rise to the question whether the new solution is reasonable.
@@
 
-\subsubsection{Error detection and appproximation}
 For this approach it is difficult to give some mathematically rigorous statement regarding the convergence of the method, because it only applies if the error of the original method causes the solution to get negative.
@@
 One could say that the main idea of the adaption of the weights alters the error in a certain way, so that the numerical solution gets 'better' in the sense that it leads to positive solutions. 
-The local error of an RKM $e(\dt) =u(t_0 + \dt) - u^1$ can be expressed using the Taylor series of the RKM and the Taylor series of the exact solution. % Hairer I P134
-
+The local error of an RKM $e(\dt) =u(t_0 + \dt) - u^1$ can be expressed using the Taylor series of the RKM and the Taylor series of the exact solution, % Hairer I P134
 \begin{align}\label{eq:Taylor_sol_ref}
-u(t_0 + \dt) &= u(t_0) + u'(t_0) \dt + \frac{u''(t_0)}{2} \dt^2 + \cdots + \frac{u^{(p)}(t_0)}{p!} \dt^p + \frac{u^{(p+1)}(t_0)}{(p+1)!} \dt^{p+1} + \cdots \\
-u^{n+1} &= u(t_0)  + v_1 \dt + \frac{v_2}{2} \dt^2 + \cdots + \frac{v_p}{p!} \dt^p + \frac{v_{p+1}}{(p+1)!} \dt^{p+1} + \cdots 
+u(t_0 + \dt) &= u(t_0) + u'(t_0) \dt + \frac{u''(t_0)}{2} \dt^2 + \cdots + \frac{u^{(p)}(t_0)}{p!} \dt^p + \frac{u^{(p+1)}(t_0)}{(p+1)!} \dt^{p+1} + \cdots, \\
+u^{n+1} &= u(t_0)  + v_1 \dt + \frac{v_2}{2} \dt^2 + \cdots + \frac{v_p}{p!} \dt^p + \frac{v_{p+1}}{(p+1)!} \dt^{p+1} + \cdots .
 \end{align}
-
 The derivatives of $u(t)$ can be solely expressed with the partial derivatives of the RHS $f(t,u)$, whereas the coefficients $v_1,v_2,\cdots$ also depend on the RKM. 
@@
 Therefore $v_1,v_2,\cdots$ are functions of $b$.
-Because the the weights are chosen to ensure the Order Conditions the coefficients $v_1,\cdots,v_p$ are fixed and equal the according derivatives. 
+Because the weights are chosen to ensure the order conditions, the coefficients $v_1,\cdots,v_p$ are fixed and equal the according derivatives.
 The only remaining degrees of freedom are the coefficients $v_{p+1},v_{p+2},\cdots$.
@@
 This approximation is not very useful for real applications. 
-To approximate the error of a new step we propose following approximation of the local error:
+To approximate the error of a new step we propose the following approximation of the local error:
 
 \begin{align}
-err = |u(t^n)-u^n| &= |u(t^n) - (u^n_{unadapted}+\dt K(\tilde{b}-b))| \\
- &\leq \underbrace{|u(t^n)-u^n_{unadapted}|}_{\approx err_T}+\underbrace{|\dt K(\tilde{b}-b)|}_{= err_{adapt}} \label{eq:Err}
+err = |u(t^{n+1})-u^{n+1}| &= |u(t^{n+1}) - (u^{n+1}_{unadapted}+\dt F(\tilde{b}-b))| \\
+ &\leq \underbrace{|u(t^{n+1})-u^{n+1}_{unadapted}|}_{\approx err_T}+\underbrace{|\dt F(\tilde{b}-b)|}_{= err_{adapt}} \label{eq:Err}
 \end{align}
@@
 
-In Figure\,\ref{fig:stab} the stabilit regions of adapted RKM are plotted. 
+In Figure\,\ref{fig:stab} the stability regions of adapted RKMs are plotted.
 In Figure\,\ref{fig:stab_ck5} the Dormand--Prince 5 method is freely adapted.
@@
 \begin{equation}\label{eq:stab_func}
-R(z) = 1 + zb^T(I - zA)^{-1}e
+R(z) = 1 + zb^T(I - zA)^{-1}e.
 \end{equation}
@@
 \begin{equation}
-P(z) = R(z)-1 = z b^T (I - zA)^{-1} e
+P(z) = R(z)-1 = z b^T (I - zA)^{-1} e.
 \end{equation}
@@
 (Note: for all explicit methods this is always true, for implicit methods we have to exclude the poles $\{z | \det(I - zA) = 0\}$ or  $\{z | \det(I - zA) > \epsilon \}$ )
+
+\todo{Note (HR)} No. They are polynomials or rational functions
+on the complex plane. Hence, they can only be bounded if they are
+constant.
+
 We define $\mathcal{P}$ as all $P(z)$ for the class of $RKM_A$. 
@@
 \begin{equation}
-R(z) = 1 + zb^T(I - zA)^{-1}e
+R(z) = 1 + zb^T(I - zA)^{-1}e.
 \end{equation}
@@
 \begin{equation}
-R(z,b) = 1 + zb^T(I - zA)^{-1}e
+R(z,b) = 1 + zb^T(I - zA)^{-1}e.
 \end{equation}
@@
 \subsubsection{Stability of convex adaptation}
+
 If the new weights are chosen as a convex adaptation of weights it is easy to prove properties of the stability region.
-The stability function $R_{\chi b_a + (1-\chi) b_b}(z)$ of a convex adaptation can be calculated by the weighted sum of the stability functions $R_{b_a}(z)$ and $R_{b_b}(z)$
-
-Proof:
-\begin{multline}
-R_{\chi b_a + (1-\chi) b_b}(z) = R(z) = 1 + (\chi b_a + (1-\chi) b_b)^Tz(I - zA)^{-1}e \\
-= \chi + (1-\chi) + \chi b_a ^Tz(I - zA)^{-1}e + (1-\chi) b_b^Tz(I - zA)^{-1}e \\
-= \chi R_{b_a}(z) + (1-\chi) R_{b_b}(z)
-\end{multline}
-
-The resulting region of absolute stability is at least the intersection of the regions of absolute stability of the methods that define the convex adaptation.
-This also means that if all the used embedded methods used to construct the new weights are A-Stable, the resulting method is also A-Stable.
-
-We want to show that the convex adaptation of two methods is absolutely stable at the points where both original methods were absolutely stable.
-This can be formulated as $(|R_{b_a}(z)|  < 1) (|R_{b_b}(z)| < 1) \Rightarrow |R_{\chi b_a +(1- \chi) b_b}(z)| < 1$ with $\chi \in [0,1]$.
-
-\begin{multline}
-|R_{\chi b_a +(1-\chi) b_b}(z)| = |\chi R_{b_a}(z) + (1-\chi) R_{b_a}(z)| \leq |\chi R_{b_a}(z)| + |(1-\chi) R_{b_a}(z)|\\
- =| \chi| \underbrace{|R_{b_a}(z)|}_{<1} + |(1-\chi)| \underbrace{|R_{b_a}(z)|}_{<1} < \chi + (1-\chi)= 1
-\end{multline}
+
+\begin{theorem}
+  The stability region of a Runge--Kutta method $(A,b)$ where
+  $b = \sum_{i} g_i b^i$ is a convex combination of
+  $b^1, \dots, b^m \in \R^s$ (i.e.\ $g_i \in [0,1]$, $\sum_i g_i = 1$),
+  contains at least the intersection of the stability regions of
+  the methods $(A,b^1), \dots, (A,b^m)$
+\end{theorem}
+\begin{proof}
+  Since the stability function is an affine-linear function of the
+  weights, $R_{\sum_i g_i b^i}(z) = \sum_i g_i R_{b^i}(z)$. Hence,
+  if $z$ is in the stability region of all methods $(A,b^1), \dots, (A,b^m)$,
+  \begin{equation}
+    | R_{\sum_i g_i b^i}(z) |
+    \le
+    \sum_i g_i | R_{b^i}(z) |
+    \le
+    1.
+  \end{equation}
+\end{proof}
+
+\begin{remark}
+  This result is particularly important for implicit methods.
+  If all the used embedded methods used to construct the new weights
+  are A-Stable, the resulting method is also A-Stable.
+\end{remark}
 
@@
 When adapting the weights we cannot change them independently but only in ways that comply with the order conditions.
-$m$ is the number of dimensions of the Kernel of the Order Condition $m = \mathrm{dim}\{\mathrm{Ker} (Q) \}$
-The $b_1,\cdots,b_m$  are the Basis vectors of $\mathrm{Ker} (Q)$. With a coordinate transformation we get the coordinate vector $\hat{b}$ with the Matrix $B = [b_1,\cdots,b_m]$ and $b = B \hat{b}$.
-
-The general stability function of all embedded RKM that satisfy the Order Conditions $Q$ is given by
+$m$ is the number of dimensions of the kernel of the order condition $m = \mathrm{dim}\{\mathrm{Ker} (Q) \}$
+The $b_1,\cdots,b_m$  are the basis vectors of $\mathrm{Ker} (Q)$. With a coordinate transformation we get the coordinate vector $\hat{b}$ with the matrix $B = [b_1,\cdots,b_m]$ and $b = B \hat{b}$.
+
+The general stability function of all embedded RKM that satisfy the order conditions $Q$ is given by
 
@@
 \begin{equation}\label{eq:f(w,a)}
-f: \R^m \times \R \rightarrow \R f(\hat{b},a) = |R(\hat{b},z_0 + n a)|^2 -1 
+f: \R^m \times \R \rightarrow \R,  f(\hat{b},a) = |R(\hat{b},z_0 + n a)|^2 -1
 \end{equation}
 If explicit methods are used this function is known to be continuously differentiable because it can be written as a polynomial of $\hat{b},u$ and $v$.
-This is a necessary condition for the use of the Implicit function theorem.
+This is a necessary condition for the use of the implicit function theorem.
 It is also known by definition, that for the point $b = b_{orig}$ and $z = z_0$ the equation $ |R(\hat{b},z_0)|^2 -1 = 0 $ is satisfied.
@@
 
+\todo{TODO: Some better theorem?} %TODO
+\begin{lemma}
+  The stability function $R_{\bt}$ of an adapted RK method
+  satisfies
+  \begin{equation}
+    | R_{\bt}(z) |
+    \le
+    | R_{b}(z) |
+    + \| \bt - b \|_1 \| z (\I - z A)^{-1} e \|_\infty.
+  \end{equation}
+\end{lemma}
+\begin{proof}
+  Compute
+  \begin{equation}
+  \begin{aligned}
+    | R_{\bt}(z) |
+    &\le
+    | R_{b}(z) |
+    + | R_{\bt}(z) - R_{b}(z) |
+    =
+    | R_{b}(z) |
+    + | z (\bt - b) (\I - z A)^{-1} e
+    \\
+    &\le
+    | R_{b}(z) |
+    + \| \bt - b \|_1 \| z (\I - z A)^{-1} e \|_\infty.
+  \end{aligned}
+  \end{equation}
+\end{proof}
+This shows why the objective $\min \| \bt - b \|_1 \|$ is a good
+choice to control the change of the region of absolute stability,
+in particular for explicit methods for which
+$\| z (\I - z A)^{-1} e \|_\infty$ can be bounded by polynomial
+in $|z|$.
+
 \section{Implementation aspects of the algorithm}\label{sec:imple}
 
-TODO: Discuss in this Section aspects that are needed for implementation (and improve the usability) but do not concern the mathematical idea of the approach.
+\todo{TODO}: Discuss in this Section aspects that are needed for implementation (and improve the usability) but do not concern the mathematical idea of the approach. %TODO
 
@@
 
-For the further discussion, we are concerned with a system of $m$ coupled ODEs. 
-In this setup $u \in \R^m$.
+For the further discussion, we are concerned with a system of
+$m$ coupled ODEs, i.e.\ $u \in \R^m$.
 Systems of ODEs with a high number of dimensions can appear when solving a PDE with the method of lines.
@@
 To ensure that the solution is positive
-
 \begin{equation}
@@
 \end{equation}
-has to be fulfilled.
-This infers $m$ positivity constraints to the optimization problem. These can be written as
-
-\begin{equation}
-u_i + \dt \sum_{j=0}^s F_{i,j}  b_j \geq 0 \quad \forall {i \in \{1, \cdots,m \}}  
-\end{equation}
-
-This causes unnecessarily many inequality constraints on the LP.
-To simplify the problem, one wants to reduce the number of positivity constraints by ignoring it for ODEs where there is no issue with positivity.
-The new positivity constraints can be formulated as
-\begin{equation}
-u_i + \dt \sum_{j=0}^s F_{i,j}  b_j  \geq 0 \quad   \forall {i \in h \subseteq \{1,\cdots,m \}} 
-\end{equation}
-For this a set $h$ that includes all indecies $i \in \{1,\cdots,m \}$ that are needed for determining the weights has to be chosen.
-
+has to be fulfilled. These $m$ positivity constraints of the
+optimization problem can be written as
+\begin{equation}
+u_i^n + \dt \sum_{j=0}^s F_{i,j}  b_j \geq 0 \quad \forall {i \in \{1, \cdots,m \}}
+\end{equation}
+
+Imposing all of these constraints in the LP solver usually results
+in unnecessarily many constraints. In many cases, positivity is
+violated only for a subset $h \subseteq \{1, \dots, m\}$ of all
+components. This reduced set of positivity constraints can be
+formulated as
+\begin{equation}
+u_i^n + \dt \sum_{j=0}^s F_{i,j}  b_j  \geq 0 \quad \forall i \in h \subseteq \{1,\cdots,m \}.
+\end{equation}
 A reasonable approach is to set $h_0 = \{ i \in \{1,\cdots,m \} |  u_i^{n+1}  < 0 \}$. 
-After the new weights and $u^{n+1}$ is computed using $h_0$, the algorithm has to check if the positivity of the other values $u_i | i \notin h_0 \subseteq \{1,\cdots,m \}$ is still met. 
+After the new weights and $u^{n+1}$ are computed using $h_0$, the algorithm has to check if the positivity of the other values $\{u_i | i \notin h_0 \subseteq \{1,\cdots,m \}\}$ is still satisfied.
 If the condition is not met a new set $h_{a+1} = \{ i \in \{1,\cdots,m \}|  u_i^{n+1}  < 0 \} \cup h_{a}$ using the new $u^{n+1}$ is generated and the LP is solved again.
-This process is repeated until a solution is found that leads to a $u^{n+1} \geq 0$.
-The choice of $h_{a+1}$ based on the $h_{a}$ ensures that the algorithm cannot get stuck in a loop.
+This process is repeated until a solution is found that leads to $u^{n+1} \geq 0$.
+The choice of $h_{a+1}$ based on $h_{a}$ ensures that the algorithm cannot get stuck in a loop.
 In the worst case $h_a$ is growing very slowly until $h_a = \{1,\cdots,m \}$.
@@
 
-When enforcing a maximum value the number of constraints can be reduced using the same technique.
+When enforcing a maximum value, the number of constraints can be reduced using the same technique.
 When enforcing both maximum and minimum values two separate sets of active constraints $k_{min}$ and $k_{max}$ are used.
@@
 \subsection{Implemention}
-(Todo: Maybe a flowchart and a description on how we stitch everything together.)
-
-At fist the stagevaules and stagedeerivatives are calculated.
+(\todo{TODO}: Maybe a flowchart and a description on how we stitch everything together.) %TODO
+
+At fist the stage values $y_i$ and stage derivatives $f_i$ are calculated.
 After this $u^{n+1}$ is computed using the original weights. If the positivity is not violated, this solution is used to update the solution. 
 If $u^{n+1}$ contains negative values, the weights are adapted using the following algorithm.
-For this different orders are tried, stariting with the highest order that has degrees of Freedom for $\tilde{b}$. 
-At first the Order Conditions are constructed.
-On these a LP-Solver is invoiced.
+For this different orders are tried, starting with the highest order that has degrees of freedom for $\tilde{b}$.
+At first the order conditions are constructed.
+On these an LP solver is invoiced.
 If the LP is infeasible the algorithm immediately reduces the order.
 If the LP is feasible we test if the solution is acceptable.
-At first it is tested weather the new $u^{n+1}$ indeed complies with the constraints.
+At first it is tested whether the new $u^{n+1}$ indeed complies with the constraints.
 This is necessary because sometimes the used LP-solvers incorrectly identify the problem as feasible.
@@
 Afterwards the adaption error is calculated using \eqref{eq:Err}. If the error is bigger than a set maximum error the $\tilde{b}$ is also rejected.
-If the new set of weights is requected in one of the tests the order is reduced.
+If the new set of weights is rejected in one of the tests the order is reduced.
 As soon as the algorithm finds suitable weights these are used to update the solution.
-If all acceptable orders are tried and no weights were found the step hast to be requected.
-If the reduction of stepsize is used this process is done first for $\theta = 1$. If no weights were found, the process is repeated for a smaler $\theta$.
-
-\section{Results of experiments}\label{sec:Numeric_Results}
-
-(Note: Is here a good place to give a note on where supplementary code can be found?)
+If all acceptable orders are tried and no weights were found the step hast to be rejected.
+If the reduction of stepsize is used this process is done first for $\theta = 1$. If no weights were found, the process is repeated for a smaller $\theta$.
+
+\section{Results of numerical experiments}\label{sec:Numeric_Results}
+
+The implementation of the algorithms described above and code to
+reproduce the numerical examples reported here can be found in
+\cite{repository}.
 

### 2020-03-30 latex/overview.tex

@@
 
+\todo{Note (HR)} I think this should be moved to the beginning.
+
 \subsection{Explicit methods}\label{sec:Ex_expl}
 First we want to test the adaptive RKM used with explicit methods.  
-When used with explicit methods the cost of solving the LP significant because the computation of the stagederivatives only requires $s$ evaluations of the RHS. 
+When used with explicit methods the cost of solving the LP is significant because the computation of the stage derivatives only requires $s$ evaluations of the RHS.
 For most of the linear test problems tried, the explicit methods yield to positive results.  
 When increasing the step size, issues with stability occur before getting negative values.  
-An example therefore the ODE in in Section~\ref{sec:example_lin}.
+An example for this is the ODE in Section~\ref{sec:example_lin}.
 Some nonlinear RHS may require very small timesteps to preserve positivity.  
@@
 An example is the reaction equation solved in Section~\ref{sec:example_reac}.
-The positivity of stage values is not ensured, therefore the right hand side has to be defined for negative values. If this is not the case, the right hand side has to be defined for negative values. 
-%All the following testproblems did not nead this modification.
+Since the positivity of the stage values is not ensured, the right hand side has to be defined for negative values. If this is not the case, the right hand side has to be adapted accordingly.
 
@@
 \begin{equation}
-\frac{\partial u}{\partial t} = -a \frac{\partial u}{\partial z} - K (u)
+\frac{\partial u}{\partial t} = -a \frac{\partial u}{\partial x} - K (u)
 \end{equation}
 is simulated with the method of lines on the domain $x = [0,1]$ with $N =100$ points.
-The problem consists of an Advection part and an Exponential decay.
+The problem consists of an advection part and an exponential decay.
 At $x=0$ a inflow boundary with $u(0,t)=1$ is used.
@@
 \centering
-\begin{subfigure}[b]{0.45\textwidth}
+\begin{subfigure}[t]{0.45\textwidth}
+\centering
 \includegraphics[width=1\textwidth]{plots/Advection_Decay.pdf}
-\caption{Solution of Advection Decay Problem}
+\caption{Numerical solution.}
 \label{fig:sol_AdDe}
-\end{subfigure}
-\\
-\begin{subfigure}[b]{0.45\textwidth}
+\end{subfigure}%
+~
+\begin{subfigure}[t]{0.45\textwidth}
+\centering
 \includegraphics[width=1\textwidth]{plots/b_Advection_Decay.pdf}\\
-\caption{Weights used for Advecion Decay Problem}
+\caption{Adaptation of the weights used for }
 \label{fig:weights_AdDe}
 \end{subfigure}
+\caption{Numerical results for the advection decay problem.}
 \end{figure}
@@
 A very interesting class of methods are the implicit extrapolation methods.
-These are Diagonally implicit and $p < s$.
+These are diagonally implicit and $p < s$.
 This means that they allow for changes of the weights without a reduction of the order.
 Also, all stage values are computed using the BE method. This means that all intermediate stages are positive.  
-Also, an embedded BE Euler step is included. This ensures that a positive solution always exists, even if it is of first order.  
+Also, an embedded BE step is included. This ensures that a positive solution always exists, even if it is of first order.
 We test the adaptive RKM on the diffusion equation 
 \begin{equation}
-\frac{\partial }{\partial t} = D \frac{\partial^2}{\partial x^2}
+\frac{\partial }{\partial t} u = D \frac{\partial^2}{\partial x^2} u
 \end{equation}
@@
 \end{equation}
-As initial condition $u^0 = (0,\cdots,0,1,0,\cdots,0)^T$ is used. The diffusion coefficient is $D=1$ 
+As initial condition $u^0 = (0,\cdots,0,1,0,\cdots,0)^T$ is used. The diffusion coefficient is $D=1$.
 The ODE is solved using the implicit EX3 method.  
-For large (Note: better word for 'not small enough') $\dt$ the method computes negative values for $u^1$. 
+For large (\todo{TODO}: better word for 'not small enough') $\dt$ the method computes negative values for $u^1$.
 These can be corrected by adapting the weights.  
@@
 \end{subequations}
-
-With the parameters $a=\num{1e-2} $ and $ d=\num{1e-6}$. 
+with the parameters $a=\num{1e-2} $ and $ d=\num{1e-6}$.
 The PDE is simulated on the domain $x = [0,1]$ with $N=100$ points. Periodic boundary conditions are used. 
@@
 The computation is done using the EX3 method with free adaptation. 
-As step size control a PI-Control from\,\cite{hairer_solving_1996} is used. The Error was estimated using \eqref{eq:Err}. The Tolerance was set to $Tol = 0.01$.  
+As step size control a PI-control from\,\cite{hairer_solving_1996} is used. The error was estimated using \eqref{eq:Err}. The tolerance was set to $Tol = 0.01$.
 The final time is $t_{end} = 50$. 
@@
         \includegraphics[width=\textwidth]{plots/ADP_sol_18.pdf}
-        \caption{$t=$}
+        \caption{$t=18$}
         \label{fig:sol_ADP18}
@@
         \includegraphics[width=\textwidth]{plots/ADP_sol_27.pdf}
-        \caption{$t=$}
+        \caption{$t=27$}
         \label{fig:sol_ADP27}
@@
         \includegraphics[width=\textwidth]{plots/ADP_sol_50.pdf}
-        \caption{$t=$}
+        \caption{$t=50$}
         \label{fig:sol_ADP50}
     \end{subfigure}
-
-    \caption{Solution of Advection-Diffusion-Reaction Problems at different times}\label{fig:Sol_ADP}
+    \caption{Solution of advection-diffusion-reaction problem at different times.}\label{fig:Sol_ADP}
 \end{figure}
@@
 \includegraphics[width=0.85\textwidth]{plots/ADP_stepsize,b.pdf}
-\caption{Different statistics of computation of ADPD-Problem}
+\caption{Different statistics of computation of the advection-diffusion-reaction problem.}
 \label{fig:Stats_ADP}
@@
 
+\todo{Note (HR)} Should the minimum of $u$ in Figure~\ref{fig:Stats_ADP} be negative before the adaptation?
+
 \section{Conclusion} \label{sec:conclusion}
 It is possible to adapt the weights to enforce positivity for RKMs that are not positivity preserving. 
-One main limitation is that the resulting Order has to be lower than the number of stages.  
+One main limitation is that the resulting order has to be lower than the number of stages.
 An error approximation for this method was given.  
@@
 Because the time step size is limited by the stability it is only useful for a small interval of time steps. 
-The adaptive method mainly interesting for diagonally implicit methods.  
+The adaptive method is mainly interesting for diagonally implicit methods.
 The times step size is not limited by stability. 

### 2020-03-31 latex/overview.tex
references in tables of DOFs and layout

@@
 \addbibresource{Bib_RK_optim.bib}
+
+\usepackage{booktabs}
+\usepackage{multirow}
 
@@
 \centering    %Generated below============ 
-\begin{tabular}{|l c|c c c c c c |} 
- \hline 
-Order & &1&2&3&4&5&6 \\ 
- \hline Classical RK4& s=4&3&2&0&0& - & -  \\ 
- SSPRK(10,4)& s=10&9&8&6&4& - & -  \\ 
- Cash-Karp RK5(4)6& s=6&5&4&2&1&0& -  \\ 
- Dormand-Prince RK5(4)7& s=7&6&5&3&1&0& -  \\ 
- \hline 
- \end{tabular}
+% \begin{tabular}{|l c|c c c c c c |}
+%  \hline
+% Order & &1&2&3&4&5&6 \\
+%  \hline Classical RK4& s=4&3&2&0&0& - & -  \\
+%  SSPRK(10,4)& s=10&9&8&6&4& - & -  \\
+%  Cash-Karp RK5(4)6& s=6&5&4&2&1&0& -  \\
+%  Dormand-Prince RK5(4)7& s=7&6&5&3&1&0& -  \\
+%  \hline
+%  \end{tabular}
+  \begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}lr*6c@{}}
+    \toprule
+    Method & $s$ & \multicolumn{6}{c}{Order $\tilde p$} \\
+    & & 1 & 2 & 3 & 4 & 5 & 6 \\
+    \midrule
+    Classical RK4 \cite{kutta1901beitrag} & 4 & 3 & 2 & 0 & 0 & --- & --- \\
+    SSPRK(10,4) \cite{ketcheson2008highly} & 10&9&8&6&4& --- & ---\\
+    Cash--Karp RK5(4)6 \cite{cash1990variable} & 6&5&4&2&1&0& --- \\
+    Dormand--Prince RK5(4)7 \cite{prince1981high}& 7&6&5&3&1&0& --- \\
+    \bottomrule
+  \end{tabular*}
   \caption{Degrees of freedom for the choice of the weights for some explicit methods.} %Generated above============
@@
 \centering   %Generated below============ 
-\begin{tabular}{|l c|c c c c c c |} 
- \hline 
-Order & &1&2&3&4&5&6 \\ 
- \hline Implicit Euler& s=1&0& - & - & - & - & -  \\ 
- Lobatto~IIIC4& s=4&3&2&1&0&0&0 \\
- Radau~IIA3& s=3&2&1&0&0&0& -  \\
- SDIRK 54& s=5&4&3&1&0& - & -  \\ 
- TR-BDF2& s=3&2&1& - & - & - & -  \\ 
- Im-Euler 2& s=3&2&1& - & - & - & -  \\ 
- Im-Euler 3& s=6&5&4&2& - & - & -  \\ 
- Im-Euler 4& s=10&9&8&6&3& - & -  \\ 
- \hline 
- \end{tabular}
+%   \begin{tabular}{|l c|c c c c c c |}
+%   \hline
+%   Order & &1&2&3&4&5&6 \\
+%   \hline Implicit Euler& s=1&0& - & - & - & - & -  \\
+%   Lobatto~IIIC4& s=4&3&2&1&0&0&0 \\
+%   Radau~IIA3& s=3&2&1&0&0&0& -  \\
+%   SDIRK(5,4)& s=5&4&3&1&0& - & -  \\
+%   TR-BDF2& s=3&2&1& - & - & - & -  \\
+%   Im-Euler 2& s=3&2&1& - & - & - & -  \\
+%   Im-Euler 3& s=6&5&4&2& - & - & -  \\
+%   Im-Euler 4& s=10&9&8&6&3& - & -  \\
+%   \hline
+%   \end{tabular}
+   \begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}lr*6c@{}}
+    \toprule
+    Method & $s$ & \multicolumn{6}{c}{Order $\tilde p$} \\
+    & & 1 & 2 & 3 & 4 & 5 & 6 \\
+    \midrule
+    Implicit Euler& 1&0& --- & --- & --- & --- & ---  \\
+    Lobatto~IIIC4 \cite{chipman1971stable} & 4&3&2&1&0&0&0 \\
+    Radau~IIA3 \cite{ehle1969pade} & 3&2&1&0&0&0& ---  \\
+    SDIRK(5,4) \cite[eq. (6.18)]{hairer_solving_1996}& 5&4&3&1&0& --- & ---  \\
+    TR-BDF2 \cite{bank1985transient} & 3&2&1& --- & --- & --- & ---  \\
+    Extrapolation Im-Euler 2& 3&2&1& --- & --- & --- & ---  \\
+    Extrapolation Im-Euler 3& 6&5&4&2& --- & --- & ---  \\
+    Extrapolation Im-Euler 4& 10&9&8&6&3& --- & ---  \\
+    \bottomrule
+  \end{tabular*}
   \caption{Degrees of freedom for the choice of the weights for some implicit methods.} %Generated above============
@@
 We can see that the fully implicit methods Lobatto~IIIC4 and Radau~IIA3 require a drastic reduction of the order as expected.
-The diagonally implicit SDIRK 54 method only requires an order reduction of one to get one degree of freedom for the weights.
+The diagonally implicit SDIRK(5,4) method only requires an order reduction of one to get one degree of freedom for the weights.
 The TR-BDF2 method even allows adaptations without reducing the order.  

### 2020-03-31 latex/overview.tex
revisions of Section 4

@@
 %   Im-Euler 2& s=3&2&1& - & - & - & -  \\
-%   Im-Euler 3& s=6&5&4&2& - & - & -  \\
+%   Im-Euler~3& s=6&5&4&2& - & - & -  \\
 %   Im-Euler 4& s=10&9&8&6&3& - & -  \\
@@
     Extrapolation Im-Euler 2& 3&2&1& --- & --- & --- & ---  \\
-    Extrapolation Im-Euler 3& 6&5&4&2& --- & --- & ---  \\
+    Extrapolation Im-Euler~3& 6&5&4&2& --- & --- & ---  \\
     Extrapolation Im-Euler 4& 10&9&8&6&3& --- & ---  \\
@@
 The TR-BDF2 method even allows adaptations without reducing the order.  
-The Implicit-Euler extrapolation methods also exhibit degrees of freedom without a reduction of the order.
+The implicit Euler extrapolation methods also exhibit degrees of freedom without a reduction of the order.
 
@@
 This corresponds to the question whether there is an embedded first order method that ensures positivity.
-This is particularly interesting for implicit methods, because the $\dt$ is not limited by the stability.
+This is particularly interesting for implicit methods, because $\dt$ can be unrestricted by stability requirements.
 If we have an embedded first order method, we can ensure that there is always a positive solution, which might be of first order.
@@
 When changing the weights the solution $u^{n+1}$ is changed. 
-By incorporating the order conditions for a certain order the space of possible weights that can be chosen is limited to a linear subspace.
+By incorporating the order conditions for a certain order the space of possible weights that can be chosen is limited to an affine subspace.
 Now it is important to know if the new solution $u^{n+1}$ still resembles the solution of the ODE $u(t_{n+1})$. 
@@
 Both errors are added to get an approximation of the total error $err = err_T + err_{adapt}$.
-The truncation error and the adaption error never have the same direction.  
-Therefore, the triangle inequality becomes a strict inequality and the real error is always smaller than the estimate. (Note: as long as the Truncation error estimator is right) 
+% The truncation error and the adaption error never have the same direction.
+% Therefore, the triangle inequality becomes a strict inequality and the real error is always smaller than the estimate. (Note: as long as the Truncation error estimator is right)
 %\subsubsection{Steppsize Control}
@@
 
-By adapting the $b$ the used RK method is changed. This also alters the stability function and therefore changes the region of absolute stability.
-It depends on the solved problem, the RKM and the used $\dt$ by how many steps the stability function is altered.
-For problems where only a small number of steps are affected a bigger change in the stability function would be acceptable.
-For problems which require an adaptation of the weights for every step one needs to make sure that the resulting method is stable.
-
-In Figure\,\ref{fig:stab} the stability regions of adapted RKMs are plotted.
-In Figure\,\ref{fig:stab_ck5} the Dormand--Prince 5 method is freely adapted.
-The weights are taken form the example in Section~\ref{sec:Ex_expl}.
-
-In Figure\,\ref{fig:stab_ex3} the stability regions of the BE-Extraplotaion 3 method and the embedded chain of three BE steps is plotted.
-Additionally the stability regions of convex combinations of these two methods are plotted.
+Adapting the weights $b$ changes the RK method. Hence, the stability
+function is altered and the region of absolute stability varies.
+How often and how much the stability function is changed depends
+of course on the problem, the RKM, and the timestep $\dt$.
+For problems where only a small number of steps are affected, a bigger
+change in the stability function might be acceptable. For problems
+that require an adaptation of the weights for every step, one needs
+to make sure that the resulting method is stable.
+
+The stability regions of adapted RKMs are visualized in
+Figure\,\ref{fig:stab}.
+In Figure\,\ref{fig:stab_dp5} the Dormand--Prince RK5 method is freely adapted.
+The weights are taken from the example in Section~\ref{sec:Ex_expl}.
+In Figure\,\ref{fig:stab_ex3} the stability regions of the Im-Euler~3
+extrapolation method and the embedded chain of three BE steps are
+plotted. Additionally the stability regions of convex combinations
+of these two methods are shown.
 
@@
          \includegraphics[width=\textwidth]{plots/stab_dp5.pdf}
-         \caption{Dormand--Prince 5\\
-         with $\tilde{b}$ form \ref{fig:weights_AdDe}
-         }
-         \label{fig:stab_ck5}
+         \caption{Dormand--Prince RK5 with free adaptation
+                   of the weights $\bt$ as in the example in Figure~\ref{fig:weights_AdDe}.}
+         \label{fig:stab_dp5}
      \end{subfigure}
@@
          \includegraphics[width=\textwidth]{plots/stab_ex3.pdf}
-         \caption{Implicit Extrapolation 3 and embedded chain of three BE steps with convex combinations of both weights}
+         \caption{Im-Euler~3 extrapolation method and embedded chain
+                  of three BE steps with convex combinations of both
+                  weights.}
          \label{fig:stab_ex3}
      \end{subfigure}
-        \caption{Change of stability region for free adaptation(a) and for convex adaptation(b)}
+        \caption{Change of stability region for the free adaptation and
+                 the convex adaptation of the weights.}
         \label{fig:stab}
@@
 
-\paragraph*{Version 1} $ $\\
-
-The stability function is defined as 
-\begin{equation}\label{eq:stab_func}
-R(z) = 1 + zb^T(I - zA)^{-1}e.
-\end{equation}
-We can define a function $P(z)$
-\begin{equation}
-P(z) = R(z)-1 = z b^T (I - zA)^{-1} e.
-\end{equation}
-These rational functions are $\in \mathcal{L}^\infty$.
-(Note: for all explicit methods this is always true, for implicit methods we have to exclude the poles $\{z | \det(I - zA) = 0\}$ or  $\{z | \det(I - zA) > \epsilon \}$ )
-
-\todo{Note (HR)} No. They are polynomials or rational functions
-on the complex plane. Hence, they can only be bounded if they are
-constant.
-
-We define $\mathcal{P}$ as all $P(z)$ for the class of $RKM_A$. 
-This is a vector space.
-Proof:
-\begin{multline}
-\alpha P_a(z) + \gamma P_c(z) =  \alpha z a^T (I - zA)^{-1} e + \gamma z c^T (I - zA)^{-1} e=\\
-z (\alpha a+ \gamma c)^T (I - zA)^{-1} e= P_{\alpha a + \gamma c}(z) \in \mathcal{P}
-\end{multline}
-
-The $P(z)$ is a linear combination of the rational functions
-\begin{subequations}
-\begin{align}
-p_1(z) &= z e_1^T (I - zA)^{-1}\\
-\vdots \\
-p_s(z) &= z e_s^T (I - zA)^{-1}
-\end{align}
-\end{subequations}
-whereas $e_1,\cdots,e_s$ is the standard basis of $\R^s$.
-These span $\mathcal{P}$. If $p_{e_1}(z),\cdots,p_{e_s}(z)$ are linearly independent $p_{e_1}(z),\cdots,p_{e_s}(z)$ are a basis of $\mathcal{P}$.
-
-With this results the stability function $R_b(z)$ of any RKM in $RKM_{(A)}$ can be calculated with 
-\begin{equation}
-R_b(z) = 1 + P_b(z) = 1 + \sum_{j = 1}^s b_s p_{e_1}
-\end{equation}
-
-\paragraph*{Version 2} $ $\\
 The stability function is defined as a function $\CN \to \CN$ with
 \begin{equation}
-R(z) = 1 + zb^T(I - zA)^{-1}e.
-\end{equation}
-We can add $b$ as a second parameter and get a function $\CN \times \R^s \to \CN$ with
-\begin{equation}
-R(z,b) = 1 + zb^T(I - zA)^{-1}e.
-\end{equation}
-The stability function is a linear function in the weights.
-
-\paragraph*{Version 3} $ $\\
-(Ignore this topic, because the following Sections also can work without it)
-
-%(Todo: It is not straightforward if there are no pole zero cancellations in the stability function.
+R(z) = 1 + zb^T(I - zA)^{-1}e,
+\end{equation}
+where $e = (1, \dots, 1)^T \in \R^s$.
+We can add $b$ as a second parameter and get a function
+$R\colon \R^s \times \CN  \to \CN$ with
+\begin{equation}
+R_b(z) = 1 + zb^T(I - zA)^{-1}e.
+\end{equation}
+The stability function is an affine-linear function of the weights.
+
+%TODO: It is not straightforward if there are no pole zero cancellations in the stability function.
 For explicit methods, the stability function is a polynomial o degree $\leq s$ (Hundsdorfer).
@@
 
-If the new weights are chosen as a convex adaptation of weights it is easy to prove properties of the stability region.
-
+If the new weights are chosen by convex adaptation of given weights,
+it is easy to prove some properties of the stability region.
 \begin{theorem}
@@
   contains at least the intersection of the stability regions of
-  the methods $(A,b^1), \dots, (A,b^m)$
+  the methods $(A,b^1), \dots, (A,b^m)$.
 \end{theorem}
@@
 
-\begin{remark}
 This result is particularly important for implicit methods.
-  If all the used embedded methods used to construct the new weights
-  are A-Stable, the resulting method is also A-Stable.
-\end{remark}
+If all the embedded methods used to construct the new weights
+are A-stable, the resulting method is also A-stable.
 
@@
 in particular for explicit methods for which
-$\| z (\I - z A)^{-1} e \|_\infty$ can be bounded by polynomial
+$\| z (\I - z A)^{-1} e \|_\infty$ can be bounded by a polynomial
 in $|z|$.

### 2020-03-31 latex/overview.tex
revisions of Sections 5 and 6

@@
 
-\todo{TODO}: Discuss in this Section aspects that are needed for implementation (and improve the usability) but do not concern the mathematical idea of the approach. %TODO
+Having described the mathematical idea of our approach,
+we discuss some practical implementation details of the
+proposed algorithm in the following.
 
@@
 
-At fist the stage values $y_i$ and stage derivatives $f_i$ are calculated.
-After this $u^{n+1}$ is computed using the original weights. If the positivity is not violated, this solution is used to update the solution. 
+At first, the stage values $y_i$ and stage derivatives $f_i$ are calculated.
+Afterwards, $u^{n+1}$ is computed using the original weights $b$.
+If the positivity is not violated, this solution is used to update the solution.
 If $u^{n+1}$ contains negative values, the weights are adapted using the following algorithm.
-For this different orders are tried, starting with the highest order that has degrees of freedom for $\tilde{b}$.
-At first the order conditions are constructed.
-On these an LP solver is invoiced.
-If the LP is infeasible the algorithm immediately reduces the order.
-If the LP is feasible we test if the solution is acceptable.
-At first it is tested whether the new $u^{n+1}$ indeed complies with the constraints.
-This is necessary because sometimes the used LP-solvers incorrectly identify the problem as feasible.
-%(Note: because of the chosen objective function it cannot be unbounded)
-Afterwards the adaption error is calculated using \eqref{eq:Err}. If the error is bigger than a set maximum error the $\tilde{b}$ is also rejected.
-If the new set of weights is rejected in one of the tests the order is reduced.
-As soon as the algorithm finds suitable weights these are used to update the solution.
-If all acceptable orders are tried and no weights were found the step hast to be rejected.
+
+Start with the highest order $\tilde p$ such that there are degrees
+of freedom for the weights $\bt$. Apply an LP solver to the optimization
+problem with accuracy and positivity constraints.
+If the LP is infeasible, reduce the order $\tilde p$ by one.
+If the LP is feasible, test whether the solution is acceptable.
+At first, test whether the new $u^{n+1}$ satisfies the constraints.
+This is necessary because sometimes the applied LP solvers incorrectly
+identify the problem as feasible.
+Afterwards the adaption error is calculated using \eqref{eq:Err}.
+If the error is bigger than a given maximum error tolerance, the
+new weights $\bt$ are also rejected.
+If the new set of weights is rejected in one of the tests the order
+$\tilde p$ is reduced by one.
+As soon as the algorithm finds suitable weights, these are used to update the solution.
+If all acceptable orders are tried and no weights were found, the step hast to be rejected.
+
 If the reduction of stepsize is used this process is done first for $\theta = 1$. If no weights were found, the process is repeated for a smaller $\theta$.
@@
 reproduce the numerical examples reported here can be found in
-\cite{repository}.
-
-(Note: Should we discuss which solver is used. Probably it is not worth using space for some plot etc. on this... Maybe a short note?)
+\cite{repository}. We have used CVXPY to solve the LPs
+\cite{cvxpy, cvxpy_rewriting}.
+\todo{TODO: Which solver of CVXPY?} %TODO
 
@@
 \begin{equation}
-u_i=0 \Rightarrow f_i(t,u_1,\cdots,u_i,\cdots,u_n) \geq 0  \forall {u_c \geq 0} \forall {t}
+u_i=0 \implies f_i(t,u_1,\cdots,u_i,\cdots,u_n) \geq 0 \quad \forall {u_c \geq 0}\; \forall {t}
 \end{equation}
@@
 
-As test problem the PDE
-\begin{equation}
-\frac{\partial u}{\partial t} = -a \frac{\partial u}{\partial x} - K (u)
-\end{equation}
-is simulated with the method of lines on the domain $x = [0,1]$ with $N =100$ points.
-The problem consists of an advection part and an exponential decay.
-At $x=0$ a inflow boundary with $u(0,t)=1$ is used.
-The initial condition is $u(x,0)=0$.
-As semidiscretization an upwinding scheme is used. This leads to positivity preserving ODE
-\begin{equation}
-\frac{\mathrm d}{\mathrm d t} u_i = \frac{a}{\Delta x} \left( u_{i-1} - u_i \right) - K u_{i-1}
+A test problem similar to \cite{shampine_non-negative_2005} is the PDE
+\begin{equation}
+\begin{aligned}
+  \frac{\partial u(t,x)}{\partial t}
+  &=
+  -a \frac{\partial u(t,x)}{\partial x} - K u(t,x),
+  && x \in (0, 1), t \in (0,1),
+  \\
+  u(t,0) &= 1,
+  && t \in (0,1],
+  \\
+  u(0,x) &= 0,
+  && x \in [0,1],
+\end{aligned}
+\end{equation}
+which consists of an advection part and an exponential decay.
+The numerical approximation uses the method of lines and a first order
+upwind finite difference semidiscretization with $N = 100$ points.
+This leads to the positivity preserving ODE
+\begin{equation}
+\frac{\mathrm d}{\mathrm d t} u_i = \frac{a}{\Delta x} \left( u_{i-1} - u_i \right) - K u_{i-1}.
 \end{equation}
 The parameters are set to $a=1$ and $K=1$.
-We use the Dormand--Prince 5 RKM and adapt the weights using the free adaptation.
+We use the Dormand--Prince RK5 method and adapt the weights using the free adaptation.
 In Figure\,\ref{fig:sol_AdDe} the results for $\Delta t = 0.02$ are plotted for different timesteps. We can see that the solution approaches an exponential function with $t \rightarrow \infty$.
@@
 \includegraphics[width=1\textwidth]{plots/Advection_Decay.pdf}
-\caption{Numerical solution.}
+\caption{Numerical solution at different times.}
 \label{fig:sol_AdDe}
@@
 \includegraphics[width=1\textwidth]{plots/b_Advection_Decay.pdf}\\
-\caption{Adaptation of the weights used for }
+\caption{Adaptation of the weights.}
 \label{fig:weights_AdDe}
@@
 For a larger $\dt$ the original method leads to negative values and the weights are altered.
-For $\dt >0.016$ the dp5 is no longer stable.
+For $\dt >0.016$ the baseline method is no longer stable.
 For the ODE, the reference solution can be computed using the matrix exponential. 
@@
 \centering
-\includegraphics[scale=0.75]{plots/conv_adde.pdf}
-\caption{Convergence adapted dp5}
+\includegraphics[scale=0.6]{plots/conv_adde.pdf}
+\caption{Convergence results of the adapted Dormand--Prince RK5.}
 \label{fig:conv_expl}
@@
 Implicit methods seem like an advantageous choice for a couple of reasons. 
-Fistly the cost of solving the LP is relatively small compared to the cost of solving the stage equations. 
-Secondly the timestep is not limited by the stability of the method. 
-Therefore, it is possible to use larger timesteps, that are more likely to lead to negative values. 
+Firstly, the cost of solving the LP is relatively small compared to the cost of solving the stage equations.
+Secondly, the timestep is not limited by the stability of the method.
+Therefore, it is possible to use larger timesteps that are more likely to lead to negative values.
+
 A very interesting class of methods are the implicit extrapolation methods.
@@
 This means that they allow for changes of the weights without a reduction of the order.
-Also, all stage values are computed using the BE method. This means that all intermediate stages are positive.  
+Also, all stage values are computed using the BE method. Hence, all intermediate stages are positive.
 Also, an embedded BE step is included. This ensures that a positive solution always exists, even if it is of first order.
-We test the adaptive RKM on the diffusion equation 
+
+We test the proposed adaptation algorithm on the diffusion equation
 \begin{equation}
@@
 \end{equation}
-with homogeneous Dirichlet boundary conditions on the domain $x = [-0.5,0.5]$ with $N=100$ points. The equation is semidiscreticed using the 3-point-scheme
+with homogeneous Dirichlet boundary conditions on the domain $x = [-0.5,0.5]$ with $N=100$ points. The equation is semidiscretized using the 3-point-scheme
 \begin{equation}
@@
 As initial condition $u^0 = (0,\cdots,0,1,0,\cdots,0)^T$ is used. The diffusion coefficient is $D=1$.
-The ODE is solved using the implicit EX3 method.  
-For large (\todo{TODO}: better word for 'not small enough') $\dt$ the method computes negative values for $u^1$.
+
+The ODE is solved using the Im-Euler~3 extrapolation method.
+For large $\dt$ the method computes negative values for $u^1$.
 These can be corrected by adapting the weights.  
 The solutions are computed using the free adaptation and convex adaptation for $\dt = \num{1e-3}$.
-The results for the free adaptation are plotted in Figure\,\ref{fig:sol_Diff_a}. The used weights are plotted in Figure\,\ref{fig:weights_Diff_a}. 
+The results for the free adaptation are plotted in Figure\,\ref{fig:sol_Diff_a} and the corresponding change of the weights
+is shown in Figure\,\ref{fig:weights_Diff_a}.
 The original solution for the first step is negative. Therefore, the weights have to be changed. 
@@
 To prevent this glitch from happening we choose the weights based on a convex adaptation. 
-A first order embedded method is added. The solution is shown in Figure\,\ref{fig:sol_Diff_c} and the weights are plotted in Figure\,\ref{fig:weights_Diff_c}. 
+A first order embedded method is added. The solution is shown in Figure\,\ref{fig:sol_Diff_c} and the weights are visualized in Figure\,\ref{fig:weights_Diff_c}.
 The weights for the first step are altered again. 
@@
 \begin{subfigure}[b]{0.45\textwidth}
-\includegraphics[width=1\textwidth]{plots/Diff_Direct.pdf}\\
-\caption{Solution of Diffusion for free adaptation \newline $ $}
+\centering
+\includegraphics[width=1\textwidth]{plots/Diff_Direct.pdf}
+\caption{Solution using the free adaptation.}
 \label{fig:sol_Diff_a}
@@
 \begin{subfigure}[b]{0.45\textwidth}
-\includegraphics[width=1\textwidth]{plots/Diff_Convex.pdf}\\
-\caption{Solution of Diffusion for convex adaptation}
+\centering
+\includegraphics[width=1\textwidth]{plots/Diff_Convex.pdf}
+\caption{Solution using the convex adaptation.}
 \label{fig:sol_Diff_c}
 \end{subfigure}
-
+\\
 \begin{subfigure}[b]{0.45\textwidth}
-\includegraphics[width=1\textwidth]{plots/b_Diff_Direct.pdf}\\
-\caption{Weights used for free adaptation}
+\centering
+\includegraphics[width=1\textwidth]{plots/b_Diff_Direct.pdf}
+\caption{Change of weights for free adaptation.}
 \label{fig:weights_Diff_a}
@@
 \begin{subfigure}[b]{0.45\textwidth}
-\includegraphics[width=1\textwidth]{plots/b_Diff_Convex.pdf}\\
-\caption{Weights used for convex adaptation}
+\centering
+\includegraphics[width=1\textwidth]{plots/b_Diff_Convex.pdf}
+\caption{Change of weights for convex adaptation.}
 \label{fig:weights_Diff_c}
 \end{subfigure}
+\caption{Numerical results for the diffusion problem and the
+         adapted Im-Euler~3 extrapolation method.}
 \end{figure}
 
-In Figure\,\ref{fig:conv_impl} the convergence is shown for the unaltered EX3 method, that potentially leads to negative values, the adaptive method with free adaptation (EX3a) and the adapted method using convex adaptations (EX3c).
+In Figure\,\ref{fig:conv_impl}, the convergence is shown for the unaltered Im-Euler~3 extrapolation method (ex3, potentially resulting in negative values), the adaptive method with free adaptation (ex3a) and the adapted method using convex adaptations (ex3c).
 Also, the BE method is plotted. It is only of 1st order but preserves positivity for all $\dt$. 
 For $\dt < \num{3e-5} $ the standard weights yield to a positive result. For larger $\dt$ the weights have to be adapted to ensure positivity. 
-The EX3a shows similar convergence properties as the original method. 
+The ex3a shows similar convergence properties as the original method.
 This can be expected, because the adapted method is still of 3rd order. 
-The EX3c has larger errors than the EX3a but leads to physical solutions for all time-steps.  
-This is no surprise because the adapted RKM used for the first step is only of 1st order. 
+The ex3c has larger errors than the ex3a but leads to physical solutions for all time-steps.
+This is no surprise because the adapted RKM used for the first step is only of first order.
 But the adaptive method still outperforms the BE, even when accounting for the higher cost per step.
@@
 \includegraphics[scale=0.75]{plots/conv_heat.pdf}
-\caption{Convergence adapted ex3}
+\caption{Convergence result of baseline and adapted Im-Euler~3
+         extrapolation methods for the diffusion problem.}
 \label{fig:conv_impl}
@@
 \end{subequations}
-with the parameters $a=\num{1e-2} $ and $ d=\num{1e-6}$.
+with parameters $a=\num{1e-2} $ and $ d=\num{1e-6}$.
 The PDE is simulated on the domain $x = [0,1]$ with $N=100$ points. Periodic boundary conditions are used. 
-The advection part is semidiscreticed using a first order upwind scheme and the diffusion part is semidiscreticed using a 3-point-scheme. This leads to a positivity preserving system of ODEs.  
-The computation is done using the EX3 method with free adaptation. 
+The advection part is semidiscretized using a first order upwind scheme and the diffusion part is semidiscretized using a 3-point-scheme. This leads to a positivity preserving system of ODEs.
+The computation is done using the Im-Euler~3 extrapolation
+method with free adaptation.
 As step size control a PI-control from\,\cite{hairer_solving_1996} is used. The error was estimated using \eqref{eq:Err}. The tolerance was set to $Tol = 0.01$.
 The final time is $t_{end} = 50$.
+
 The simulation required 264 steps. Of these, 72 required an adaptation of the weights. 
@@
 In Figure\,\ref{fig:sol_ADP18} it can be seen that the reaction occurs in a small interfaces. 
-Outside of this regions quantities are close to $0$. Therefore, it is very likely that negative values occur in the numerical approximation. 
+Outside of this regions quantities are close to zero. Therefore, it is very likely that negative values occur in the numerical approximation.
 In Figure\,\ref{fig:sol_ADP27} it can be seen that at $T=27$ the two reaction interfaces merged.
@@
 In the first subplot the step size is plotted. For $t<25$ the timesteps are small. After $t = 30$ the step size increases, because the solution only evolves slowly afterwards. 
-In the second subplot the $min(u_1,u_2,u_3,u_4)$ is plotted for all timesteps that initially lead to negative values. This value is computed before and after adapting the weights.  
+In the second subplot the minimum of $u_1,u_2,u_3,u_4$ is plotted for all timesteps that initially lead to negative values. This value is computed before and after adapting the weights.
 We can see that relatively large negative values occurred at some timesteps. 
@@
         \includegraphics[width=\textwidth]{plots/ADP_sol_00.pdf}
-        \caption{$t=0$}
+        \caption{$t=0$.}
         \label{fig:sol_ADP00}
@@
         \includegraphics[width=\textwidth]{plots/ADP_sol_18.pdf}
-        \caption{$t=18$}
+        \caption{$t=18$.}
         \label{fig:sol_ADP18}
@@
         \includegraphics[width=\textwidth]{plots/ADP_sol_27.pdf}
-        \caption{$t=27$}
+        \caption{$t=27$.}
         \label{fig:sol_ADP27}
@@
         \includegraphics[width=\textwidth]{plots/ADP_sol_50.pdf}
-        \caption{$t=50$}
+        \caption{$t=50$.}
         \label{fig:sol_ADP50}
@@
 
-\todo{Note (HR)} Should the minimum of $u$ in Figure~\ref{fig:Stats_ADP} be negative before the adaptation?
+\todo{Note (HR)} Should the minimum of $u$ in Figure~\ref{fig:Stats_ADP} be negative before the adaptation? %TODO
 

### 2020-03-31 nusslein_vsrp/latex/overview.tex

--- 2020-03-31 latex/overview.tex
@@
 important advantage compared to orthogonal projection methods
-\cite{ranocha2020relaxationHamiltonian}.
+\cite{ranocha2020relaxationHamiltonian}. Of course, it is also
+possible to enforce the preservation of linear invariants in projection
+methods, but these invariants have to be known explicitly
+\cite{sandu2001positive}.
 
@@
 
+In contrast to other projection methods
+\cite{shampine1986conservation,sandu2001positive},
+minimizing the deviation of the weights instead of the deviation
+of the projected solution is computationally much more efficient
+for large systems, arising for example in the discretization of
+PDEs.
+
 \subsection{Example I}\label{sec:example_reac}
@@
 is it worth exploring this or is it better to just leave this open for further research it may be also interesting regardless the study of positivity)
+
+\todo{TODO} Is this approach used in the numerical experiments?
+If not, we could/should probably remove this part%TODO
 
@@
 reproduce the numerical examples reported here can be found in
-\cite{repository}. We have used CVXPY to solve the LPs
-\cite{cvxpy, cvxpy_rewriting}.
+\cite{repository}. \todo{TODO: New reopsitory with DOI} %TODO
+We have used CVXPY to solve the LPs \cite{cvxpy, cvxpy_rewriting}.
 \todo{TODO: Which solver of CVXPY?} %TODO

### 2020-03-31 nusslein_vsrp/latex/overview.tex
additional software references

@@
 \cite{repository}. \todo{TODO: New reopsitory with DOI} %TODO
+The methods are implemented in Python using NumPy/SciPy
+\cite{virtanen2019scipy}, NodePy \cite{nodepy08}, and
+Matplotlib \cite{hunter2007matplotlib} for the visualizations.
 We have used CVXPY to solve the LPs \cite{cvxpy, cvxpy_rewriting}.

### 2020-04-01 nusslein_vsrp/latex/overview.tex
adapt discussion of Example II

--- 2020-03-31 nusslein_vsrp/latex/overview.tex
@@
 
-(Note: The primary goal here is to put the idea of changing the b into perspective.
-We have shown the general idea and how it works, now we want to discuss on what it does in practice.
-At first, we state it is good to preserve the order...
-Get to the Example ...
-and show that there is more to it)
-
-The main goal is to choose a method $(A,b)$ in such a way that $u^{n+1}$ resembles the solution of the ODE $u(t_{n+1})$.
-An obvious objective is to use a $(A,b)$ with a high order, but this is not enough.
-To get a better understanding for the method we consider the behavior of a simple problem.
-
-We take the linear, positivity preserving ODE from \cite{kopecz_unconditionally_2018} with $u = [u_1,u_2]^T$ with the initial conditions $u_0 = [1,0]^T$ and
-
-\begin{equation}
-u'(t) = L u(t) \qquad L = \left[\begin{matrix}- 5 & 1\\5 & -1\end{matrix}\right] 
-\end{equation}
-and use the three stage and third order SSP method (SSP33)
+The main goal is to choose a method $(A,b)$ such that $u^{n+1}$
+approximates the solution of the ODE $u(t_{n+1})$.
+An obvious objective while modifying the Runge--Kutta coefficients
+is to retain a high order of accuracy, but this is not enough.
+To get a better understanding for the method we consider the
+behavior for a simple problem.
+
+We take the linear, positivity preserving ODE
+\cite{kopecz_unconditionally_2018}
+\begin{equation}
+  u'(t) = L u(t),
+  \quad
+  u(0) = \begin{pmatrix} 1 \\ 0 \end{pmatrix},
+  \qquad
+  L = \begin{pmatrix} -5 & 1 \\ 5 & -1\end{pmatrix},
+\end{equation}
+and use the three stage, third order SSP method SSP(3,3)
 \begin{align}
@@
 \end{align}
-The matrix $L$ has the eigenvalues $\lambda_1 = -6,\lambda_2=0$.
-The real stability interval of the SSP33 method includes $[-2.5,0]$.
-The method is absolut stable for this problem with $\dt = \frac{1}{3}$.
-The stage derivatives are
-\begin{equation}
-f(y_1)=\left(\begin{matrix}-5\\5\end{matrix}\right),\qquad f(y_2)=\left(\begin{matrix}5\\-5\end{matrix}\right),\qquad f(y_3)=\left(\begin{matrix}-5\\5\end{matrix}\right).
+of \cite{shu1988efficient}.
+The matrix $L$ has the eigenvalues zero and $-6$ and its operator
+norm is $2 \sqrt{13}$.
+The real stability interval of SSP33 includes $[-2.5,0]$. Hence,
+it seems reasonable\footnote{A more careful analysis following
+\cite{ranocha2018L2stability} shows that the method is not
+absolutely stable for this problem and choice of $\dt$ since
+$L$ is not normal.} to choose $\dt = \nicefrac{1}{3}$.
+The corresponding stage derivatives are
+\begin{equation}
+  f(y_1) = \begin{pmatrix} -5 \\ 5 \end{pmatrix},\quad
+  f(y_2) = \begin{pmatrix} 5 \\ -5 \end{pmatrix},\quad
+  f(y_3) = \begin{pmatrix} -5 \\ 5 \end{pmatrix}.
 \end{equation}
@@
 \begin{equation}
-u_n=\left(\begin{matrix} \nicefrac{-1}{9}\\ \nicefrac{10}{9}\end{matrix}\right)
-\end{equation}
-The new solution contains negative values. 
-Therefore, we want to adapt the weights to ensure positivity. 
-All weights that comply with the order condition for the first and second order can be expressed as
-\begin{equation}
-\bt=\begin{pmatrix}\nicefrac{1}{6}\\ \nicefrac{1}{6}\\ \nicefrac{2}{3}\end{pmatrix}+\alpha \begin{pmatrix}\nicefrac{1}{2}\\\nicefrac{1}{2}\\-1\end{pmatrix}.
+  u^1 = \begin{pmatrix} \nicefrac{-1}{9} \\ \nicefrac{10}{9} \end{pmatrix}.
+\end{equation}
+Since the first component of the new solution is negative,
+we want to adapt the weights to ensure positivity.
+All weights that comply with the constraints for first and second
+order of accuracy can be expressed as
+\begin{equation}
+  \bt =
+  \begin{pmatrix}
+    \nicefrac{1}{6} \\
+    \nicefrac{1}{6} \\
+    \nicefrac{2}{3}
+  \end{pmatrix}
+  + \alpha \begin{pmatrix}
+    \nicefrac{1}{2} \\
+    \nicefrac{1}{2} \\
+    -1
+  \end{pmatrix},
+  \qquad
+  \alpha \in \R.
 \end{equation}
@@
 parameterized by $\alpha$.
-If the general expression for the weights is inserted in \eqref{eq:Combination} the general solution is 
-\begin{equation}
-u^{1} = u^0 + \dt  \left(f_1, f_2, f_3\right) \bt =
-\begin{pmatrix} \nicefrac{-1}{9}\\\nicefrac{10}{9}\end{pmatrix}
-+\alpha \begin{pmatrix}\nicefrac{5}{3}\\ \nicefrac{-5}{3}\end{pmatrix}
-\end{equation}
-By changing the parameter $\alpha$, the weights and the solution are altered.
-With a suitable choice of $\alpha \in \left[\nicefrac{1}{15},\nicefrac{2}{3}\right]$ any $u$ that complies with mass conservation and positivity can be reached.
-By adding additional constraints on the weights, the choice of $\alpha$ can be narrowed down. 
+If the general expression for the weights is inserted in
+\eqref{eq:Combination} the general solution is
+\begin{equation}
+  u^{1}
+  =
+  u^0 + \dt \left(f_1, f_2, f_3\right) \bt
+  =
+  \begin{pmatrix}
+    \nicefrac{-1}{9} \\
+    \nicefrac{10}{9}
+  \end{pmatrix}
+  +\alpha \begin{pmatrix}
+    5 \\
+    -5
+  \end{pmatrix}.
+\end{equation}
+By changing the parameter $\alpha$, the weights and the new
+solution are altered. With a suitable choice of
+$\alpha \in \left[ \nicefrac{1}{45}, \nicefrac{2}{9} \right]$,
+any $u$ that complies with mass conservation and positivity can
+be reached.
+By adding additional constraints on the weights, the choice of
+$\alpha$ can be narrowed down.
 An objective function is also needed to make the choice unique. This should be designed in a way to prefer weights that are close to the original weights.
-
-%This gives raise to the question whether the new solution is reasonable.
-%This also poses the problem on developing a system that can decide if the new solution is has to be rejected.
-
-\todo{Note (HR)} %TODO
-This example has to be fixed.
-Since the matrix is not normal, eigenvalue analysis
-and classical linear stability theory do not apply. By a direct
-computation, we can see that SSP33 is absolutely stable for this
-initial condition and the first step for $\dt \lesssim 0.309$
-but not for $\dt = \nicefrac{1}{3}$.
 

### 2020-04-01 latex/overview.tex
additional TODO note
Changed Plots according to https://ranocha.de/blog/colors/#python-code, Added Order information

--- 2020-03-31 latex/overview.tex
@@
 \addbibresource{Bib_RK_optim.bib}
-
-\usepackage{booktabs}
-\usepackage{multirow}
 
@@
 \centering    %Generated below============ 
-% \begin{tabular}{|l c|c c c c c c |}
-%  \hline
-% Order & &1&2&3&4&5&6 \\
-%  \hline Classical RK4& s=4&3&2&0&0& - & -  \\
-%  SSPRK(10,4)& s=10&9&8&6&4& - & -  \\
-%  Cash-Karp RK5(4)6& s=6&5&4&2&1&0& -  \\
-%  Dormand-Prince RK5(4)7& s=7&6&5&3&1&0& -  \\
-%  \hline
-%  \end{tabular}
-  \begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}lr*6c@{}}
-    \toprule
-    Method & $s$ & \multicolumn{6}{c}{Order $\tilde p$} \\
-    & & 1 & 2 & 3 & 4 & 5 & 6 \\
-    \midrule
-    Classical RK4 \cite{kutta1901beitrag} & 4 & 3 & 2 & 0 & 0 & --- & --- \\
-    SSPRK(10,4) \cite{ketcheson2008highly} & 10&9&8&6&4& --- & ---\\
-    Cash--Karp RK5(4)6 \cite{cash1990variable} & 6&5&4&2&1&0& --- \\
-    Dormand--Prince RK5(4)7 \cite{prince1981high}& 7&6&5&3&1&0& --- \\
-    \bottomrule
-  \end{tabular*}
+\begin{tabular}{|l c|c c c c c c |} 
+ \hline 
+Order & &1&2&3&4&5&6 \\ 
+ \hline Classical RK4& s=4&3&2&0&0& - & -  \\ 
+ SSPRK(10,4)& s=10&9&8&6&4& - & -  \\ 
+ Cash-Karp RK5(4)6& s=6&5&4&2&1&0& -  \\ 
+ Dormand-Prince RK5(4)7& s=7&6&5&3&1&0& -  \\ 
+ \hline 
+ \end{tabular}
  \caption{Degrees of freedom for the choice of the weights for some explicit methods.} %Generated above============
@@
 \centering   %Generated below============ 
-%   \begin{tabular}{|l c|c c c c c c |}
-%   \hline
-%   Order & &1&2&3&4&5&6 \\
-%   \hline Implicit Euler& s=1&0& - & - & - & - & -  \\
-%   Lobatto~IIIC4& s=4&3&2&1&0&0&0 \\
-%   Radau~IIA3& s=3&2&1&0&0&0& -  \\
-%   SDIRK(5,4)& s=5&4&3&1&0& - & -  \\
-%   TR-BDF2& s=3&2&1& - & - & - & -  \\
-%   Im-Euler 2& s=3&2&1& - & - & - & -  \\
-%   Im-Euler 3& s=6&5&4&2& - & - & -  \\
-%   Im-Euler 4& s=10&9&8&6&3& - & -  \\
-%   \hline
-%   \end{tabular}
-   \begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}lr*6c@{}}
-    \toprule
-    Method & $s$ & \multicolumn{6}{c}{Order $\tilde p$} \\
-    & & 1 & 2 & 3 & 4 & 5 & 6 \\
-    \midrule
-    Implicit Euler& 1&0& --- & --- & --- & --- & ---  \\
-    Lobatto~IIIC4 \cite{chipman1971stable} & 4&3&2&1&0&0&0 \\
-    Radau~IIA3 \cite{ehle1969pade} & 3&2&1&0&0&0& ---  \\
-    SDIRK(5,4) \cite[eq. (6.18)]{hairer_solving_1996}& 5&4&3&1&0& --- & ---  \\
-    TR-BDF2 \cite{bank1985transient} & 3&2&1& --- & --- & --- & ---  \\
-    Extrapolation Im-Euler 2& 3&2&1& --- & --- & --- & ---  \\
-    Extrapolation Im-Euler 3& 6&5&4&2& --- & --- & ---  \\
-    Extrapolation Im-Euler 4& 10&9&8&6&3& --- & ---  \\
-    \bottomrule
-  \end{tabular*}
+\begin{tabular}{|l c|c c c c c c |} 
+ \hline 
+Order & &1&2&3&4&5&6 \\ 
+ \hline Implicit Euler& s=1&0& - & - & - & - & -  \\ 
+ Lobatto~IIIC4& s=4&3&2&1&0&0&0 \\
+ Radau~IIA3& s=3&2&1&0&0&0& -  \\
+ SDIRK 54& s=5&4&3&1&0& - & -  \\ 
+ TR-BDF2& s=3&2&1& - & - & - & -  \\ 
+ Im-Euler 2& s=3&2&1& - & - & - & -  \\ 
+ Im-Euler 3& s=6&5&4&2& - & - & -  \\ 
+ Im-Euler 4& s=10&9&8&6&3& - & -  \\ 
+ \hline 
+ \end{tabular}
  \caption{Degrees of freedom for the choice of the weights for some implicit methods.} %Generated above============
@@
 We can see that the fully implicit methods Lobatto~IIIC4 and Radau~IIA3 require a drastic reduction of the order as expected.
-The diagonally implicit SDIRK(5,4) method only requires an order reduction of one to get one degree of freedom for the weights.
+The diagonally implicit SDIRK 54 method only requires an order reduction of one to get one degree of freedom for the weights.
 The TR-BDF2 method even allows adaptations without reducing the order.  
@@
 \centering
-\includegraphics[scale=0.75]{plots/conv_adde.pdf}
+\includegraphics[width=0.85\textwidth]{plots/conv_adde.pdf}
 \caption{Convergence adapted dp5}
@@
 \centering
-\includegraphics[scale=0.75]{plots/conv_heat.pdf}
+\includegraphics[width=0.85\textwidth]{plots/conv_heat.pdf}
 \caption{Convergence adapted ex3}
@@
 
-\todo{Note (HR)} Should the minimum of $u$ in Figure~\ref{fig:Stats_ADP} be negative before the adaptation?
-
 \section{Conclusion} \label{sec:conclusion}

### 2020-04-01 utils.py

@@ -12,30 +12,32 @@ import matplotlib.pyplot as plt
 from RKimple import RK_integrate
 
 
-def plot_convergence(problem,solver,dt,refernce,step=1,error='abs',dx='1',Norm = 2,Params = {}):
+def plot_convergence(problem,solver,dt,refernce,step=1,error='abs',dx='1',Norm = 2,Params = {},get_order=False):
     """"
     Parameters:
-    problem:            Problem Objects 
+    problem:            Problem Objects
     solver:             Solver object
     dt:                 dt array with dts
-    reference:          Array with reference solutions to compare the computet solution against 
+    reference:          Array with reference solutions to compare the computet solution against
     error:              Definition of error computation, one of 'abs','rel','grid'
     dx:                 Discretisation for grid norm
     Norm:               Norm to use for Error calculation ||u-u'||_Norm
     Params:             Parameters for time integrator
-    
+    get_order:          If True: also returns list with lowest orders
+
     Return:
     sol:                Array with the solutions used for calculationg the errors
     err:                Array with errors
     change:             Array, True if b was changed False otherwise
 
     """
-    
+
     err = np.zeros_like(dt)
     change = np.zeros_like(dt,dtype=bool)
+    order = np.zeros_like(dt)
     sol = []
-    
-    
+
+
     for i in range(dt.size):
         print('dt='+str(dt[i]))
         solver.dt = dt[i]
@@ -53,29 +55,39 @@ def plot_convergence(problem,solver,dt,refernce,step=1,error='abs',dx='1',Norm =
                 print('Error not defined')
                 print(error)
                 raise ValueError
+
+            orders = np.array(status['order'])[np.array(status['order'])!=None]
+            if len(orders)>0:
+                order[i]=np.min(orders)
+            else:
+                order[i] = np.nan
         else:
             err[i] = np.nan
+            order[i] = np.nan
         sol.append(u[step])
-        
+
     plt.plot(dt,err,'o-')
     plt.yscale('log')
     plt.xscale('log')
     plt.grid()
     plt.ylabel('Error')
     plt.xlabel('dt')
-    
-    return sol,err,change
-    
-        
+
+    if get_order:
+        return sol,err,change,order
+    else:
+        return sol,err,change
+
+
 
[…]

### 2020-04-01 nusslein_vsrp/latex/overview.tex
update scipy reference
cite mosek

--- 2020-04-01 nusslein_vsrp/latex/overview.tex
@@
 reproduce the numerical examples reported here can be found in
-\cite{repository}. \todo{TODO: New reopsitory with DOI} %TODO
-We have used CVXPY to solve the LPs \cite{cvxpy, cvxpy_rewriting}.
-\todo{TODO: Which solver of CVXPY?} %TODO
+\cite{repository}. \todo{TODO: New repository with DOI} %TODO
+The methods are implemented in Python using NumPy/SciPy
+\cite{virtanen2020scipy}, NodePy \cite{nodepy08}, and
+Matplotlib \cite{hunter2007matplotlib} for the visualizations.
+We have used MOSEK \cite{mosek} via CVXPY to solve the LPs
+\cite{cvxpy, cvxpy_rewriting}.
 

### 2020-04-01 nusslein_vsrp/latex/overview.tex
Cleanup Merge

--- 2020-04-01 nusslein_vsrp/latex/overview.tex
@@
 
-The main goal is to choose a method $(A,b)$ such that $u^{n+1}$
-approximates the solution of the ODE $u(t_{n+1})$.
-An obvious objective while modifying the Runge--Kutta coefficients
-is to retain a high order of accuracy, but this is not enough.
-To get a better understanding for the method we consider the
-behavior for a simple problem.
-
-We take the linear, positivity preserving ODE
-\cite{kopecz_unconditionally_2018}
-\begin{equation}
-  u'(t) = L u(t),
-  \quad
-  u(0) = \begin{pmatrix} 1 \\ 0 \end{pmatrix},
-  \qquad
-  L = \begin{pmatrix} -5 & 1 \\ 5 & -1\end{pmatrix},
-\end{equation}
-and use the three stage, third order SSP method SSP(3,3)
+(Note: The primary goal here is to put the idea of changing the b into perspective.
+We have shown the general idea and how it works, now we want to discuss on what it does in practice.
+At first, we state it is good to preserve the order...
+Get to the Example ...
+and show that there is more to it)
+
+The main goal is to choose a method $(A,b)$ in such a way that $u^{n+1}$ resembles the solution of the ODE $u(t_{n+1})$.
+An obvious objective is to use a $(A,b)$ with a high order, but this is not enough.
+To get a better understanding for the method we consider the behavior of a simple problem.
+
+We take the linear, positivity preserving ODE from \cite{kopecz_unconditionally_2018} with $u = [u_1,u_2]^T$ with the initial conditions $u_0 = [1,0]^T$ and
+
+\begin{equation}
+u'(t) = L u(t) \qquad L = \left[\begin{matrix}- 5 & 1\\5 & -1\end{matrix}\right] 
+\end{equation}
+and use the three stage and third order SSP method (SSP33)
 \begin{align}
@@
 \end{align}
-of \cite{shu1988efficient}.
-The matrix $L$ has the eigenvalues zero and $-6$ and its operator
-norm is $2 \sqrt{13}$.
-The real stability interval of SSP33 includes $[-2.5,0]$. Hence,
-it seems reasonable\footnote{A more careful analysis following
-\cite{ranocha2018L2stability} shows that the method is not
-absolutely stable for this problem and choice of $\dt$ since
-$L$ is not normal.} to choose $\dt = \nicefrac{1}{3}$.
-The corresponding stage derivatives are
-\begin{equation}
-  f(y_1) = \begin{pmatrix} -5 \\ 5 \end{pmatrix},\quad
-  f(y_2) = \begin{pmatrix} 5 \\ -5 \end{pmatrix},\quad
-  f(y_3) = \begin{pmatrix} -5 \\ 5 \end{pmatrix}.
+The matrix $L$ has the eigenvalues $\lambda_1 = -6,\lambda_2=0$.
+The real stability interval of the SSP33 method includes $[-2.5,0]$.
+The method is absolut stable for this problem with $\dt = \frac{1}{3}$.
+The stage derivatives are
+\begin{equation}
+f(y_1)=\left(\begin{matrix}-5\\5\end{matrix}\right),\qquad f(y_2)=\left(\begin{matrix}5\\-5\end{matrix}\right),\qquad f(y_3)=\left(\begin{matrix}-5\\5\end{matrix}\right).
 \end{equation}
@@
 \begin{equation}
-  u^1 = \begin{pmatrix} \nicefrac{-1}{9} \\ \nicefrac{10}{9} \end{pmatrix}.
-\end{equation}
-Since the first component of the new solution is negative,
-we want to adapt the weights to ensure positivity.
-All weights that comply with the constraints for first and second
-order of accuracy can be expressed as
-\begin{equation}
-  \bt =
-  \begin{pmatrix}
-    \nicefrac{1}{6} \\
-    \nicefrac{1}{6} \\
-    \nicefrac{2}{3}
-  \end{pmatrix}
-  + \alpha \begin{pmatrix}
-    \nicefrac{1}{2} \\
-    \nicefrac{1}{2} \\
-    -1
-  \end{pmatrix},
-  \qquad
-  \alpha \in \R.
+u_n=\left(\begin{matrix} \nicefrac{-1}{9}\\ \nicefrac{10}{9}\end{matrix}\right)
+\end{equation}
+The new solution contains negative values. 
+Therefore, we want to adapt the weights to ensure positivity. 
+All weights that comply with the order condition for the first and second order can be expressed as
+\begin{equation}
+\bt=\begin{pmatrix}\nicefrac{1}{6}\\ \nicefrac{1}{6}\\ \nicefrac{2}{3}\end{pmatrix}+\alpha \begin{pmatrix}\nicefrac{1}{2}\\\nicefrac{1}{2}\\-1\end{pmatrix}.
 \end{equation}
@@
 parameterized by $\alpha$.
-If the general expression for the weights is inserted in
-\eqref{eq:Combination} the general solution is
-\begin{equation}
-  u^{1}
-  =
-  u^0 + \dt \left(f_1, f_2, f_3\right) \bt
-  =
-  \begin{pmatrix}
-    \nicefrac{-1}{9} \\
-    \nicefrac{10}{9}
-  \end{pmatrix}
-  +\alpha \begin{pmatrix}
-    5 \\
-    -5
-  \end{pmatrix}.
-\end{equation}
-By changing the parameter $\alpha$, the weights and the new
-solution are altered. With a suitable choice of
-$\alpha \in \left[ \nicefrac{1}{45}, \nicefrac{2}{9} \right]$,
-any $u$ that complies with mass conservation and positivity can
-be reached.
-By adding additional constraints on the weights, the choice of
-$\alpha$ can be narrowed down.
+If the general expression for the weights is inserted in \eqref{eq:Combination} the general solution is 
+\begin{equation}
+u^{1} = u^0 + \dt  \left(f_1, f_2, f_3\right) \bt =
+\begin{pmatrix} \nicefrac{-1}{9}\\\nicefrac{10}{9}\end{pmatrix}
++\alpha \begin{pmatrix}\nicefrac{5}{3}\\ \nicefrac{-5}{3}\end{pmatrix}
+\end{equation}
+By changing the parameter $\alpha$, the weights and the solution are altered.
+With a suitable choice of $\alpha \in \left[\nicefrac{1}{15},\nicefrac{2}{3}\right]$ any $u$ that complies with mass conservation and positivity can be reached.
+By adding additional constraints on the weights, the choice of $\alpha$ can be narrowed down. 
 An objective function is also needed to make the choice unique. This should be designed in a way to prefer weights that are close to the original weights.
+
+%This gives raise to the question whether the new solution is reasonable.
+%This also poses the problem on developing a system that can decide if the new solution is has to be rejected.
+
+\todo{Note (HR)} %TODO
+This example has to be fixed.
+Since the matrix is not normal, eigenvalue analysis
+and classical linear stability theory do not apply. By a direct
+computation, we can see that SSP33 is absolutely stable for this
+initial condition and the first step for $\dt \lesssim 0.309$
+but not for $\dt = \nicefrac{1}{3}$.
 
@@
 \centering
-\includegraphics[scale=0.6]{plots/conv_adde.pdf}
+\includegraphics[width=0.85\textwidth]{plots/conv_adde.pdf}
 \caption{Convergence results of the adapted Dormand--Prince RK5.}
@@
 \centering
-\includegraphics[scale=0.75]{plots/conv_heat.pdf}
+\includegraphics[width=0.85\textwidth]{plots/conv_heat.pdf}
 \caption{Convergence result of baseline and adapted Im-Euler~3
@@
 
-\todo{Note (HR)} Should the minimum of $u$ in Figure~\ref{fig:Stats_ADP} be negative before the adaptation? %TODO
-
 \section{Conclusion} \label{sec:conclusion}

### 2020-04-02 nusslein_vsrp/latex/overview.tex
Revisions when reading the text,
Order Conditions 
Sum constraint in LP
Text hatching; closes #2
Remove TODO flowchart

--- 2020-04-01 nusslein_vsrp/latex/overview.tex
@@
 We investigate different approaches to choosing the weights by considering
-different objective functions.  We also provide some analysis of the properties
+adding further constraints.
+We also provide some analysis of the properties
 of Runge--Kutta methods with perturbed weights.  Numerical examples demonstrate
@@
 linear in $b$ and can be written as $Q_p b = r_p$, where
-and $Q_p\in{\mathbb R}^{s\times s}, r_p\in{\mathbb R}^s$ represent the set of all conditions up to
-and including order $p$.  It may not be possible to find modified weights
+$Q_p\in{\mathbb R}^{v\times s}, r_p\in{\mathbb R}^v$ represent the set of all conditions up to
+and including order $p$. Here $v$ is the number of order conditions. It may not be possible to find modified weights
 that also satisfy the conditions of order $p$ and yield positivity, so
@@
 0 & \le g_i \le 1, \\
+\sum_i g_i &= 1 \\
 u^{n+1}&=u^n+\dt F \tilde{b} \geq 0.
@@
 If the classical RK4 method is used the order has to be reduced more because the RK4 method does not have embedded methods of order~3.
-In contrast to this, methods with $s > p$ have degrees of freedom even without reducing the order.
+In contrast to this, some methods with $s > p$ have degrees of freedom even without reducing the order.
+An example of this is SSPRK(10,4), that has 4 degrees of freedom for $\pt = p$.
 For Cash--Karp RK5 and Dormand--Prince RK5 the order has to be reduced to get degrees of freedom even though the number of stages is higher than the order.
@@
         \caption{Change of stability region for the free adaptation and
-                 the convex adaptation of the weights.}
+                 the convex adaptation of the weights. 
+                 Regions with $|R(z)| \leq 1$ are hatched for the original methods.}
         \label{fig:stab}
@@
 \subsection{Implemention}
-(\todo{TODO}: Maybe a flowchart and a description on how we stitch everything together.) %TODO
 
@@
 
-\todo{Note (HR)} I think this should be moved to the beginning.
-
 \subsection{Explicit methods}\label{sec:Ex_expl}
@@
 \centering
-\includegraphics[scale=0.6]{plots/conv_adde.pdf}
+\includegraphics[width=0.8\textwidth]{plots/conv_adde.pdf}
 \caption{Convergence results of the adapted Dormand--Prince RK5.}
@@
 \centering
-\includegraphics[scale=0.75]{plots/conv_heat.pdf}
+\includegraphics[width=0.8\textwidth]{plots/conv_heat.pdf}
 \caption{Convergence result of baseline and adapted Im-Euler~3

### 2020-04-07 nusslein_vsrp/latex/overview.tex
comment of @ketch and some minor fixes

@@
 \cite{ranocha2018L2stability} shows that the method is not
-absolutely stable for this problem and choice of $\dt$ since
+monotone (energy stable) for this problem and choice of $\dt$ since
 $L$ is not normal.} to choose $\dt = \nicefrac{1}{3}$.
@@
 In the worst case $h_a$ is growing very slowly until $h_a = \{1,\cdots,m \}$.
-This worst case did not occur when runing the testproblems. 
+This worst case did not occur when running the test problems.
 %(Note: A approach on this would be to make a educated guess which indices could get a problem with positivity. A possible estimate would be $\frac{u_i^{n+1}}{max(K_{(i,0)}, \cdots ,K_{(i,0)})} $ )
@@
 For problems where the exact solution is positive for certain $u_0$ but positivity is not preserved for all $u_0 \geq 0$ tests did not show promising results.
-Additionaly it is not certain if the computed solutions would be reasonable.
+Additionally, it is not certain whether the computed solutions would be reasonable.
 

### 2020-04-08 nusslein_vsrp/latex/overview.tex
Pseudocode for algorithm

@@
 \newcommand{\todo}[1]{{\Large{\color{red}{#1}}}}
+
+%pseudocode
+\usepackage[boxed]{algorithm}
+\usepackage{algpseudocode}
 
@@
 
+In this section an algorithm to update the solution $u^n$ using the adaptive RKM is proposed.
+The pseudocode is given in Algorithm\,\ref{alg:Adaption}.
+t
 At first, the stage values $y_i$ and stage derivatives $f_i$ are calculated.
@@
 
-Start with the highest order $\tilde p$ such that there are degrees
-of freedom for the weights $\bt$. Apply an LP solver to the optimization
+Start with $\tilde{p}$ equal to the highest order $p_{start}$ such that there are degrees of freedom for the weights $\bt$. 
+Apply an LP solver to the optimization
 problem with accuracy and positivity constraints.
@@
 If the reduction of stepsize is used this process is done first for $\theta = 1$. If no weights were found, the process is repeated for a smaller $\theta$.
+
+\begin{algorithm}
+\begin{algorithmic}[1]
+\State Caclculate Stagedrivatives
+\State $u^{n+1}=\Delta t Fb$
+\If {$u^{n+1} \geq 0$} 
+        \State Update and End
+\Else
+	\ForAll{$\tilde p \in [p_{start}:p_{min}]$} 
+		\State Solve LP
+		\If {LP Feasible}
+			\State caclulate $err_{adapt}$
+			\If {$err_{adapt} < tol$}
+				\State $u^{n+1}=\Delta t F \tilde b$
+				\State Update and End
+			\EndIf 
+		\EndIf
+	\EndFor
+	\State Reject step and End
+\EndIf 
+\end{algorithmic}
+\caption{Adaption of weights}
+\label{alg:Adaption}
+\end{algorithm}
 

### 2020-04-26 nusslein_vsrp/latex/overview.tex
Changes from meeting
Corrected mutliple minor things from issue #17
Added the mass conservation for the ADP-Problem

@@
 \newcommand{\pt}{{\tilde{p}}}
+\newcommand{\perturb}{perturb}
 \newcommand{\todo}[1]{{\Large{\color{red}{#1}}}}
@@
 \end{align}
+A sufficient condition for positivity of an ODE is 
+\begin{equation} \label{eq:condition_ODE_pos}
+u_i=0 \implies f_i(t,u_1,\cdots,u_i,\cdots,u_n) \geq 0 \quad \forall {u_c \geq 0}\; \forall {t} 
+\end{equation} 
+For such ODEs a step using the forward Euler method will lead to a positive result for small $\dt$.
+A step using the backward Euler method will lead to a positive result for any $\dt$.
 Any RKM (or in fact any general linear method) that is unconditionally positivity preserving for all positive ODEs
@@
 Another class of methods that adapt the weights at the end of an RK step
-are the relaxation Runge–Kutta (RRK) methods. 
+are the relaxation Runge--Kutta (RRK) methods. 
 In these, the weights are scaled by a scalar relaxation parameter in order to guarantee
@@
 the stage values $y_1,\cdots,y_s$ are computed according to
-\begin{equation}
+\begin{equation}\label{eq:stagevalues}
 y_j =  u^n + \dt \sum_{k = 1}^{s} a_{jk} f(t^n + \dt c_k,y_k),  \quad j = 1,\cdots,s.
@@
 where the $j$th column of $F$ is $f_j$.
-
 We wish to impose the discrete analog of \eqref{continuous-positivity}; i.e.
@@
 stability, small error coefficients, and so forth.
-Replacing these carefully-chosen weights $b$ with arbitary weights
+Replacing these carefully-chosen weights $b$ with arbitrary weights
 $\bt$ could lead to the loss of these desirable properties.
@@
         and restrict the choice of $\bt$ to convex combinations of this set.
-    \item Require that $\|\bt - b\|_1$ be small, rejecting the step if it is not.
+    \item Require that the perturbation $\|\tilde u - u\|$ is small and reject the step if it is not.
 \end{enumerate}
 
-We discuss the first strategy here; the second is deferred to a later section.
+We discuss the first strategy here; the second is deferred to section\,\ref{sec:error}.
 Ideally every element of the set of potential weight vectors would correspond to
@@
 
-\subsection{Adaption of the timestep using the dense output formula}
-
-{\bf Need to also discuss relaxing the order conditions}
-
-If the original step $\dt$ was too large and there is no suitable set of weights, the LP is infeasible.
-In this case we would like to calculate $u^{n+\theta}$ using the stage values instead of rejecting the step.
-The time step is reduced after computing the stage values.
-This can be done using the order conditions for dense output \cite[Section~II.6]{hairer_solving_1993}.
-We introduce $\theta \in (0,1]$. The new step taken has the length $\theta \dt$, where $\dt$ is the time step used when calculating the stage values.
-To calculate a reduced step the order conditions have to be adapted and a new $b(\theta)$ for the objective function has to be generated. 
-When adapting the order condition to a new $\theta$ the order condition matrix $Q$ stays unchanged whereas the right hand side $r(\theta)$ depends on $\theta$.
-
-The dependency of $r(\theta)$ is nonlinear. This means that the maximum $\theta$ with suitable weights cannot be calculated by solving a LP. This means that different $\theta$ have to be tested. 
-%(Note: we could include some algorithm to search for it)
-
-\subsubsection{Construction of embedded methods}
-
-As a first order method, the weights that correspond to a series of backward Euler steps are added. These are known to yield a positive result for any step size.
-It is also possible to add embedded methods of higher order.
-An important property of the embedded methods is that they should show the desired stability characteristics.
-
-A reduction of the step size using the order conditions for dense output would also be possible.
-For this a set of embedded methods $b_1(\theta),\cdots,b_d(\theta)$ has to be constructed.
-This was not the aim of this research.
-
-(Note: Doing dense Output here is a bit complicated, but it would not help ensure positivity only (maybe) increase exactness.
-is it worth exploring this or is it better to just leave this open for further research it may be also interesting regardless the study of positivity)
-
-\todo{TODO} Is this approach used in the numerical experiments?
-If not, we could/should probably remove this part%TODO
-
 \section{Properties of adaptive RKM} \label{sec:integration}
@@
 In the previous sections an algorithm for choosing positivity preserving weights $\tilde{b}$ has been presented.
-Here, we explain how this algorithm can be integrated into the framework of existing solvers.
+In the next section properties of the adaptive RKM are discussed.
 
@@
 
-\subsection{Error detection and approximation}
-When changing the weights the solution $u^{n+1}$ is changed. 
-By incorporating the order conditions for a certain order the space of possible weights that can be chosen is limited to an affine subspace.
-Now it is important to know if the new solution $u^{n+1}$ still resembles the solution of the ODE $u(t_{n+1})$. 
-This can be seen in the example in Section~\ref{sec:example_lin}.
-This gives rise to the question whether the new solution is reasonable.
-This also poses the problem on developing a system that can decide whether the new solution is acceptable or has to be rejected.
-
-For this approach it is difficult to give some mathematically rigorous statement regarding the convergence of the method, because it only applies if the error of the original method causes the solution to get negative.
-This can only happen if the truncation error is large enough.
-This requires $\dt \gg 0$ where lower coefficients of the Taylor series are no longer sufficient to approximate the behavior.
-One could say that the main idea of the adaption of the weights alters the error in a certain way, so that the numerical solution gets 'better' in the sense that it leads to positive solutions. 
-The local error of an RKM $e(\dt) =u(t_0 + \dt) - u^1$ can be expressed using the Taylor series of the RKM and the Taylor series of the exact solution, % Hairer I P134
-\begin{align}\label{eq:Taylor_sol_ref}
-u(t_0 + \dt) &= u(t_0) + u'(t_0) \dt + \frac{u''(t_0)}{2} \dt^2 + \cdots + \frac{u^{(p)}(t_0)}{p!} \dt^p + \frac{u^{(p+1)}(t_0)}{(p+1)!} \dt^{p+1} + \cdots, \\
-u^{n+1} &= u(t_0)  + v_1 \dt + \frac{v_2}{2} \dt^2 + \cdots + \frac{v_p}{p!} \dt^p + \frac{v_{p+1}}{(p+1)!} \dt^{p+1} + \cdots .
-\end{align}
-The derivatives of $u(t)$ can be solely expressed with the partial derivatives of the RHS $f(t,u)$, whereas the coefficients $v_1,v_2,\cdots$ also depend on the RKM. 
-In this case the RKM is fixed apart from the weights. 
-Therefore $v_1,v_2,\cdots$ are functions of $b$.
-Because the weights are chosen to ensure the order conditions, the coefficients $v_1,\cdots,v_p$ are fixed and equal the according derivatives.
-The only remaining degrees of freedom are the coefficients $v_{p+1},v_{p+2},\cdots$.
-These can be adapted by changing the weights.
-
+\subsection{Error detection and approximation}\label{sec:error}
+For this approach it is difficult to give some mathematically rigorous statement regarding the convergence of the method, because the RKM is only adapted if the step sizes are large. 
+For sufficiently small step sizes the solution of the unadapted RKM is positive. Therefore the original weights are used.
+We have to describe the behavior outside the asymptotic regime.
 %The main way to ensure that the RKM with adapted b converges to the correct solution is to ensure that the weights approaches $b_{orig}$ for $\dt \to 0$.
@@
 %\end{equation}
-This approximation is not very useful for real applications. 
 To approximate the error of a new step we propose the following approximation of the local error:
@@
 \begin{align}
-err = |u(t^{n+1})-u^{n+1}| &= |u(t^{n+1}) - (u^{n+1}_{unadapted}+\dt F(\tilde{b}-b))| \\
- &\leq \underbrace{|u(t^{n+1})-u^{n+1}_{unadapted}|}_{\approx err_T}+\underbrace{|\dt F(\tilde{b}-b)|}_{= err_{adapt}} \label{eq:Err}
+err = \|u(t^{n+1})-\tilde u^{n+1}\| &= \|u(t^{n+1}) - (u^{n+1}+\dt F(\tilde{b}-b))\| \\
+ &\leq \underbrace{\|u(t^{n+1})-u^{n+1}\|}_{\approx err_T}+\underbrace{\|\dt F(\tilde{b}-b)\|}_{= perturb} \label{eq:Err}
 \end{align}
 
-The total error is split up in the truncation error $err_T$ and the adaption error $err_{adapt}$ using the triangle inequality. 
-The truncation error can be estimated using the standard error estimators $err_T = | u^{n}_{b_{orig}} - u^{n}_{\hat{b}} |$. 
-After adapting the weights, the adaption error is calculated. If the adaption error is larger than the tolerance the weights are rejected.
-Both errors are added to get an approximation of the total error $err = err_T + err_{adapt}$.
+The total error is split up in the truncation error $err_T$ and the perturbation $perturb$ using the triangle inequality. 
+The truncation error can be estimated using the standard error estimators $err_T = | u^{n}_{b} - u^{n}_{\hat{b}} |$. 
+After adapting the weights, the perturbation is calculated. If the perturbation is larger than the tolerance the weights are rejected.
+Both values are added to get an approximation of the total error $err = err_T + perturb$.
 % The truncation error and the adaption error never have the same direction.
@@
 function is altered and the region of absolute stability varies.
-How often and how much the stability function is changed depends
-of course on the problem, the RKM, and the timestep $\dt$.
-For problems where only a small number of steps are affected, a bigger
-change in the stability function might be acceptable. For problems
-that require an adaptation of the weights for every step, one needs
-to make sure that the resulting method is stable.
-
-The stability regions of adapted RKMs are visualized in
+
+An an example, the stability regions of adapted RKMs are visualized in
 Figure\,\ref{fig:stab}.
@@
 \begin{theorem}
-  The stability region of a Runge--Kutta method $(A,b)$ where
-  $b = \sum_{i} g_i b^i$ is a convex combination of
+  The stability region of a Runge--Kutta method $(A,\bt)$ where
+  $\bt = \sum_{i} g_i b^i$ is a convex combination of
   $b^1, \dots, b^m \in \R^s$ (i.e.\ $g_i \in [0,1]$, $\sum_i g_i = 1$),
@@
 If the weights is adapted freely there is no such proof.
-A useful property of the resulting stability region is that it is similar to the stability region of the baseline method. 
-(Note: is it possible (or needed at all) to give some statement like: If the stability regions of the RKM used for multiple steps are similar, then the overall method is stable for the intersection of the stability regions.)
-
-For most methods, a small change of the weights does not alter the stability function dramatically.
-(TODO: I am missing a piece here from ‘the derivative is finite’ to ‘is similar’)
-
-We would like to calculate the derivative of the border in respect to a change of the weights.
-
-The stability function depending on the weights is expressed. 
-When adapting the weights we cannot change them independently but only in ways that comply with the order conditions.
-$m$ is the number of dimensions of the kernel of the order condition $m = \mathrm{dim}\{\mathrm{Ker} (Q) \}$
-The $b_1,\cdots,b_m$  are the basis vectors of $\mathrm{Ker} (Q)$. With a coordinate transformation we get the coordinate vector $\hat{b}$ with the matrix $B = [b_1,\cdots,b_m]$ and $b = B \hat{b}$.
-
-The general stability function of all embedded RKM that satisfy the order conditions $Q$ is given by
-
-\begin{equation}\label{eq:gen_stabilityf}
-R(\hat{b},z) = 1 +  (b_{orig} +B \hat{b})^T z(I - zA)^{-1}
-\end{equation}
-The border is defined by the implicit function 
-\begin{equation}\label{eq:border}
-|R(\hat{b},u+iv)|^2 -1 = 0
-\end{equation}
-This function can be viewed either as $\R^m \times  \CN \rightarrow \R$ or $\R^m \times  \R^2 \rightarrow \R$.
-We are only concerned with the expansion or contraction. Because of this we are only interested in the change perpendicular to the border of the stability region.
-
-For a point $z_0= u_0 +i v_0 $ 
-with $ |R(w,z_0)|^2 -1 = 0 $ the gradient of
- $|R(w,u,v)|^2$ is calculated. 
-If the gradient does not vanish at $z = z_0$ \eqref{eq:border} can be transformed in a function $\R^m \times \R \rightarrow \R$ depending on $\hat{b}$ and a single variable $a$ by setting $z = z_0 + n a$.
-The complex number $n$ denotes the normal vector $\vec{n} = \frac{\nabla |R(\hat{b},u,v))|^2}{\left| \nabla |R(\hat{b},u,v))|^2 \right|}$ of the border of the stability region written in complex notation.
-This defines the function $a(\hat{b})$ with $|R(\hat{b},z_0 + n a(\hat{b}))|^2 -1 = 0$.
-This implies that this approach is only applicable if the gradient does not vanish on the border of the stability function. This is not true in general but for the most common RKM. 
-The derivative $\frac{\mathrm d}{\mathrm d \hat{b}} (a(\hat{b}))$ can be calculated using the implicit function theorem. The defining function is 
-\begin{equation}\label{eq:f(w,a)}
-f: \R^m \times \R \rightarrow \R,  f(\hat{b},a) = |R(\hat{b},z_0 + n a)|^2 -1
-\end{equation}
-If explicit methods are used this function is known to be continuously differentiable because it can be written as a polynomial of $\hat{b},u$ and $v$.
-This is a necessary condition for the use of the implicit function theorem.
-It is also known by definition, that for the point $b = b_{orig}$ and $z = z_0$ the equation $ |R(\hat{b},z_0)|^2 -1 = 0 $ is satisfied.
-For simplicity only the formulas for the derivative at the value $b = b_{orig}$ are given. This corresponds to $\hat{b} = 0$.
-The wanted derivative is given by
-
-\begin{equation}
- \frac{\partial a}{\partial \hat{b}_j} (w) =
- - \left[ \frac{\partial f}{\partial a} (\hat{b},a(\hat{b}))  \right] ^{-1} 
-   \left[ \frac{\partial f}{\partial \hat{b}_j}(\hat{b},a(\hat{b})) \right]
-\end{equation}
-as long as $\frac{\partial f}{\partial a} \neq 0$ on the border of the region of absolute stability. This is true in most cases as it has been discussed above.
-
-The derivative of \eqref{eq:f(w,a)} in respect to $a$ evaluated at $\hat{b}=0$ can be calculated using the directional derivative
-\begin{align*}
- \frac{\partial f}{\partial a} \Big|_{\hat{b}=0} =& 
- \frac{\partial f}{\partial a} (\hat{b},a(\hat{b})) \Big|_{\hat{b}=0} = 
- \nabla |R(\hat{b},a(\hat{b})|^2 \vec{n} \Big|_{\hat{b}=0} \\
-=& \left| \nabla|R(0,z_0)|^2 \right| = \left| \nabla|R_{b_{orig}}(z_0)|^2 \right|
-\end{align*}
-
-The derivative in respect to $\hat{b}_j$ is 
-$ \frac{\partial f}{\partial \hat{b}_j}(\hat{b},a(\hat{b})) \Big|_{\hat{b}=0}$.
-This simplifies to 
-\begin{multline}\label{eq:derivative_to_b}
- \frac{\partial f}{\partial \hat{b}_j}(\hat{b},a(\hat{b})) \Big|_{\hat{b}=0} = 
- \frac{\partial }{\partial \hat{b}}(R(\hat{b},z)R(\hat{b},z^*)) \\
- = (B\hat{b}_j)^T (I-zA)^{-1} R_{b_{orig}}(z^*) + (B\hat{b}_j)^T (I-z^*A)^{-1} R_{b_{orig}}(z)
-\end{multline}
-using the fact that $z_0$ is on the border of the stability region.
-The derivative $(B\hat{b})^T (I-zA)^{-1} R_{b_{orig}}(z^*) (B\hat{b})^T (I-z^*A)^{-1} R_{b_{orig}}(z)$ is known to be finite for explicit methods, because it is a polynomial and $|z_0| < \infty$
-The same process can be done for all points on the border of the stability region. 
-As long as $\nabla|R(\hat{b},u+iv)|^2 \neq 0 \forall \hat{b} < \epsilon  \forall z$ the derivative $\frac{\partial a}{\partial \hat{b}_j} (\hat{b})$ is finite.
-This means that a small change of the weights only leads to a small change of the stability region.
-
-\todo{TODO: Some better theorem?} %TODO
+A useful property of the resulting stability function is that it is similar to the stability function of the baseline method.
+
 \begin{lemma}
@@
     | R_{b}(z) |
-    + | z (\bt - b) (\I - z A)^{-1} e
+    + | z (\bt - b) (\I - z A)^{-1} e|
     \\
@@
 \end{proof}
-This shows why the objective $\min \| \bt - b \|_1 \|$ is a good
+This shows why the objective $\min \| \bt - b \|_1 $ is a good
 choice to control the change of the region of absolute stability,
@@
 
-\subsection{Positivity}
+\subsection{Simplification of positivity constraints}
 
@@
 Systems of ODEs with a high number of dimensions can appear when solving a PDE with the method of lines.
-
-To ensure that the solution is positive
-\begin{equation}
- u_i^{n+1} \geq 0   \quad   \forall {i \in \{1, \cdots,m \}}  
-\end{equation}
-has to be fulfilled. These $m$ positivity constraints of the
-optimization problem can be written as
-\begin{equation}
-u_i^n + \dt \sum_{j=0}^s F_{i,j}  b_j \geq 0 \quad \forall {i \in \{1, \cdots,m \}}
-\end{equation}
-
+To ensure that the solution is positive \eqref{eq:ut} has to be fulfilled. 
+This causes $m$ scalar positivity constraints.
 Imposing all of these constraints in the LP solver usually results
@@
 
-When enforcing a maximum value, the number of constraints can be reduced using the same technique.
-When enforcing both maximum and minimum values two separate sets of active constraints $k_{min}$ and $k_{max}$ are used.
-The set $k_{max}$ denotes the active maximum constraints and the set $k_{min}$ denotes the active minimum constraints.
+When enforcing a maximum value, the number of constraints can be reduced using the same technique. When enforcing both maximum and minimum values two separate sets of active constraints are used. 
 In this case it is important to update these sets simultaneously.
 
-\subsection{Implemention}
-
-In this section an algorithm to update the solution $u^n$ using the adaptive RKM is proposed.
+\subsection{Implementation}
+
+In this section an algorithm to calculate the solution of an ODE using the adaptive RKM is proposed.
 The pseudocode is given in Algorithm\,\ref{alg:Adaption}.
-t
-At first, the stage values $y_i$ and stage derivatives $f_i$ are calculated.
+
+To calculate the next solution $u^{n+1}$ the following is algorithm is executed. This is repeated until the final time $t_{end}$ is reached.
+
+At first, an $\dt$ is selected using a step size control algorithm or a fixed step size is used.
+With this $\dt$ the stage values $y_i$ and stage derivatives $f_i$ are calculated.
 Afterwards, $u^{n+1}$ is computed using the original weights $b$.
@@
 Apply an LP solver to the optimization
-problem with accuracy and positivity constraints.
+problem.
 If the LP is infeasible, reduce the order $\tilde p$ by one.
@@
 identify the problem as feasible.
-Afterwards the adaption error is calculated using \eqref{eq:Err}.
-If the error is bigger than a given maximum error tolerance, the
+Afterwards the perturbation is calculated using \eqref{eq:Err}.
+If the perturbation is bigger than a given tolerance, the
 new weights $\bt$ are also rejected.
@@
 
-If the reduction of stepsize is used this process is done first for $\theta = 1$. If no weights were found, the process is repeated for a smaller $\theta$.
-
-\begin{algorithm}
+\begin{algorithm}[ht]
 \begin{algorithmic}[1]
-\State Caclculate Stagedrivatives
-\State $u^{n+1}=\Delta t Fb$
+\While{$t<t_{end}$}
+\State Choose $\Delta t$
+\State Calculate $F$ using \eqref{eq:stagevalues}
+\State $u^{n+1}=u^n + \Delta t Fb$
 \If {$u^{n+1} \geq 0$} 
-        \State Update and End
+	\State $success = True$
 \Else
-	\ForAll{$\tilde p \in [p_{start}:p_{min}]$} 
+	\State{$\tilde p = p_{start}$}
+	\While{$\tilde p \geq p_{min}$} 
 		\State Solve LP
-		\If {LP Feasible}
-			\State caclulate $err_{adapt}$
-			\If {$err_{adapt} < tol$}
-				\State $u^{n+1}=\Delta t F \tilde b$
-				\State Update and End
+		\If {LP Feasible \textbf{and} result correct}
+			\State $perturb = \| \Delta t F(\tilde{b}-b)\|$
+			\If {$perturb < tol$}
+				\State $u^{n+1}=u^n+\Delta t F \tilde b$
+				\State $success = True$ 
+				\State \textbf{break}
 			\EndIf 
 		\EndIf
-	\EndFor
-	\State Reject step and End
+		\If {$\tilde p = p_{end}$}
+			\State $sucess=False$			
 		\EndIf
+		\State $\tilde p = \tilde p -1$
+	\EndWhile
+\EndIf 
+\If {$sucess$}
+	\State $u^n=u^{n+1},t^n=t^{n+1}+\Delta t,n=n+1$
+\EndIf
+\EndWhile
 \end{algorithmic}
@@
 
-\subsection{Applicable problems}\label{sec:app_problem}
-The approach can be used with ODE systems $u' = f(t,u)$ that ensure  $u(t) \geq 0 \forall t \forall {  u_0 \geq 0}$ 
-A sufficient condition is  
-\begin{equation}
-u_i=0 \implies f_i(t,u_1,\cdots,u_i,\cdots,u_n) \geq 0 \quad \forall {u_c \geq 0}\; \forall {t}
-\end{equation}
-For problems where the exact solution is positive for certain $u_0$ but positivity is not preserved for all $u_0 \geq 0$ tests did not show promising results.
+The adaptive RKM can be used with ODEs that satisfy \eqref{eq:condition_ODE_pos}.
+For problems where the exact solution is positive for certain $u_0$ but do not satisfy \eqref{eq:condition_ODE_pos} tests did not show promising results.
 Additionally, it is not certain whether the computed solutions would be reasonable.
 
-\subsection{Explicit methods}\label{sec:Ex_expl}
-First we want to test the adaptive RKM used with explicit methods.  
+\subsection{Nonstiff problem with fixed stepsize}\label{sec:Ex_expl}
+First the adaptive RKM are tested on non stiff problems with a fixed step size. For this a explicit methods are used.
 When used with explicit methods the cost of solving the LP is significant because the computation of the stage derivatives only requires $s$ evaluations of the RHS.
@@
 An example for this is the ODE in Section~\ref{sec:example_lin}.
-Some nonlinear RHS may require very small timesteps to preserve positivity.  
+Some nonlinear RHS may require very small time steps to preserve positivity.  
 For these, adapting the weights could be a possible way to solve them. 
@@
 We use the Dormand--Prince RK5 method and adapt the weights using the free adaptation.
-In Figure\,\ref{fig:sol_AdDe} the results for $\Delta t = 0.02$ are plotted for different timesteps. We can see that the solution approaches an exponential function with $t \rightarrow \infty$.
+In Figure\,\ref{fig:sol_AdDe} the results for $\Delta t = 0.02$ are plotted for different time steps. We can see that the solution approaches an exponential function with $t \rightarrow \infty$.
 In Figure\,\ref{fig:weights_AdDe} the used weights are plotted. For $t\leq 0.6$ the weights are altered. For $t > 0.6$ the original weights lead to a positive solution.
@@
 
-Next, different timesteps are used.
+Next, different time steps are used.
 For $\dt \leq 0.0082$ the unaltered method leads to positive solutions.
@@
 
-\subsection{Implicit methods}
-Implicit methods seem like an advantageous choice for a couple of reasons. 
+\subsection{Stiff problem with fixed step size}
+Next the adaptive RKM is tested on stiff problems. Here implicit methods are used.
+Implicit methods are an advantageous choice for a couple of reasons. 
 Firstly, the cost of solving the LP is relatively small compared to the cost of solving the stage equations.
-Secondly, the timestep is not limited by the stability of the method.
-Therefore, it is possible to use larger timesteps that are more likely to lead to negative values.
+Secondly, the time step is not limited by the stability of the method.
+Therefore, it is possible to use larger time steps that are more likely to lead to negative values.
 
@@
 The original solution for the first step is negative. Therefore, the weights have to be changed. 
-If we take a look at the solution after the first timestep at $t=0.001$ we can see that at $x=0$ the solution is smaller than the solution at the surrounding points.  
+If we take a look at the solution after the first time step at $t=0.001$ we can see that at $x=0$ the solution is smaller than the solution at the surrounding points.  
 This is not physical.  
-The next timesteps lead to physical solutions again. 
+The next time steps lead to physical solutions again. 
 To prevent this glitch from happening we choose the weights based on a convex adaptation. 
@@
 
-\subsection{Stepsize control}
+\subsection{Stiff problem with adaptive step size}
 
@@
 \frac{\partial u_3}{\partial t} &=-a \frac{\partial u_3}{\partial x} + d\frac{\partial^2 u_3}{\partial x ^2} + 0.5(1-\exp(-1.21u_2^2)) u_3 - 0.01 u_3 -0.02 u_3 \\ 
-\frac{\partial u_4}{\partial t} &=-a \frac{\partial u_4}{\partial x} + d\frac{\partial^2 u_4}{\partial x ^2} + 0.05 u_2 + 0.02 u_3 + 0.003u_4 
+\frac{\partial u_4}{\partial t} &=-a \frac{\partial u_4}{\partial x} + d\frac{\partial^2 u_4}{\partial x ^2} + 0.05 u_2 + 0.02 u_3 - 0.003u_4 
 \end{align}
@@
 The advection part is semidiscretized using a first order upwind scheme and the diffusion part is semidiscretized using a 3-point-scheme. This leads to a positivity preserving system of ODEs.
+The ODE is also mass conservative.
 The computation is done using the Im-Euler~3 extrapolation
@@
 In Figure\,\ref{fig:Stats_ADP} different values are plotted.  
-In the first subplot the step size is plotted. For $t<25$ the timesteps are small. After $t = 30$ the step size increases, because the solution only evolves slowly afterwards. 
-In the second subplot the minimum of $u_1,u_2,u_3,u_4$ is plotted for all timesteps that initially lead to negative values. This value is computed before and after adapting the weights.
-We can see that relatively large negative values occurred at some timesteps. 
+In the first subplot the step size is plotted. For $t<25$ the time steps are small. After $t = 30$ the step size increases, because the solution only evolves slowly afterwards. 
+In the second subplot the minimum of $u_1,u_2,u_3,u_4$ is plotted for all time steps that initially lead to negative values. This value is computed before and after adapting the weights.
+We can see that relatively large negative values occurred at some time steps. 
 After the adaption of the weights, all values are close to $0$. Therefore, the adaption of weights successfully preserved positivity. 
-In the third subplot the approximated truncation error $err_T$ and the adaption errro $err_{adapt}$ are plotted. 
+In the third subplot the approximated truncation error $err_T$ and the adaption error $err_{adapt}$ are plotted. 
 We can see that the adaption error is of a similar magnitude as the truncation error. Therefore, the total error of the method is not increased drastically.  
-In the last subplots the weights are plotted.  
+In the next subplots objective function is plotted.  
 We can see that the changes to the weights are only very small. The adapted RKM is still very close to the original RKM. 
+In the last subplot the deviation of the sum over $u_1,u_2,u_3,u_4$ from the initial sum is plotted. 
+The mass is conserved within roundoff error.
 

### 2020-04-28 nusslein_vsrp/latex/Stratospheric_Reac.tex
Added Reaction Problem form Sandu
Tex in seperate file to clean it up independently

\subsection{Stiff problem with adaptive step size II}

As next problem we consider the the Statospheric reaction Problem form \cite{sandu2001positive}.
This Modell describes atmospheric Reaction between the substances in the concentration vector $u=[O^{1D},O,O_3,O_2,'NO,NO_2]$. 
The Problem is time dependent and has two linear invariants.

The problem describes as follows:
\begin{align}
    T &= ((t/3600)+t_start) \bmod 24 
    \qquad  Tr = 4.5
    \qquad  Ts = 19.5 \\
    \sigma(T) &= 
     \begin{cases}
    0.5+0.5 \cos(\pi\left|\frac{(2T-Tr-Ts)}{(Ts-Tr)}\right| \left[\frac{(2T-Tr-Ts)}{(Ts-Tr)}\right]       & \quad \text{if } Tr \leq T \leq Ts\\
    0  & \quad \text{else } 
    \end{cases}
\end{align}

\begin{align}
r_1 &= O_2\,k_1       &  k_1 &= \num{2.643e-10} \sigma^3 \\
r_2 &= O\,O_2 k_2     &  k_2 &= \num{8.018e-17} \\
r_3 &= O_3\,k_3       &  k_3 &= \num{6.120e-4}\sigma \\
r_4 &= O_3\,O\,k_4     &  k_4 &= \num{1.567e-15} \\
r_5 &= O_3\,k_5       &  k_5 &= \num{1.070e-3}\sigma^2 \\
r_6&= O^{1D}\,M\,k_6     &  k_6 &= \num{7.110e-11} \\
r_7 &= O^{1D}\,O_3\,k_7   &  k_7 &= \num{1.200e-10} \\
r_8 &= O_3\,NO\,k_8    &  k_8 &= \num{6.062e-15} \\
r_9 &= NO_2\,O\,k_9    &  k_9 &= \num{1.069e-11} \\
r_{10}&=NO_2\,k_{10}  &  k_{10}&=\num{1.289e-2}\sigma \\
r_{11}&=NO\,O\,k_{11} &  k_{11}&=\num{1.0e-8}
\end{align}

with $M  = 8.120e16$

\begin{align}
\frac{\mathrm d}{\mathrm d t}O^{1D}  &=   +r_5 -r_6 -  r_7 \\
\frac{\mathrm d}{\mathrm d t}dO   &=  +2*r_1 -r_2 +r_3 -  r_4  +r_6 -r_9 +r_{10} -r_{11} \\
\frac{\mathrm d}{\mathrm d t}dO_3  &=  +r_2 -r_3 -  r_4 -r_5  -r_7 -r_8 \\
\frac{\mathrm d}{\mathrm d t}dO_2  &=  -r_1 -r_2 +r_3 +2r_4 +r_5 +2r_7 +r_8 +r_9 \\
\frac{\mathrm d}{\mathrm d t}dNO  &=  -r_8 +r_9 +r_{10} -r_{11} \\
\frac{\mathrm d}{\mathrm d t}dNO_2 &=  +r_8 -r_9 -r_{10} +r_{11} 
\end{align}

The two linear invariants are 
\begin{align}
m_O = [1,1,3,2,1,2]^T & \qquad m_O^Tu=\text{const} \\
m_N = [0,0,0,0,1,1]^T & \qquad m_N^Tu=\text{const}
\end{align}

The initial conditions are $u(0)=[\num{9.906e+1},\num{6.624e08},\num{5.326e11},\num{1.697e16},\num{4.000e6},\num{1.093e9}]$
The system was normalized internally such that $u_n(0)=1$ to achieve an suitable error estimation. After the computation the results are scaled back.
The system is solved in the time from $t_{start}=12h$ to $t_{end}=84h$. Using the adapted implicit is done using the Im-Euler~3 extrapolation method with free adaptation. 
Also the same step size control is used. 
The results are plotted in figure\,\ref{fig:Stratospheric_time}. We can see that the results of the adapted solution is close to the reference solution obtained with the unadapted Im-Euler~3 extrapolation method and a higher accuracy.

\begin{figure}
\centering
\includegraphics[width=0.9\textwidth]{plots/Stratospheric_time.pdf}
\caption{Numerical approximation of stratospheric reaction system. A reference solution is also plotted}
\label{fig:Stratospheric_time}
\end{figure}

To compute the solution 249 steps were computed, two of these were rejected due to a violation of the error bound.
More details are plotted in figure\,\ref{fig:Stats_Strat}. The rejected steps are drawn with thick crosses.
In the first subplot it can be seen that the step size undergoes multiple sudden changes due to the timedependence of the problem.
The minimum values before and after the adaptation are also drawn for all steps where the initial values where negative. 
This was only the case for some time intervals. 25 Steps exhibited negative values.
Almost all of them very close to 0 and the adaptation did only show a small improvement. 
The smallest value of the solution is $min(u) = \num{-1.59e-11}$.
The used weights are also very close to the original weights, except of the steps that were rejected anyway due to an violation of the tolerance.
The change of the two linear invariantes are plotted in the last two subplots. Both are preserved within roundoff error.
Due to the internal normalization the round off error is amplified to the order of $10$ for $m_O^Tu$.

\begin{figure}
\centering
\includegraphics[width=0.85\textwidth]{plots/Stratospheric_stepsize,b.pdf}
\caption{Different statistics of computation of the stratospheric reaction.}
\label{fig:Stats_Strat}
\end{figure}

### 2020-04-29 nusslein_vsrp/latex/overview.tex
move revised results for stratospheric reaction system into the main file

@@
 
+\subsection{Stiff problem with adaptive step size II}
+
+Here, we consider the stratospheric reaction problem of
+\cite{sandu2001positive}, which models the reactions of the substances
+in the concentration vector $u = [O^{1D}, O, O_3, O_2, NO, NO_2]$.
+This ODE has two linear invariants
+\begin{align}
+  m_O^T u &= \text{const}, & m_O &= [1,1,3,2,1,2]^T, \\
+  m_N^T u &= \text{const}, & m_N &= [0,0,0,0,1,1]^T,
+\end{align}
+which describe the conservation of the total mass of oxygen and nitrogen,
+respectively.
+
+The reaction system
+\begin{equation}
+\label{eq:stratospheric}
+\begin{aligned}
+  \frac{\mathrm d}{\mathrm d t} O^{1D} &= r_5 - r_6 - r_7, \\
+  \frac{\mathrm d}{\mathrm d t} O      &= 2 r_1 - r_2 + r_3 - r_4 + r_6 - r_9 + r_{10} - r_{11}, \\
+  \frac{\mathrm d}{\mathrm d t} O_3    &= r_2 - r_3 - r_4 - r_5 - r_7 - r_8, \\
+  \frac{\mathrm d}{\mathrm d t} O_2    &= -r_1 - r_2 + r_3 + 2 r_4 + r_5 + 2r_7 + r_8 + r_9, \\
+  \frac{\mathrm d}{\mathrm d t} NO     &= -r_8 + r_9 + r_{10} - r_{11}, \\
+  \frac{\mathrm d}{\mathrm d t} NO_2   &= r_8 - r_9 - r_{10} + r_{11},
+\end{aligned}
+\end{equation}
+with time $t$ in seconds is given by the reaction rates
+\begin{equation}
+\begin{aligned}
+  r_{1}  &= k_1\, O_2,          &  k_{1}  &= \num{2.643e-10} \sigma^3, &
+  r_{2}  &= k_2\, O\, O_2,      &  k_{2}  &= \num{8.018e-17}, \\
+  r_{3}  &= k_3\, O_3,          &  k_{3}  &= \num{6.120e-4} \sigma, &
+  r_{4}  &= k_4\, O_3\, O,      &  k_{4}  &= \num{1.567e-15}, \\
+  r_{5}  &= k_5\, O_3,          &  k_{5}  &= \num{1.070e-3} \sigma^2, &
+  r_{6}  &= k_6\, M\, O^{1D},   &  k_{6}  &= \num{7.110e-11}, \\
+  r_{7}  &= k_7\, O^{1D}\, O_3, &  k_{7}  &= \num{1.200e-10}, &
+  r_{8}  &= k_8\, O_3\, NO,     &  k_{8}  &= \num{6.062e-15}, \\
+  r_{9}  &= k_9\, NO_2\, O,     &  k_{9}  &= \num{1.069e-11}, &
+  r_{10} &= k_{10}\, NO_2,      &  k_{10} &= \num{1.289e-2} \sigma, \\
+% \stepcounter{equation}\tag{\theequation}
+  r_{11} &= k_{11}\, NO\, O,    &  k_{11} &= \num{1.0e-8},
+\end{aligned}
+\end{equation}
+where $M  = \num{8.120e16}$ and
+\begin{align}
+    T &= (t/3600) \bmod 24,
+    \quad T_r = 4.5,
+    \quad T_s = 19.5 \\
+    \sigma(T) &=
+    \begin{cases}
+    0.5+0.5 \cos\Bigl(\pi \Bigl|\frac{(2T-T_r-T_s)}{(T_s-T_r)}\Bigr| \frac{(2T-T_r-T_s)}{(T_s-T_r)} \Bigr)       & \quad \text{if } T_r \leq T \leq T_s,\\
+    0  & \quad \text{otherwise}.
+    \end{cases}
+\end{align}
+
+The initial conditions are
+\begin{equation}
+  u(t_0) = [
+    \num{9.906e+1},
+    \num{6.624e08},
+    \num{5.326e11},
+    \num{1.697e16},
+    \num{4.000e6},
+    \num{1.093e9}
+  ]^T.
+\end{equation}
+The system was normalized internally such that $\forall n\colon u_n(t_0) = 1$
+for the computation to achieve a suitable error estimation.
+The system is solved in the time from $t_{0} = \SI{12}{h}$ to $t = \SI{84}{h}$
+using the Im-Euler~3 extrapolation method with free adaptation and step size
+control.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.9\textwidth]{plots/Stratospheric_time.pdf}
+\caption{Numerical approximation of stratospheric reaction system
+         \eqref{eq:stratospheric}.}
+\label{fig:Stratospheric_time}
+\end{figure}
+
+The results are shown in Figure~\ref{fig:Stratospheric_time}.
+The adapted solution is close to the reference solution obtained
+with the unadapted Im-Euler~3 extrapolation method and a higher accuracy.
+For this solution, 249 steps were computed; two of these were rejected due to
+a violation of the error bound.
+More details are shown in Figure~\ref{fig:Stats_Strat}. The rejected steps are drawn with thick crosses.
+The step size $\dt$ undergoes multiple sudden changes due to the explicit dependence
+on time of the problem.
+The minimum values before and after the adaptation are also shown for all steps where the initial values were negative.
+This was only the case for some time intervals. 25 steps exhibited negative values.
+Almost all of them were very close to zero and the adaptation did only show a small improvement.
+The smallest value of the solution is $\min(u) = \num{-1.59e-11}$.
+The used weights are also very close to the original weights, except of the steps that were rejected anyway due to a violation of the tolerance.
+The change of the two linear invariants are shwon in the last two subplots. Both are preserved within roundoff error.
+Due to the internal normalization the round off error is amplified to the order of $10$ for $m_O^Tu$.
+
+\todo{TODO: Plot relative change $(m_O^T u - m_O^T u(t_0)) / m_O^T u(t_0)$
+instead of the absolute change of the linear invariants?}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.85\textwidth]{plots/Stratospheric_stepsize,b.pdf}
+\caption{Different statistics of computation of the stratospheric reaction system
+         \eqref{eq:stratospheric}.}
+\label{fig:Stats_Strat}
+\end{figure}
+
 \section{Conclusion} \label{sec:conclusion}
+
 It is possible to adapt the weights to enforce positivity for RKMs that are not positivity preserving.

### 2020-04-29 nusslein_vsrp/latex/overview.tex
additional revisions

@@
 
-%It is possible to semi-discretize PDEs in ways that preserve positivity\cite{kopecz_comparison_2019,kopecz_unconditionally_2018}.
 We say an ODE
@@
 \end{align}
-is positive if
+in $\R^m$ is positive if
 \begin{align} \label{continuous-positivity}
@@
 reaches zero and then proceed in some special way. This approach is implemented in the MATLAB ODE Suite.
+Another approach to ensure positivity in a practical application is shown in
+\cite{shampine_non-negative_2005}, where the ODE is redefined outside the
+physical domain.
 If positivity is preserved under a forward Euler step (with
@@
 $\dt \leq {\mathcal C} \dt_\text{FE}$, where ${\mathcal C}$ depends on the SSPRK method.
-The modified Patankar--Runge--Kutta (MPRK) methods represent another approach to ensuring
-positivity, for specific classes of ODEs.  MPRK methods introduce multiplicative
+Modified Patankar--Runge--Kutta (MPRK) methods represent another approach to ensuring
+positivity for specific classes of ODEs. MPRK methods introduce multiplicative
 factors within the Runge--Kutta stages to ensure positivity, but require the solution
 of a linear algebraic system; see e.g.\ \cite{kopecz_comparison_2019} and references therein.
-Yet another approach to ensure positivity in a practical application is shown in \cite{shampine_non-negative_2005},
-where the ODE is redefined outside the physical domain.
 Finally, we mention diagonally split Runge--Kutta (DSRK) methods, which can be unconditionally
-positive and have order higher than one (they avoid the restriction mentioned above because
-they are not general linear methods \cite{horvath_positivity_1998}.  However, in practice such
+positive and have order higher than one. As MPRK schemes, DSRK methods avoid
+the restriction mentioned above because
+they are not general linear methods \cite{horvath_positivity_1998}.  However, in practice DSRK
 methods are less accurate than backward Euler for large step sizes \cite{macdonald2007}.
@@
 that is positivity preserving, at least for a single step.
-The main idea is to adaptively choose the weights $b$ of the RKM, after the
+The main idea is to adaptively choose the weights $b$ of the RKM, after
 the stage values are known, in a way that ensures positivity.
@@
 
-The paper unfolds as follows. In Section~\ref{sec:main_idea} the main idea is explained. How the method can be written as a LP is explained in Section~\ref{sec:LP}.
-Section~\ref{sec:integration} describes how the new approach can be used with different RKM, how it can be integrated in a step size control and how the region of absolute stability can be approximated.
+The paper unfolds as follows. In Section~\ref{sec:main_idea} the main idea is explained. How the method can be written as an LP is explained in Section~\ref{sec:LP}.
+Section~\ref{sec:integration} describes how the new approach can be used with different RKMs, how it can be integrated in a step size control, and how the region of absolute stability can be approximated.
 In Section~\ref{sec:imple} further details that were used when implementing the algorithm are described.
@@
 \end{equation}
-Based on these values the next solution $u^{n+1}$ is computed by
+Based on these values, the next solution $u^{n+1}$ is computed as
 \begin{equation} \label{eq:rkstep}
@@
 \begin{equation}\label{eq:Combination}
-u^{n+1} = u^n + \dt F b
+u^{n+1} = u^n + \dt F b,
 \end{equation}
@@
 
-To illustrate the usage of the method we consider the following reaction equation from\,\cite{kopecz_comparison_2019}.
-
+To illustrate the usage of the method we consider the reaction system
+\cite{kopecz_comparison_2019}
 \begin{subequations}
@@
 \begin{align}
-u_1' &= 0.01u_2 + 0.01 u_3 +0.003u_4 - \frac{u_1 u_2}{0.01+u_1} \\
-u_2' &= \frac{u_1u_2}{0.01+u_1}-0.01 u_2-0.5(1-\exp(-1.21 u_2^2)) u_3 -0.05 u_2 \\
-u_3' &= 0.5(1-\exp(-1.21u_2^2)) u_3 - 0.01 u_3 -0.02 u_3 \\
-u_4' &=0.05 u_2 + 0.02 u_3 + 0.003u_4
+u_1' &= 0.01u_2 + 0.01 u_3 +0.003u_4 - \frac{u_1 u_2}{0.01+u_1}, \\
+u_2' &= \frac{u_1u_2}{0.01+u_1}-0.01 u_2-0.5(1-\exp(-1.21 u_2^2)) u_3 -0.05 u_2, \\
+u_3' &= 0.5(1-\exp(-1.21u_2^2)) u_3 - 0.01 u_3 -0.02 u_3, \\
+u_4' &=0.05 u_2 + 0.02 u_3 + 0.003u_4,
 \end{align}
 \end{subequations}
-and the initial conditions
+with initial conditions
 \begin{equation}
@@
 \end{equation}
-When solved with the Cash-Karp (CK5) method and $\dt = 0.005$ the approximated solution contains negative values. This causes qualitatively wrong solutions to the problem.
+When solved with the Cash-Karp (CK5) \cite{cash1990variable} method and $\dt = 0.005$ the approximated solution contains negative values. This causes qualitatively wrong solutions to the problem.
 In Figure\,\ref{fig:exampleI} the obtained results are plotted with dashed lines.
@@
     \includegraphics[width=0.75\textwidth]{plots/exampleI.pdf}
-    \caption{Numerical approximation of the reaction problem from \cite{kopecz_comparison_2019} computed with CK5 and $\dt = 0.005$.
+    \caption{Numerical approximation of the reaction problem \eqref{eq:Reaction} computed with CK5 and $\dt = 0.005$.
 The dashed lines are the approximations obtained without the adaption of the weights.
@@
 \LPblock{
-(Free adaptation) Given $F, p$, and $b$, find $\bt$ that minimizes $||\tilde{b} - b||_1$ subject to
+(Free adaptation) Given $F$, $\pt$, and $b$, find $\bt$ that minimizes $\|\bt - b\|_1$ subject to
 \begin{subequations}
+\label{eq:free-adaptation}
 \begin{align}
-u^{n+1}&=u^n+\dt F \tilde{b} \geq 0, \label{eq:direct_pos}\\
-Q_\pt\tilde{b} &=r_\pt \label{eq:direct_Order}.
+u^{n+1}&=u^n+\dt F \bt \geq 0, \label{eq:direct_pos}\\
+Q_\pt\bt &=r_\pt \label{eq:direct_Order}.
 \end{align}
@@
 \LPblock{
-(Convex adaptation) Given $F, p$, and $b$, find $g$ that minimizes $||\tilde{b} - b||_1$ subject to
+(Convex adaptation) Given $F$, $B$, and $b$, find $g$ that minimizes $\|\bt - b\|_1$ subject to
 \begin{subequations}
@@
 0 & \le g_i \le 1, \\
-\sum_i g_i &= 1 \\
-u^{n+1}&=u^n+\dt F \tilde{b} \geq 0.
+\sum_i g_i &= 1, \\
+u^{n+1}&=u^n+\dt F \bt \geq 0.
 \end{align}
@@
 
-\section{Properties of adaptive RKM} \label{sec:integration}
-
-In the previous sections an algorithm for choosing positivity preserving weights $\tilde{b}$ has been presented.
-In the next section properties of the adaptive RKM are discussed.
+\section{Properties of adaptive RKMs} \label{sec:integration}
+
+In the previous sections an algorithm for choosing positivity preserving weights $\bt$ has been presented.
+In the next section properties of the adaptive RKMs are discussed.
 
 \subsection{Choice of baseline method}
-An important property of the baseline method is the existence of embedded methods and the degrees of freedom for the weights $b$.
+An important property of the baseline method is the existence of embedded methods and the degrees of freedom for the weights $\bt$.
 As noted in Section~\ref{sec:OrderCond} the number of stages has to be higher than the order.
@@
 Another important property is the number of degrees of freedom for the choice of the new weights.
-These can be calculated using $s-\mathrm{rank}(Q)$. %$\mathrm{dim}(\mathrm{ker}(Q))$.
+These can be calculated using $s-\mathrm{rank}(Q_{\pt})$. %$\mathrm{dim}(\mathrm{ker}(Q))$.
 The number of degrees of freedom for different explicit methods are shown in Table\,\ref{table:DOF_exp}.
@@
 
-%(Note: up to this point I do not understand why some methods perform better than other. We only see it occur depending on the Problem-RKM combination)
-
 A third property is the existence of a positive solution.
@@
 \begin{align}
-err = \|u(t^{n+1})-\tilde u^{n+1}\| &= \|u(t^{n+1}) - (u^{n+1}+\dt F(\tilde{b}-b))\| \\
- &\leq \underbrace{\|u(t^{n+1})-u^{n+1}\|}_{\approx err_T}+\underbrace{\|\dt F(\tilde{b}-b)\|}_{= perturb} \label{eq:Err}
+err = \|u(t^{n+1})-\tilde u^{n+1}\| &= \|u(t^{n+1}) - (u^{n+1}+\dt F(\bt-b))\| \\
+ &\leq \underbrace{\|u(t^{n+1})-u^{n+1}\|}_{\approx err_T}+\underbrace{\|\dt F(\bt-b)\|}_{= perturb} \label{eq:Err}
 \end{align}
@@
 The total error is split up in the truncation error $err_T$ and the perturbation $perturb$ using the triangle inequality.
-The truncation error can be estimated using the standard error estimators $err_T = | u^{n}_{b} - u^{n}_{\hat{b}} |$.
+The truncation error can be estimated using the standard error estimators $err_T = \| u^{n}_{b} - u^{n}_{\hat{b}} \|$.
 After adapting the weights, the perturbation is calculated. If the perturbation is larger than the tolerance the weights are rejected.
@@
 
-An an example, the stability regions of adapted RKMs are visualized in
+As an example, the stability regions of adapted RKMs are visualized in
 Figure\,\ref{fig:stab}.
@@
 \begin{equation}
-u_i^n + \dt \sum_{j=0}^s F_{i,j}  b_j  \geq 0 \quad \forall i \in h \subseteq \{1,\cdots,m \}.
-\end{equation}
-A reasonable approach is to set $h_0 = \{ i \in \{1,\cdots,m \} |  u_i^{n+1}  < 0 \}$.
-After the new weights and $u^{n+1}$ are computed using $h_0$, the algorithm has to check if the positivity of the other values $\{u_i | i \notin h_0 \subseteq \{1,\cdots,m \}\}$ is still satisfied.
-If the condition is not met a new set $h_{a+1} = \{ i \in \{1,\cdots,m \}|  u_i^{n+1}  < 0 \} \cup h_{a}$ using the new $u^{n+1}$ is generated and the LP is solved again.
+u_i^n + \dt \sum_{j=0}^s F_{i,j}  b_j  \geq 0 \quad \forall i \in h \subseteq \{1,\dots,m \}.
+\end{equation}
+A reasonable approach is to set $h_0 = \{ i \in \{1,\dots,m \} |  u_i^{n+1}  < 0 \}$.
+After the new weights and $u^{n+1}$ are computed using $h_0$, the algorithm has to check if the positivity of the other values $\{u_i | i \notin h_0 \subseteq \{1,\dots,m \}\}$ is still satisfied.
+If the condition is not met a new set $h_{a+1} = \{ i \in \{1,\dots,m \}|  u_i^{n+1}  < 0 \} \cup h_{a}$ using the new $u^{n+1}$ is generated and the LP is solved again.
 This process is repeated until a solution is found that leads to $u^{n+1} \geq 0$.
 The choice of $h_{a+1}$ based on $h_{a}$ ensures that the algorithm cannot get stuck in a loop.
-In the worst case $h_a$ is growing very slowly until $h_a = \{1,\cdots,m \}$.
+In the worst case $h_a$ is growing very slowly until $h_a = \{1,\dots,m \}$.
 This worst case did not occur when running the test problems.
@@
 
-At first, an $\dt$ is selected using a step size control algorithm or a fixed step size is used.
+At first, a step size $\dt$ is selected using a step size control algorithm or a fixed step size is used.
 With this $\dt$ the stage values $y_i$ and stage derivatives $f_i$ are calculated.
@@
 
-Start with $\tilde{p}$ equal to the highest order $p_{start}$ such that there are degrees of freedom for the weights $\bt$.
+Start with $\pt$ equal to the highest order $p_{start}$ such that there are degrees of freedom for the weights $\bt$.
 Apply an LP solver to the optimization
@@
 \State Choose $\Delta t$
-\State Calculate $F$ using \eqref{eq:stagevalues}
-\State $u^{n+1}=u^n + \Delta t Fb$
+\State Calculate $F = (f_1, \dots, f_s)$ \eqref{eq:stagevalues}
+\State Set $u^{n+1}=u^n + \Delta t Fb$ \eqref{eq:Combination}
 \If {$u^{n+1} \geq 0$}
@@
 	\While{$\tilde p \geq p_{min}$}
-		\State Solve LP
+		\State Solve LP \eqref{eq:free-adaptation}
 		\If {LP Feasible \textbf{and} result correct}
-			\State $perturb = \| \Delta t F(\tilde{b}-b)\|$
+			\State $perturb = \| \Delta t F(\bt-b)\|$
 			\If {$perturb < tol$}
@@
 		\EndIf
-		\If {$\tilde p = p_{end}$}
+		\If {$\tilde p = p_{min}$}
 			\State $sucess=False$
@@
 \If {$sucess$}
-	\State $u^n=u^{n+1},t^n=t^{n+1}+\Delta t,n=n+1$
+% 	\State $u^n=u^{n+1},t^n=t^{n+1}+\Delta t,n=n+1$
+	\State Set $n=n+1$ and continue
 \EndIf
@@
 \end{algorithmic}
-\caption{Adaption of weights}
+\caption{Pseudocode for the algorithm using a free adaption of weights.}
 \label{alg:Adaption}
@@
 
-\subsection{Nonstiff problem with fixed stepsize}\label{sec:Ex_expl}
-First the adaptive RKM are tested on non stiff problems with a fixed step size. For this a explicit methods are used.
+\subsection{Non-stiff problem with fixed stepsize}\label{sec:Ex_expl}
+First, adaptive RKMs based on explicit methods are tested on non-stiff problems with a fixed step size.
 When used with explicit methods the cost of solving the LP is significant because the computation of the stage derivatives only requires $s$ evaluations of the RHS.
@@
 \begin{equation}
+\label{eq:advection-decay}
 \begin{aligned}
@@
 \end{subfigure}
-\caption{Numerical results for the advection decay problem.}
+\caption{Numerical results for the advection decay problem \eqref{eq:advection-decay}.}
 \end{figure}
@@
 \includegraphics[width=0.8\textwidth]{plots/conv_adde.pdf}
-\caption{Convergence results of the adapted Dormand--Prince RK5.}
+\caption{Convergence results of the adapted Dormand--Prince RK5 for the advection decay problem \eqref{eq:advection-decay}.}
 \label{fig:conv_expl}
@@
 \subsection{Stiff problem with fixed step size}
-Next the adaptive RKM is tested on stiff problems. Here implicit methods are used.
+Next, adaptive RKMs based on implicit methods are tested on stiff problems.
 Implicit methods are an advantageous choice for a couple of reasons.
@@
 This means that they allow for changes of the weights without a reduction of the order.
-Also, all stage values are computed using the BE method. Hence, all intermediate stages are positive.
-Also, an embedded BE step is included. This ensures that a positive solution always exists, even if it is of first order.
+Moreover, all stage values are computed using the BE method. Hence, all intermediate stages are positive.
+Furthermore, an embedded BE step is included. This ensures that a positive solution always exists, even if it is of first order.
 
@@
 \begin{equation}
+\label{eq:diffusion}
 \frac{\partial }{\partial t} u = D \frac{\partial^2}{\partial x^2} u
@@
 \begin{equation}
-\frac{\mathrm d}{\mathrm d t} u_i = \frac{d}{\Delta x^2} \left( u_{i-1} - 2u_i + u_{i+1} \right)
+\frac{\mathrm d}{\mathrm d t} u_i = \frac{d}{\Delta x^2} \left( u_{i-1} - 2u_i + u_{i+1} \right).
 \end{equation}
@@
 \end{subfigure}
-\caption{Numerical results for the diffusion problem and the
+\caption{Numerical results for the diffusion problem \eqref{eq:diffusion} and the
          adapted Im-Euler~3 extrapolation method.}
@@
 \caption{Convergence result of baseline and adapted Im-Euler~3
-         extrapolation methods for the diffusion problem.}
+         extrapolation methods for the diffusion problem \eqref{eq:diffusion}.}
 \label{fig:conv_impl}
@@
 Next we test the adaptive RKM on a more complex problem.
-For this, we consider the advection-diffusion-production-destruction systems from \cite{kopecz_comparison_2019}. The system is defined by
+For this, we consider the advection-diffusion-production-destruction system  \cite{kopecz_comparison_2019}
 \begin{subequations}
@@
 \begin{align}
-\frac{\partial u_1}{\partial t} &=-a \frac{\partial u_1}{\partial x} + d\frac{\partial^2 u_1}{\partial x ^2} + 0.01u_2 + 0.01 u_3 +0.003u_4 - \frac{u_1 u_2}{0.01+u_1} \\
-\frac{\partial u_2}{\partial t} &=-a \frac{\partial u_2}{\partial x} + d\frac{\partial^2 u_2}{\partial x ^2} + \frac{u_1u_2}{0.01+u_1} -0.01 u_2-0.5(1-\exp(-1.21 u_2^2)) u_3 -0.05 u_2 \\
-\frac{\partial u_3}{\partial t} &=-a \frac{\partial u_3}{\partial x} + d\frac{\partial^2 u_3}{\partial x ^2} + 0.5(1-\exp(-1.21u_2^2)) u_3 - 0.01 u_3 -0.02 u_3 \\
-\frac{\partial u_4}{\partial t} &=-a \frac{\partial u_4}{\partial x} + d\frac{\partial^2 u_4}{\partial x ^2} + 0.05 u_2 + 0.02 u_3 - 0.003u_4
+\frac{\partial u_1}{\partial t} &=-a \frac{\partial u_1}{\partial x} + d\frac{\partial^2 u_1}{\partial x ^2} + 0.01u_2 + 0.01 u_3 +0.003u_4 - \frac{u_1 u_2}{0.01+u_1}, \\
+\frac{\partial u_2}{\partial t} &=-a \frac{\partial u_2}{\partial x} + d\frac{\partial^2 u_2}{\partial x ^2} + \frac{u_1u_2}{0.01+u_1} -0.01 u_2-0.5(1-\exp(-1.21 u_2^2)) u_3 -0.05 u_2, \\
+\frac{\partial u_3}{\partial t} &=-a \frac{\partial u_3}{\partial x} + d\frac{\partial^2 u_3}{\partial x ^2} + 0.5(1-\exp(-1.21u_2^2)) u_3 - 0.01 u_3 -0.02 u_3, \\
+\frac{\partial u_4}{\partial t} &=-a \frac{\partial u_4}{\partial x} + d\frac{\partial^2 u_4}{\partial x ^2} + 0.05 u_2 + 0.02 u_3 - 0.003u_4,
 \end{align}
@@
 with parameters $a=\num{1e-2} $ and $ d=\num{1e-6}$.
-The PDE is simulated on the domain $x = [0,1]$ with $N=100$ points. Periodic boundary conditions are used.
-The advection part is semidiscretized using a first order upwind scheme and the diffusion part is semidiscretized using a 3-point-scheme. This leads to a positivity preserving system of ODEs.
-The ODE is also mass conservative.
+The PDE is simulated on the domain $x = [0,1]$ with $N=100$ points and periodic boundary conditions.
+The advection part is semidiscretized using a first order upwind scheme and the diffusion part is semidiscretized using a central 3-point-scheme.
+This leads to a positivity preserving system of ODEs which conserves the total mass $\sum u$.
 The computation is done using the Im-Euler~3 extrapolation
@@
 After the adaption of the weights, all values are close to $0$. Therefore, the adaption of weights successfully preserved positivity.
-In the third subplot the approximated truncation error $err_T$ and the adaption error $err_{adapt}$ are plotted.
-We can see that the adaption error is of a similar magnitude as the truncation error. Therefore, the total error of the method is not increased drastically.
+In the third subplot the approximated truncation error $err_T$ and the perturbation $\perturb$ are plotted.
+We can see that $\perturb$ is of a similar magnitude as the truncation error. Therefore, the total error of the method is not increased drastically.
 In the next subplots objective function is plotted.
@@
     \end{subfigure}
-    \caption{Solution of advection-diffusion-reaction problem at different times.}\label{fig:Sol_ADP}
+    \caption{Solution of advection-diffusion-reaction problem \eqref{eq:ADR} at different times.}\label{fig:Sol_ADP}
 \end{figure}
@@
 \includegraphics[width=0.85\textwidth]{plots/ADP_stepsize,b.pdf}
-\caption{Different statistics of computation of the advection-diffusion-reaction problem.}
+\caption{Statistics of computation of the advection-diffusion-reaction problem \eqref{eq:ADR}.}
 \label{fig:Stats_ADP}
@@
   r_{10} &= k_{10}\, NO_2,      &  k_{10} &= \num{1.289e-2} \sigma, \\
-% \stepcounter{equation}\tag{\theequation}
   r_{11} &= k_{11}\, NO\, O,    &  k_{11} &= \num{1.0e-8},
@@
 for the computation to achieve a suitable error estimation.
-The system is solved in the time from $t_{0} = \SI{12}{h}$ to $t = \SI{84}{h}$
+The system is solved in the time from $t_{0} = \SI{12}{h}$ to $t_{end} = \SI{84}{h}$
 using the Im-Euler~3 extrapolation method with free adaptation and step size
@@
 \includegraphics[width=0.85\textwidth]{plots/Stratospheric_stepsize,b.pdf}
-\caption{Different statistics of computation of the stratospheric reaction system
+\caption{Statistics of computation of the stratospheric reaction system
          \eqref{eq:stratospheric}.}

### 2020-04-30 nusslein_vsrp/latex/overview.tex
Revise remark on Example II step size.

@@
 norm is $2 \sqrt{13}$.
-The real stability interval of SSP33 includes $[-2.5,0]$. Hence,
-it seems reasonable\footnote{A more careful analysis following
-\cite{ranocha2018L2stability} shows that the method is not
-monotone (energy stable) for this problem and choice of $\dt$ since
-$L$ is not normal.} to choose $\dt = \nicefrac{1}{3}$.
+The real-axis stability interval of SSP33 includes the interval $[-2.5,0]$.
+We take $\dt = \nicefrac{1}{3}$, which satisfies the spectral
+condition and guarantees boundedness (though not monotonicity)
+of the solution.
 The corresponding stage derivatives are

### 2020-04-30 nusslein_vsrp/RKimple.py
Scaled linear invariants in stratospheric reaction plot

@@ -799,20 +799,24 @@ def RK_integrate(solver = [], problem = [],stepsize_control = None, dumpK=False,
             #else:
             #    sc = stepsize_control.a_tol+u*stepsize_control.r_tol
             #error = np.linalg.norm((K@(solver.rkm.b-solver.rkm.bhat))/sc)/len(u)
-            error = dt*np.linalg.norm((K@(solver.rkm.b-solver.rkm.bhat)))
+            error = dt_old*np.linalg.norm((K@(solver.rkm.b-solver.rkm.bhat)))
             status['error'].append(error)
             if verbose: print('Error:',error)
             if error > stepsize_control.tol_reqect:
                 tol_met = False
                 if verbose: print('tol not met')
                 status['message'] += 'tol not met'
-                status['sc'].append('r')
+                if dt_old <= stepsize_control.dt_min+1e-15:
+                    if verbose: print('dt_min reached, using it anyway')
+                    status['sc'].append('m')
+                else:
+                    status['sc'].append('r')
             else:
                 tol_met = True
                 status['sc'].append('m')
 
 
-        if success and tol_met:
+        if success and (tol_met or dt_old <= stepsize_control.dt_min+1e-15):
             if verbose: print('advancing t')
             t += dt
             u = u_n

### 2020-04-30 nusslein_vsrp/latex/overview.tex
Scaled linear invariants in stratospheric reaction plot
Revise discussion of simplifying positivity constraints.

--- 2020-04-29 nusslein_vsrp/latex/overview.tex
@@
 \subsection{Simplification of positivity constraints}
-
-For the further discussion, we are concerned with a system of
-$m$ coupled ODEs, i.e.\ $u \in \R^m$.
-Systems of ODEs with a high number of dimensions can appear when solving a PDE with the method of lines.
-To ensure that the solution is positive \eqref{eq:ut} has to be fulfilled.
-This causes $m$ scalar positivity constraints.
-Imposing all of these constraints in the LP solver usually results
-in unnecessarily many constraints. In many cases, positivity is
-violated only for a subset $h \subseteq \{1, \dots, m\}$ of all
-components. This reduced set of positivity constraints can be
-formulated as
+The number of positivity constraints implied by \eqref{eq:ut}
+is equal to $m$, the number of ODEs being solved.  This number
+may be very large, for instance if the system is a semi-discretization
+of a PDE.  This makes solution of the LP very costly.
+But in most cases, positivity is violated only for a very small subset $h \subseteq \{1, \dots, m\}$ of
+the solution components.  We can solve a much less expensive LP
+by replacing \eqref{eq:ut} with
 \begin{equation}
@@
 \end{equation}
-A reasonable approach is to set $h_0 = \{ i \in \{1,\dots,m \} |  u_i^{n+1}  < 0 \}$.
-After the new weights and $u^{n+1}$ are computed using $h_0$, the algorithm has to check if the positivity of the other values $\{u_i | i \notin h_0 \subseteq \{1,\dots,m \}\}$ is still satisfied.
-If the condition is not met a new set $h_{a+1} = \{ i \in \{1,\dots,m \}|  u_i^{n+1}  < 0 \} \cup h_{a}$ using the new $u^{n+1}$ is generated and the LP is solved again.
-This process is repeated until a solution is found that leads to $u^{n+1} \geq 0$.
-The choice of $h_{a+1}$ based on $h_{a}$ ensures that the algorithm cannot get stuck in a loop.
-In the worst case $h_a$ is growing very slowly until $h_a = \{1,\dots,m \}$.
-This worst case did not occur when running the test problems.
+Of course, it must be checked that the solution of the resulting LP still
+satisfies the full set of constraints \eqref{eq:ut}.  In practice, we
+have found the following approach to be effective.
+First, set
+$$h_0 = \{ i \in \{1,\dots,m \} \ | \  u_i^{n+1}  < 0 \}.$$
+Solve the LP and let $\ut^{n+1}$ denote the new solution.
+If $\ut^{n+1}$ satisfies \eqref{eq:ut}, accept this as the new
+solution; otherwise, repeatedly take
+$$h_{a+1} = \{ i \in \{1,\dots,m \}|  \ut_i^{n+1}  < 0 \} \cup h_{a}$$
+until $\ut^{n+1}$ is found to satisfy \eqref{eq:ut}.
+In the examples we have studied, this approach was found to always converge in
+at most X iterations.
 %(Note: A approach on this would be to make a educated guess which indices could get a problem with positivity. A possible estimate would be $\frac{u_i^{n+1}}{max(K_{(i,0)}, \cdots ,K_{(i,0)})} $ )

### 2020-04-30 nusslein_vsrp/latex/overview.tex

@@
 The change of the two linear invariants are shwon in the last two subplots. Both are preserved within roundoff error.
-Due to the internal normalization the round off error is amplified to the order of $10$ for $m_O^Tu$.
-
-\todo{TODO: Plot relative change $(m_O^T u - m_O^T u(t_0)) / m_O^T u(t_0)$
-instead of the absolute change of the linear invariants?}
 

### 2020-05-01 nusslein_vsrp/RKimple.py
Changed implementation to new objective function for convex optimisation, Max number of iterations
Rerun all plot functions and measure max number of iterations of LP-Solve
Added a new notebook for the function to get a matrix vector expression of convex hull

@@ -7,6 +7,7 @@ import scipy.optimize as opt
 from OrderCondition import *
 import cvxpy as cp
 from scipy.optimize import linprog
+from scipy.spatial import ConvexHull, convex_hull_plot_2d
 
 
 
@@ -65,9 +66,9 @@ def solve_LP(solver,O,rhs,rkm,b_orig,u,K,dt,reduce = False,verbose_LP = False,mi
     if reduce: #We do not need it otherwise
         u_ = u +dt*K@b_orig
 
-
+    cnt_iter_h = 0
     while True: #We check at the end for break conditions
-
+        cnt_iter_h += 1
         #Reduce
         if reduce:
             i_min[i_n] = i_min[i_n] | (u_[i_n] < minval[i_n])
@@ -163,7 +164,7 @@ def solve_LP(solver,O,rhs,rkm,b_orig,u,K,dt,reduce = False,verbose_LP = False,mi
                 b = ap_op.value - an_op.value + b_orig
             except:
                 if verbose_LP: print('Solver crashed')
-                return (5,l,None)
+                return (5,l,None,0)
 
             if prob.status == cp.OPTIMAL:
                 status = 0
@@ -182,10 +183,82 @@ def solve_LP(solver,O,rhs,rkm,b_orig,u,K,dt,reduce = False,verbose_LP = False,mi
             #if there are no new negative values appearing
             break
 
-    return (status,l,b)
+    return (status,l,b,cnt_iter_h)
+
+def Eqsys_line(B):
+    N = np.array([[-1.,1.]])
+    a = np.array([[np.min(B)],[-np.max(B)]])
+    if a[0]+a[1] == 0:
+        print('only single Point, not able to compute convex hull')
+        raise ValueError
+    return N,a
+
+def Eqsys_chull(B):
+    """
+    Function to compute the equarionsystem defining a c_hull of the points b_1,...,b_l
+
+    Parameters:
+    B = [b_1,...,b_l] Points that span the chull
+
+    Returns:
+    Two sets of constriants
+        N_hull^T x     + a_hull    <= 0
+        N_subspace^T x + a_subspace = 0
+
+    N_hull = [n_1,...,n_k] Matrix with normal vectors of the inequality constraints
+    a_hull = [a_r,...,a_k] Vector containig the offsets of the inequality constraints
[…]

### 2020-05-01 nusslein_vsrp/utils.py

@@ -286,6 +286,13 @@ def plot_times(methods,dt,effective = False,title = ''):
     print(labels)
 
 
+def get_max_iter_h(status):
+    max_iter = 0
+    for lp_stat in status['LP_stat']:
+        if lp_stat:
+            max_iter = max(max_iter,lp_stat['cnt_iter_h_max'])
+    return max_iter
+
 def show_status(status):
     #print success
     print('Succesfull:'); print(status['success'])

### 2020-05-03 nusslein_vsrp/latex/overview.tex
Changed implementation to new objective function for convex optimisation, Max number of iterations
Rerun all plot functions and measure max number of iterations of LP-Solve
Added a new notebook for the function to get a matrix vector expression of convex hull
Major revision of the paper.

@@
 of Runge--Kutta methods with perturbed weights.  Numerical examples demonstrate
-the effectiveness of the approach, including application to both explicit and
-implicit methods.
+the effectiveness of the approach, including application to both stiff and
+non-stiff problems.
 \end{abstract}
@@
 
-We say an ODE
+We say an initial value problem
 \begin{align} \label{ode}
-    u'(t) & = f(t,u)
+    u'(t) & = f(t,u) \\
+    u(0) & = u_0
 \end{align}
-in $\R^m$ is positive if
+where $u: [0,T] \to \R^m$ is positive if
 \begin{align} \label{continuous-positivity}
-    u(0)\ge 0 \implies u(t) \ge 0 \text{ for all } t>0.
+    u(0)\ge 0 \implies u(t) \ge 0 \text{ for all } t \in [0,T].
 \end{align}
-A sufficient condition for positivity of an ODE is
+A sufficient condition for positivity of \eqref{ode} is
 \begin{equation} \label{eq:condition_ODE_pos}
-u_i=0 \implies f_i(t,u_1,\cdots,u_i,\cdots,u_n) \geq 0 \quad \forall {u_c \geq 0}\; \forall {t}
-\end{equation}
-For such ODEs a step using the forward Euler method will lead to a positive result for small $\dt$.
-A step using the backward Euler method will lead to a positive result for any $\dt$.
+u_i=0 \implies f_i(t,u_1,\cdots,u_i,\cdots,u_n) \geq 0 \quad \forall {u \geq 0,}\; \ \  \forall {t\in[0,T]}.
+\end{equation}
+For such ODEs, the backward Euler method is guaranteed to preserve positivity
+under any step size, while the forward Euler will preserve positivity for small enough $\dt$ \cite{hundsdorfer_numerical_2003}.
 Any RKM (or in fact any general linear method) that is unconditionally positivity preserving for all positive ODEs
-must have order $\le 1$ \cite{bolley_conservation_1978}. Notably, the (first-order)
-backward Euler method is unconditionally positivity preserving.
+must have order $\le 1$ \cite{bolley_conservation_1978}.
 For any higher order method, we expect positivity only under some restriction
@@
 Several approaches to ensuring numerical positivity exist in the literature.
-One technique is to use event finding methods in order to stop when some quantity
-reaches zero and then proceed in some special way. This approach is implemented in the MATLAB ODE Suite.
-Another approach to ensure positivity in a practical application is shown in
-\cite{shampine_non-negative_2005}, where the ODE is redefined outside the
-physical domain.
+The most basic approach is orthogonal projection onto the positive orthant,
+which means simply setting negative values to zero \cite{shampine1986conservation}.  This approach is often problematic;
+for instance, it will violate linear invariants such as mass conservation.
+As another approach, one may use event finding methods in order to stop when any solution component
+reaches zero, and then proceed in some special way
+\cite{shampine_non-negative_2005}. This approach is implemented in the MATLAB
+ODE Suite along with the idea of redefining the ODE outside the positive orthant (usually
+by evaluating at the nearest point on the boundary of the positive orthant).
 If positivity is preserved under a forward Euler step (with
@@
 Finally, we mention diagonally split Runge--Kutta (DSRK) methods, which can be unconditionally
-positive and have order higher than one. As MPRK schemes, DSRK methods avoid
+positive and have order higher than one. Like MPRK schemes, DSRK methods avoid
 the restriction mentioned above because
-they are not general linear methods \cite{horvath_positivity_1998}.  However, in practice DSRK
-methods are less accurate than backward Euler for large step sizes \cite{macdonald2007}.
+they are not general linear methods \cite{horvath_positivity_1998}.  However, in practice 
+unconditionally positive DSRK methods are less accurate than backward Euler for
+large step sizes \cite{macdonald2007}.
 
@@
 step for which the numerical solution would otherwise be non-positive.
-This is a significant cost, but may in fact be an economical alternative to rejecting
+This is a significant cost, but may in some cases be an economical alternative to rejecting
 a step or using excessively small step sizes.
@@
 The idea of adapting the weights after calculating the stage values has also been used,
-for instance in \cite{ketcheson_spatially_2013}.
-In this case it is used to adapt the properties of the time integrator for a method of lines solution of a PDE.
-With this approach, it is possible to have different properties of the RKM at different parts of the domain.
-Another class of methods that adapt the weights at the end of an RK step
-are the relaxation Runge--Kutta (RRK) methods.
-In these, the weights are scaled by a scalar relaxation parameter in order to guarantee
-conservation or monotonicity of a desired functional; e.g. to conserve or
-dissipate energy or entropy
+for instance in \cite{ketcheson_spatially_2013}.  In this case it is used to
+adapt the properties of the time integrator for a method of lines solution of a
+PDE.  Another class of methods that adapt the weights at the end of an RK step
+are the relaxation Runge--Kutta (RRK) methods.  In these, the weights are
+scaled by a scalar relaxation parameter in order to guarantee conservation or
+monotonicity of a desired functional; e.g. to conserve or dissipate energy or
+entropy
 \cite{ketcheson_relaxation_2019,ranocha_relaxation_2019,ranocha2020general}.
@@
 approach, where the numerical solution is adapted to satisfy the
-positivity constraint at the end of each step. In contrast to the
-usual orthogonal projection, which has also been proposed to deal
+positivity constraint at the end of each step. In contrast to
+simple orthogonal projection, which has also been proposed to deal
 with positivity constraints \cite{shampine1986conservation},
@@
 invariants can be very important. e.g.\ the total mass for a transport
-problem or in reaction systems. The property to preserve all linear
-invariants is shared by RRK methods and has been shown to be an
-important advantage compared to orthogonal projection methods
+problem or in reaction systems. Preserve of linear
+invariants has been shown to be an important advantage of RRK methods
+over orthogonal projection methods
 \cite{ranocha2020relaxationHamiltonian}. Of course, it is also
 possible to enforce the preservation of linear invariants in projection
-methods, but these invariants have to be known explicitly
+methods, but the invariants have to be known explicitly
 \cite{sandu2001positive}.
 
-The paper unfolds as follows. In Section~\ref{sec:main_idea} the main idea is explained. How the method can be written as an LP is explained in Section~\ref{sec:LP}.
-Section~\ref{sec:integration} describes how the new approach can be used with different RKMs, how it can be integrated in a step size control, and how the region of absolute stability can be approximated.
-In Section~\ref{sec:imple} further details that were used when implementing the algorithm are described.
+The paper unfolds as follows. In Section~\ref{sec:main_idea} the main idea is
+explained. Section~\ref{sec:LP} contains the formulation of the linear program
+for selection of the weights at each step.
+Section~\ref{sec:integration} describes how the new approach can be used with
+different RKMs, how it can be combined adaptive error control, and how the
+region of absolute stability can be approximated.
+Section~\ref{sec:imple} contains some details regarding efficient implementation 
+of the algorithm.
 In Section~\ref{sec:Numeric_Results} numerical results are given for multiple test problems.
@@
 \end{align}
-the stage values $y_1,\cdots,y_s$ are computed according to
+the stage values are computed according to
 \begin{equation}\label{eq:stagevalues}
@@
 $$
-Qb = r
+Q\bt = r
 $$
@@
 
-We see that the choice of $\bt$ is subject to linear equality and inequality
-constraints.  If we choose a linear objective function, the resulting problem
-for finding the modified weights is a linear program, which can be efficiently
-solved by standard algorithms.  A natural choice of objective function is
-$$
-\text{minimize } \|\bt - b\|_1.
-$$
-The resulting problem can be phrased as an LP by using slack variables.
-In general, this LP may not have a solution; we can relax the constraints
-by requiring a lower order of consistency than the design order of the
-method.  These choices and alternatives will be considered in Section~\ref{sec:LP}.
-
-In contrast to other projection methods
-\cite{shampine1986conservation,sandu2001positive},
-minimizing the deviation of the weights instead of the deviation
-of the projected solution is computationally much more efficient
-for large systems, arising for example in the discretization of
-PDEs.
-
-\subsection{Example I}\label{sec:example_reac}
-
-To illustrate the usage of the method we consider the reaction system
-\cite{kopecz_comparison_2019}
-\begin{subequations}
-\label{eq:Reaction}
-\begin{align}
-u_1' &= 0.01u_2 + 0.01 u_3 +0.003u_4 - \frac{u_1 u_2}{0.01+u_1}, \\
-u_2' &= \frac{u_1u_2}{0.01+u_1}-0.01 u_2-0.5(1-\exp(-1.21 u_2^2)) u_3 -0.05 u_2, \\
-u_3' &= 0.5(1-\exp(-1.21u_2^2)) u_3 - 0.01 u_3 -0.02 u_3, \\
-u_4' &=0.05 u_2 + 0.02 u_3 + 0.003u_4,
-\end{align}
-\end{subequations}
-with initial conditions
-\begin{equation}
-u(0) = (8,2,1,4)^T.
-\end{equation}
-When solved with the Cash-Karp (CK5) \cite{cash1990variable} method and $\dt = 0.005$ the approximated solution contains negative values. This causes qualitatively wrong solutions to the problem.
-In Figure\,\ref{fig:exampleI} the obtained results are plotted with dashed lines.
-At $t=1.905$ the value of $u_1$ gets negative. This leads to a diverging solution.
-
-Now the weights are adapted. The adapted weights are of 4th order. The results are also plotted in Figure\,\ref{fig:exampleI}, with solid lines.
-The positivity constraint is now fulfilled. A qualitatively correct solution is obtained.
-
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=0.75\textwidth]{plots/exampleI.pdf}
-    \caption{Numerical approximation of the reaction problem \eqref{eq:Reaction} computed with CK5 and $\dt = 0.005$.
-The dashed lines are the approximations obtained without the adaption of the weights.
-The lower plot shows the adaptation of the weights.
-}
-    \label{fig:exampleI}
-\end{figure}
-
-The used weights are also plotted in Figure\,\ref{fig:exampleI}.
-The used weights are equal to the original weights for $t<1.905$.
-At $t=1.905$ the weights are first adapted to ensure the positivity of the solution.
-For $t>2.63$ the original set of weights lead to a positive solution.
-Because of this, the used weights equal the original weights again.
-
-\section{Selection of modified weights}\label{sec:LP}
-
-In this section, we consider the formulation of the LP to choose
-the modified weights $\bt$.  In particular, we focus on the choice
-of objective function and how to relax the constraints to ensure that
-a feasible solution exists.
-
-\subsection{Example II}\label{sec:example_lin}
+\subsection{Example I}\label{sec:example_lin}
 
@@
 An obvious objective while modifying the Runge--Kutta coefficients
-is to retain a high order of accuracy, but this is not enough.
+is to retain a high order of accuracy, but this does not fully
+determine the choice of weights in general.
 To get a better understanding for the method we consider the
@@
 \end{equation}
-The values of the next step using the standard weights is
+The value of the next step using the standard weights is
 \begin{equation}
@@
 $\alpha$ can be narrowed down.
-An objective function is also needed to make the choice unique. This should be designed in a way to prefer weights that are close to the original weights.
+An objective function is also needed to make the choice unique. This should be
+designed in a way to prefer weights that are close to the original weights.
+
+We see that the choice of $\bt$ is subject to linear equality and inequality
+constraints.  If we choose a linear objective function, the resulting problem
+for finding the modified weights is a linear program, which can be efficiently
+solved by standard algorithms.  A natural choice of objective function is
+$$
+\text{minimize } \|\bt - b\|_1.
+$$
+The resulting problem can be phrased as an LP by using slack variables.
+In general, this LP may not have a solution; we can relax the constraints
+by requiring a lower order of consistency than the design order of the
+method.  These choices and alternatives will be considered in Section~\ref{sec:LP}.
+
+In contrast to other projection methods
+\cite{shampine1986conservation,sandu2001positive},
+minimizing the deviation of the weights instead of the deviation
+of the projected solution is computationally much more efficient
+for large systems, arising for example in the discretization of
+PDEs.
+
+\subsection{Example II}\label{sec:example_reac}
+
+To illustrate the usage of the method we consider the reaction system
+\cite{kopecz_comparison_2019}
+\begin{subequations}
+\label{eq:Reaction}
+\begin{align}
+u_1' &= 0.01u_2 + 0.01 u_3 +0.003u_4 - \frac{u_1 u_2}{0.01+u_1}, \\
+u_2' &= \frac{u_1u_2}{0.01+u_1}-0.01 u_2-0.5(1-\exp(-1.21 u_2^2)) u_3 -0.05 u_2, \\
+u_3' &= 0.5(1-\exp(-1.21u_2^2)) u_3 - 0.01 u_3 -0.02 u_3, \\
+u_4' &=0.05 u_2 + 0.02 u_3 + 0.003u_4,
+\end{align}
+\end{subequations}
+with initial conditions
+\begin{equation}
+u(0) = (8,2,1,4)^T.
+\end{equation}
+When solved with the Cash-Karp (CK5) \cite{cash1990variable} method and $\dt = 0.005$ the approximated solution contains negative values. This causes qualitatively wrong solutions to the problem.
+In Figure\,\ref{fig:exampleI} the obtained results are plotted with dashed lines.
+At $t=1.905$ the value of $u_1$ gets negative. This leads to a diverging solution.
+
+Now the weights are adapted. The adapted weights are of 4th order. The results are also plotted in Figure\,\ref{fig:exampleI}, with solid lines.
+The positivity constraint is now fulfilled. A qualitatively correct solution is obtained.
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=0.75\textwidth]{plots/exampleI.pdf}
+    \caption{Numerical approximation of the reaction problem \eqref{eq:Reaction} computed with CK5 and $\dt = 0.005$.
+The dashed lines are the approximations obtained without the adaption of the weights.
+The lower plot shows the adaptation of the weights.
+}
+    \label{fig:exampleI}
+\end{figure}
+
+The difference $\|\bt-b\|$ is also plotted in Figure\,\ref{fig:exampleI}.
+No modification of the weights is required for $t<1.905$. At $t=1.905$ the
+weights are first adapted to ensure the positivity of the solution. For
+$t>2.63$ the original set of weights again lead to a positive solution,
+and no further modification is necessary.
+
+\section{Selection of modified weights}\label{sec:LP}
+
+In this section, we consider further the formulation of the LP to choose
+the modified weights $\bt$.  In particular, we focus on the choice
+of objective function and how to relax the constraints to ensure that
+a feasible solution exists.
 
@@
 $\bt$ could lead to the loss of these desirable properties.
-In order to preserve good properties of the method, we use as
+In order to preserve as much as possible the good properties of the method, we use as
 objective function $\|\bt - b\|_1$.  This has the additional benefit
@@
 
-\section{Properties of adaptive RKMs} \label{sec:integration}
+\subsection{Reduction of number of positivity constraints}
+The number of positivity constraints implied by \eqref{eq:ut}
+is equal to $m$, the number of ODEs being solved.  This number
+may be very large, for instance if the system is a semi-discretization
+of a PDE.  This makes solution of the LP very costly.
+But in most cases, positivity is violated only for a very small subset $h \subseteq \{1, \dots, m\}$ of
+the solution components.  We can solve a much less expensive LP
+by replacing \eqref{eq:ut} with
+\begin{equation}
+u_i^n + \dt \sum_{j=0}^s F_{i,j}  b_j  \geq 0 \quad \forall i \in h \subseteq \{1,\dots,m \}.
+\end{equation}
+Of course, it must be checked that the solution of the resulting LP still
+satisfies the full set of constraints \eqref{eq:ut}.  In practice, we
+have found the following approach to be effective.
+First, set
+$$h_0 = \{ i \in \{1,\dots,m \} \ | \  u_i^{n+1}  < 0 \}.$$
+Solve the LP and let $\ut^{n+1}$ denote the new solution.
+If $\ut^{n+1}$ satisfies \eqref{eq:ut}, accept this as the new
+solution; otherwise, repeatedly take
+$$h_{a+1} = \{ i \in \{1,\dots,m \}|  \ut_i^{n+1}  < 0 \} \cup h_{a}$$
+until $\ut^{n+1}$ is found to satisfy \eqref{eq:ut}.
+In the examples we have studied, this approach was found to always converge in
+at most 2 iterations.
+%(Note: A approach on this would be to make a educated guess which indices could get a problem with positivity. A possible estimate would be $\frac{u_i^{n+1}}{max(K_{(i,0)}, \cdots ,K_{(i,0)})} $ )
+
+When enforcing a maximum value, the number of constraints can be reduced using the same technique. When enforcing both maximum and minimum values two separate sets of active constraints are used.
+In this case it is important to update these sets simultaneously.
+
+\subsection{Summary of the algorithm}
+
+Our proposed method is summarized in Algorithm\,\ref{alg:Adaption}.
+
+%To calculate the next solution $u^{n+1}$ the following is algorithm is executed. This is repeated until the final time $t_{end}$ is reached.
+%
+%At first, a step size $\dt$ is selected using a step size control algorithm or a fixed step size is used.
+%With this $\dt$ the stage values $y_i$ and stage derivatives $f_i$ are calculated.
+%Afterwards, $u^{n+1}$ is computed using the original weights $b$.
+%If the positivity is not violated, this solution is used to update the solution.
+%If $u^{n+1}$ contains negative values, the weights are adapted using the following algorithm.
+%
+%Start with $\pt$ equal to the highest order $p_{start}$ such that there are degrees of freedom for the weights $\bt$.
+%Apply an LP solver to the optimization
+%problem.
+%If the LP is infeasible, reduce the order $\tilde p$ by one.
+%If the LP is feasible, test whether the solution is acceptable.
+%At first, test whether the new $u^{n+1}$ satisfies the constraints.
+%This is necessary because sometimes the applied LP solvers incorrectly
+%identify the problem as feasible.
+%Afterwards the perturbation is calculated using \eqref{eq:Err}.
+%If the perturbation is bigger than a given tolerance, the
+%new weights $\bt$ are also rejected.
+%If the new set of weights is rejected in one of the tests the order
+%$\tilde p$ is reduced by one.
+%As soon as the algorithm finds suitable weights, these are used to update the solution.
+%If all acceptable orders are tried and no weights were found, the step hast to be rejected.
+
+\begin{algorithm}[ht]
+\begin{algorithmic}[1]
+\While{$t<t_{end}$}
+\State Choose $\Delta t$
+\State Calculate $F = (f_1, \dots, f_s)$ \eqref{eq:stagevalues}
+\State Set $u^{n+1}=u^n + \Delta t Fb$ \eqref{eq:Combination}
+\If {$u^{n+1} \geq 0$}
+	\State $success = True$
+\Else
+	\State{$\tilde p = p_{start}$}
+	\While{$\tilde p \geq p_{min}$}
+		\State Solve LP \eqref{eq:free-adaptation}
+		\If {LP Feasible \textbf{and} result correct}
+			\State $\delta = \| \Delta t F(\bt-b)\|$
+			\If {$\delta < tol$}
+				\State $u^{n+1}=u^n+\Delta t F \tilde b$
+				\State $success = True$
+				\State \textbf{break}
+			\EndIf
+		\EndIf
+		\If {$\tilde p = p_{min}$}
+			\State $sucess=False$
+		\EndIf
+		\State $\tilde p = \tilde p -1$
+	\EndWhile
+\EndIf
+\If {$sucess$}
+% 	\State $u^n=u^{n+1},t^n=t^{n+1}+\Delta t,n=n+1$
+	\State Set $n=n+1$ and continue
+\EndIf
+\EndWhile
+\end{algorithmic}
+\caption{Pseudocode for the algorithm using a free adaption of weights.}
+\label{alg:Adaption}
+\end{algorithm}
+
+\section{Properties of adaptive RKMs}
 
@@
 
-\subsection{Choice of baseline method}
+\subsection{Choice of baseline method} \label{sec:integration}
 An important property of the baseline method is the existence of embedded methods and the degrees of freedom for the weights $\bt$.
 As noted in Section~\ref{sec:OrderCond} the number of stages has to be higher than the order.
-Therefore, explicit RKMs are interesting, because an explicit RKM of order $p$ requires at least $p$ stages.
-Fully implicit RKMs can achieve higher orders.
-To fulfill $p<s$ the order has to be reduced drastically. Therefore, these methods are not good candidates for the adaptation.
-As stated in \cite{norsett_attainable_1977} the highest achievable order of a diagonally implicit RKM is $s+1$. This condition is far closer to the condition required for the adaptation.
-Another important property is the number of degrees of freedom for the choice of the new weights.
-These can be calculated using $s-\mathrm{rank}(Q_{\pt})$. %$\mathrm{dim}(\mathrm{ker}(Q))$.
-The number of degrees of freedom for different explicit methods are shown in Table\,\ref{table:DOF_exp}.
+It is natural to use explicit and diagonally implicit methods, both for their efficiency and because
+the order need not be reduced as much in order to satisfy the condition $\pt<s$.
+For a given method and reduced order $\pt$, the number of degrees of freedom for the choice of the new weights
+is given by $s-\mathrm{rank}(Q_{\pt})$. %$\mathrm{dim}(\mathrm{ker}(Q))$.
+The resulting number of degrees of freedom is shown in Table\,\ref{table:DOF_exp}
+for some explicit methods and in Table\,\ref{table:DOF_imp} for several implicit
+methods.
 
@@
 
-For all methods with the number of stages equal to the order of the method, there are no degrees of freedom without reducing the order.
-If the classical RK4 method is used the order has to be reduced more because the RK4 method does not have embedded methods of order~3.
-In contrast to this, some methods with $s > p$ have degrees of freedom even without reducing the order.
+For explicit methods with the number of stages equal to the order of the
+method, the order must be reduced in order to allow any freedom in the weights.
+If the classical RK4 method is used the order has to be reduced more because
+the RK4 method does not have embedded methods of order~3.
+In contrast to this, some methods with $s > p$ admit changes to the weights without reducing the order.
 An example of this is SSPRK(10,4), that has 4 degrees of freedom for $\pt = p$.
-For Cash--Karp RK5 and Dormand--Prince RK5 the order has to be reduced to get degrees of freedom even though the number of stages is higher than the order.
-
-In Table\,\ref{table:DOF_imp}, the same properties of some implicit
-methods are reported.
-We can see that the fully implicit methods Lobatto~IIIC4 and Radau~IIA3 require a drastic reduction of the order as expected.
+For Cash--Karp RK5 and Dormand--Prince RK5, even though the number of stages is higher than the order,
+the order must be reduced in order to allow any modification of the weights.
+
+Regarding implicit methods,
+we can see that the fully implicit methods Lobatto~IIIC4 and Radau~IIA3 require a drastic reduction of the order, as expected.
 The diagonally implicit SDIRK(5,4) method only requires an order reduction of one to get one degree of freedom for the weights.
@@
 
-Another important property is the stability region.
-The adaption of the weights does not solve issues with stability when used with a too big $\dt$. Therefore, the RK method has to be stable for large $\dt$.
-
-A third property is the existence of a positive solution.
-This corresponds to the question whether there is an embedded first order method that ensures positivity.
-This is particularly interesting for implicit methods, because $\dt$ can be unrestricted by stability requirements.
-If we have an embedded first order method, we can ensure that there is always a positive solution, which might be of first order.
-This is the case for the Im-Euler extrapolation methods, because they contain a chain of BE steps.
-For explicit methods the FE method is always embedded, but a reduction of the step size might be required.
+It is also desirable that the baseline method have a large stability region.
+
+Note that for any diagonally implicit method, by allowing the order to be reduced to one
+we can guarantee the existence of a solution to the LP, since the backward Euler method
+is unconditionally positive.  For explicit methods, reducing the order to one is guaranteed
+to yield a solution of the LP only if the step size is small enough.
 
 \subsection{Error detection and approximation}\label{sec:error}
-For this approach it is difficult to give some mathematically rigorous statement regarding the convergence of the method, because the RKM is only adapted if the step sizes are large.
-For sufficiently small step sizes the solution of the unadapted RKM is positive. Therefore the original weights are used.
-We have to describe the behavior outside the asymptotic regime.
-%The main way to ensure that the RKM with adapted b converges to the correct solution is to ensure that the weights approaches $b_{orig}$ for $\dt \to 0$.
-For this the weights it is already known that the numeric solution converges to the exact solution.
-%This is made sure by the property of the Objective function that
-%\begin{equation}
-%\mathrm{argmin}(f_{optim}(b)) = b_{orig}
-%\end{equation}
+Stability analysis for the proposed approach is very challenging, since in principle
+a different method may be used at every step.  At the same time, as long as the exact
+solution is positive, we expect that as the step size goes to zero, eventually no modification
+of the weights will be required and the convergence of the unmodified method will be observed.
+This holds true in the examples shown in Section \ref{sec:Numeric_Results}.  We are thus more concerned with
+the behavior of the modified method outside the asymptotic convergence regime.
+
 To approximate the error of a new step we propose the following approximation of the local error:
-
 \begin{align}
 err = \|u(t^{n+1})-\tilde u^{n+1}\| &= \|u(t^{n+1}) - (u^{n+1}+\dt F(\bt-b))\| \\
- &\leq \underbrace{\|u(t^{n+1})-u^{n+1}\|}_{\approx err_T}+\underbrace{\|\dt F(\bt-b)\|}_{= perturb} \label{eq:Err}
+ &\leq \underbrace{\|u(t^{n+1})-u^{n+1}\|}_{\approx err_T}+\underbrace{\|\dt F(\bt-b)\|}_{= \delta} \label{eq:Err}
 \end{align}
 
-The total error is split up in the truncation error $err_T$ and the perturbation $perturb$ using the triangle inequality.
+The total error is split up in the truncation error $err_T$ and the perturbation $\delta$ using the triangle inequality.
 The truncation error can be estimated using the standard error estimators $err_T = \| u^{n}_{b} - u^{n}_{\hat{b}} \|$.
-After adapting the weights, the perturbation is calculated. If the perturbation is larger than the tolerance the weights are rejected.
-Both values are added to get an approximation of the total error $err = err_T + perturb$.
-% The truncation error and the adaption error never have the same direction.
-% Therefore, the triangle inequality becomes a strict inequality and the real error is always smaller than the estimate. (Note: as long as the Truncation error estimator is right)
-%\subsubsection{Steppsize Control}
-%An important part of an RKM method is the ability to approximate the error of the solution to use it to control the step size.
-%The truncation error $|u(t^n)-u^n_{b_{orig}}|$ is usually approximated with an standard error estimator $err_T = | u^{n}_{b_{orig}} - u^{n}_{\hat{b}} |$.
- %The parameter $w_a \leq 1$ is used to account for the fact that the real error is smaller than $err = err_T +err_{adapt}$.
-The factor also improves the stability of the step size control because $err_{adapt}$ is not as smooth as $err_T$, especially if negative values only occur for a small number of steps.
-%If the LP was infeasible and no weights were found the $err_{approx}$ is set to a custom value.
-This value should be chosen sufficiently large to ensure that the next step take is smaller, but still not too large to keep the step size control stable.
-This type of error estimation is easy to implement because it can be easily incorporated in an existing step size control and takes advantage of the standard error approximation.
-
-%To make sure that the new adapted solution is still reasonable we measure the deviation from the Original method.
-In the worst case the errors add up.
-(todo: here some math formula...
-(We know that cannot be the case because we already know that we will reduce the error for the quantity we will make positive, still we can use it as a upper bound.))
+After adapting the weights, the perturbation is calculated. If the perturbation is larger than the tolerance, the weights are rejected.
+The two values are added to get an approximation of the total error $err = err_T + \delta$.
+This type of error estimation is easy to implement because it can be easily
+incorporated in an existing step size control and takes advantage of the
+standard error approximation.
 
@@
 
-The stability function is defined as a function $\CN \to \CN$ with
-\begin{equation}
-R(z) = 1 + zb^T(I - zA)^{-1}e,
+Let the stability function be denoted by $R(z): \CN \to \CN$.
+%\begin{equation}
+%R(z) = 1 + zb^T(I - zA)^{-1}e,
+%\end{equation}
+Since we intend to vary the weights, we view $R(z)$
+as a function parameterized by the weight vector $b$:
+\begin{equation}
+R_b(z) = 1 + zb^T(I - zA)^{-1}e,
 \end{equation}
 where $e = (1, \dots, 1)^T \in \R^s$.
-We can add $b$ as a second parameter and get a function
-$R\colon \R^s \times \CN  \to \CN$ with
-\begin{equation}
-R_b(z) = 1 + zb^T(I - zA)^{-1}e.
-\end{equation}
-The stability function is an affine-linear function of the weights.
+The stability function is an affine function of the weights.
 
@@
 it is easy to prove some properties of the stability region.
-\begin{theorem}
+\begin{theorem} \label{thm:stability-region}
   The stability region of a Runge--Kutta method $(A,\bt)$ where
@@
   $b^1, \dots, b^m \in \R^s$ (i.e.\ $g_i \in [0,1]$, $\sum_i g_i = 1$),
-  contains at least the intersection of the stability regions of
+  contains the intersection of the stability regions of
   the methods $(A,b^1), \dots, (A,b^m)$.
@@
 \subsubsection{Stability of free adaptation}
-If the weights is adapted freely there is no such proof.
-A useful property of the resulting stability function is that it is similar to the stability function of the baseline method.
+If the weights are adapted freely, in general we have no result like Theorem \ref{thm:stability-region}.
+Still, if the change in the weights is small then the resulting stability
+function is by some measure similar to the stability function of the baseline
+method.
 
@@
 \end{proof}
-This shows why the objective $\min \| \bt - b \|_1 $ is a good
-choice to control the change of the region of absolute stability,
+This result suggests that the objective $\min \| \bt - b \|_1 $ is an
+appropriate choice to control the change of the region of absolute stability,
 in particular for explicit methods for which
@@
 in $|z|$.
-
-\section{Implementation aspects of the algorithm}\label{sec:imple}
-
-Having described the mathematical idea of our approach,
-we discuss some practical implementation details of the
-proposed algorithm in the following.
-
-\subsection{Simplification of positivity constraints}
-The number of positivity constraints implied by \eqref{eq:ut}
-is equal to $m$, the number of ODEs being solved.  This number
-may be very large, for instance if the system is a semi-discretization
-of a PDE.  This makes solution of the LP very costly.
-But in most cases, positivity is violated only for a very small subset $h \subseteq \{1, \dots, m\}$ of
-the solution components.  We can solve a much less expensive LP
-by replacing \eqref{eq:ut} with
-\begin{equation}
-u_i^n + \dt \sum_{j=0}^s F_{i,j}  b_j  \geq 0 \quad \forall i \in h \subseteq \{1,\dots,m \}.
-\end{equation}
-Of course, it must be checked that the solution of the resulting LP still
-satisfies the full set of constraints \eqref{eq:ut}.  In practice, we
-have found the following approach to be effective.
-First, set
-$$h_0 = \{ i \in \{1,\dots,m \} \ | \  u_i^{n+1}  < 0 \}.$$
-Solve the LP and let $\ut^{n+1}$ denote the new solution.
-If $\ut^{n+1}$ satisfies \eqref{eq:ut}, accept this as the new
-solution; otherwise, repeatedly take
-$$h_{a+1} = \{ i \in \{1,\dots,m \}|  \ut_i^{n+1}  < 0 \} \cup h_{a}$$
-until $\ut^{n+1}$ is found to satisfy \eqref{eq:ut}.
-In the examples we have studied, this approach was found to always converge in
-at most X iterations.
-%(Note: A approach on this would be to make a educated guess which indices could get a problem with positivity. A possible estimate would be $\frac{u_i^{n+1}}{max(K_{(i,0)}, \cdots ,K_{(i,0)})} $ )
-
-When enforcing a maximum value, the number of constraints can be reduced using the same technique. When enforcing both maximum and minimum values two separate sets of active constraints are used.
-In this case it is important to update these sets simultaneously.
-
-\subsection{Implementation}
-
-In this section an algorithm to calculate the solution of an ODE using the adaptive RKM is proposed.
-The pseudocode is given in Algorithm\,\ref{alg:Adaption}.
-
-To calculate the next solution $u^{n+1}$ the following is algorithm is executed. This is repeated until the final time $t_{end}$ is reached.
-
-At first, a step size $\dt$ is selected using a step size control algorithm or a fixed step size is used.
-With this $\dt$ the stage values $y_i$ and stage derivatives $f_i$ are calculated.
-Afterwards, $u^{n+1}$ is computed using the original weights $b$.
-If the positivity is not violated, this solution is used to update the solution.
-If $u^{n+1}$ contains negative values, the weights are adapted using the following algorithm.
-
-Start with $\pt$ equal to the highest order $p_{start}$ such that there are degrees of freedom for the weights $\bt$.
-Apply an LP solver to the optimization
-problem.
-If the LP is infeasible, reduce the order $\tilde p$ by one.
-If the LP is feasible, test whether the solution is acceptable.
-At first, test whether the new $u^{n+1}$ satisfies the constraints.
-This is necessary because sometimes the applied LP solvers incorrectly
-identify the problem as feasible.
-Afterwards the perturbation is calculated using \eqref{eq:Err}.
-If the perturbation is bigger than a given tolerance, the
-new weights $\bt$ are also rejected.
-If the new set of weights is rejected in one of the tests the order
-$\tilde p$ is reduced by one.
-As soon as the algorithm finds suitable weights, these are used to update the solution.
-If all acceptable orders are tried and no weights were found, the step hast to be rejected.
-
-\begin{algorithm}[ht]
-\begin{algorithmic}[1]
-\While{$t<t_{end}$}
-\State Choose $\Delta t$
-\State Calculate $F = (f_1, \dots, f_s)$ \eqref{eq:stagevalues}
-\State Set $u^{n+1}=u^n + \Delta t Fb$ \eqref{eq:Combination}
-\If {$u^{n+1} \geq 0$}
-	\State $success = True$
-\Else
-	\State{$\tilde p = p_{start}$}
-	\While{$\tilde p \geq p_{min}$}
-		\State Solve LP \eqref{eq:free-adaptation}
-		\If {LP Feasible \textbf{and} result correct}
-			\State $perturb = \| \Delta t F(\bt-b)\|$
-			\If {$perturb < tol$}
-				\State $u^{n+1}=u^n+\Delta t F \tilde b$
-				\State $success = True$
-				\State \textbf{break}
-			\EndIf
-		\EndIf
-		\If {$\tilde p = p_{min}$}
-			\State $sucess=False$
-		\EndIf
-		\State $\tilde p = \tilde p -1$
-	\EndWhile
-\EndIf
-\If {$sucess$}
-% 	\State $u^n=u^{n+1},t^n=t^{n+1}+\Delta t,n=n+1$
-	\State Set $n=n+1$ and continue
-\EndIf
-\EndWhile
-\end{algorithmic}
-\caption{Pseudocode for the algorithm using a free adaption of weights.}
-\label{alg:Adaption}
-\end{algorithm}
 

### 2020-05-04 nusslein_vsrp/latex/overview.tex
Remove reference to removed section.
Fix a few mistakes.

@@
 invariants can be very important. e.g.\ the total mass for a transport
-problem or in reaction systems. Preserve of linear
+problem or in reaction systems. Preservation of linear
 invariants has been shown to be an important advantage of RRK methods
@@
 Section~\ref{sec:integration} describes how the new approach can be used with
-different RKMs, how it can be combined adaptive error control, and how the
+different RKMs, how it can be combined with adaptive error control, and how the
 region of absolute stability can be approximated.
-Section~\ref{sec:imple} contains some details regarding efficient implementation 
-of the algorithm.
 In Section~\ref{sec:Numeric_Results} numerical results are given for multiple test problems.
@@
 
-The difference $\|\bt-b\|$ is also plotted in Figure\,\ref{fig:exampleI}.
+The difference $\|\bt-b\|_1$ is also plotted in Figure\,\ref{fig:exampleI}.
 No modification of the weights is required for $t<1.905$. At $t=1.905$ the
@@
 
-Note that for any diagonally implicit method, by allowing the order to be reduced to one
+Note that for many diagonally implicit methods, the first stage is a scaled backward Euler step.
+For such methods, by allowing the order to be reduced to one
 we can guarantee the existence of a solution to the LP, since the backward Euler method

### 2020-05-04 nusslein_vsrp/latex/overview.tex
WIP: further revision

@@
 \end{align}
-where $u: [0,T] \to \R^m$ is positive if
+where $u\colon [0,T] \to \R^m$ is positive if
 \begin{align} \label{continuous-positivity}
@@
 \end{align}
+Here and in the following, inequalities like $u \geq 0$ are meant componentwise.
 A sufficient condition for positivity of \eqref{ode} is
@@
 scaled by a scalar relaxation parameter in order to guarantee conservation or
-monotonicity of a desired functional; e.g. to conserve or dissipate energy or
+monotonicity of a desired functional; e.g.\ to conserve or dissipate energy or
 entropy
@@
 our approach preserves all linear invariants of the given ODE. These
-invariants can be very important. e.g.\ the total mass for a transport
+invariants can be very important, e.g.\ the total mass for a transport
 problem or in reaction systems. Preservation of linear
@@
 where the $j$th column of $F$ is $f_j$.
-We wish to impose the discrete analog of \eqref{continuous-positivity}; i.e.
+We wish to impose the discrete analog of \eqref{continuous-positivity}; i.e.\
 \begin{align} \label{positivity}
@@
 \end{equation}
-When solved with the Cash-Karp (CK5) \cite{cash1990variable} method and $\dt = 0.005$ the approximated solution contains negative values. This causes qualitatively wrong solutions to the problem.
+When solved with the Cash--Karp RK5 method \cite{cash1990variable} and $\dt = 0.005$ the approximated solution contains negative values. This causes qualitatively wrong solutions to the problem.
 In Figure\,\ref{fig:exampleI} the obtained results are plotted with dashed lines.
@@
     \includegraphics[width=0.75\textwidth]{plots/exampleI.pdf}
-    \caption{Numerical approximation of the reaction problem \eqref{eq:Reaction} computed with CK5 and $\dt = 0.005$.
+    \caption{Numerical approximation of the reaction problem \eqref{eq:Reaction} computed with Cash--Karp RK5 and $\dt = 0.005$.
 The dashed lines are the approximations obtained without the adaption of the weights.
@@
 
-The order conditions for an $s$-stage, order $p$ RKM are a set of equations depending on $A$, $b$ and
+The order conditions for an $s$-stage, order $p$ RKM are a set of equations depending on $A$, $b$, and
 $c$.  As mentioned already, if $A$ and $c$ are given, the order conditions are
@@
 \begin{enumerate}
-    \item Select in advance a set of desirable weight vectors $b^1, b^2, \dots, b^m$
+    \item Select in advance a set of desirable weight vectors $b^1, b^2, \dots, b^K$
         corresponding to known good methods,
@@
 We can formulate an LP using the first strategy (restriction to a convex set)
-as follows.  Let $B$ denote the matrix with columns $b^1, b^2, \dots, b^m$ and let
-$g\in{\mathbb R}^m$.  The LP is then as follows:
+as follows.  Let $B$ denote the matrix with columns $b^1, b^2, \dots, b^K$ and let
+$g\in \R^K$.  The LP is then as follows:
 
@@
 \bt & = Bg, \\
-0 & \le g_i \le 1, \\
-\sum_i g_i &= 1, \\
+0 & \le g_k \le 1, \\
+\sum_{k=1}^K g_k &= 1, \\
 u^{n+1}&=u^n+\dt F \bt \geq 0.
@@
 for some explicit methods and in Table\,\ref{table:DOF_imp} for several implicit
-methods.
+methods. The backward Euler extrapolation methods use the harmonic sequence
+as described in \cite[Section~II.9]{hairer_solving_1993} and
+\cite[Section~IV.9]{hairer_solving_1996}.
 
@@
 %  SSPRK(10,4)& s=10&9&8&6&4& - & -  \\
-%  Cash-Karp RK5(4)6& s=6&5&4&2&1&0& -  \\
+%  Cash--Karp RK5(4)6& s=6&5&4&2&1&0& -  \\
 %  Dormand-Prince RK5(4)7& s=7&6&5&3&1&0& -  \\
@@
 %   Im-Euler 2& s=3&2&1& - & - & - & -  \\
-%   Im-Euler~3& s=6&5&4&2& - & - & -  \\
+%   Im-Euler 3& s=6&5&4&2& - & - & -  \\
 %   Im-Euler 4& s=10&9&8&6&3& - & -  \\
@@
     \midrule
-    Implicit Euler& 1&0& --- & --- & --- & --- & ---  \\
+    Backward Euler& 1&0& --- & --- & --- & --- & ---  \\
     Lobatto~IIIC4 \cite{chipman1971stable} & 4&3&2&1&0&0&0 \\
@@
     TR-BDF2 \cite{bank1985transient} & 3&2&1& --- & --- & --- & ---  \\
-    Extrapolation Im-Euler 2& 3&2&1& --- & --- & --- & ---  \\
-    Extrapolation Im-Euler~3& 6&5&4&2& --- & --- & ---  \\
-    Extrapolation Im-Euler 4& 10&9&8&6&3& --- & ---  \\
+    Extrapolation BE~2 \cite[Sec.~II.9]{hairer_solving_1993} & 3&2&1& --- & --- & --- & ---  \\
+    Extrapolation BE~3 \cite[Sec.~II.9]{hairer_solving_1993} & 6&5&4&2& --- & --- & ---  \\
+    Extrapolation BE~4 \cite[Sec.~II.9]{hairer_solving_1993} & 10&9&8&6&3& --- & ---  \\
     \bottomrule
@@
 The TR-BDF2 method even allows adaptations without reducing the order.
-The implicit Euler extrapolation methods also exhibit degrees of freedom without a reduction of the order.
+The backward Euler extrapolation methods also exhibit degrees of freedom without a reduction of the order.
 
@@
 err = \|u(t^{n+1})-\tilde u^{n+1}\| &= \|u(t^{n+1}) - (u^{n+1}+\dt F(\bt-b))\| \\
- &\leq \underbrace{\|u(t^{n+1})-u^{n+1}\|}_{\approx err_T}+\underbrace{\|\dt F(\bt-b)\|}_{= \delta} \label{eq:Err}
+ &\leq \underbrace{\|u(t^{n+1})-u^{n+1}\|}_{\approx err_T}+\underbrace{\|\dt F(\bt-b)\|}_{= \delta}. \label{eq:Err}
 \end{align}
@@
 The weights are taken from the example in Section~\ref{sec:Ex_expl}.
-In Figure\,\ref{fig:stab_ex3} the stability regions of the Im-Euler~3
-extrapolation method and the embedded chain of three BE steps are
+In Figure\,\ref{fig:stab_ex3} the stability regions of the BE~3
+extrapolation method and the embedded chain of three BE steps with
+time step $\dt/3$ are
 plotted. Additionally the stability regions of convex combinations
@@
          \includegraphics[width=\textwidth]{plots/stab_ex3.pdf}
-         \caption{Im-Euler~3 extrapolation method and embedded chain
+         \caption{BE~3 extrapolation method and embedded chain
                   of three BE steps with convex combinations of both
@@
 
-The ODE is solved using the Im-Euler~3 extrapolation method.
+The ODE is solved using the BE~3 extrapolation method.
 For large $\dt$ the method computes negative values for $u^1$.
@@
 \caption{Numerical results for the diffusion problem \eqref{eq:diffusion} and the
-         adapted Im-Euler~3 extrapolation method.}
+         adapted BE~3 extrapolation method.}
 \end{figure}
 
-In Figure\,\ref{fig:conv_impl}, the convergence is shown for the unaltered Im-Euler~3 extrapolation method (ex3, potentially resulting in negative values), the adaptive method with free adaptation (ex3a) and the adapted method using convex adaptations (ex3c).
+In Figure\,\ref{fig:conv_impl}, the convergence is shown for the unaltered BE~3 extrapolation method (ex3, potentially resulting in negative values), the adaptive method with free adaptation (ex3a) and the adapted method using convex adaptations (ex3c).
 Also, the BE method is plotted. It is only of 1st order but preserves positivity for all $\dt$.
@@
 \includegraphics[width=0.8\textwidth]{plots/conv_heat.pdf}
-\caption{Convergence result of baseline and adapted Im-Euler~3
+\caption{Convergence result of baseline and adapted BE~3
          extrapolation methods for the diffusion problem \eqref{eq:diffusion}.}
@@
 This leads to a positivity preserving system of ODEs which conserves the total mass $\sum u$.
-The computation is done using the Im-Euler~3 extrapolation
+The computation is done using the BE~3 extrapolation
 method with free adaptation.
@@
 The system is solved in the time from $t_{0} = \SI{12}{h}$ to $t_{end} = \SI{84}{h}$
-using the Im-Euler~3 extrapolation method with free adaptation and step size
+using the BE~3 extrapolation method with free adaptation and step size
 control.
@@
 The adapted solution is close to the reference solution obtained
-with the unadapted Im-Euler~3 extrapolation method and a higher accuracy.
+with the unadapted BE~3 extrapolation method and a higher accuracy.
 For this solution, 249 steps were computed; two of these were rejected due to

### 2020-05-04 nusslein_vsrp/latex/overview.tex

@@
 
-Our proposed method is summarized in Algorithm\,\ref{alg:Adaption}.
+Our proposed method to solve a positive initial value problem \eqref{ode}
+is summarized in Algorithm\,\ref{alg:Adaption}.
 
@@
 \begin{algorithmic}[1]
+\State Initialize $n \leftarrow 0$, $u^n \leftarrow u_0$, $t \leftarrow 0$
 \While{$t<t_{end}$}
-\State Choose $\Delta t$
-\State Calculate $F = (f_1, \dots, f_s)$ \eqref{eq:stagevalues}
-\State Set $u^{n+1}=u^n + \Delta t Fb$ \eqref{eq:Combination}
+\State Choose $\Delta t$ (fixed or via an adaptive stepsize control)
+\State Calculate $F = (f_1, \dots, f_s)$, cf.\ \eqref{eq:stagevalues}
+\State $u^{n+1} \leftarrow u^n + \Delta t Fb$, cf.\ \eqref{eq:Combination}
 \If {$u^{n+1} \geq 0$}
-	\State $success = True$
+	\State $success \leftarrow True$
 \Else
-	\State{$\tilde p = p_{start}$}
-	\While{$\tilde p \geq p_{min}$}
+	\State{$\pt \leftarrow p_{start}$}
+	\While{$\pt \geq p_{min}$}
 		\State Solve LP \eqref{eq:free-adaptation}
-		\If {LP Feasible \textbf{and} result correct}
-			\State $\delta = \| \Delta t F(\bt-b)\|$
+		\If {LP is feasible \textbf{and} result is correct} \todo{TODO: define these}
+			\State $\delta \leftarrow \| \Delta t F(\bt-b)\|$
 			\If {$\delta < tol$}
-				\State $u^{n+1}=u^n+\Delta t F \tilde b$
-				\State $success = True$
+				\State $u^{n+1} \leftarrow u^n+\Delta t F \tilde b$
+				\State $success \leftarrow True$
 				\State \textbf{break}
@@
 		\EndIf
-		\If {$\tilde p = p_{min}$}
-			\State $sucess=False$
+		\If {$\pt = p_{min}$}
+			\State $sucess \leftarrow False$
 		\EndIf
-		\State $\tilde p = \tilde p -1$
+		\State $\pt \leftarrow \pt - 1$
 	\EndWhile
@@
 \If {$sucess$}
-% 	\State $u^n=u^{n+1},t^n=t^{n+1}+\Delta t,n=n+1$
-	\State Set $n=n+1$ and continue
+	\State $t \leftarrow t + \dt$, $n \leftarrow n+1$ and continue
+\Else
+  \State \todo{TODO}
 \EndIf

### 2020-05-05 nusslein_vsrp/latex/overview.tex
Changed Pseudocode cf. #22
additional minor revision

@@
 \State Choose $\Delta t$ (fixed or via an adaptive stepsize control)
-\State Calculate $F = (f_1, \dots, f_s)$, cf.\ \eqref{eq:stagevalues}
+\State Calculate $F = (f_1, \dots, f_s)$, cf.\ \eqref{eq:stagevalues} \label{line:calc_stge}
 \State $u^{n+1} \leftarrow u^n + \Delta t Fb$, cf.\ \eqref{eq:Combination}
 \If {$u^{n+1} \geq 0$}
-	\State $success \leftarrow True$
+	\State  \textbf{GOTO} line\,\ref{line:update}
 \Else
@@
 		\State Solve LP \eqref{eq:free-adaptation}
-		\If {LP is feasible \textbf{and} result is correct} \todo{TODO: define these}
+		\If {LP is feasible} 
 			\State $\delta \leftarrow \| \Delta t F(\bt-b)\|$
-			\If {$\delta < tol$}
+			\If {$\delta < tol_\delta$}
 				\State $u^{n+1} \leftarrow u^n+\Delta t F \tilde b$
-				\State $success \leftarrow True$
-				\State \textbf{break}
+				\State  \textbf{GOTO} line\,\ref{line:update} 
 			\EndIf
-		\EndIf
-		\If {$\pt = p_{min}$}
-			\State $sucess \leftarrow False$
 		\EndIf
@@
 	\EndWhile
+	\State Reduce $\dt$
+	\State \textbf{GOTO} line\,\ref{line:calc_stge} 
 \EndIf
-\If {$sucess$}
-	\State $t \leftarrow t + \dt$, $n \leftarrow n+1$ and continue
+\State Calculate $error$ cf. \ref{eq:Err} \label{line:update} 
+\If {$error \leq tol_{error}$}
+	\State $t \leftarrow t + \dt$, $n \leftarrow n+1$ 
 \Else
-  \State \todo{TODO}
+	\State Reduce $\dt$
+	\State \textbf{GOTO} line\,\ref{line:calc_stge} 
 \EndIf
@@
     \end{subfigure}
-    \caption{Solution of advection-diffusion-reaction problem \eqref{eq:ADR} at different times.}\label{fig:Sol_ADP}
+    \caption{Numerical solution of the advection-diffusion-reaction problem \eqref{eq:ADR} at different times.}\label{fig:Sol_ADP}
 \end{figure}

### 2020-05-07 nusslein_vsrp/latex/overview.tex
additional minor revision of the pseudocode
update legends to remove ex3; closes #23

@@
 \State Choose $\Delta t$ (fixed or via an adaptive stepsize control)
-\State Calculate $F = (f_1, \dots, f_s)$, cf.\ \eqref{eq:stagevalues} \label{line:calc_stge}
+\State Calculate $F = (f_1, \dots, f_s)$ according to \eqref{eq:stagevalues} \label{line:calc_stge}
 \State $u^{n+1} \leftarrow u^n + \Delta t Fb$, cf.\ \eqref{eq:Combination}
@@
 \EndIf
-\State Calculate $error$ cf. \ref{eq:Err} \label{line:update} 
+\State Estimate $error$ according to \eqref{eq:Err} \label{line:update}
 \If {$error \leq tol_{error}$}
@@
 
-In Figure\,\ref{fig:conv_impl}, the convergence is shown for the unaltered BE~3 extrapolation method (ex3, potentially resulting in negative values), the adaptive method with free adaptation (ex3a) and the adapted method using convex adaptations (ex3c).
-Also, the BE method is plotted. It is only of 1st order but preserves positivity for all $\dt$.
+In Figure\,\ref{fig:conv_impl}, the convergence is shown for the unaltered BE~3 extrapolation method (potentially resulting in negative values), the adaptive method with free adaptation, and the adapted method using convex adaptation.
+Additionally, results for the BE method are plotted. It is only of 1st order but preserves positivity for all $\dt$.
 For $\dt < \num{3e-5} $ the standard weights yield to a positive result. For larger $\dt$ the weights have to be adapted to ensure positivity.
-The ex3a shows similar convergence properties as the original method.
+The free adaptation results in similar convergence properties as the original method.
 This can be expected, because the adapted method is still of 3rd order.
-The ex3c has larger errors than the ex3a but leads to physical solutions for all time-steps.
+The convex adaptation yields larger errors than the free adaptation but leads to physical solutions for all time-steps.
 This is no surprise because the adapted RKM used for the first step is only of first order.
@@
 \centering
-\includegraphics[width=0.8\textwidth]{plots/conv_heat.pdf}
+\includegraphics[width=\textwidth]{plots/conv_heat.pdf}
 \caption{Convergence result of baseline and adapted BE~3

### 2020-05-07 nusslein_vsrp/utils.py
clean up reproducibility notebooks

@@ -298,8 +298,8 @@ def show_status(status):
     print('Succesfull:'); print(status['success'])
     #number of adaptations
     print('number of adaptations:');print(status['b'].count('c'))
-    #Number of step reqects
-    print('Number of step reqects:');print(status['b'].count('r'))
+    #Number of step rejects
+    print('Number of step rejects:');print(status['b'].count('r'))
 
     fig, axs = plt.subplots(5,1)
 
@@ -308,11 +308,11 @@ def show_status(status):
     axs[0].eventplot(np.nonzero(np.array(status['b'])=='c'), colors='blue', lineoffsets=0.5,
                     linelengths=0.5)
 
-    #plot the step reqects
+    #plot the step rejects
     axs[0].eventplot(np.nonzero(np.array(status['b'])=='r'), colors='red', lineoffsets=-0.5,
                     linelengths=0.5)
 
-    #plot step reqect caused by tol
+    #plot step reject caused by tol
     axs[0].eventplot(np.nonzero(np.array(status['sc'])=='r'), colors='red', lineoffsets=-1.5,
                     linelengths=0.5)
 

### 2020-05-10 nusslein_vsrp/latex/overview.tex
clean up reproducibility notebooks

@@
 \begin{equation} \label{eq:condition_ODE_pos}
-u_i=0 \implies f_i(t,u_1,\cdots,u_i,\cdots,u_n) \geq 0 \quad \forall {u \geq 0,}\; \ \  \forall {t\in[0,T]}.
+u_i=0 \implies f_i(t,[u_1,\cdots,u_i,\cdots,u_n]^T) \geq 0 \quad \forall {u \geq 0,}\; \ \  \forall {t\in[0,T]}.
 \end{equation}
@@
 \State Calculate $F = (f_1, \dots, f_s)$ according to \eqref{eq:stagevalues} \label{line:calc_stge}
-\State $u^{n+1} \leftarrow u^n + \Delta t Fb$, cf.\ \eqref{eq:Combination}
+\State $u^{n+1} \leftarrow u^n + \Delta t Fb$, according to \eqref{eq:Combination}
 \If {$u^{n+1} \geq 0$}
@@
 
-The total error is split up in the truncation error $err_T$ and the perturbation $\delta$ using the triangle inequality.
+The total error is split up in the truncation error and the perturbation $\delta$ using the triangle inequality.
 The truncation error can be estimated using the standard error estimators $err_T = \| u^{n}_{b} - u^{n}_{\hat{b}} \|$.
@@
 
-A very interesting class of methods are the implicit extrapolation methods.
-These are diagonally implicit and $p < s$.
-This means that they allow for changes of the weights without a reduction of the order.
+A very interesting class of methods are the implicit extrapolation methods. These allow changes of the weights without a reduction of the order, as discussed in Section\,\ref{sec:integration}.
 Moreover, all stage values are computed using the BE method. Hence, all intermediate stages are positive.
@@
 \centering
-\includegraphics[width=0.85\textwidth]{plots/ADP_stepsize,b.pdf}
+\includegraphics[width=0.85\textwidth]{plots/ADP_stepsize_b.pdf}
 \caption{Statistics of computation of the advection-diffusion-reaction problem \eqref{eq:ADR}.}
@@
 \centering
-\includegraphics[width=0.85\textwidth]{plots/Stratospheric_stepsize,b.pdf}
+\includegraphics[width=0.85\textwidth]{plots/Stratospheric_stepsize_b.pdf}
 \caption{Statistics of computation of the stratospheric reaction system

### 2020-05-11 nusslein_vsrp/latex/overview.tex
add address etc.

@@
 
+\newcommand{\orcid}[1]{ORCID:~\href{https://orcid.org/#1}{#1}}
+\usepackage{authblk}
+\renewcommand\Affilfont{\fontsize{8}{10.8}\selectfont}
+
+\newenvironment{keywords}{\par\textbf{Key words.}}{\par}
+\newenvironment{AMS}{\par\textbf{AMS subject classification.}}{\par}
+
 \title{Positivity-Preserving Adaptive Runge--Kutta Methods}
-\author{Stephan Nüßlein \and Hendrik Ranocha \and David I. Ketcheson}
+\author[1]{Stephan Nüßlein\thanks{\orcid{0000-0002-2455-4222}}}
+\author[1]{Hendrik Ranocha\thanks{\orcid{0000-0002-3456-2277}}}
+\author[1]{David I. Ketcheson\thanks{\orcid{0000-0002-1212-126X}}}
+\affil[1]{%
+King Abdullah University of Science and Technology (KAUST),
+Extreme Computing Research Center (ECRC),
+Computer Electrical and Mathematical Science and Engineering Division (CEMSE),
+Thuwal, 23955-6900, Saudi Arabia}
+% \date{May 11, 2020} %TODO
 
 \makeatletter
+\hypersetup{pdfauthor={Stephan Nüßlein, Hendrik Ranocha, David I. Ketcheson}}
+\hypersetup{pdftitle={\@title}}
 \makeatother
@@
 \end{abstract}
+
+\begin{keywords}
+  positivity preserving,
+  bound preserving,
+  Runge--Kutta methods,
+  linear programming
+\end{keywords}
+
+\begin{AMS}
+  65L06,  % NA, ODEs: Multistep, Runge-Kutta and extrapolation methods
+  65L20,  % NA, ODEs: Stability and convergence of numerical methods
+  65M12   % NA, PDEs, IVPs, IBVPs: Stability and convergence of numerical methods
+\end{AMS}
 

### 2020-05-11 nusslein_vsrp/latex/overview.tex
remove ECRC

@@
 King Abdullah University of Science and Technology (KAUST),
-Extreme Computing Research Center (ECRC),
 Computer Electrical and Mathematical Science and Engineering Division (CEMSE),

### 2020-05-11 nusslein_vsrp/latex/overview.tex
Changed dt of Advection-decay equation
The original dt was set to dt = 0.02 = 2*dx which was too high to expect stability. I have changed it to dt = 0.015 = 1.5*dx which is inside the stability interval

@@
 We use the Dormand--Prince RK5 method and adapt the weights using the free adaptation.
-In Figure\,\ref{fig:sol_AdDe} the results for $\Delta t = 0.02$ are plotted for different time steps. We can see that the solution approaches an exponential function with $t \rightarrow \infty$.
-In Figure\,\ref{fig:weights_AdDe} the used weights are plotted. For $t\leq 0.6$ the weights are altered. For $t > 0.6$ the original weights lead to a positive solution.
+In Figure\,\ref{fig:sol_AdDe} the results for $\Delta t = 0.015$ are plotted for different time steps. We can see that the solution approaches an exponential function with $t \rightarrow \infty$.
+In Figure\,\ref{fig:weights_AdDe} the used weights are plotted. For $t\leq 0.375$ the weights are altered. For $t > 0.375$ the original weights lead to a positive solution.
 \begin{figure}

### 2021-01-30 nusslein_vsrp/latex/response_to_reviewers.tex
prepare response to reviewer

\documentclass[11pt]{letter}
\usepackage[USenglish]{babel}
\usepackage[utf8]{luainputenc}
\usepackage[a4paper]{geometry}
\usepackage[plainpages=false,pdfpagelabels,hidelinks,unicode]{hyperref}
\hypersetup{pdftitle={Response to Reviewers}}
\usepackage{enumitem}

\usepackage{xr} % for references to another document
\externaldocument{overview}

\usepackage{color}
\newcommand{\todo}[1]{{\Large \textcolor{red}{#1}}}

\newcommand{\revA}[1]{{\color{red}#1}}
\newcommand{\revB}[1]{{\color{blue}#1}}
\newcommand{\revC}[1]{{\color{cyan}#1}}

\usepackage{amssymb}
\usepackage{commath}

\usepackage{xspace}
\newcommand{\cf}[0]{{cf.\@}\xspace}
\newcommand{\eg}[0]{{e.g.\@}\xspace}
\newcommand{\ie}[0]{{i.e.\@}\xspace}
\newcommand{\etc}[0]{{etc.\@}\xspace}

% fonts
\usepackage{ifluatex}
\ifluatex
  \usepackage[no-math]{fontspec}
\else
  \usepackage[T1]{fontenc}
\fi
\usepackage{newpxtext,newpxmath}

% \date{June 16, 2020} %TODO: date

\begin{document}
\begin{letter}{}

\textbf{Response to Reviewers}

\opening{Dear Editor, Dear Reviewer,}

we are very grateful for the helpful comments on the manuscript
``Positivity-Preserving Adaptive Runge--Kutta Methods''.
We have revised the manuscript in accordance with the reviewer's suggestions.
Please find below a point-by-point description of these changes and reply to
the reviewer's comments.

\textbf{\large Reviewer \#1 \revA{(manuscript changes highlighted in red)}}
% TODO: Reviewer 1

\emph{%
This manuscript studies the efficiency of a new kind of Positivity-Preserving
Runge-Kutta Methods. The main idea in the new approach consists in adaptively
choosing the weights b of the RKM in a way that ensures positivity, or other
bound constraints, for problems satisfying the sufficient condition (1.4).
One of the main benefits of this new approach is that all linear invariants of
the problem are preserved. Other projection methods also preserve invariants,
but only if you explicitly know these invariants. If any component of the
numerical solution is negative, a linear program (LP) must be solved at
each step in order to choose the new weights, and this could be an extra time
consuming process. Nevertheless, the idea of minimizing the deviation of the
weights is simple and in some cases may be much more efficient, particularly
for large systems. The numerical experiments are well presented and quite
thorough including different stiff and non-stiff problems, with fixed and
adaptive step size implementations.
\\
Positivity-Preserving methods have been studied quite a lot in recent decades
and the authors have summarized really well the state of art in the introduction
by referring to Projection methods, SSPRK methods, MPRK methods and DSRK positive
methods. In the new schemes proposed the weights of some baseline RKM are adapted
in order to ensure positivity or other bounds. These new methods are valid both
for stiff and non-stiff problems satisfying the sufficient condition for
positivity (1.4). In addition, the authors clarify that for some positive
problems not satisfying (1.4) numerical tests did not show promising results.
I personally value the algorithm given in Section 3. Theorem 4.1 about the
stability region of the new adaptive Runge-Kutta method with convex adaptation
is quite interesting, particularly for A-stable methods. Lemma 4.2 with the
bound for the stability function of the adapted method is interesting too. The
paper is clear and well written, thus I would recommend this work to be published
provided that the authors address properly the points listed below.
}

\begin{enumerate}[label=\arabic*.]
  \setcounter{enumi}{1}
  \item Bound-preserving adaptive Runge–Kutta methods
  \begin{enumerate}[label=\alph*)]
    \item \emph{%
    In equation (2.14b), it is weird to read $-0.01 u_2$ and $- 0.05 u_2$
    in the same line. Is it an error? The same thing happens in equations (2.14c)
    with $u_3$, in (5.5b) with $u_2$, and in (5.5c) with $u_3$.
    }

    \revA{TODO} %TODO

    \item \emph{%
    In Figure 1, in the difference $\| \tilde b - b \|$, a tilde is missing.
    }

    \revA{TODO} %TODO
  \end{enumerate}

  \item Selection of the modified weights
  \begin{enumerate}[label=\alph*)]
    \item \emph{%
    In section 3.2, two different approaches to adapt the weights are considered:
    Free adaptation and Convex adaptation. At the same time, it is explained that
    ``...we can additionally impose either or both of the following strategies \dots''.
    I think that the first strategy is nothing but the second approach (Convex
    adaptation). I suggest making this part clearer in order to avoid confusion
    between approaches and strategies.
    }

    \revA{TODO} %TODO
  \end{enumerate}

  \item Properties of adaptive RKMs
  \begin{enumerate}[label=\alph*)]
    \item \emph{%
    Figure 3a, with free adaptation of the weights $\tilde b$, does not give enough
    information about the stability regions of the adapted methods. In my opinion,
    it would be clearer if each stability curve was labeled with the value of
    the difference $\| \tilde b - b \|$, accordingly with the values in Figure 4b.
    In this way we could see the progress of the stability regions during the
    numerical integration process.
    }

    \revA{TODO} %TODO

    \item \emph{%
    In the first line of formula (4.6) the vector $(\tilde b - b)$ should be
     $(\tilde b - b)^T$.
    }

    \revA{TODO} %TODO
  \end{enumerate}

  \item Numerical experiments
  \begin{enumerate}[label=\alph*)]
    \item \emph{%
    In the third line of Section 5.1, RHS appears for the first time without
    explaining its meaning.
    }

    \revA{TODO} %TODO

    \item \emph{%
    In section 5.1, just before the formula (5.1), it is said ``Since the
    positivity of the stage values is not ensured, the right hand side has to be
    defined for negative values. If this is not the case, the right hand side
    has to to be adapted accordingly''. This is a bit confusing, particularly
    the text ``...has to be defined for negative values...''. I suggest that
    you make this part clearer.
    }

    \revA{TODO} %TODO

    \item \emph{%
    In formula (5.2), the last $u_{i-1}$, shouldn’t it be $u_i$?
    }

    \revA{TODO} %TODO

    \item \emph{%
    After formula (5.2) ``...In Figure 4a the results for $\Delta t = 0.015$
    are plotted for different time steps...'' should be ``...In Figure 4a the
    results for $\Delta t = 0.015$ are plotted for different values of time $t$...''
    }

    \revA{TODO} %TODO

    \item \emph{%
    Figure 5, about the non-stiff problem with fixed stepsize, is quite interesting
    and gives information about the convergence of the adapted mehtod. For
    $\Delta t = 0.0082$ the unaltered method leads to positive solutions. For
    $\Delta t = 0.015$ the original method leads to negative values in the
    interval $[0, 0.375]$ and the weights are altered by solving an LP problem.
    Is this extra time consuming a significant cost or could it be an economical
    alternative to using excessively small step sizes? A further explanation
    about this issue should be added.
    }

    \revA{TODO} %TODO

    \item \emph{%
    In Figures 6c and 6d, in the difference $\| \tilde b - b \|$ and also in the
    legend, a tilde is missing.
    }

    \revA{TODO} %TODO

    \item \emph{%
    In the third subplot of Figure 9, in the difference $\| \tilde b - b \|$,
    a tilde is missing.
    }

    \revA{TODO} %TODO

    \item \emph{%
    In the fourth subplot of Figure 11, in the difference $\| \tilde b - b \|$,
    a tilde is missing.
    }

    \revA{TODO} %TODO
  \end{enumerate}
\end{enumerate}

\end{letter}
\end{document}

### 2021-01-30 nusslein_vsrp/latex/response_to_reviewers.tex
WIP: response to reviewer 1

@@
 
-\textbf{Response to Reviewers}
+\textbf{Response to Reviewer}
 
@@
 
-we are very grateful for the helpful comments on the manuscript
+we are very grateful for the thorough review and helpful comments on our manuscript
 ``Positivity-Preserving Adaptive Runge--Kutta Methods''.
-We have revised the manuscript in accordance with the reviewer's suggestions.
+We revised the manuscript in accordance with the reviewer's suggestions.
 Please find below a point-by-point description of these changes and reply to
@@
 
-    \revA{TODO} %TODO
+    We fixed equation \eqref{eq:proof-of-lemma-free-adaptation}.
   \end{enumerate}
@@
 
-    \revA{TODO} %TODO
+    We changed this sentence and introduced the abbreviation RHS
+    [line~\ref{line:RHS}].
 
@@
 
-    \revA{TODO} %TODO
+    We fixed equation \eqref{eq:advection-decay-semidiscrete}.
 
@@
 
-    \revA{TODO} %TODO
+    We fixed this sentence [line~\ref{line:time-vs-dt}].
 

### 2021-01-30 nusslein_vsrp/latex/overview.tex
set date
prepare response to reviewer
WIP: response to reviewer 1

@@
 \fi
+
+% line numbering
+% see http://phaseportrait.blogspot.de/2007/08/lineno-and-amsmath-compatibility.html
+\usepackage{lineno}
+\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
+  \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
+  \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
+  \renewenvironment{#1}%
+     {\linenomath\csname old#1\endcsname}%
+     {\csname oldend#1\endcsname\endlinenomath}}%
+\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
+  \patchAmsMathEnvironmentForLineno{#1}%
+  \patchAmsMathEnvironmentForLineno{#1*}}%
+\AtBeginDocument{%
+\patchBothAmsMathEnvironmentsForLineno{equation}%
+\patchBothAmsMathEnvironmentsForLineno{align}%
+\patchBothAmsMathEnvironmentsForLineno{flalign}%
+\patchBothAmsMathEnvironmentsForLineno{alignat}%
+\patchBothAmsMathEnvironmentsForLineno{gather}%
+\patchBothAmsMathEnvironmentsForLineno{multline}%
+}
+\linenumbers %TODO: remove for arXiv version
 
@@
 \newcommand{\todo}[1]{{\Large{\color{red}{#1}}}}
+
+\newcommand{\revA}[1]{{\color{red}#1}}
+\newcommand{\revB}[1]{{\color{blue}#1}}
+\newcommand{\revC}[1]{{\color{cyan}#1}}
 
@@
 Thuwal, 23955-6900, Saudi Arabia}
-% \date{May 11, 2020} %TODO
+% \date{May 13, 2020} %TODO: date
 
@@
 A conclusion is given in Section~\ref{sec:conclusion}.
-
-%(TODO: We know that we cannot work around the limitation from \cite{hundsdorfer_numerical_2003}, Where would be a good place to put this?)
 
@@
 
-%TODO: It is not straightforward if there are no pole zero cancellations in the stability function.
-For explicit methods, the stability function is a polynomial o degree $\leq s$ (Hundsdorfer).
-Pole zero cancellations cannot happen for these.
-For implicit functions that are A-stable the poles have to be in the right halve plane.
-Here is starts to get interesting.
-Because the A-Stability of one method with a certain $A$ is not a sufficient condition that there are no Poles for all RKM with this A-Matrix.
-But it is possible (or easy) to prove that if $a_{ii} > 0 \forall i = 1,\cdots , s$ then there are only poles in the right halve plane where they do not matter for the further considerations.)
-
 \subsubsection{Stability of convex adaptation}
@@
   \begin{equation}
+  \label{eq:proof-of-lemma-free-adaptation}
   \begin{aligned}
@@
     | R_{b}(z) |
-    + | z (\bt - b) (\I - z A)^{-1} e|
+    + | z (\bt - b)\revA{^T} (\I - z A)^{-1} e|
     \\
@@
 reproduce the numerical examples reported here can be found in
-\cite{repository}. \todo{TODO: New repository with DOI} %TODO
+\cite{nusslein2020positivityRepro}.
 The methods are implemented in Python using NumPy/SciPy
@@
 First, adaptive RKMs based on explicit methods are tested on non-stiff problems with a fixed step size.
-When used with explicit methods the cost of solving the LP is significant because the computation of the stage derivatives only requires $s$ evaluations of the RHS.
+When used with explicit methods the cost of solving the LP is significant because the computation of the stage derivatives only requires $s$ evaluations of the
+\linelabel{line:RHS}\revA{right-hand side (RHS)}.
 For most of the linear test problems tried, the explicit methods yield to positive results.
@@
 \begin{equation}
-\frac{\mathrm d}{\mathrm d t} u_i = \frac{a}{\Delta x} \left( u_{i-1} - u_i \right) - K u_{i-1}.
+\label{eq:advection-decay-semidiscrete}
+\frac{\mathrm d}{\mathrm d t} u_i = \frac{a}{\Delta x} \left( u_{i-1} - u_i \right) - K \revA{u_{i}}.
 \end{equation}
@@
 We use the Dormand--Prince RK5 method and adapt the weights using the free adaptation.
-In Figure\,\ref{fig:sol_AdDe} the results for $\Delta t = 0.015$ are plotted for different time steps. We can see that the solution approaches an exponential function with $t \rightarrow \infty$.
+In Figure\,\ref{fig:sol_AdDe} the results for $\Delta t = 0.015$ are plotted for different \linelabel{line:time-vs-dt}\revA{of time $t$}.
+We can see that the solution approaches an exponential function with $t \rightarrow \infty$.
 In Figure\,\ref{fig:weights_AdDe} the used weights are plotted. For $t\leq 0.375$ the weights are altered. For $t > 0.375$ the original weights lead to a positive solution.
@@
 
-%also if exactnes is not the limiting factor but positivity
-
 \printbibliography

### 2021-01-30 nusslein_vsrp/latex/overview.tex
finished comments of reviewer 1 not requiring changes of the code

@@
 \end{equation}
-When solved with the Cash--Karp RK5 method \cite{cash1990variable} and $\dt = 0.005$ the approximated solution contains negative values. This causes qualitatively wrong solutions to the problem.
+\linelabel{line:prod-dest}\revA{Note that we wrote \eqref{eq:Reaction} as in
+\cite{kopecz_comparison_2019}, sometimes using multiple terms containing the same
+variables but with different constants, e.g.\ $-0.01 u_2 -0.05 u_2$ in the
+time derivative of $u_2$. This notation is useful to see the structure of
+a production-destruction system which is exploited for positivity-preserving
+(modified) Patankar--Runge--Kutta methods as in \cite{kopecz_comparison_2019}.
+We will use the same notation also later in this article.}
+
+Using the Cash--Karp RK5 method \cite{cash1990variable} and $\dt = 0.005$ to
+solve \eqref{eq:Reaction}, the approximated solution contains negative values.
+This causes qualitatively wrong solutions to the problem.
 In Figure\,\ref{fig:exampleI} the obtained results are plotted with dashed lines.
@@
 In order to avoid issues that might be caused by poor weights, we can
-additionally impose either or both of the following strategies:
-
-\begin{enumerate}
-    \item Select in advance a set of desirable weight vectors $b^1, b^2, \dots, b^K$
+\linelabel{line:strategies}\revA{additionally use either or both of the following ideas:
+
+\begin{itemize}
+    \item Convex adaptation:
+      Select in advance a set of desirable weight vectors $b^1, b^2, \dots, b^K$
       corresponding to known good methods,
       and restrict the choice of $\bt$ to convex combinations of this set.
-    \item Require that the perturbation $\|\tilde u - u\|$ is small and reject the step if it is not.
-\end{enumerate}
-
-We discuss the first strategy here; the second is deferred to section\,\ref{sec:error}.
+    \item Stepsize control:
+      Require that the perturbation $\|\tilde u - u\|$ is small and reject
+      the step if it is not.
+\end{itemize}
+
+We discuss the first idea here; the second is deferred to section\,\ref{sec:error}.
 Ideally every element of the set of potential weight vectors would correspond to
@@
 and conditionally for forward Euler).
-We can formulate an LP using the first strategy (restriction to a convex set)
-as follows.  Let $B$ denote the matrix with columns $b^1, b^2, \dots, b^K$ and let
-$g\in \R^K$.  The LP is then as follows:
+We can formulate an LP using the approach of convex adaptation as follows.
+Let $B$ denote the matrix with columns $b^1, b^2, \dots, b^K$ and let
+$g\in \R^K$.  The LP is then as follows:}
 
@@
 An example is the reaction equation solved in Section~\ref{sec:example_reac}.
-Since the positivity of the stage values is not ensured, the right hand side has to be defined for negative values. If this is not the case, the right hand side has to be adapted accordingly.
+\linelabel{line:RHS-negative-values}\revA{%
+Since the positivity of the stage values is not ensured, the RHS has to be defined
+for negative values. If this is not the case, it to be adapted accordingly.
+For example, an RHS involving a square root can throw an error (forcing the
+numerical algorithm to terminate) or return ``not a number'' (\texttt{NaN})
+when called with negative values, depending on the implementation.
+In both cases, further (useful) computations are not possible. A possible fix
+allowing our mechanism to continue the integration can be to restrict the arguments
+to the square root to non-negative values by clipping them accordingly, e.g.\
+to replace \texttt{sqrt(u)} by \texttt{sqrt(max(u, 0))} or something similar.}
 

### 2021-01-31 nusslein_vsrp/latex/response_to_reviewers.tex

@@
 
-    \revA{TODO} %TODO
+    We added an explanation why we use this seemingly weird notation of the
+    equations [lines~{line:prod-dest}ff.].
 
@@
 
-    \revA{TODO} %TODO
+    We modified this part to be more clear [lines~\ref{line:strategies}ff.].
   \end{enumerate}
@@
 
-    \revA{TODO} %TODO
+    We tried to clarify this part by giving an example and adding further
+    explanations [lines~\ref{line:RHS-negative-values}ff.].
 

### 2021-02-10 nusslein_vsrp/latex/response_to_reviewers.tex
Updated response
Added some notes regardion the plots. 
- Missing \tilde
- Changes in stability region and the plot settings

@@
 \usepackage[USenglish]{babel}
-\usepackage[utf8]{luainputenc}
+\usepackage[utf8]{inputenc}
 \usepackage[a4paper]{geometry}
@@
 
-    \revA{TODO} %TODO
+    We added the tilde to the plot.
   \end{enumerate}
@@
 
-    \revA{TODO} %TODO
+    \revA{Review the following:} %TODO
+    We found an additional issue: The step size used for this plot was not $\Delta t = 0.015$ as for Figure 4a.
+We corrected the step size.
+Therefore the shape of the plotted stability regions changed.
+	To indicate the value of $\| \tilde b - b \|$ for the different stability regions the boundaries were colored using a colormap.  
+	We removed the filling pattern to improve clarity.
 
@@
 
-    \revA{TODO} %TODO
+    We added the tilde to the plot.
 

### 2021-02-28 nusslein_vsrp/latex/response_to_reviewers.tex
update some TODOs in the response

@@
     We added an explanation why we use this seemingly weird notation of the
-    equations [lines~{line:prod-dest}ff.].
+    equations [lines~\ref{line:prod-dest}ff.].
 
@@
 
-    \revA{Review the following:} %TODO
+    \revA{TODO: Review the following:} %TODO
     We found an additional issue: The step size used for this plot was not $\Delta t = 0.015$ as for Figure 4a.
@@
 
-    \revA{TODO} %TODO
+    \revA{TODO: Stephan wants to run some numerical experiments} %TODO
 

### 2021-03-01 nusslein_vsrp/latex/overview.tex
fix missing values
Added some comment on the cost of solving the LP for the non stiff case

@@
 We use the Dormand--Prince RK5 method and adapt the weights using the free adaptation.
-In Figure\,\ref{fig:sol_AdDe} the results for $\Delta t = 0.015$ are plotted for different \linelabel{line:time-vs-dt}\revA{of time $t$}.
+In Figure\,\ref{fig:sol_AdDe} the results for $\Delta t = 0.015$ are plotted for different \linelabel{line:time-vs-dt}\revA{values of time $t$}.
 We can see that the solution approaches an exponential function with $t \rightarrow \infty$.
@@
 Even though the order is reduced, most errors are still close to the error of the unaltered method.
+\linelabel{line:cost}\revA{
+In this example the cost of computing the stage values is small. Additionally a small reduction of $\dt$ is enough to obtain positive results.  
+If we compare the computation time we can see that using a smaller $\dt$ is faster than the adaption. 
+Therefore solving the LP is no economical alternative to reducing the step size.}
 

### 2021-03-01 nusslein_vsrp/latex/overview.tex
Rephrase explanation about efficiency.

--- 2021-01-30 nusslein_vsrp/latex/overview.tex
@@
 Even though the order is reduced, most errors are still close to the error of the unaltered method.
+
+\linelabel{line:cost}\revA{
+It is natural to ask whether the adaptation of the weights through the algorithm proposed
+here is more efficient than simply using a smaller step size.  For this problem that can
+be discretized explicitly with a right-hand-side that is relatively cheap to evaluate,
+using a smaller step size is
+generally more efficient, at least with the current un-optimized implementation of the LP
+solution.  For problems where implicit time integration is necessary, or with a more computationally
+expensive derivative evaluation, the LP approach may be more efficient, especially if
+the implementation were optimized.  This kind of optimization is the subject of future work.
+}
 

### 2021-03-01 nusslein_vsrp/latex/response_to_reviewers.tex
Added some comment on the cost of solving the LP for the non stiff case
Rephrase explanation about efficiency.
Rephrase some of the revisions.

--- 2021-01-31 nusslein_vsrp/latex/response_to_reviewers.tex
@@
 
-we are very grateful for the thorough review and helpful comments on our manuscript
+We are very grateful for the thorough review and helpful comments on our manuscript
 ``Positivity-Preserving Adaptive Runge--Kutta Methods''.
 We revised the manuscript in accordance with the reviewer's suggestions.
-Please find below a point-by-point description of these changes and reply to
+Please find below a point-by-point description of these changes and our reply to
 the reviewer's comments.
@@
 
-    We added an explanation why we use this seemingly weird notation of the
-    equations [lines~{line:prod-dest}ff.].
+    We added an explanation of we use this unusual form of the
+    equations [lines~\ref{line:prod-dest}ff.].
 
@@
 
-    \revA{TODO} %TODO
+    To indicate the value of $\| \tilde b - b \|$ for the different stability
+    regions, we now color the boundaries using a colormap and provide a colorbar.
+	We also removed the filling pattern to improve clarity.
+
+    Additionally, we discovered that the step size used for this plot was not
+    the correct value ($\Delta t = 0.015$) corresponding to Figure 4a. We
+    corrected the step size. Because of this,
+    the shape of the plotted stability regions changed in the revision.
 
@@
 
-    \revA{TODO} %TODO
+    We investigated this and added a remark at the end of Section 5.1, explaining
+    that the LP approach is more expensive in this case but would hopefully be
+    more efficient for certain more challenging problems.  [line~\ref{line:cost}]
 

### 2021-03-01 nusslein_vsrp/latex/overview.tex
Rephrase some of the revisions.

@@
 \linelabel{line:RHS-negative-values}\revA{%
-Since the positivity of the stage values is not ensured, the RHS has to be defined
-for negative values. If this is not the case, it to be adapted accordingly.
-For example, an RHS involving a square root can throw an error (forcing the
-numerical algorithm to terminate) or return ``not a number'' (\texttt{NaN})
-when called with negative values, depending on the implementation.
-In both cases, further (useful) computations are not possible. A possible fix
-allowing our mechanism to continue the integration can be to restrict the arguments
-to the square root to non-negative values by clipping them accordingly, e.g.\
-to replace \texttt{sqrt(u)} by \texttt{sqrt(max(u, 0))} or something similar.}
+Since the stage values are not guaranteed to be positive, the RHS must be
+defined also
+for negative values. If there is not a natural definition for negative values,
+one can instead extend the function in a smooth way or simply replace negative
+stage values by zero,  e.g.\ replace \texttt{sqrt(u)} by \texttt{sqrt(max(u, 0))}.}
 

### 2021-03-04 nusslein_vsrp/latex/response_to_reviewers.tex
Mention efficiency for stiff problem.

@@
     We investigated this and added a remark at the end of Section 5.1, explaining
-    that the LP approach is more expensive in this case but would hopefully be
-    more efficient for certain more challenging problems.  [line~\ref{line:cost}]
+    that the LP approach is more expensive for the problem in Section 5.1 but more
+    economical for the problem in Section 5.2.  One should not extrapolate too
+    much from these results, but our approach would be even
+    more efficient for more challenging problems where evaluation of $f$ or
+    solution of the stage equations is more costly.  [line~\ref{line:cost}]
 

### 2021-03-04 nusslein_vsrp/latex/overview.tex
update references
Mention efficiency for stiff problem.

@@
 generally more efficient, at least with the current un-optimized implementation of the LP
-solution.  For problems where implicit time integration is necessary, or with a more computationally
-expensive derivative evaluation, the LP approach may be more efficient, especially if
-the implementation were optimized.  This kind of optimization is the subject of future work.
+solution.  For the problem considered in the next section, where an implicit
+integrator is used, it is more efficient to maintain positivity with our proposed
+approach instead of reducing the step size.
+Adaptation of the weights could be made even more efficient with
+an optimized implementation of the LP setup and solve; this is the subject of future work.
 }
